---
title: "GP DO Regimes"
author: "CLB"
format: html
editor: visual
---

```{r}
#| include: false
library(data.table)
library(tidyverse, warn.conflicts = F)
options(readr.show_col_types = F)
library(broom)
library(googlesheets4)
library(ggthemes)
library(ggpmisc)
library(NatParksPalettes)
library(ggpubr)
library(patchwork)
library(gridExtra)
library(plotly)
library(here)
library(styler)
library(streamMetabolizer)
library(rstan)
options(mc.cores = parallel::detectCores())
library(dataRetrieval)
library(scales)
library(knitr)
library(lubridate)
library(zoo)
library(parallel)
library(doParallel)
library(foreach)
library(nasapower)
library(imputeTS)
library(rgbif)
library(moments)
library(inflection)
library(corrplot)
library(factoextra)
library(tidymodels)
library(bestNormalize)
#library(rstatix)
library(emmeans)
#library(infer)
library(cluster)
library(mgcv)  # For GAMs
source(here("Code/priceDrying.R"))
source(here("Code/Functions/exoPipeline.R"))
```

# Site Locations

```{r}
siteInfo <- tibble(Site = c("KMZ", "SHN", "YMR", "OKA", "GBJ", "DRC", "JDR", "TAL", "WHR", "PRF"),
                   Lat = c(39.092353, 39.115220, 37.56452, 34.437642, 42.78483889, 43.7184, 43.1226, 33.7621789, 32.98463, 34.968669),
                   Long = c(-96.587238, -96.558380, -96.49102, -96.652997, -112.44484722, -116.137, -116.776, -85.595521, -88.012269, -86.165439))
	
siteElev <-  elevation(latitude = siteInfo$Lat, longitude = siteInfo$Long, username = "conbro") %>% 
  select(Elev = elevation_geonames)
siteInfo$Elev_m <- siteElev$Elev
```

# Rating Curves

```{r}
# GP Rating Curve
gpRC <- read_sheet(ss = "1DkF2Ugr6tEgnZHOGJ_UY2kW9JWH4ce8xbyzSyyA6-CI", col_types = "cDcddddddddcccc") %>%
  filter(Quality %in% c("Maybe", "Good") &
           SiteName %in% c("SFM01", "ENM01", "SHM01")) %>%
  mutate(dateTime = ymd_hms(paste(Date, InjectionTime, sep = ":"))) %>% 
    select(Site = SiteName, dateTime, NaCl_g, UpstreamWidth_m, DownstreamWidth_m, ReachLength_m, Q_Ls, SensorStage_ft,
    ManualStage_ft, Offset, Quality) %>% 
  mutate(Site = case_when(Site == "ENM01" ~ "YMR",
                          Site == "SFM01" ~ "KMZ",
                          Site == "SHM01" ~ "SHN"))

# PT vs Stage
ptStage <- read_sheet(ss = "1DkF2Ugr6tEgnZHOGJ_UY2kW9JWH4ce8xbyzSyyA6-CI",
                   sheet = "PT vs Stage",
                   col_types = "cDccdd") %>% 
    mutate(dateTime = ymd_hms(paste(Date, Time, sep = ":"))) %>% 
  select(-VisitType, -Date, -Time) %>% 
  mutate(Site = case_when(Site == "ENM01" ~ "YMR",
                          Site == "SFM01" ~ "KMZ",
                          Site == "SHM01" ~ "SHN")) %>% 
  bind_rows(gpRC %>% select(Site, dateTime, SensorStage_ft, ManualStage_ft))

# YMR manual stage from sensor stage
ymrStage <- ptStage %>%
  filter(Site %in% c("YMR")) %>%
  lm(ManualStage_ft ~ SensorStage_ft, data = .) %>% 
  coef()

# SHN sensor to manual stage
shnStage <- ptStage %>%
  filter(Site %in% c("SHN")) %>%
  lm(SensorStage_ft ~ ManualStage_ft, data = .) %>% 
  coef()

# Apply the coefficients to calculate manual stage for YMR
gpRC <- gpRC %>%
  mutate(ManualStage_ft = case_when(
    Site == "YMR" & is.na(ManualStage_ft) ~ ymrStage["(Intercept)"] + ymrStage["SensorStage_ft"] * SensorStage_ft,
    TRUE ~ ManualStage_ft
  ))

```

```{r}

ptStage %>%
  ggplot(aes(ManualStage_ft, SensorStage_ft)) +
  geom_point() +
  labs(x = "Manual Stage (ft)",
       y = "Sensor Stage (ft)")+
  geom_smooth(method='lm') +
  stat_regline_equation(aes(label = ..rr.label..))+
  theme_few() +
  facet_wrap(~Site, ncol = 1, scales = "free")

gpRC %>%
  filter(Q_Ls > 0) %>% 
  ggplot(aes(SensorStage_ft, Q_Ls)) +
  geom_point() +
  labs(x = "Sensor Stage (ft)",
       y = "Discharge (L/s)")+
  geom_smooth(method='lm') +
  stat_regline_equation(aes(label = ..rr.label..))+
  theme_few() +
  facet_wrap(~Site, ncol = 1, scales = "free")

gpRC %>%
  drop_na(ManualStage_ft) %>% 
  filter(Q_Ls > 0) %>% 
  ggplot(aes(ManualStage_ft, Q_Ls)) +
  geom_point() +
  labs(x = "Manual Stage (ft)",
       y = "Discharge (L/s)")+
  geom_smooth(method='lm') +
  stat_regline_equation(aes(label = ..rr.label..))+
  theme_few() +
  facet_wrap(~Site, ncol = 1, scales = "free")

gpRC %>%
  drop_na(ManualStage_ft) %>% 
  filter(Q_Ls > 0) %>% 
  ggplot(aes(exp(log(ManualStage_ft)), exp(log1p(Q_Ls)))) +
  geom_point() +
  labs(x = "log(Manual Stage (ft))",
       y = "log(Discharge (L/s))")+
  geom_smooth(method='lm') +
  stat_regline_equation(aes(label = ..rr.label..))+
  theme_few() +
  facet_wrap(~Site, ncol = 1, scales = "free")

gpRC %>%
  drop_na(ManualStage_ft) %>% 
  filter(Q_Ls > 0) %>% 
  ggplot(aes(exp(log(ManualStage_ft)), Q_Ls)) +
  geom_point() +
  labs(x = "log(Manual Stage (ft))",
       y = "Discharge (L/s)")+
  geom_smooth(method='lm') +
  stat_regline_equation(aes(label = ..rr.label..))+
  theme_few() +
  facet_wrap(~Site, ncol = 1, scales = "free")

gpRC %>%
  drop_na(ManualStage_ft) %>% 
  filter(Q_Ls > 0) %>% 
  ggplot(aes(ManualStage_ft, exp(log(Q_Ls)))) +
  geom_point() +
  labs(x = "Manual Stage (ft)",
       y = "log(Discharge (L/s))")+
  geom_smooth(method='lm') +
  stat_regline_equation(aes(label = ..rr.label..))+
  theme_few() +
  facet_wrap(~Site, ncol = 1, scales = "free")

```

## Calc Discharge Rating Curve for Each Site

```{r}
#KMZ
kmzRC <- gpRC %>%
  filter(Site %in% c("KMZ")) %>%
  drop_na(ManualStage_ft) %>%
  lm(log1p(Q_Ls) ~ log(ManualStage_ft), data = .) %>% 
  coef()

#SHN
shnRC <- gpRC %>% 
  filter(Site %in% c("SHN")) %>% 
  drop_na(SensorStage_ft) %>% 
  lm(Q_Ls ~ SensorStage_ft, data = .) %>%
  coef()

#YMR
ymrRC <- gpRC %>% 
  filter(Site %in% c("YMR")) %>% 
  drop_na(ManualStage_ft) %>% 
  lm(Q_Ls ~ ManualStage_ft, data = .) %>%
  coef()
```

# Get Sensor Data

```{r}
# Path to QAQC pipeline
qaqcPath <- here("Data/EXO_QAQC/01_data/04_postQaqc/")
# Site to load data
#Sites <- c("SHN", "KMZ", "YMR", "OKA")
Sites <- c("SHN", "KMZ", "YMR")


# Read data for each site and combine
rawGP <- rbindlist(lapply(Sites, read_QAQCdata), fill = TRUE) %>%
  filter(alphaQf == 0 & betaQfSd == 0) %>%
  select(Site, dateTime = timestamp, parameterName, value, wetQf) %>%
  filter(parameterName %in% c("Temp_EXO_C", "SpCond_uS_cm", "Conductivity_uS_cm", "ODO_mg_L", "SW_Temp_PT_C", "SW_Level_ft",  "Temp_MiniDOT_C", "ODO_MiniDOT_mg_L", "Temp_s-can_C")) %>% 
  pivot_wider(names_from = parameterName, values_from = value, values_fn = mean) %>%
  mutate(
    dateTime = parse_date_time(dateTime, orders = c("ymd HMS"), tz = "UTC"),
    ODO_mg_L = ifelse(ODO_mg_L == 0, NA, ODO_mg_L)
  ) %>% 
  filter(!(Site == "YMR" & dateTime < "2021-07-18")) %>% 
  filter(!(Site == "SHN" & dateTime < "2023-03-24")) %>% 
  filter(!(Site == "KMZ" & dateTime < "2021-06-10"))
  

#Get all sites to 10 minute timestep
rawGP <- rawGP %>% 
  group_by(Site) %>%
  complete(dateTime = seq(floor_date(min(dateTime), "10 min"), floor_date(max(dateTime), "10 min"), by = "10 min")) %>%
  arrange(dateTime) %>%
  mutate(across(c(Temp_EXO_C:ODO_MiniDOT_mg_L), ~ zoo::na.approx(.x, na.rm = F, maxgap = 10))) %>%
  filter(minute(dateTime) %% 10 == 0) %>%
  ungroup() %>% 
  mutate(DO.obs = coalesce(ODO_mg_L, ODO_MiniDOT_mg_L),
         temp.water = coalesce(Temp_EXO_C, `Temp_s-can_C`, SW_Temp_PT_C, Temp_MiniDOT_C),
         SW_Level_cm = SW_Level_ft * 30.47999902) %>% 
  select(-ODO_mg_L, -ODO_MiniDOT_mg_L, -Temp_EXO_C, -`Temp_s-can_C`, -Temp_MiniDOT_C)

#Calc SpCond if missing
condMissing <- rawGP %>%
  filter((!is.na(SpCond_uS_cm) & is.na(Conductivity_uS_cm)) | (is.na(SpCond_uS_cm) & !is.na(Conductivity_uS_cm)))

rawGP <- rawGP %>% 
  mutate(SpCondCalc = Conductivity_uS_cm / (1 + 0.02 * (temp.water - 25)),
         CondCalc = SpCond_uS_cm * (1 + 0.02 * (temp.water - 25)),
         SpCond = coalesce(SpCond_uS_cm, SpCondCalc),
         Cond = coalesce(Conductivity_uS_cm, CondCalc)) %>% 
  select(-SpCond_uS_cm, -Conductivity_uS_cm, -SpCondCalc, -CondCalc)

# Wet Dry Flagging
rawGP <- rawGP %>% 
      mutate(wetQf = case_when(
      SW_Level_cm < 1 ~ 1,
      SW_Level_cm < 2 & SpCond < 10 ~ 1,
      Site == "KMZ" & SW_Level_cm > 3 ~ 0,
      Site == "KMZ" & SpCond < 10 ~ 1,
      Site == "KMZ" & (month(dateTime) == 12) | (month(dateTime) == 1) | (month(dateTime) == 2) ~ 1,
      Site == "KMZ" & (year(dateTime) == "2022") & (month(dateTime) == 9) | (month(dateTime) == 10) | (month(dateTime) == 11) | (month(dateTime) == 12) ~ 1,
      is.na(SW_Level_cm) & is.na(SpCond) & lag(wetQf, default = 0) == 1 ~ 1,
      is.na(SW_Level_cm) & SpCond == 0 & lag(wetQf, default = 0) == 1 ~ 1,
      Site == "YMR" & SW_Level_cm < 1.82 & SpCond < 300 ~ 1,
      Site == "YMR" & (year(dateTime) == "2023") & (month(dateTime) == 8) | (month(dateTime) == 9) | (month(dateTime) == 10) | (month(dateTime) == 11) ~ 1,
      Site == "SHN" & is.na(SpCond) | SpCond == 0 & SW_Level_cm < 3.1 & lag(wetQf, default = 1) == 1 ~ 1,
      Site == "SHN" & SW_Level_cm < 1 & temp.water < 0 ~ 1,
      Site == "SHN" & SW_Level_cm < 25 ~ 1,
      Site == "KMZ" & SW_Level_cm < 10 ~ 1,
      TRUE ~ 0
    ))

rawGP <- rawGP %>% 
  filter(!wetQf == 1) %>% 
  group_by(Site) %>%
  complete(dateTime = seq(floor_date(min(dateTime), "10 min"), floor_date(max(dateTime), "10 min"), by = "10 min")) %>%
  arrange(dateTime) %>%
  mutate(across(c(SW_Temp_PT_C:Cond), ~ zoo::na.approx(.x, na.rm = F, maxgap = 10))) %>%
  filter(minute(dateTime) %% 10 == 0) %>%
  ungroup()
```

## Calc Q from Sensor Data

```{r}
gpDat <- rawGP %>%
  mutate(
    Q_Ls = case_when(
      Site == "KMZ" ~ kmzRC["(Intercept)"] + kmzRC["log(ManualStage_ft)"] * SW_Level_ft,
      Site == "SHN" ~ shnRC["(Intercept)"] + shnRC["SensorStage_ft"] * SW_Level_ft,
      Site == "YMR" ~ ymrRC["(Intercept)"] + ymrRC["ManualStage_ft"] * SW_Level_ft, 
      TRUE ~ NA_real_))
```

## Check Out Data

```{r}
gpDatLong <- gpDat %>%
  select(-SW_Temp_PT_C, -SW_Level_ft) %>% 
  pivot_longer(cols = c(Q_Ls, SW_Level_cm, DO.obs, Cond, SpCond), names_to = "parameter", values_to = "value")

# Create plots for each site
gpDatLong %>%
  group_by(Site) %>%
  do(plot = ggplot(data = ., aes(x = dateTime, y = value, color = parameter)) +
       geom_line() +
       labs(title = unique(.$Site), x = "DateTime", y = "Value") +
       theme_few() +
        scale_x_datetime(labels = date_format("%b '%y"), breaks = "2 months") +
       scale_color_manual(values = c("Q_Ls" = "#0084a8", "SW_Level_cm" = "black", "DO.obs" = "#5d782e", "Cond" = "black", "SpCond" = "black"))+
     facet_wrap(~parameter, ncol = 1, scales = "free_y")+
       theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")) -> plt
# Print plots to the console
for(i in 1:length(plt$Site)) {
  print(plt$plot[[i]])
  datPlt <- plt$plot[[i]]
#  ggsave(paste0(plt$Site[[i]], ".png"), datPlt, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")
}
```

# Download NASA Data

```{r}
selSites <- siteInfo %>% 
  filter(Site %in% unique(gpDat$Site))

#Download barometric presssure from NASAPower for each site
nasaDat <- list()
for (i in 1:nrow(selSites)) {
  cat(paste("Downloading data for", selSites$Site[i], "\n"))
  tempDat <- gpDat %>% 
    filter(Site == selSites$Site[i])
  sDate <- as_date(min(tempDat$dateTime, na.rm = T))
  eDate <- as_date(max(tempDat$dateTime, na.rm = T))
  
  nasaDat[[i]] <- get_power(
    community = "sb",
    pars = c("PSC", "ALLSKY_SFC_SW_DWN", "PRECTOTCORR"),
    temporal_api = "hourly",
    lonlat = c(selSites$Long[i], selSites$Lat[i]),
    dates = c(sDate, eDate),
    site_elevation = selSites$Elev_m[i],
    time_standard = "UTC"
  )
  nasaDat[[i]]$Site <- selSites$Site[i]
  cat(paste("Finished downloading data for", selSites$Site[i], "\n"))
  #Break loop on last iteration
  if (i == nrow(selSites)) {
    break
  }
}

nasaDat <- do.call(rbind, nasaDat)
nasaDat <- nasaDat %>% 
  select(Site, YYYYMMDD, HR, PSC, ALLSKY_SFC_SW_DWN, PPT_mm = PRECTOTCORR) %>% 
  mutate(Hour = str_pad(HR, 2, side = "left", pad = 0)) %>% 
  unite(dateTime, YYYYMMDD, Hour, sep = " ") %>% 
  mutate(dateTime = ymd_hms(paste0(dateTime, ":00:00")),
         PSC = PSC*10, #Convert kilopascals to millibar
         PAR.obs = convert_SW_to_PAR(ALLSKY_SFC_SW_DWN)) %>% 
  select(-HR, -ALLSKY_SFC_SW_DWN)

gpDat <- gpDat %>% 
  full_join(nasaDat, by = c("Site", "dateTime")) %>% 
  left_join(selSites, by = "Site") %>% 
  group_by(Site) %>% 
  mutate(solar.time = convert_UTC_to_solartime(dateTime, longitude = Long, time.type = "mean solar"),
         isDaytime = calc_is_daytime(solar.time, Lat),
         PSC = na.approx(PSC, maxgap = 6, na.rm = F),
         PAR.obs = na.approx(PAR.obs, maxgap = 6, na.rm = F),
         PPT_mm = na.approx(PPT_mm, maxgap = 6, na.rm = F),
         DO.sat = calc_DO_sat(temp.water = temp.water, PSC, model = "garcia-benson"),
         DO.pctsat = 100 * (DO.obs/ DO.sat),
         WaterYear = calcWaterYear(dateTime))
```

## Model Missing Light Data

```{r}
# Set the number of cores to use
numCores <- detectCores() - 1  # Use one less than the number of available cores
cl <- makeCluster(numCores)
registerDoParallel(cl)

# Parallel processing with foreach
gpDat <- foreach(i = unique(gpDat$Site), .combine = rbind, .packages = c("dplyr", "data.table", "streamMetabolizer")) %dopar% {
  tempDat <- gpDat %>% filter(Site == i)
  # Get max observed light value
  maxLight <- max(tempDat$PAR.obs, na.rm = TRUE)
  tempDat$PAR.mod <- calc_light(tempDat$solar.time, latitude = tempDat$Lat, longitude =  tempDat$Long, max.PAR = maxLight)
  # Merge calc and obs light
  tempDat$light <- fcoalesce(tempDat$PAR.obs, tempDat$PAR.mod)
  tempDat
}

# Stop the cluster
stopCluster(cl)
rm(cl)

```

### PPT Exceedence Probability

```{r}
#Download barometric presssure from NASAPower for each site
pptDat <- list()
for (i in 1:nrow(selSites)) {
  cat(paste("Downloading data for", selSites$Site[i], "\n"))

  
  pptDat[[i]] <- get_power(
    community = "sb",
    pars = c("PRECTOTCORR"),
    temporal_api = "daily",
    lonlat = c(selSites$Long[i], selSites$Lat[i]),
    dates = c("1981-01-01", "2024-07-01"),
    site_elevation = selSites$Elev_m[i],
    time_standard = "UTC"
  )
  pptDat[[i]]$Site <- selSites$Site[i]
  cat(paste("Finished downloading data for", selSites$Site[i], "\n"))
  #Break loop on last iteration
  if (i == nrow(selSites)) {
    break
  }
}

pptDat <- do.call(rbind, pptDat)
pptDat <- pptDat %>% 
  select(Site, Date = YYYYMMDD,  PPT_mm = PRECTOTCORR) %>% 
  mutate(Year = year(Date),
         Period = case_when(
           Year < 2021 ~ "1981-2020",
           Year == 2021 ~ "2021",
           Year == 2022 ~ "2022",
           Year == 2023 ~ "2023",
           Year == 2024 ~ "2024")) %>% 
  group_by(Period, Site) %>%
  arrange(desc(PPT_mm)) %>%
  mutate(
    Rank = row_number(),
    PPT_ExceProb = (Rank / (n() + 1)) * 100)

  

pptDat %>% 
  ggplot(aes(x = PPT_ExceProb, y = PPT_mm, color = Period)) +
  geom_line() +
  scale_y_log10() +
  scale_color_manual(values = c("1981-2020" = "#000000", "2021" = "#1B9E77", "2022" = "#D95F02", "2023" = "#7570B3", "2024" = "#E7298A")) +
  labs(x = "Exceedance Probability (%)",
       y = "Precipitation (mm/d)") +
  theme_few()+
  facet_wrap(~Site)

pptDat %>% 
  filter(Year >= "2021") %>% 
  ggplot(aes(Date, PPT_mm)) +
  geom_col(color = "#0084a8") +
  ylab("Precipitation (mm)") +
  theme_few() +
  scale_y_reverse()+
  scale_x_date(labels = date_format("%b '%y")) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~Site, ncol = 1, scales = "free")

pptMonthAvg <- pptDat %>%
  mutate(Month = month(Date)) %>% 
  group_by(Site, Period, Year, Month) %>%
  summarise(PPT_mm = mean(PPT_mm, na.rm = T))
pptMonthAvg %>% 
  ggplot(aes(x = as.factor(Period), y = PPT_mm, fill = Site)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Dark2", type = "div") +
  facet_wrap(~ Site, scales = "free") +
  scale_y_continuous(labels = label_number())+
  #scale_y_log10(labels = label_number()) +
  labs(title = "Monthly Average Precipitation",
       x = "",
       y = "PPT (mm)") +
  theme_few() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position = "none") -> pptPlt
pptPlt
ggsave("pptPlt.png", pptPlt, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")
```

### PPT Stats

```{r}
#get average for each year
pptYrAvg <- pptDat %>% 
  group_by(Site, Year) %>%
  summarise(PPT_mm = mean(PPT_mm, na.rm = T))

# Get long term average ppt
pptLongAvg <- pptDat %>% 
  group_by(Site, Period) %>%
  filter(Period == "1981-2020") %>% 
  summarise(longTermPPT = mean(PPT_mm, na.rm = T)) %>% 
  dplyr::select(-Period)

# Group by Site and perform the paired t-test
pptYrAvg <- pptYrAvg %>% 
  left_join(pptLongAvg)

#t-test for each site's annual average against its long-term average
pptTtest <- pptYrAvg %>%
  group_by(Site) %>%
  filter(Year > 2020) %>% 
  summarize(t_test = list(t.test(PPT_mm, longTermPPT, paired = TRUE)))

#t-test results
pptTtest <- pptTtest %>%
  mutate(
    p_value = sapply(t_test, function(x) x$p.value),
    statistic = sapply(t_test, function(x) x$statistic),
    conf_low = sapply(t_test, function(x) x$conf.int[1]),
    conf_high = sapply(t_test, function(x) x$conf.int[2])
  )

# Convert Year to a factor for ANOVA
pptYrAvg <- pptYrAvg %>% 
  filter(Year > "2020" & Year < "2024") %>% 
  mutate(Year = as.factor(Year))

# Perform two-way ANOVA
pptAnova <- aov(PPT_mm ~ Site + Year, data = pptYrAvg)
summary(pptAnova)
# Summary of the ANOVA
TukeyHSD(pptAnova, "Site")
TukeyHSD(pptAnova, "Year")

######
#get average for each year
pptYrAvg <- pptDat %>% 
  group_by(Site, Period) %>%
  summarise(PPT_mm = mean(PPT_mm, na.rm = T)) %>% 
  mutate(Period = factor(Period))

pptAnova <- aov(PPT_mm ~ Site + Period, data = pptYrAvg)
summary(pptAnova)
TukeyHSD(pptAnova, "Site")
TukeyHSD(pptAnova, "Period")

```

```{r}
gpDat %>%
  ggplot(aes(dateTime, DO.pctsat)) +
  geom_line(color = "#5d782e") + 
  ylab("DO % Sat") +
  theme_few() +
  scale_x_datetime(labels = date_format("%b '%y"), breaks = "6 months") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~Site, ncol = 2, scales = "free")

gpDat %>%
  ggplot(aes(dateTime, light)) +
  geom_line(color = "darkgoldenrod1") +
  ylab(expression(PAR~(mu*"mol m"^-2*"s"^-1))) +
  theme_few() +
  scale_x_datetime(labels = date_format("%b '%y"), breaks = "6 months") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~Site, ncol = 2, scales = "free")

gpDat %>%
  ggplot(aes(dateTime, Q_Ls)) +
  geom_line(color = "#0084a8") +
  ylab("Discharge (L/s)") +
  theme_few() +
  scale_x_datetime(labels = date_format("%b '%y"), breaks = "6 months") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~Site, ncol = 2, scales = "free")
```

## Check \# Days

```{r}
gpDatDays <- gpDat %>% 
  mutate(Date = as.Date(dateTime),
         Year = year(dateTime)) %>% 
  group_by(Site, Year, Date) %>%
  summarise(completeDay = all(!is.na(wetQf)) & 
                            all(!is.na(temp.water)) & 
                            all(!is.na(SpCond)) & 
                            all(!is.na(Cond)) & 
                            all(!is.na(DO.obs)) & 
                            all(!is.na(SW_Temp_PT_C)) & 
                            all(!is.na(SW_Level_ft)),
            .groups = 'drop') %>% 
  group_by(Site, Year) %>%
  summarise(complete_days_count = sum(completeDay),
            .groups = 'drop')


gpDatDays <- gpDat %>% 
  mutate(Date = as.Date(dateTime),
         Year = year(dateTime)) %>% 
  group_by(Site, Year, Date) %>%
  summarise(total_obs = n(),
            across(wetQf:Cond, ~ sum(!is.na(.))),
            .groups = 'drop') %>% 
  mutate(at_least_50_percent = (wetQf / total_obs >= 0.25) &
                                (temp.water / total_obs >= 0.25) &
                                (SpCond / total_obs >= 0.25) &
                                (Cond / total_obs >= 0.25) &
                                (DO.obs / total_obs >= 0.25) &
                                (SW_Temp_PT_C / total_obs >= 0.25) &
                                (SW_Level_ft / total_obs >= 0.25)) %>% 
  group_by(Site, Year) %>%
  summarise(days_with_50_percent_obs = sum(at_least_50_percent),
            .groups = 'drop')

```

## Calc Dry vs. Wet Days

```{r}
aimsWetDry <- gpDat %>% 
  select(Site, dateTime, wetQf, temp.water, SpCond, Cond, DO.obs, SW_Level_cm, SW_Level_ft) %>% 
  mutate(Date = as.Date(dateTime)) %>%
  group_by(Site, Date) %>%
  summarize(across(where(is.numeric), \(x) mean(x, na.rm = TRUE)), .groups = "drop") %>% 
  mutate(wetQf = if_else(is.na(wetQf) | wetQf != 0 | SW_Level_ft < 0.09, 1, wetQf))


# Read data for each site and combine
aimsWetDry <- rawGP %>% 
  select(Site, dateTime, wetQf, SW_Level_ft, SW_Temp_PT_C, temp.water, SpCond, Cond) %>% 
  mutate(Date = as.Date(dateTime)) %>%
  group_by(Site, Date) %>%
  summarize(across(where(is.numeric), \(x) mean(x, na.rm = TRUE)), .groups = "drop") %>% 
  mutate(temp.water = na.approx(temp.water, na.rm = F, maxgap = 4),
         SpCond = na.approx(SpCond, na.rm = F, maxgap = 4),
         Cond = na.approx(Cond, na.rm = F, maxgap = 4),
         SW_Temp_PT_C = na.approx(SW_Temp_PT_C, na.rm = F),
         SW_Level_ft = na.approx(SW_Level_ft, na.rm = F),
         wetQf = replace_na(wetQf, 1)
         #wetQfInterp = na.approx(wetQf, na.rm = F),
         ) %>%
  mutate(wetQf2 = if_else(wetQf > 0 | SW_Level_ft < 0.05 | SW_Temp_PT_C < 5, 1, 0, wetQf)) %>% 
  # Set wetQf to 1 if (SW_Level_ft < 0.02 and Conductivity_uS_cm < 5 or NA) or SW_Temp_PT_C < 0 or (SW_Level_ft < 0.02 & SW_Temp_PT_C < 0)
  mutate(wetQf3 = if_else((SW_Level_ft < 0.02 & (is.na(Cond) | Cond < 5)) | SW_Temp_PT_C < 0 | (SW_Level_ft < 0.02 & SW_Temp_PT_C < 0), 1, wetQf2))

aimsWetDryDays <- aimsWetDry %>% 
  mutate(Year = year(Date)) %>%
  filter(Year < "2024") %>% 
  group_by(Site, Year) %>%
  summarize(
    DryDays = sum(wetQf2 == 1),
    WetDays = sum(wetQf2 == 0),
    TotalDays = n(),
    PercWetPeriod = round((WetDays/TotalDays)*100),
    PercDryPeriod = round((DryDays/TotalDays)*100),
    .groups = 'drop'  # To ungroup after summarizing
  )

```

## Get Season

```{r}
# Function to determine the season
calcSeason <- function(date) {
  m <- lubridate::month(date)
  d <- lubridate::day(date)
  if ((m == 12 && d >= 21) || (m %in% c(1, 2)) || (m == 3 && d < 21)) {
    return("Winter")
  } else if ((m == 3 && d >= 21) || (m %in% c(4, 5)) || (m == 6 && d < 21)) {
    return("Spring")
  } else if ((m == 6 && d >= 21) || (m %in% c(7, 8)) || (m == 9 && d < 21)) {
    return("Summer")
  } else if ((m == 9 && d >= 21) || (m %in% c(10, 11)) || (m == 12 && d < 21)) {
    return("Fall")
  }
  return(NA)
}

# Add season column
# gpDat <- gpDat %>%
#   mutate(date = as.POSIXct(dateTime, format = "%Y-%m-%d %H:%M:%S"),
#          season = sapply(date, calcSeason))

```

# Calculate DO Metrics

```{r}
# Daily metrics
dailyDO <- gpDat %>% 
  mutate(Date = floor_date(solar.time, "1 day")) %>%
  group_by(Site, Date) %>%
  summarize(
    minDO.obs = min(DO.obs, na.rm = T),
    maxDO.obs = max(DO.obs, na.rm = T),
    minDO.pctsat = min(DO.pctsat, na.rm = T),
    maxDO.pctsat = max(DO.pctsat, na.rm = T),
    avgDO.obs = mean(DO.obs, na.rm = T),
    avgDO.pctsat = mean(DO.pctsat, na.rm = T),
    DOSat.hyp = sum(DO.pctsat <= 50, na.rm = T),
    DO.hyp = sum(DO.obs <= 2, na.rm = T),
    totalDO = n()
    ) %>% 
  mutate(across(where(is.numeric), ~ ifelse(is.finite(.), ., NA))) %>% 
  mutate(across(where(is.numeric), ~ ifelse(if_any(everything(), is.na), NA, .))) %>% 
  mutate(across(minDO.obs:avgDO.pctsat, ~ ifelse(. == 0, NA, .))) %>% 
  mutate(DO.amp = maxDO.obs - minDO.obs,
         DOpctsat.amp = maxDO.pctsat - minDO.pctsat,
         probHyp = DO.hyp / totalDO,
         probHypSat = DOSat.hyp / totalDO,
         WaterYear = calcWaterYear(Date),
         season = sapply(Date, calcSeason),
         Year = year(Date))

dailyDO <- gpDat %>% 
  mutate(Date = floor_date(solar.time, "1 day")) %>%
  group_by(Site, Date) %>%
  summarize(
    minDO.obs = min(DO.obs, na.rm = TRUE),
    maxDO.obs = max(DO.obs, na.rm = TRUE),
    minDO.pctsat = min(DO.pctsat, na.rm = TRUE),
    maxDO.pctsat = max(DO.pctsat, na.rm = TRUE),
    avgDO.obs = mean(DO.obs, na.rm = TRUE),
    avgDO.pctsat = mean(DO.pctsat, na.rm = TRUE),
    sdDO.obs = sd(DO.obs, na.rm = TRUE),
    sdDO.pctsat = sd(DO.pctsat, na.rm = TRUE),
    DOSat.hyp = sum(DO.pctsat <= 50, na.rm = TRUE),
    DO.hyp = sum(DO.obs <= 2, na.rm = TRUE),
    totalDO = n()
  ) %>% 
  mutate(across(where(is.numeric), ~ ifelse(is.finite(.), ., NA))) %>% 
  mutate(across(where(is.numeric), ~ ifelse(if_any(everything(), is.na), NA, .))) %>% 
  mutate(across(minDO.obs:avgDO.pctsat, ~ ifelse(. == 0, NA, .))) %>% 
  mutate(DO.amp = maxDO.obs - minDO.obs,
         DOpctsat.amp = maxDO.pctsat - minDO.pctsat,
         probHyp = DO.hyp / totalDO,
         probHypSat = DOSat.hyp / totalDO,
         WaterYear = calcWaterYear(Date),
         season = sapply(Date, calcSeason),
         Year = year(Date),
         cvDO.obs = (sdDO.obs / avgDO.obs) * 100,
         cvDO.pctsat = (sdDO.pctsat / avgDO.pctsat) * 100) %>% 
  select(-sdDO.obs, -sdDO.pctsat) 

#nighttime hypoxia ratio
nightHypRat <- gpDat %>% 
  mutate(across(c("DO.obs", "DO.pctsat", "DO.sat"), ~ ifelse(. == 0, NA, .))) %>% 
  select(Site, dateTime, solar.time, DO.obs, DO.pctsat, DO.sat, isDaytime) %>% 
  mutate(Date = floor_date(solar.time, "1 day")) %>%
  group_by(Date, Site, isDaytime) %>% 
  summarise(DO.hyp.obs = sum(DO.obs <= 2, na.rm = TRUE),
            DO.hyp.sat = sum(DO.pctsat <= 50, na.rm = TRUE), .groups = 'drop') %>% 
  pivot_wider(names_from = isDaytime, 
              values_from = c(DO.hyp.obs, DO.hyp.sat), 
              names_prefix = "DO_hyp_", 
              names_sep = "_") %>% 
  rename(DayHyp_obs = DO.hyp.obs_DO_hyp_TRUE, 
         NightHyp_obs = DO.hyp.obs_DO_hyp_FALSE,
         DayHyp_sat = DO.hyp.sat_DO_hyp_TRUE, 
         NightHyp_sat = DO.hyp.sat_DO_hyp_FALSE) %>% 
  mutate(NHR_obs = NightHyp_obs / DayHyp_obs,
         NHR_sat = NightHyp_sat / DayHyp_sat,
         WaterYear = calcWaterYear(Date),
         Year = year(Date)) %>% 
  mutate(across(where(is.numeric), ~ ifelse(is.finite(.), ., NA))) %>% 
  select(Date, Site, NHR_obs, NHR_sat)


# Calculate probabilities for daytime
dayHypProb <- gpDat %>% 
  mutate(across(c("DO.obs", "DO.pctsat", "DO.sat"), ~ ifelse(. == 0, NA, .))) %>% 
  select(Site, dateTime, solar.time, DO.obs, DO.pctsat, DO.sat, isDaytime) %>% 
  mutate(Date = floor_date(solar.time, "1 day")) %>%
  filter(isDaytime) %>%
  group_by(Site, Date) %>%
  summarise(
    dayProbConc = mean(DO.obs < 2, na.rm = TRUE),
    dayProbSat = mean(DO.pctsat < 50, na.rm = TRUE),
    .groups = 'drop'
  )

# Calculate probabilities for nighttime
nightHypProb <- gpDat %>%
  mutate(across(c("DO.obs", "DO.pctsat", "DO.sat"), ~ ifelse(. == 0, NA, .))) %>% 
  select(Site, dateTime, solar.time, DO.obs, DO.pctsat, DO.sat, isDaytime) %>% 
  mutate(Date = floor_date(solar.time, "1 day")) %>%
  filter(!isDaytime) %>%
  group_by(Site, Date) %>%
  summarise(
    nightProbConc = mean(DO.obs < 2, na.rm = TRUE),
    nightProbSat = mean(DO.pctsat < 50, na.rm = TRUE),
    .groups = 'drop'
  )

# Join the daytime and nighttime probabilities
nightHypRat <- dayHypProb %>%
  inner_join(nightHypProb, by = c("Site", "Date")) %>% 
  mutate(NHR_obs = nightProbConc/dayProbConc,
         NHR_sat = nightProbSat/dayProbSat) %>% 
  mutate(across(where(is.numeric), ~ ifelse(is.finite(.), ., NA))) %>% 
  mutate(WaterYear = calcWaterYear(Date),
         Year = year(Date))




#Merge DO Metrics
doMet <- dailyDO %>% 
  full_join(nightHypRat) %>% 
  arrange(Site, Date, WaterYear)
```

## DO Duration

```{r}
# Calculate the exceedance values
doDur <- dailyDO %>%
  filter(!Site == "SHN") %>% 
  select(Site, WaterYear, avgDO.pctsat) %>% 
  group_by(Site, WaterYear) %>%
  filter(!is.na(avgDO.pctsat)) %>%
  mutate(exceedance = 100 * (rank(-avgDO.pctsat) / (length(avgDO.pctsat) + 1)))

# Calculate undersaturation
underSat <- doDur %>%
  group_by(Site, WaterYear) %>%
  summarise(
    modalUndersat = 100 - as.numeric(names(sort(table(avgDO.pctsat), decreasing = TRUE))[1]),
    meanUndersat = mean(avgDO.pctsat)
  )

doDur <- doDur %>%
  left_join(underSat)

doDur %>% 
  ggplot(aes(x = exceedance, y = avgDO.pctsat, color = factor(WaterYear))) +
  geom_line(linewidth = 1.5) +
  geom_hline(yintercept = 50, linetype = "dashed", alpha = 0.5)+
  #scale_y_log10() +
  scale_color_manual(values = c("2021" = "#1B9E77", "2022" = "#D95F02", "2023" = "#7570B3", "2024" = "#E7298A")) +
  labs(title = "",
       color = "Year",
       x = "Exceedance Probability (%)",
       y = expression(Daily ~ DO["%Sat"])) +
  theme_few()+
  theme(legend.position = "bottom", legend.title = element_blank(),
        text = element_text(size = 25),       # Increase overall text size
    axis.title = element_text(size = 25), # Increase axis titles size
    axis.text = element_text(size = 25),  # Increase axis text size
    strip.text = element_text(size = 25),  # Increase facet labels size
    legend.text = element_text(size = 25)
    )+
  facet_wrap(~ Site) -> doExce
doExce
ggsave("doExce.png", doExce, here("Figs/"), device = "png", width = 11, height = 6, units = "in", dpi = "retina")

doDur %>% 
  ggplot(aes(x = exceedance, y = avgDO.pctsat, color = factor(WaterYear))) +
  geom_line() +
  geom_hline(data = doDur, aes(yintercept = modalUndersat, color = factor(WaterYear), linetype = "Modal")) +
  geom_hline(data = doDur, aes(yintercept = meanUndersat, color = factor(WaterYear), linetype = "Mean")) +
  scale_color_manual(values = c("2021" = "#1B9E77", "2022" = "#D95F02", "2023" = "#7570B3", "2024" = "#E7298A")) +
  scale_linetype_manual(name = "Undersaturation", values = c("Modal" = "dashed", "Mean" = "dotted")) +
  labs(title = "",
       color = "Year",
       x = "Exceedance Probability (%)",
       y = expression(Daily ~ DO["%Sat"])) +
  theme_few() +
  facet_wrap(~ Site)

doDur %>% 
  filter(!Site == "SHN") %>% 
  ggplot(aes(x = exceedance, y = avgDO.pctsat, color = Site)) +
  geom_line() +
  geom_hline(yintercept = 50, linetype = "dashed", alpha = 0.5)+
  scale_color_manual(values = c("KMZ" = "#1B9E77", "YMR" = "#D95F02", "SHN" = "#7570B3")) +
  labs(title = "",
       color = "Site",
       x = "Exceedance Probability (%)",
       y = expression(Daily ~ DO["%Sat"])) +
  theme_few() +
  facet_wrap(~ WaterYear)-> doExceWY
doExceWY
ggsave("doExceWY.png", doExceWY, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")

doDurAllTime <- dailyDO %>%
  select(Site, WaterYear, avgDO.pctsat) %>% 
  group_by(Site) %>%
  filter(!is.na(avgDO.pctsat)) %>%
  mutate(exceedance = 100 * (rank(-avgDO.pctsat) / (length(avgDO.pctsat) + 1)))

doDurAllTime %>% 
  ggplot(aes(x = exceedance, y = avgDO.pctsat, color = Site)) +
  geom_line() +
  geom_hline(yintercept = 50, linetype = "dashed", alpha = 0.5)+
  scale_color_manual(values = c("KMZ" = "#1B9E77", "YMR" = "#D95F02", "SHN" = "#7570B3")) +
  labs(title = "",
       color = "Site",
       x = "Exceedance Probability (%)",
       y = expression(Daily ~ DO["%Sat"])) +
  theme_few() -> doExceAll
doExceAll
ggsave("doExceAll.png", doExceAll, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")

```

## DO Metric Figures

```{r}
gpDat %>%
  ggplot(aes(x = DO.pctsat, fill = isDaytime, after_stat(scaled))) +
  geom_density(alpha = 0.7) +
  scale_fill_manual(breaks = c("TRUE", "FALSE"),
                    labels = c("Daytime", "Nighttime"),
                    values = c("#FFA726", "#0D47A1")) +
  geom_vline(xintercept = 50, linetype = "dashed") +
  labs(x = expression(DO[sat]~"(%)" ), y = "Density") +
  theme_few() +
  theme(legend.title = element_blank(), legend.position = "bottom") +
  facet_grid(WaterYear~Site)

dailyDO %>%
  ggplot(aes(x = avgDO.obs, fill = Site, after_stat(scaled))) +
  geom_density(alpha = 0.5) +
  scale_fill_brewer(palette = "Dark2") +
  geom_vline(xintercept = 2, linetype = "dashed") +
  labs(x = "DO (mg/L)", y = "Density") +
  theme_few() +
  theme(legend.title = element_blank(), legend.position = "bottom")+
  facet_wrap(~WaterYear, ncol = 1, scale = "free_y") -> concDen
concDen
ggsave("concDen.png", concDen, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")

dailyDO %>%
  filter(!Site == "SHN") %>% 
  ggplot(aes(x = avgDO.pctsat, fill = Site, after_stat(scaled))) +
  geom_density(alpha = 0.8) +
  scale_fill_manual(values = c("KMZ" = "#127088", "YMR" = "#DE3C37"))+
  geom_vline(xintercept = 50, linetype = "dashed") +
  labs(x = expression(DO["%sat"]), y = "Density") +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 3)) +
  theme_few() +
  theme(legend.position = "bottom", legend.title = element_blank(),
        text = element_text(size = 25),       # Increase overall text size
    axis.title = element_text(size = 25), # Increase axis titles size
    axis.text = element_text(size = 25),  # Increase axis text size
    strip.text = element_text(size = 25),  # Increase facet labels size
    legend.text = element_text(size = 25)
    )+
  facet_wrap(~WaterYear, ncol = 1, scale = "free_y") -> satDen
satDen
ggsave("satDen.png", satDen, here("Figs/"), device = "png", width = 11, height = 6, units = "in", dpi = "retina")

dailyDO %>%
  mutate(WaterYear = as.factor(WaterYear)) %>%
  ggplot(aes(x = WaterYear, y = avgDO.pctsat, color = Site)) +  # Add color = Site
  geom_boxplot(outlier.color = NA) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.75), alpha = 0.5) +
  scale_fill_brewer(palette = "Dark2") +
  scale_color_brewer(palette = "Dark2") +  # Add this line
  geom_hline(yintercept = 50, linetype = "dashed") +
  labs(x = "Water Year", y = expression(Daily ~ DO["%Sat"])) +
  theme_few() +
  theme(legend.title = element_blank(), legend.position = "bottom") -> satBox
satBox
ggsave("satBox.png", satBox, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")


dailyDO %>%
  filter(!Site == "SHN") %>% 
  ggplot(aes(x = DOpctsat.amp, fill = Site, after_stat(scaled))) +
  geom_density(alpha = 0.5) +
  scale_fill_brewer(palette = "Dark2") +
  #geom_vline(xintercept = 50, linetype = "dashed") +
  labs(x = expression(Daily ~ DO["%Sat"] ~ "Amplitude"),
       y = "Density")+
  theme_few() +
  theme(legend.title = element_blank(), legend.position = "bottom")+
  facet_wrap(~WaterYear, ncol = 1, scale = "free_y") -> satAmpDen
satAmpDen
ggsave("satAmpDen.png", satAmpDen, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")

dailyDO %>%
  filter(!Site == "SHN") %>% 
  ggplot(aes(x = probHypSat, fill = Site, after_stat(scaled))) +
  geom_density(alpha = 0.5) +
  #xlim(-1.5, 1.5)+
  scale_fill_brewer(palette = "Dark2") +
  labs(x = expression(DO["%Sat"] ~ "Probability of Hypoxia"), y = "Density") +
  theme_few() +
  theme(legend.position = "bottom", legend.title = element_blank(),
        text = element_text(size = 25),       # Increase overall text size
        axis.title = element_text(size = 25), # Increase axis titles size
        axis.text = element_text(size = 25),  # Increase axis text size
        strip.text = element_text(size = 25)  # Increase facet labels size
    )+
  facet_wrap(~WaterYear, ncol = 1, scale = "free_y") -> probDen
probDen
ggsave("probDen.png", probDen, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")

doMet %>%
  ggplot(aes(Date, probHypSat, color = Site)) +
  geom_point(shape = 4, alpha = 0.8) +
  theme_few() +
  labs(y = expression(DO["%Sat"] ~ "Probability of Hypoxia"))+
  scale_x_datetime(labels = date_format("%b"), breaks = "1 months") +
  scale_color_brewer(palette = "Dark2")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_blank()) +
  facet_wrap(~WaterYear, ncol = 1, scales = "free") -> probHyp
probHyp
ggsave("probHyp.png", probHyp, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")


doMet %>%
  ggplot(aes(Date, DOpctsat.amp, color = Site)) +
  geom_point(shape = 4, alpha = 0.8) +
  theme_few() +
  labs(y = expression(Daily ~ DO["%Sat"] ~ "Amplitude"))+
  scale_x_datetime(labels = date_format("%b"), breaks = "1 months") +
  scale_color_brewer(palette = "Dark2")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_blank()) +
  facet_wrap(~WaterYear, ncol = 1, scales = "free") -> satAmpSiteYr
satAmpSiteYr
ggsave("satAmpSiteYr.png", satAmpSiteYr, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")

doMet %>%
  ggplot(aes(Date, NHR_obs, color = Site)) +
  geom_point(shape = 4, alpha = 0.8) +
  theme_few() +
  labs(y = expression(DO["mg/L"] ~ "Night Hypoxia Ratio"))+
  scale_x_datetime(labels = date_format("%b"), breaks = "1 months") +
  scale_color_brewer(palette = "Dark2")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_blank()) +
  facet_wrap(~WaterYear, ncol = 1, scales = "free") -> nhrConc
nhrConc
ggsave("nhrConc.png", nhrConc, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")


doMet %>%
  ggplot(aes(Date, NHR_sat, color = Site)) +
  geom_point(shape = 4, alpha = 0.8) +
  theme_few() +
  scale_x_datetime(labels = date_format("%b"), breaks = "1 months") +
  scale_color_brewer(palette = "Dark2")+
  labs(y = expression(DO["%Sat"] ~ "Night Hypoxia Ratio"))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_blank()) +
  facet_wrap(~WaterYear, ncol = 1, scales = "free") -> nhrSat
nhrSat
ggsave("nhrSat.png", nhrSat, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")

doMet %>%
  ggplot(aes(x = cvDO.pctsat, fill = Site, after_stat(scaled))) +
  geom_density(alpha = 0.5) +
  #xlim(-1.5, 1.5)+
  scale_fill_brewer(palette = "Dark2") +
  labs(x = expression(DO["%Sat"] ~ "Probability of Hypoxia"), y = "Density") +
  theme_few() +
  theme(legend.title = element_blank(), legend.position = "bottom")+
  facet_wrap(~WaterYear, ncol = 1, scale = "free_y") -> cvSat
cvSat
```

## Daily DO Sat Stats

```{r}
doPctSat <- dailyDO %>% 
  filter(!Site == "SHN") %>% 
  dplyr::select(Site, Date, WaterYear, avgDO.pctsat) %>% 
  drop_na() %>% 
  mutate(Site = as.factor(Site),
         WaterYear = as.factor(WaterYear),
         logDO.pctsat = log1p(avgDO.pctsat))

# Perform two-way ANOVA
doSatAnova <- aov(avgDO.pctsat ~ Site * WaterYear, data = doPctSat)
summary(doSatAnova)

doSatAnova_repeated <- nlme::lme(avgDO.pctsat ~ Site * WaterYear, random = ~1|Site/WaterYear, data = doPctSat)
summary(doSatAnova_repeated)

#Check assumptions
# Normality of Residuals
plot(doSatAnova, 2) # Q-Q plot
shapiro.test(residuals(doSatAnova))

# Homogeneity of Variances
plot(doSatAnova, 1) # Residuals vs Fitted plot

# Tukey's HSD post-hoc test
tukey_results <- TukeyHSD(doSatAnova)

# Extracting the Tukey HSD results for the main effect or interaction of interest
tukey_sites <- tukey_results$`Site:WaterYear`

# Generate compact letter displays (CLDs)
cld <- multcompView::multcompLetters4(doSatAnova, tukey_results)

# Extract the CLDs for plotting
cld_df <- data.frame(SiteWaterYear = names(cld$`Site:WaterYear`$Letters), 
                     Letters = cld$`Site:WaterYear`$Letters)

# Merging the CLDs with the original data for plotting
plot_data <- doPctSat %>%
  unite("SiteWaterYear", Site, WaterYear, remove = FALSE, sep = ":")

plot_data <- merge(plot_data, cld_df, by = "SiteWaterYear")

# Creating the boxplot with CLDs
plot_data %>%
  ggplot(aes(x = Site, y = avgDO.pctsat, color = WaterYear)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.75), alpha = 0.5) +  
  scale_color_brewer(palette = "Dark2") +
  geom_hline(yintercept = 50, linetype = "dashed") +
  geom_text(aes(label = Letters, y = max(avgDO.pctsat) + 2.5), 
            position = position_dodge(width = 0.75), 
            vjust = 0.2, 
            size = 5) +
  labs(x = "Water Year", y = expression(Daily ~ DO["%Sat"])) +
  theme_few() +
  theme(legend.title = element_blank(), legend.position = "bottom") -> satBoxAnova
satBoxAnova
ggsave("satBoxAnova.png", satBoxAnova, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")


# Perform mixed-effects model
library(lme4)
doSatMixedModel <- lmer(logDO.pctsat ~ Site * WaterYear + (1|Site),
                              data = doPctSat,
                              control = lmerControl(optimizer = "bobyqa",
                                                    optCtrl = list(maxfun = 100000)))
summary(doSatMixedModel)
t <- performance::check_model(doSatMixedModel)
plot(t)
```

```{r}
gpQ <- gpDat %>% 
  select(Site, dateTime, Q_Ls, SW_Level_ft, SW_Level_cm) %>% 
  mutate(Date = as_date(dateTime)) %>% 
  group_by(Site, Date) %>% 
  summarise(
    Q_Ls = mean(Q_Ls, na.rm = TRUE),
    SW_Level_ft = mean(SW_Level_ft, na.rm = TRUE),
    SW_Level_cm = mean(SW_Level_cm, na.rm = TRUE)
  )

doMet <- doMet %>% 
  full_join(gpQ, by = c("Site", "Date")) %>% 
  mutate(Date = as.Date(Date))
```

```{r}
# Convert Date to Date format if not already
doMet$Date <- as.Date(doMet$Date)

# Split the data by Site and WaterYear
split_data <- split(doMet, list(doMet$Site, doMet$WaterYear))

# Restore the default plotting area
par(mfrow = c(1, 1))
# Loop over each subset of the data
for (i in 1:length(split_data)) {
    data <- split_data[[i]]
    
    # Check if data is not empty
    if (nrow(data) > 0) {
        # Create a filename for the plot
        filename <- paste0(here("Figs/"), "DO_Amp_", names(split_data)[i], ".png")
        
        # Open a PNG device
        png(filename, width = 960, height = 540)
        
        # Set up the plotting area
        par(mfrow = c(1, 1), mar = c(6, 4, 2, 4), xpd = TRUE)
        
        # Plot the first y-axis (DOpctsat.amp)
        plot(data$Date, data$DOpctsat.amp, type = "b", col = "#1B9E77", pch = 16,
             ylab = "", xlab = "", main = paste(names(split_data)[i]),
             ylim = range(data$DOpctsat.amp, na.rm = TRUE), xaxt = "n")
        axis(side = 2)
        mtext("Daily DO%Sat Amplitude", side = 2, line = 2, col = 1)
        
        # Add the x-axis with monthly breaks
        axis.Date(1, at = seq(min(data$Date), max(data$Date), by = "month"), format = "%b")
        
        # Add the second y-axis (SW_Level_cm)
        par(new = TRUE)
        plot(data$Date, data$SW_Level_cm, type = "b", col = "black", axes = FALSE, xlab = "", ylab = "",
             ylim = range(data$SW_Level_cm, na.rm = TRUE))
        axis(side = 4)
        mtext("SW Level (cm)", side = 4, line = 2, col = 1)
        
        # Add a legend
        legend("bottom", legend = c("DO % Saturation", "SW Level (cm)"), col = c("#1B9E77", "black"), pch = c(16, NA), lty = c(NA, 1), inset = c(0, -0.15), horiz = TRUE, bty = "n")
        
        # Close the PNG device
        dev.off()
    }
}

#"#1B9E77" "#D95F02" "#7570B3" "#E7298A" "#66A61E" "#E6AB02" "#A6761D" "#666666"
# Restore the default plotting area
par(mfrow = c(1, 1))
# Loop over each subset of the data
for (i in 1:length(split_data)) {
    data <- split_data[[i]]
    
    # Check if data is not empty
    if (nrow(data) > 0) {
        # Create a filename for the plot
        filename <- paste0(here("Figs/"), "ProbHyp_", names(split_data)[i], ".png")
        
        # Open a PNG device
        png(filename, width = 960, height = 540)
        
        # Set up the plotting area
        par(mfrow = c(1, 1), mar = c(6, 4, 2, 4), xpd = TRUE)
        
        # Plot the first y-axis (DOpctsat.amp)
        plot(data$Date, data$probHypSat, type = "b", col = "#D95F02", pch = 16,
             ylab = "", xlab = "", main = paste(names(split_data)[i]),
             ylim = range(data$probHypSat, na.rm = TRUE), xaxt = "n")
        axis(side = 2)
        mtext("Probability of Hypoxia", side = 2, line = 2, col = 1)
        
        # Add the x-axis with monthly breaks
        axis.Date(1, at = seq(min(data$Date), max(data$Date), by = "month"), format = "%b")
        
        # Add the second y-axis (SW_Level_cm)
        par(new = TRUE)
        plot(data$Date, data$SW_Level_cm, type = "b", col = "black", axes = FALSE, xlab = "", ylab = "",
             ylim = range(data$SW_Level_cm, na.rm = TRUE))
        axis(side = 4)
        mtext("SW Level (cm)", side = 4, line = 2, col = 1)
        
        # Add a legend
        legend("bottom", legend = c("Probability of Hypoxia", "SW Level (cm)"), col = c("#D95F02", "black"), pch = c(16, NA), lty = c(NA, 1), inset = c(0, -0.15), horiz = TRUE, bty = "n")
        
        # Close the PNG device
        dev.off()
    }
}

# Restore the default plotting area
par(mfrow = c(1, 1))
# Loop over each subset of the data
for (i in 1:length(split_data)) {
    data <- split_data[[i]]
     # Filter out rows with non-finite values in NHR_sat or SW_Level_cm
    #data <- data[is.finite(data$avgDO.pctsat) , ]
    # Check if data is not empty
    if (nrow(data) > 0) {
        # Create a filename for the plot
        filename <- paste0(here("Figs/"), "DOSat_", names(split_data)[i], ".png")
        
        # Open a PNG device
        png(filename, width = 960, height = 540)
        
        # Set up the plotting area
        par(mfrow = c(1, 1), mar = c(6, 4, 2, 4), xpd = TRUE)
        
        # Plot the first y-axis (DOpctsat.amp)
        plot(data$Date, data$avgDO.pctsat, type = "b", col = "#66A61E", pch = 16,
             ylab = "", xlab = "", main = paste(names(split_data)[i]),
             ylim = range(data$avgDO.pctsat, na.rm = TRUE), xaxt = "n")
        axis(side = 2)
        mtext("DO % Saturation", side = 2, line = 2, col = 1)
        
        # Add the x-axis with monthly breaks
        axis.Date(1, at = seq(min(data$Date), max(data$Date), by = "month"), format = "%b")
        
        # Add the second y-axis (SW_Level_cm)
        par(new = TRUE)
        plot(data$Date, data$SW_Level_cm, type = "b", col = "black", axes = FALSE, xlab = "", ylab = "",
             ylim = range(data$SW_Level_cm, na.rm = TRUE))
        axis(side = 4)
        mtext("SW Level (cm)", side = 4, line = 2, col = 1)
        
        # Add a legend
        legend("bottom", legend = c("DO % Saturation", "SW Level (cm)"), col = c("#66A61E", "black"), pch = c(16, NA), lty = c(NA, 1), inset = c(0, -0.15), horiz = TRUE, bty = "n")
        
        # Close the PNG device
        dev.off()
    }
}

# Restore the default plotting area
par(mfrow = c(1, 1))
# Loop over each subset of the data
for (i in 1:length(split_data)) {
    data <- split_data[[i]]
    
    # Check if data is not empty
    if (nrow(data) > 0) {
        # Create a filename for the plot
        filename <- paste0(here("Figs/"), "cvDO_", names(split_data)[i], ".png")
        
        # Open a PNG device
        png(filename, width = 960, height = 540)
        
        # Set up the plotting area
        par(mfrow = c(1, 1), mar = c(6, 4, 2, 4), xpd = TRUE)
        
        # Plot the first y-axis (DOpctsat.amp)
        plot(data$Date, data$cvDO.pctsat, type = "b", col = "#7570B3", pch = 16,
             ylab = "", xlab = "", main = paste(names(split_data)[i]),
             ylim = range(data$cvDO.pctsat, na.rm = TRUE), xaxt = "n")
        axis(side = 2)
        mtext("DO % Saturation Variability", side = 2, line = 2, col = 1)
        
        # Add the x-axis with monthly breaks
        axis.Date(1, at = seq(min(data$Date), max(data$Date), by = "month"), format = "%b")
        
        # Add the second y-axis (SW_Level_cm)
        par(new = TRUE)
        plot(data$Date, data$SW_Level_cm, type = "b", col = "black", axes = FALSE, xlab = "", ylab = "",
             ylim = range(data$SW_Level_cm, na.rm = TRUE))
        axis(side = 4)
        mtext("SW Level (cm)", side = 4, line = 2, col = 1)
        
        # Add a legend
        legend("bottom", legend = c("DO % Saturation Variability", "SW Level (cm)"), col = c("#7570B3", "black"), pch = c(16, NA), lty = c(NA, 1), inset = c(0, -0.15), horiz = TRUE, bty = "n")
        
        # Close the PNG device
        dev.off()
    }
}


```

## DO and SW Level

```{r}
# Convert Date to Date format if not already
doMet$Date <- as.Date(doMet$Date)

# Split the data by Site and WaterYear
split_data <- split(doMet, list(doMet$Site, doMet$WaterYear))

# Restore the default plotting area
par(mfrow = c(1, 1))
# Loop over each subset of the data
for (i in 1:length(split_data)) {
    data <- split_data[[i]]
    
    # Check if data is not empty
    if (nrow(data) > 0) {
        # Create a filename for the plot
        filename <- paste0(here("Figs/"), "DO_Amp_Q_", names(split_data)[i], ".png")
        
        # Open a PNG device
        png(filename, width = 960, height = 540)
        
        # Set up the plotting area
        par(mfrow = c(1, 1), mar = c(6, 4, 2, 4), xpd = TRUE)
        
        # Plot the first y-axis (DOpctsat.amp)
        plot(data$Date, data$DOpctsat.amp, type = "b", col = "#1B9E77", pch = 16,
             ylab = "", xlab = "", main = paste(names(split_data)[i]),
             ylim = range(data$DOpctsat.amp, na.rm = TRUE), xaxt = "n")
        axis(side = 2)
        mtext("Daily DO%Sat Amplitude", side = 2, line = 2, col = 1)
        
        # Add the x-axis with monthly breaks
        axis.Date(1, at = seq(min(data$Date), max(data$Date), by = "month"), format = "%b")
        
        # Add the second y-axis (SW_Level_cm)
        par(new = TRUE)
       plot(data$Date[data$Q_Ls>=0], data$Q_Ls[data$Q_Ls>=0], type = "l", col = "black", axes = FALSE, xlab = "", ylab = "",
             ylim = c(0, max(data$Q_Ls[data$Q_Ls>=0], na.rm = TRUE)))
        axis(side = 4)
        mtext("Discharge (L/s)", side = 4, line = 2, col = 1)
        
        # Add a legend
        legend("bottom", legend = c("DO % Saturation", "Discharge (L/s)"), col = c("#1B9E77", "black"), pch = c(16, NA), lty = c(NA, 1), inset = c(0, -0.15), horiz = TRUE, bty = "n")
        
        # Close the PNG device
        dev.off()
    }
}

# Restore the default plotting area
par(mfrow = c(1, 1))
# Loop over each subset of the data
for (i in 1:length(split_data)) {
    data <- split_data[[i]]
    
    # Check if data is not empty
    if (nrow(data) > 0) {
        # Create a filename for the plot
        filename <- paste0(here("Figs/"), "ProbHyp_Q_", names(split_data)[i], ".png")
        
        # Open a PNG device
        png(filename, width = 960, height = 540)
        
        # Set up the plotting area
        par(mfrow = c(1, 1), mar = c(6, 4, 2, 4), xpd = TRUE)
        
        # Plot the first y-axis (DOpctsat.amp)
        plot(data$Date, data$probHypSat, type = "b", col = "#D95F02", pch = 16,
             ylab = "", xlab = "", main = paste(names(split_data)[i]),
             ylim = range(data$probHypSat, na.rm = TRUE), xaxt = "n")
        axis(side = 2)
        mtext("Probability of Hypoxia", side = 2, line = 2, col = 1)
        
        # Add the x-axis with monthly breaks
        axis.Date(1, at = seq(min(data$Date), max(data$Date), by = "month"), format = "%b")
        
        # Add the second y-axis (SW_Level_cm)
        par(new = TRUE)
        plot(data$Date[data$Q_Ls>=0], data$Q_Ls[data$Q_Ls>=0], type = "l", col = "black", axes = FALSE, xlab = "", ylab = "",
             ylim = c(0, max(data$Q_Ls[data$Q_Ls>=0], na.rm = TRUE)))
        axis(side = 4)
        mtext("Discharge (L/s)", side = 4, line = 2, col = 1)
        
        # Add a legend
        legend("bottom", legend = c("Probability of Hypoxia", "Discharge (L/s)"), col = c("#D95F02", "black"), pch = c(16, NA), lty = c(NA, 1), inset = c(0, -0.15), horiz = TRUE, bty = "n")
        
        # Close the PNG device
        dev.off()
    }
}

# Restore the default plotting area
par(mfrow = c(1, 1))
# Loop over each subset of the data
for (i in 1:length(split_data)) {
    data <- split_data[[i]]
     # Filter out rows with non-finite values in NHR_sat or SW_Level_cm
    #data <- data[is.finite(data$avgDO.pctsat) , ]
    # Check if data is not empty
    if (nrow(data) > 0) {
        # Create a filename for the plot
        filename <- paste0(here("Figs/"), "DOSat_Q_", names(split_data)[i], ".png")
        
        # Open a PNG device
        png(filename, width = 960, height = 540)
        
        # Set up the plotting area
        par(mfrow = c(1, 1), mar = c(6, 4, 2, 4), xpd = TRUE)
        
        # Plot the first y-axis (DOpctsat.amp)
        plot(data$Date, data$avgDO.pctsat, type = "b", col = "#66A61E", pch = 16,
             ylab = "", xlab = "", main = paste(names(split_data)[i]),
             ylim = range(data$avgDO.pctsat, na.rm = TRUE), xaxt = "n")
        axis(side = 2)
        mtext("DO % Saturation", side = 2, line = 2, col = 1)
        
        # Add the x-axis with monthly breaks
        axis.Date(1, at = seq(min(data$Date), max(data$Date), by = "month"), format = "%b")
        
        # Add the second y-axis (SW_Level_cm)
        par(new = TRUE)
         plot(data$Date[data$Q_Ls>=0], data$Q_Ls[data$Q_Ls>=0], type = "l", col = "black", axes = FALSE, xlab = "", ylab = "",
             ylim = c(0, max(data$Q_Ls[data$Q_Ls>=0], na.rm = TRUE)))
        axis(side = 4)
        mtext("Discharge (L/s)", side = 4, line = 2, col = 1)
        
        # Add a legend
        legend("bottom", legend = c("DO % Saturation", "Discharge (L/s)"), col = c("#66A61E", "black"), pch = c(16, NA), lty = c(NA, 1), inset = c(0, -0.15), horiz = TRUE, bty = "n")
        
        # Close the PNG device
        dev.off()
    }
}

# Restore the default plotting area
par(mfrow = c(1, 1))
# Loop over each subset of the data
for (i in 1:length(split_data)) {
    data <- split_data[[i]]
    
    # Check if data is not empty
    if (nrow(data) > 0) {
        # Create a filename for the plot
        filename <- paste0(here("Figs/"), "cvDO_Q_", names(split_data)[i], ".png")
        
        # Open a PNG device
        png(filename, width = 960, height = 540)
        
        # Set up the plotting area
        par(mfrow = c(1, 1), mar = c(6, 4, 2, 4), xpd = TRUE)
        
        # Plot the first y-axis (DOpctsat.amp)
        plot(data$Date, data$cvDO.pctsat, type = "b", col = "#7570B3", pch = 16,
             ylab = "", xlab = "", main = paste(names(split_data)[i]),
             ylim = range(data$cvDO.pctsat, na.rm = TRUE), xaxt = "n")
        axis(side = 2)
        mtext("DO % Saturation Variability", side = 2, line = 2, col = 1)
        
        # Add the x-axis with monthly breaks
        axis.Date(1, at = seq(min(data$Date), max(data$Date), by = "month"), format = "%b")
        
        # Add the second y-axis (SW_Level_cm)
        par(new = TRUE)
        plot(data$Date[data$Q_Ls>=0], data$Q_Ls[data$Q_Ls>=0], type = "l", col = "black", axes = FALSE, xlab = "", ylab = "",
             ylim = c(0, max(data$Q_Ls[data$Q_Ls>=0], na.rm = TRUE)))
        axis(side = 4)
        mtext("Discharge (L/s)", side = 4, line = 2, col = 1)
        
        # Add a legend
        legend("bottom", legend = c("DO % Saturation Variability", "Discharge (L/s)"), col = c("#7570B3", "black"), pch = c(16, NA), lty = c(NA, 1), inset = c(0, -0.15), horiz = TRUE, bty = "n")
        
        # Close the PNG device
        dev.off()
    }
}
```

```{r}
doMet <- doMet %>%
  mutate(Year = factor(Year),
         season = factor(season, levels = c("Spring", "Summer", "Fall", "Winter"), ordered = T))

doMet %>%
  #filter(season %in% c("Spring", "Summer")) %>% 
  #filter(!Site == "SHN") %>% 
  ggplot(aes(Year, DOpctsat.amp, color = Year)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.75)) +
  geom_jitter(position = position_jitterdodge(), alpha = 0.6) +
  theme_few() +
  labs(y = expression(Daily ~ DO["%Sat"] ~ "Amplitude"))+
  scale_color_brewer(palette = "Dark2")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_blank()) +
  facet_wrap(~Site, ncol = 1, scales = "free") -> doAmpBox
doAmpBox
ggsave("doAmpBox.png", doAmpBox, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")


doMet %>%
  #filter(season %in% c("Spring", "Summer")) %>% 
  #filter(!Site == "SHN") %>% 
  ggplot(aes(Year, probHypSat, color = Year)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.75)) +
  geom_jitter(position = position_jitterdodge(), alpha = 0.6) +
  theme_few() +
  labs(y = expression(DO["%Sat"] ~ "Probability of Hypoxia"))+
  scale_color_brewer(palette = "Dark2")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_blank()) +
  facet_wrap(~Site, ncol = 1, scales = "free") -> probHypBox
probHypBox
ggsave("probHypBox.png", probHypBox, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")

doMet %>%
  filter(season %in% c("Spring", "Summer")) %>% 
  ggplot(aes(season, avgDO.pctsat, color = season)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.75)) +
  geom_jitter(position = position_jitterdodge(), alpha = 0.6) +
  theme_few() +
  scale_color_brewer(palette = "Dark2")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_blank()) +
  facet_grid(WaterYear~Site, scales = "free")

doMet %>%
  filter(season %in% c("Spring", "Summer")) %>% 
  ggplot(aes(season, DOpctsat.amp, color = season)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.75)) +
  geom_jitter(position = position_jitterdodge(), alpha = 0.6) +
  theme_few() +
  labs(y = expression(Daily ~ DO["%Sat"] ~ "Amplitude"))+ 
  scale_color_manual(values = c("Spring" = "#275E4D", "Summer" = "#FEB424"))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_blank(),
        legend.position = "none", 
        text = element_text(size = 25),       # Increase overall text size
        axis.title = element_text(size = 25), # Increase axis titles size
        axis.text = element_text(size = 25),  # Increase axis text size
        strip.text = element_text(size = 25),  # Increase facet labels size
    ) +
  facet_grid(WaterYear~Site, scales = "fixed") -> ampBox
ampBox
ggsave("ampBox.png", ampBox, here("Figs/"), device = "png", width = 11, height = 6, units = "in", dpi = "retina")


doMet %>%
  filter(season %in% c("Spring", "Summer")) %>% 
  ggplot(aes(season, minDO.pctsat, color = season)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.75)) +
  geom_jitter(position = position_jitterdodge(), alpha = 0.6) +
  theme_few() +
  labs(x = "",
       y = expression(Daily ~ Minimum ~ DO["%Sat"]))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.x = element_blank(),
        legend.position = "none",
        text = element_text(size = 25),       # Increase overall text size
        axis.title = element_text(size = 25), # Increase axis titles size
        axis.text = element_text(size = 25),  # Increase axis text size
        strip.text = element_text(size = 25),  # Increase facet labels size
        )+
  scale_y_continuous(breaks = scales::pretty_breaks(n = 3))+
  scale_color_manual(values = c("Spring" = "#275E4D", "Summer" = "#FEB424"))+
  scale_y_continuous(breaks = scales::pretty_breaks(n = 3))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_blank()) +
  facet_grid(WaterYear~Site, scales = "fixed") -> minBox
minBox
ggsave("minBox.png", minBox, here("Figs/"), device = "png", width = 11, height = 6, units = "in", dpi = "retina")

doMet %>%
  filter(season %in% c("Spring", "Summer")) %>% 
  ggplot(aes(probHypSat, fill = season, after_stat(scaled))) +
  geom_density(alpha = 0.5) +
  theme_few() +
  scale_fill_brewer(palette = "Dark2")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_blank()) +
  facet_grid(WaterYear~Site, scales = "free")

dailyDO %>%
  filter(season %in% c("Spring", "Summer")) %>% 
  ggplot(aes(x = probHypSat, fill = season, after_stat(scaled))) +
  geom_density(alpha = 0.8) +
  #xlim(-1.5, 1.5)+"
  scale_fill_manual(values = c("Spring" = "#275E4D", "Summer" = "#FEB424"))+
  labs(x = expression(DO["%Sat"] ~ "Probability of Hypoxia"), y = "Density") +
  theme_few() +
  scale_x_continuous(labels = scales::number_format(accuracy = 0.1)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 3))+
  theme(legend.position = "bottom", legend.title = element_blank(),
        text = element_text(size = 25),       # Increase overall text size
        axis.title = element_text(size = 25), # Increase axis titles size
        axis.text = element_text(size = 25),  # Increase axis text size
        strip.text = element_text(size = 25),  # Increase facet labels size
        legend.text = element_text(size = 25)
    )+
  facet_grid(WaterYear~Site) -> probHypSeason
probHypSeason
ggsave("probHypSeason.png", probHypSeason, here("Figs/"), device = "png",  width = 16, height = 8.5, units = "in", dpi = "retina")
```

```{r}
gpDOmet <- doMet %>% 
  drop_na() %>% 
  ungroup()
#PCA
doPca <- gpDOmet %>%
  select(minDO.pctsat, maxDO.pctsat, avgDO.pctsat, DOpctsat.amp, probHypSat,  NHR_sat) %>% 
  #m dayProbSat, nightProbSat,
  select(where(is.numeric)) %>% # retain only numeric columns
  prcomp(center = T, scale = T) # do PCA on scaled data

#extract loadings
doPcaLoad <- doPca %>%
  tidy(matrix = "loadings")
doPcaLoad
#Variance explained
doPcaEig <- doPca %>%
  tidy(matrix = "eigenvalues")
doPcaEig
#Scree plot
fviz_eig(doPca, addlabels = TRUE)

#Qual of representation
fviz_cos2(doPca, choice = "var", axes = 1)
fviz_cos2(doPca, choice = "var", axes = 2)

#Biplot
fviz_pca_biplot(doPca, 
                col.ind = gpDOmet$Site,
                palette =  RColorBrewer::brewer.pal(3, "Dark2"), 
                addEllipses = F,
                label = "var",
                col.var = "black",
                repel = TRUE,
                legend.title = "")+
  theme_few()
```

# Kings Long Term Data

```{r}
#Get discharge data from NWIS
rawKingsQ <- readNWISdv(siteNumbers = "06879650",
                     parameterCd = "00060",
                     startDate = "1979-04-11", 
                     endDate = today()) 
```

## Exceedance Probability by Year

```{r}
kingsQ <- renameNWISColumns(rawKingsQ) %>% 
  select(-agency_cd, -Flow_cd) %>% 
  rename(FlowCFS = Flow) %>% #Label discharge units (cubic feet per second)
  mutate(
    FlowCMS = round(FlowCFS * 0.028316847, 2),#cfs to cms
    Year = year(Date),
    Period = case_when(Year <= 2020 ~ "1980-2020",
                       Year == 2021 ~ "2021",
                       Year == 2022 ~ "2022",
                       Year == 2023 ~ "2023",
                       Year == 2024 ~ "2024")) %>% 
  group_by(Period) %>%
  arrange(desc(FlowCMS),, .by_group = TRUE) %>%
  mutate(
    Rank = row_number(),
    ExceedanceProbability = (Rank / (n() + 1)) * 100)

kingsQ %>%
  ggplot(aes(Date, FlowCMS)) +
  geom_line(color = "#0084a8") +
  geom_rect(data = kingsQ %>% filter(FlowCMS <= 0),
            aes(xmin = Date, xmax = Date + 1),
            ymin = -Inf, ymax = Inf,
            fill = "#ff7f50",
            alpha = 0.1,
            inherit.aes = FALSE) +
  labs(title = "Kings Creek",
    x = "",
       y = expression(Discharge~("m"^3*"s"^{-1}))) +
  theme_few() +
  scale_x_date(labels = date_format("%b '%y")) +
  scale_y_continuous(trans = "log1p") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)) -> kingsDry
kingsDry
ggsave("kingsDry.png", kingsDry, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")

kingsQ %>% 
  ggplot(aes(x = ExceedanceProbability, y = FlowCMS, color = Period)) +
  geom_line() +
  scale_y_log10(labels = label_number()) +
  scale_color_manual(values = c("1980-2020" = "black", "2021" = "#1B9E77", "2022" = "#D95F02", "2023" = "#7570B3", "2024" = "#E7298A")) +
  labs(title = "Kings Creek",
       x = "Exceedance Probability (%)",
       y = expression(Discharge~("m"^3*"s"^{-1}))) +
  theme_few()+
  theme(legend.position = "bottom", legend.title = element_blank()) -> excProb
excProb
ggsave("excProb.png", excProb, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")
```

```{r}
kingsQ <- renameNWISColumns(rawKingsQ) %>% 
  select(-agency_cd, -Flow_cd) %>% 
  rename(FlowCFS = Flow) %>% 
  mutate(
    FlowCMS = round(FlowCFS * 0.028316847, 2),
    Year = year(Date),
    DOY = yday(Date),
    Month = month(Date, label = TRUE, abbr = TRUE),
    DateForPlot = as.Date(paste("2020", DOY), format = "%Y %j")
  ) %>% 
  group_by(Year) %>%
  mutate(
    TotalYearlyQ = sum(FlowCMS, na.rm = TRUE),
    CumulativeQ = cumsum(FlowCMS),
    CumulativeProp = CumulativeQ / TotalYearlyQ
  ) %>% 
  ungroup() %>%
  group_by(DOY) %>%
  mutate(MeanQ = mean(FlowCMS, na.rm = TRUE),
         MedianQ = median(FlowCMS, na.rm = T)) %>%
  ungroup()

# Calculate low flow statistics
kingsLowFlow <- kingsQ %>%
  mutate(WaterYear = calcWaterYear(Date)) %>% 
  group_by(WaterYear) %>%
  summarize(
    WetDays = sum(FlowCMS > 0, na.rm = TRUE),
    DryDays = sum(FlowCMS == 0, na.rm = TRUE)
  ) %>%
  ungroup()

# Calculate the cumulative discharge for each day of the year, the total discharge for each year, merge the total discharge with the original data, calculate the relative cumulative discharge, and find the date of the center of mass (50% cumulative discharge) all in one go.

q50perc <- kingsQ %>%
  mutate(
    CumulativeQ = cumsum(FlowCMS),
    TotalQ = sum(FlowCMS, na.rm = TRUE),
    RelativeCumulativeQ = CumulativeQ / TotalQ
  ) %>%
  summarize(CentroidDOY = DOY[which.min(abs(RelativeCumulativeQ - 0.5))])

q50percAIMS <- kingsQ %>%
  group_by(Year) %>%
  filter(Year >= 2021) %>% 
  mutate(
    CumulativeQ = cumsum(FlowCMS),
    TotalQ = sum(FlowCMS, na.rm = TRUE),
    RelativeCumulativeQ = CumulativeQ / TotalQ
  ) %>%
  ungroup() %>%
  group_by(Year) %>%
  summarize(CentroidDOY = DOY[which.min(abs(RelativeCumulativeQ - 0.5))]) %>%
  ungroup() %>% 
  mutate(CentroidDate = as.Date(paste("2020", CentroidDOY), format = "%Y %j"))
```

```{r}
ggplot() +
  geom_line(data = kingsQ %>% filter(Year != 2024), aes(x = DateForPlot, y = FlowCMS, group = Year), color = "grey", alpha = 0.5) +
  geom_line(data = kingsQ %>% filter(Year == 2024), aes(x = DateForPlot, y = FlowCMS, group = Year), color = "black") +
  geom_line(data = kingsQ, aes(x = DateForPlot, y = MeanQ), color = "red", alpha = 0.5) +
  geom_line(data = kingsQ, aes(x = DateForPlot, y = MedianQ), color = "blue", alpha = 0.5) +
  scale_x_date(date_labels = "%b", date_breaks = "1 month") +
  labs(x = "",
       y = expression(Discharge~("m"^3*"s"^{-1}))) +
  theme_few() -> longTermQ
longTermQ
ggsave("longTermQ.png", longTermQ, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")

yearColor <- c("2021" = "#1B9E77", "2022" = "#D95F02", "2023" = "#7570B3", "2024" = "#E7298A", "1980-2020" = "black")
ggplot() +
  geom_line(data = kingsQ %>% filter(Year == 2021), aes(x = DateForPlot, y = CumulativeProp, color = factor(Year)), linetype = "solid") +
  geom_line(data = kingsQ %>% filter(Year == 2022), aes(x = DateForPlot, y = CumulativeProp, color = factor(Year)), linetype = "solid") +
  geom_line(data = kingsQ %>% filter(Year == 2023), aes(x = DateForPlot, y = CumulativeProp, color = factor(Year)), linetype = "solid") +
  #geom_line(data = kingsQ %>% filter(Year == 2024), aes(x = DateForPlot, y = CumulativeProp, color = factor(Year)), linetype = "solid") +
  geom_line(data = kingsQ %>% group_by(DOY) %>%
              filter(Year < 2021) %>% 
              summarise(DateForPlot = first(DateForPlot), AverageCumulativeProp = mean(CumulativeProp, na.rm = TRUE)), 
            aes(x = DateForPlot, y = AverageCumulativeProp, color = "1980-2020"), linetype = "dashed") +
  labs(title = "Kings Creek", x = "", y = "Cumulative Proportion \n of Total Yearly Discharge") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month") +
  scale_color_manual(values = yearColor, labels = c("2021", "2022", "2023", "1980-2020")) +
  theme_few() +
  theme(legend.position = "bottom", legend.title = element_blank()) -> q50
q50
ggsave("q50.png", q50, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")

ggplot() +
  geom_line(data = kingsQ %>% filter(Year != 2024), aes(x = DateForPlot, y = FlowCMS, group = Year), color = "grey", alpha = 0.5) +
  geom_line(data = kingsQ %>% filter(Year == 2024), aes(x = DateForPlot, y = FlowCMS, group = Year), color = "black") +
  geom_line(data = kingsQ, aes(x = DateForPlot, y = MeanQ), color = "red", alpha = 0.5) +
  geom_vline(data = q50percAIMS, aes(xintercept = CentroidDate, color = as.factor(Year)), linetype = "dashed") +
  scale_color_brewer(palette = "Dark2", type = "div") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month") +
  labs(title = "Kings Creek",
       x = "",
       y = expression(Discharge~("m"^3*"s"^{-1})),
       color = "50% Cumulative\n    Discharge") +
  theme_few()+
  theme(legend.position = "bottom") -> q50date
ggsave("q50Date.png", q50date, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")

kingsLowFlow %>% 
  ggplot(aes(WaterYear, DryDays))+
  geom_point()+
  geom_smooth(method='lm') +
  stat_regline_equation(aes(label = ..rr.label..))+
  theme_few()
```

# LTER Weirs

```{r}
# weirDat <- read_csv(here("Data/knzWeirs_2023.csv")) %>% 
#   filter(RecYear >= 1993 & Discharge < 1000) %>% 
#   mutate(RecDay = case_when(RecDay == 0 ~ 1, TRUE ~ RecDay),
#     Date = paste(RecYear, RecMonth, RecDay, sep = "-"),
#          RecHour = parse_time(str_pad(RecHour, width = 4, side = "left", pad = 0), "%H%M"),
#          dateTime = ymd_hms(paste(Date, RecHour, sep = " "))) %>% 
#   select(Watershed, FlowCMS = Discharge, QualFlag, dateTime) %>% 
#   mutate(SqKm = case_when(
#     Watershed == "N04D" ~ 1357221.619483/1000000,
#     Watershed == "N02B" ~ 1188677.233751/1000000,
#     Watershed == "N01B" ~ 1208908.624873/1000000,
#     Watershed == "N20B" ~ 839577.609151/1000000))
# 
# # Calculate daily average discharge
# dailyWeir <- weirDat %>% 
#   mutate(Date = as.Date(dateTime)) %>% 
#   group_by(Date, Watershed, SqKm) %>% 
#   summarise(FlowCMS = mean(FlowCMS, na.rm = TRUE), .groups = 'drop') %>% 
#   mutate(Year = year(Date),
#          DOY = yday(Date),
#          Month = month(Date, label = TRUE, abbr = TRUE),
#          DateForPlot = as.Date(paste("2020", DOY), format = "%Y %j"),
#          Period = case_when(Year <= 2020 ~ "1993-2020",
#                             Year == 2021 ~ "2021",
#                             Year == 2022 ~ "2022",
#                             Year == 2023 ~ "2023",
#                             Year == 2024 ~ "2024")) %>% 
#   group_by(Watershed, Period) %>%
#   arrange(desc(FlowCMS), .by_group = TRUE) %>%
#   mutate(ExceedanceProbability = (row_number() / (n() + 1)) * 100)
# 
# 
# 
# dailyWeir %>%
#   ggplot(aes(Date, FlowCMS)) +
#   geom_line(color = "#0084a8") +
#   geom_rect(data = dailyWeir %>% filter(FlowCMS <= 0), aes(xmin = Date, xmax = Date + 1, ymin = -Inf, ymax = +Inf),
#     fill = "#ff7f50",
#     alpha = 0.1) +
#   labs(x = "",
#        y = expression(Discharge~("m"^3*"s"^{-1})))+
#   theme_few() +
#   scale_x_date(labels = date_format("%b '%y")) +
#   theme(axis.title.x = element_blank(),
#         axis.text.x = element_text(angle = 45, hjust = 1))+
#   facet_wrap(~Watershed, ncol = 2, scales = "free")
# 
# dailyWeir %>% 
#   ggplot(aes(x = ExceedanceProbability, y = FlowCMS, color = Period)) +
#   geom_line() +
#   scale_y_log10(labels = label_number()) +
#   scale_color_manual(values = c("1993-2020" = "black", "2021" = "#1B9E77", "2022" = "#D95F02", "2023" = "#7570B3", "2024" = "#E7298A")) +
#   labs(x = "Exceedance Probability (%)",
#        y = expression(Discharge~("m"^3*"s"^{-1}))) +
#   theme_few()+
#   theme(legend.position = "bottom", legend.title = element_blank())+
#   facet_wrap(~Watershed, ncol = 2, scales = "free")
```

# Elk River (YMR) Long Term Data

```{r}
#Get discharge data from NWIS
rawElkQdaily <- readNWISdv(siteNumbers = "07169800",
                     parameterCd = "00060",
                     startDate = "1991-04-27", 
                     endDate = today()) 
```

```{r}
elkQdaily <- renameNWISColumns(rawElkQdaily) %>% 
  select(-agency_cd, -Flow_cd) %>% 
  rename(FlowCFS = Flow) %>% #Label discharge units (cubic feet per second)
  mutate(
    FlowCMS = FlowCFS * 0.028316847,#cfs to cms
    Year = year(Date),
    Period = case_when(Year <= 2020 ~ "1991-2020",
                       Year == 2021 ~ "2021",
                       Year == 2022 ~ "2022",
                       Year == 2023 ~ "2023",
                       Year == 2024 ~ "2024")) %>% 
  group_by(Period) %>%
  arrange(desc(FlowCMS)) %>%
  mutate(ExceedanceProbability = (row_number() / (n() + 1)) * 100)

elkQdaily %>%
  ggplot(aes(Date, FlowCMS)) +
  geom_line(color = "#0084a8") +
  geom_rect(data = elkQdaily %>% filter(FlowCMS <= 0), aes(xmin = Date, xmax = Date + 1, ymin = -Inf, ymax = +Inf),
    fill = "#ff7f50", alpha = 0.1) +
  labs(title = "Elk River",
    x = "",
       y = expression(Discharge~("m"^3*"s"^{-1})))+
  theme_few() +
  scale_x_date(labels = date_format("%b '%y")) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)) -> elkDry
elkDry
ggsave("elkDry.png", elkDry, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")


elkQdaily %>% 
  ggplot(aes(x = ExceedanceProbability, y = FlowCMS, color = Period)) +
  geom_line() +
  scale_y_log10(labels = label_number()) +
  scale_color_manual(values = c("1991-2020" = "black", "2021" = "#1B9E77", "2022" = "#D95F02", "2023" = "#7570B3", "2024" = "#E7298A")) +
  labs(title = "Elk River",
       x = "Exceedance Probability (%)",
       y = expression(Discharge~("m"^3*"s"^{-1}))) +
  theme_few()+
  theme(legend.position = "bottom", legend.title = element_blank())-> elkExcProb
elkExcProb
ggsave("ElkExcProb.png", elkExcProb, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")
```

```{r}
elkQdaily <- renameNWISColumns(rawElkQdaily) %>% 
  select(-agency_cd, -Flow_cd) %>% 
  rename(FlowCFS = Flow) %>% 
  mutate(
    FlowCMS = round(FlowCFS * 0.028316847, 2),
    Year = year(Date),
    DOY = yday(Date),
    Month = month(Date, label = TRUE, abbr = TRUE),
    DateForPlot = as.Date(paste("2020", DOY), format = "%Y %j")
  ) %>% 
  group_by(Year) %>%
  mutate(
    TotalYearlyQ = sum(FlowCMS, na.rm = TRUE),
    CumulativeQ = cumsum(FlowCMS),
    CumulativeProp = CumulativeQ / TotalYearlyQ
  ) %>% 
  ungroup() %>%
  group_by(DOY) %>%
  mutate(MeanQ = mean(FlowCMS, na.rm = TRUE),
         MedianQ = median(FlowCMS, na.rm = T)) %>%
  ungroup()

# Calculate low flow statistics
elkLowFlow <- elkQdaily %>%
  mutate(WaterYear = calcWaterYear(Date)) %>% 
  group_by(WaterYear) %>%
  summarize(
    WetDays = sum(FlowCMS > 0, na.rm = TRUE),
    DryDays = sum(FlowCMS == 0, na.rm = TRUE)
  ) %>%
  ungroup()

# Calculate the cumulative discharge for each day of the year, the total discharge for each year, merge the total discharge with the original data, calculate the relative cumulative discharge, and find the date of the center of mass (50% cumulative discharge) all in one go.

elkQdaily50perc <- elkQdaily %>%
  mutate(
    CumulativeQ = cumsum(FlowCMS),
    TotalQ = sum(FlowCMS, na.rm = TRUE),
    RelativeCumulativeQ = CumulativeQ / TotalQ
  ) %>%
  summarize(CentroidDOY = DOY[which.min(abs(RelativeCumulativeQ - 0.5))])

elkQdaily50percAIMS <- elkQdaily %>%
  group_by(Year) %>%
  filter(Year >= 2021) %>% 
  mutate(
    CumulativeQ = cumsum(FlowCMS),
    TotalQ = sum(FlowCMS, na.rm = TRUE),
    RelativeCumulativeQ = CumulativeQ / TotalQ
  ) %>%
  ungroup() %>%
  group_by(Year) %>%
  summarize(CentroidDOY = DOY[which.min(abs(RelativeCumulativeQ - 0.5))]) %>%
  ungroup() %>% 
  mutate(CentroidDate = as.Date(paste("2020", CentroidDOY), format = "%Y %j"))

ggplot() +
  geom_line(data = elkQdaily %>% filter(Year != 2024), aes(x = DateForPlot, y = FlowCMS, group = Year), color = "grey", alpha = 0.5) +
  geom_line(data = elkQdaily %>% filter(Year == 2024), aes(x = DateForPlot, y = FlowCMS, group = Year), color = "black") +
  geom_line(data = elkQdaily, aes(x = DateForPlot, y = MeanQ), color = "red", alpha = 0.5) +
  geom_line(data = elkQdaily, aes(x = DateForPlot, y = MedianQ), color = "blue", alpha = 0.5) +
  scale_x_date(date_labels = "%b", date_breaks = "1 month") +
  labs(x = "",
       y = expression(Discharge~("m"^3*"s"^{-1}))) +
  theme_few()

yearColor <- c("2021" = "#1B9E77", "2022" = "#D95F02", "2023" = "#7570B3", "2024" = "#E7298A", "1991-2020" = "black")
ggplot() +
  geom_line(data = elkQdaily %>% filter(Year == 2021), aes(x = DateForPlot, y = CumulativeProp, color = factor(Year)), linetype = "solid") +
  geom_line(data = elkQdaily %>% filter(Year == 2022), aes(x = DateForPlot, y = CumulativeProp, color = factor(Year)), linetype = "solid") +
  geom_line(data = elkQdaily %>% filter(Year == 2023), aes(x = DateForPlot, y = CumulativeProp, color = factor(Year)), linetype = "solid") +
  #geom_line(data = elkQdaily %>% filter(Year == 2024), aes(x = DateForPlot, y = CumulativeProp, color = factor(Year)), linetype = "solid") +
  geom_line(data = elkQdaily %>% group_by(DOY) %>%
              filter(Year < 2021) %>% 
              summarise(DateForPlot = first(DateForPlot), AverageCumulativeProp = mean(CumulativeProp, na.rm = TRUE)), 
            aes(x = DateForPlot, y = AverageCumulativeProp, color = "1991-2020"), linetype = "dashed") +
  labs(title = "Elk River", x = "", y = "Cumulative Proportion \n of Total Yearly Discharge") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month") +
  scale_color_manual(values = yearColor, labels = c("2021", "2022", "2023", "1991-2020")) +
  theme_few() +
  theme(legend.position = "bottom", legend.title = element_blank()) -> eklCumuQ
eklCumuQ
ggsave("ElkCumuQ.png", eklCumuQ, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")

ggplot() +
  geom_line(data = elkQdaily %>% filter(Year != 2024), aes(x = DateForPlot, y = FlowCMS, group = Year), color = "grey", alpha = 0.5) +
  geom_line(data = elkQdaily %>% filter(Year == 2024), aes(x = DateForPlot, y = FlowCMS, group = Year), color = "black") +
  geom_line(data = elkQdaily, aes(x = DateForPlot, y = MeanQ), color = "red", alpha = 0.5) +
  geom_vline(data = elkQdaily50percAIMS, aes(xintercept = CentroidDate, color = as.factor(Year)), linetype = "dashed") +
  scale_color_brewer(palette = "Dark2", type = "div") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month") +
  labs(x = "",
       y = expression(Discharge~("m"^3*"s"^{-1})),
       color = "50% Cumulative\n    Discharge") +
  theme_few()+
  theme(legend.position = "bottom")

elkLowFlow %>% 
  ggplot(aes(WaterYear, DryDays))+
  geom_point()+
  geom_smooth(method='lm') +
  stat_regline_equation(aes(label = ..rr.label..))+
  theme_few()
```

## Elk River vs. YMR

```{r}
ymrKm2 <- 1.45
elkKm2 <- 569.797384

ymrDat <- gpDat %>% 
  select(Site, dateTime, Q_Ls) %>% 
  filter(Site == "YMR") %>%
  mutate(FlowCMS = Q_Ls * 0.001, .keep = "unused",
         areaKM2 = ymrKm2,
         areaM2 = ymrKm2 * 1e6)

#Get discharge data from NWIS
rawElkQ <- readNWISuv(siteNumbers = "07169800",
                     parameterCd = "00060",
                     startDate = min(as_date(ymrDat$dateTime), na.rm = T), 
                     endDate = max(as_date(ymrDat$dateTime), na.rm = T)) %>% 
  renameNWISColumns() %>% 
  select(-site_no, -agency_cd, -Flow_Inst_cd, -tz_cd) %>% 
  rename(FlowCFS = Flow_Inst) %>% 
  mutate(FlowCMS = FlowCFS * 0.028316847,
         areaM2 = elkKm2 * 1e6,
         areaKM2 = elkKm2,
         FlowScaled = FlowCMS * (ymrKm2/elkKm2),
         Site = "ELK", .keep = "unused")
```

\

```{r}
#Merge YMR and Elk Data
ymrElk <- ymrDat %>% 
  full_join(rawElkQ) %>% 
  group_by(Site) %>% 
  mutate(FlowCMS = na.approx(FlowCMS, maxgap = 10, na.rm = F),
         WaterYear = calcWaterYear(dateTime),
         FlowScaled = coalesce(FlowScaled, FlowCMS)) %>% 
  ungroup() %>% 
  drop_na(FlowCMS)

ymrElk %>% 
  ggplot(aes(x = dateTime, y = FlowScaled, color = Site)) +
  geom_line() +
  facet_wrap(~ WaterYear, scales = "free_x") +
  #scale_y_log10(labels = label_number())+
  scale_x_datetime()+
  scale_color_brewer(palette = "Dark2")+
  labs(x = "",
       y = expression(Discharge~("m"^3*"s"^{-1}/"km"^2))) +
  theme_few()+
    theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_blank(), legend.position = "bottom",
          legend.title = element_blank()) -> ymrElkQScaled
ymrElkQScaled
ggsave("ymrElkQ_Scaled.png", ymrElkQScaled, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")

ymrElk %>% 
  ggplot(aes(x = dateTime, y = FlowCMS, color = Site)) +
  geom_line() +
  facet_wrap(~ WaterYear, scales = "free_x") +
  scale_y_log10(labels = label_number())+
  scale_x_datetime()+
  scale_color_brewer(palette = "Dark2")+
  labs(x = "",
       y = expression(Discharge~("m"^3*"s"^{-1}))) +
  theme_few()+
    theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_blank(), legend.position = "bottom",
          legend.title = element_blank()) -> ymrElkQ
ymrElkQ
ggsave("ymrElkQ.png", ymrElkQ, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")

ymrElkWide <- ymrDat %>% 
  rename(ymrQ = FlowCMS, ymrM2 = areaM2, ymrKM2 = areaKM2) %>% 
  ungroup() %>% 
  dplyr::select(-Site) %>% 
  full_join(rawElkQ %>% rename(elkQ = FlowCMS, elkM2 = areaM2, elkKM2 = areaKM2) %>% select(-Site)) %>% 
  mutate(ymrQ = na.approx(ymrQ, maxgap = 10, na.rm = F),
         elkQ = na.approx(elkQ, maxgap = 10, na.rm = F),
         WaterYear = calcWaterYear(dateTime),
         elkM2 = na_locf(elkM2),
         elkKM2 = na_locf(elkKM2),
         elkQ_Scaled = na.approx(FlowScaled, maxgap = 10, na.rm = F)) %>% 
  filter(!is.na(ymrQ) & !is.na(elkQ) & WaterYear < "2024")

ymrElkWide %>%
  ggplot(aes(x = elkQ, y = ymrQ)) +
  geom_point() +
  labs(x = "Elk River",
       y = "Youngmeyer") +
  geom_smooth(method='lm') +
  stat_regline_equation(aes(label = ..rr.label..))+
  theme_few()+
  facet_wrap(~WaterYear, scales = "free_x") -> ymrElkQBiPlt
ymrElkQBiPlt
ggsave("ymrElkQBiPlt.png", ymrElkQBiPlt, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")

# Cross-Correlation Analysis
ccfRes <- ccf(ymrElkWide$elkQ_Scaled, ymrElkWide$ymrQ)
ccfDf <- data.frame(lag = ccfRes$lag, correlation = ccfRes$acf)
# Find the lag with the highest correlation
maxCorrLag <- ccfDf %>%
  filter(correlation == max(correlation)) %>%
  pull(lag)
cat("Lag with the highest correlation:", maxCorrLag, "\n")

# Interpretation
if (maxCorrLag > 0) {
  cat("elkQ lags behind ymrQ by", maxCorrLag, "time steps.\n")
} else if (maxCorrLag < 0) {
  cat("ymrQ lags behind elkQ by", abs(maxCorrLag), "time steps.\n")
} else {
  cat("ymrQ and elkQ are not correlated at zero lag.\n")
}


# Create a new data frame with lagged YMR discharge
ymrElkWide <- ymrElkWide %>%
    mutate(elkQadj = lag(elkQ_Scaled, maxCorrLag),
           qDiff = elkQadj - ymrQ)

# Scatter Plot with Lagged Data
ymrElkWide %>%
  ggplot(aes(x = elkQadj, y = ymrQ)) +
  geom_point() +
  labs(x = "Elk River",
       y = "Youngmeyer") +
  geom_smooth(method='lm') +
  stat_regline_equation(aes(label = ..rr.label..))+
  theme_few()

rm(ccfDf, ccfRes)
```

# Hydro Metrics

```{r}
gpDat <- gpDat %>% 
  mutate(
    WaterYear = calcWaterYear(dateTime),
    Period = case_when(WaterYear == 2021 ~ "2021",
                       WaterYear == 2022 ~ "2022",
                       WaterYear == 2023 ~ "2023",
                       WaterYear == 2024 ~ "2024")) %>% 
  group_by(Site, Period) %>%
  arrange(desc(SW_Level_cm)) %>%
  mutate(ExceedanceProbability = (row_number() / (n() + 1)) * 100)

gpDat %>% 
  ggplot(aes(x = ExceedanceProbability, y = SW_Level_cm, color = Period)) +
  geom_line() +
  scale_y_log10(labels = label_number()) +
  scale_color_manual(values = c("1991-2020" = "black", "2021" = "#1B9E77", "2022" = "#D95F02", "2023" = "#7570B3", "2024" = "#E7298A")) +
  labs(x = "Exceedance Probability (%)",
       y = "SW Level (cm)") +
  theme_few()+
  theme(legend.position = "bottom", legend.title = element_blank())+
  facet_wrap(~Site)
```

```{r}
dailyQ <- gpDat %>% 
  mutate(Date = ymd(floor_date(solar.time, "1 day"))) %>%
  group_by(Site, Date) %>%
  summarize(Flow = mean(Q_Ls, na.rm = T)) %>% 
  mutate(Flow = replace_na(Flow, 0),
         Flow = ifelse(Flow < 0, 0, Flow)) %>% 
  select(gage = Site, Date, Flow)

for (i in unique(dailyQ$gage)) {
  siteData <- filter(dailyQ, gage == i)
  fileName <- paste0(i, ".csv") # Customize the file name as needed
write.csv(siteData, paste0(here("Data/GP_Q/"), fileName), row.names = FALSE)
}

## Run sub_annual_metrics script
```

```{r}
dryingMet <- read_csv(here("Data/GP_metrics_by_event_Q.csv")) %>% 
  drop_na(event_id) %>% 
  mutate(dryDateStart = ymd(paste0(calendar_year, "-01-01")) + days(dry_date_start - 1),
         Date = dryDateStart,
         dryDateEnd = ymd(Date + days(dry_dur - 1)),
         peakDate = ymd(paste0(calendar_year, "-01-01")) + days(peak_date - 1))

#Merge hydro metrics with ts
hydroTS <- dailyQ %>% 
  full_join(dryingMet, by = c("gage", "Date"))
```

```{r}
hydroTS %>%
  ggplot(aes(Date, Flow)) +
  geom_line(color = "#0084a8") +
  ylab("Q (L/s)") +
  theme_few() +
  geom_rect(aes(xmin = dryDateStart, xmax = dryDateEnd, ymin = -Inf, ymax = +Inf),
    fill = "#d3d3d3",
    alpha = 0.5) +
    geom_rect(aes(xmin = peakDate, xmax = peakDate + 1, ymin = -Inf, ymax = +Inf),
    fill = "#ff0000",
    alpha = 0.8) +
  scale_x_date() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap( ~ gage, scales = "free") -> hydroMetQ
ggsave("hydroMetQ.png", hydroMetQ, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")

```

## Merge with DO Metrics

```{r}
# Expand dryingMet to include all dates between peak_date and dry_date_start
# dryingMet <- dryingMet %>%
#   rowwise() %>%
#   mutate(date_seq = list(seq(peakDate, dryDateStart, by = "day"))) %>%
#   unnest(date_seq) %>%
#   rename(Date = date_seq)
```

## Stage

```{r}
gpDat %>%
  ggplot(aes(dateTime, SW_Level_cm)) +
  geom_line(color = "#000000") +
  ylab("SW Level (cm)") +
  theme_few() +
  scale_x_datetime() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap( ~ Site, scales = "free")

dailyStage <- gpDat %>% 
  mutate(Date = ymd(floor_date(solar.time, "1 day"))) %>%
  group_by(Site, Date) %>%
  summarize(Flow = mean(SW_Level_cm, na.rm = T)) %>% 
  mutate(Flow = replace_na(Flow, 0),
         Flow = ifelse(Flow < 0, 0, Flow)) %>% 
  select(gage = Site, Date, Flow)

for (i in unique(dailyStage$gage)) {
  siteData <- filter(dailyStage, gage == i)
  fileName <- paste0(i, ".csv") # Customize the file name as needed
  write.csv(siteData, paste0(here("Data/GP_Stage/"), fileName), row.names = FALSE)
}
```

```{r}
dryingMetStage <- read_csv(here("Data/GP_metrics_by_event_Stage.csv")) %>% 
  drop_na(event_id) %>%
  mutate(dryDateStart = ymd(paste0(calendar_year, "-01-01")) + days(dry_date_start - 1),
         dryDateEnd = ymd(dryDateStart + days(dry_dur - 1)),
         peakDate = ymd(paste0(calendar_year, "-01-01")) + days(peak_date - 1)) %>% 
    select(-season, -meteorologic_year, -calendar_year, -peak_date, -dry_date_start, -dry_date_mean) 

#Merge hydro metrics with ts
dailyStageDrying <- dailyStage %>% 
  full_join(dryingMetStage %>% mutate(Date = peakDate), by = c("gage", "Date"))

dailyStageDrying %>%
  ggplot(aes(Date, Flow)) +
  geom_line(color = "#0084a8") +
  ylab("Stage (cm)") +
  theme_few() +
  geom_rect(aes(xmin = dryDateStart, xmax = dryDateEnd, ymin = -Inf, ymax = +Inf),
    fill = "#d3d3d3",
    alpha = 0.5) +
    geom_rect(aes(xmin = peakDate, xmax = peakDate + 1, ymin = -Inf, ymax = +Inf),
    fill = "#ff0000",
    alpha = 0.8) +
  scale_x_date() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap( ~ gage, scales = "free") -> hydroMetStage
hydroMetStage
#ggsave("hydroMetStage.png", hydroMetStage, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")

```

```{r}
gpDryMetStage <- dryingMetStage %>% 
  rename(Site = gage) %>% 
  ungroup() %>% 
  drop_na()
#PCA
dryPca <- gpDryMetStage %>% 
  ungroup() %>% 
  select(peak2zero, drying_rate, dry_dur, peak_value, peak_quantile) %>% 
  select(where(is.numeric)) %>% # retain only numeric columns
  prcomp(center = T, scale = T) # do PCA on scaled data

#extract loadings
dryPcaLoad <- dryPca %>%
  tidy(matrix = "loadings")
dryPcaLoad
#Variance explained
dryPcaEig <- dryPca %>%
  tidy(matrix = "eigenvalues")
dryPcaEig
#Scree plot
fviz_eig(dryPca, addlabels = TRUE)

#Qual of representation
fviz_cos2(dryPca, choice = "var", axes = 1)
fviz_cos2(dryPca, choice = "var", axes = 2)

#Biplot
fviz_pca_biplot(dryPca, 
                col.ind = gpDryMetStage$Site,
                palette =  RColorBrewer::brewer.pal(3, "Dark2"), 
                addEllipses = F,
                label = "var",
                col.var = "black",
                repel = TRUE,
                legend.title = "")+
  theme_few()
```

## Merge Metrics W/Stage

```{r}
dryingMetStageExp <- dryingMetStage %>%
  group_by(gage, event_id) %>% 
  rowwise() %>%
  do(data.frame(Date = seq(.$peakDate, .$dryDateStart, by = "day"), .)) %>%
  ungroup() %>%
  rename(Site = gage)

gpMetrics <- doMet %>% 
  full_join(dryingMetStageExp) %>% 
  drop_na() %>% 
  select(-WaterYear, -Year, -season, -minDO.obs, -maxDO.obs, -avgDO.obs, -totalDO, -DO.amp, -probHyp, -cvDO.obs, -dayProbConc, -nightProbConc, -NHR_obs, -SW_Level_ft, -DO.hyp, -DOSat.hyp, -dryDateStart, -dryDateEnd, -peakDate) %>% 
  mutate(WaterYear = calcWaterYear(Date))

## I use this till tidyverse implements a scale function
scale_numeric <- function(x) mutate_if(x, is.numeric, function(y) as.vector(scale(y)))

gpMetrics %>%
  ungroup() %>% 
  select(where(is.numeric)) %>% # retain only numeric columns
  scale_numeric() %>%  # center and scale
  cor(method = "pearson") %>%
  car::scatterplotMatrix()

gpMetrics %>%
  ungroup() %>% 
  select(where(is.numeric)) %>% # retain only numeric columns
  scale_numeric() %>%  # center and scale
  cor(method = "pearson") %>%
  corrplot(type = "upper")
```

```{r}
gpMetrics %>%
  filter(drying_rate < 0) %>% 
  ggplot(aes(peak2zero, avgDO.pctsat)) +
  geom_point() +
  labs(x = "Peak Stage to Zero",
       y = expression(Daily ~ Avg. ~ DO["%Sat"]))+
  geom_smooth(method='lm') +
  stat_regline_equation(aes(label = ..rr.label..))+
  theme_few() -> doDrying
doDrying
ggsave("doDryingLM.png", doDrying, here("Figs/"), device = "png", width = 25.4, height = 14.29, units = "cm", dpi = "retina")
summary(lm(avgDO.pctsat ~ drying_rate, data = gpMetrics %>% filter(drying_rate<0)))
```

```{r}
# Subset data to only include relevant columns
metrics <- gpMetrics %>%
  select(minDO.pctsat, maxDO.pctsat, avgDO.pctsat, DOpctsat.amp, probHypSat, NHR_sat,
         peak_value, peak_quantile, peak2zero, drying_rate, dry_dur) %>% 
  filter(drying_rate < 0)
# Function to create scatter plot with regression line
create_plot <- function(x_var, y_var, data, x_label, y_label) {
  ggplot(data, aes_string(x = x_var, y = y_var)) +
    geom_point() +
    geom_smooth(method = "lm") +
    stat_regline_equation(aes(label = ..rr.label..))+
    labs(x = x_label, y = y_label)
}

# List of DO metrics and drying metrics
do_metrics <- c("minDO.pctsat", "maxDO.pctsat", "avgDO.pctsat", "DOpctsat.amp", "probHypSat", "NHR_sat")
drying_metrics <- c("peak_value", "peak_quantile", "peak2zero", "drying_rate", "dry_dur")

# Initialize an empty dataframe to store significant results
significant_results <- data.frame()

# Loop through each DO metric and drying metric pair
for (do_metric in do_metrics) {
  for (drying_metric in drying_metrics) {
    # Create the formula for the linear model
    formula <- as.formula(paste(do_metric, "~", drying_metric))
    
    # Fit the linear model
    model <- lm(formula, data = metrics)
    
    # Tidy the model results
    tidied_model <- tidy(model)
    
    # Filter for significant results (p-value < 0.05)
    significant <- tidied_model %>%
      filter(term != "(Intercept)" & p.value < 0.05)
    
    # If there are significant results, add them to the dataframe
    if (nrow(significant) > 0) {
      significant <- significant %>%
        mutate(DO_metric = do_metric, Drying_metric = drying_metric)
      significant_results <- rbind(significant_results, significant)
    }
  }
}

# Generate plots for each pair of DO metric and drying metric
plot_list <- list()
for (do_metric in do_metrics) {
  for (drying_metric in drying_metrics) {
    plot_list[[length(plot_list) + 1]] <- create_plot(drying_metric, do_metric, metrics, drying_metric, do_metric)
  }
}

# Arrange plots in a grid
cursedLMs <- grid.arrange(grobs = plot_list, ncol = length(drying_metrics))
```

```{r}
# Subset data to only include relevant columns for clustering
metrics <- gpMetrics %>%
  ungroup() %>% 
  select(minDO.pctsat, maxDO.pctsat, avgDO.pctsat, DOpctsat.amp, probHypSat, NHR_sat,
         peak_value, peak_quantile, peak2zero, drying_rate, dry_dur)

# Determine optimal number of clusters using the elbow method
set.seed(123)
fviz_nbclust(metrics, kmeans, method = "wss")

# Perform K-means clustering with optimal number of clusters (e.g., k = 3)
set.seed(123)
kmeans_result <- kmeans(metrics, centers = 3, nstart = 25)

# Add the cluster assignment to the original data
metrics$cluster <- kmeans_result$cluster
cluster_summary <- metrics %>%
  group_by(cluster) %>%
  summarise(across(everything(), list(mean = mean, sd = sd), na.rm = TRUE))

print(cluster_summary)

# Visualize the clustering results
fviz_cluster(kmeans_result, data = metrics, geom = "point",
             stand = FALSE, ellipse = TRUE, show.clust.cent = TRUE) +
  labs(title = "K-means Clustering of DO and Drying Metrics")

# Reshape data for easier plotting
metrics <- metrics %>%
  mutate(cluster = as.factor(cluster))

metrics_long <- metrics %>%
  pivot_longer(cols = -cluster, names_to = "variable", values_to = "value")

# Create boxplots for each variable by cluster
ggplot(metrics_long, aes(x = cluster, y = value, fill = cluster)) +
  geom_boxplot() +
  facet_wrap(~variable, scales = "free", ncol = 3) +
  theme_minimal() +
  labs(title = "Boxplots of Variables by Cluster", x = "Cluster", y = "Value")
```

```{r}
# Subset data to only include relevant columns for clustering
metrics <- gpMetrics %>%
  ungroup() %>% 
  select(minDO.pctsat, maxDO.pctsat, avgDO.pctsat, DOpctsat.amp, probHypSat, NHR_sat,
         peak_value, peak_quantile, peak2zero, drying_rate, dry_dur)

# Remove rows with missing or infinite values
metrics <- metrics %>%
  filter_all(all_vars(!is.na(.))) %>%
  filter_all(all_vars(!is.infinite(.)))
# Scale the data
metrics_scaled <- scale(metrics)

# Compute the dissimilarity matrix
dist_matrix <- dist(metrics_scaled)

# Perform hierarchical clustering
hc_result <- hclust(dist_matrix, method = "ward.D2")

# Plot the dendrogram
plot(hc_result, labels = FALSE, hang = -1, main = "Hierarchical Clustering Dendrogram")

# Cut the dendrogram to form clusters (e.g., cut into 3 clusters)
k <- 3 # Adjust based on the dendrogram
clusters <- cutree(hc_result, k = k)

# Add the cluster assignment to the original data
metrics <- metrics %>%
  mutate(cluster = as.factor(clusters))

cluster_centroids <- metrics %>%
  group_by(cluster) %>%
  summarise(across(everything(), mean, na.rm = TRUE))

cluster_summary <- metrics %>%
  group_by(cluster) %>%
  summarise(across(everything(), list(mean = mean, sd = sd), na.rm = TRUE))

# Visualize the clustering results using a scatter plot matrix
fviz_cluster(data = metrics, cluster = clusters, geom = "point",
             stand = FALSE, ellipse = TRUE, show.clust.cent = TRUE) +
  labs(title = "Hierarchical Clustering of DO and Drying Metrics")

# Reshape data for easier plotting with ggplot2
metrics_long <- metrics %>%
  pivot_longer(cols = -cluster, names_to = "variable", values_to = "value")

# Create boxplots for each variable by cluster
ggplot(metrics_long, aes(x = cluster, y = value, fill = cluster)) +
  geom_boxplot() +
  facet_wrap(~variable, scales = "free", ncol = 3) +
  theme_minimal() +
  labs(title = "Boxplots of Variables by Cluster", x = "Cluster", y = "Value")

```

# PCA

## DO PCA

```{r}
#PCA
doPca <- gpMetrics %>%
  ungroup() %>% 
  select(minDO.pctsat, maxDO.pctsat, avgDO.pctsat, DOpctsat.amp, probHypSat,  NHR_sat) %>% 
  #m dayProbSat, nightProbSat,
  select(where(is.numeric)) %>% # retain only numeric columns
  prcomp(center = T, scale = T) # do PCA on scaled data

#extract loadings
doPcaLoad <- doPca %>%
  tidy(matrix = "loadings")
doPcaLoad
#Variance explained
doPcaEig <- doPca %>%
  tidy(matrix = "eigenvalues")
doPcaEig
#Scree plot
fviz_eig(doPca, addlabels = TRUE)

#Qual of representation
fviz_cos2(doPca, choice = "var", axes = 1)
fviz_cos2(doPca, choice = "var", axes = 2)

#Biplot
fviz_pca_biplot(doPca, 
                col.ind = gpMetrics$Site,
                palette =  RColorBrewer::brewer.pal(3, "Dark2"), 
                addEllipses = T,
                label = "var",
                col.var = "black",
                repel = TRUE,
                legend.title = "")+
  theme_few()
```

## Drying PCA

```{r}
#PCA
dryPca <- gpMetrics %>% 
  ungroup() %>% 
  select(peak2zero, drying_rate, dry_dur, peak_value, peak_quantile) %>% 
  select(where(is.numeric)) %>% # retain only numeric columns
  prcomp(center = T, scale = T) # do PCA on scaled data

#extract loadings
dryPcaLoad <- dryPca %>%
  tidy(matrix = "loadings")
dryPcaLoad
#Variance explained
dryPcaEig <- dryPca %>%
  tidy(matrix = "eigenvalues")
dryPcaEig
#Scree plot
fviz_eig(dryPca, addlabels = TRUE)

#Qual of representation
fviz_cos2(dryPca, choice = "var", axes = 1)
fviz_cos2(dryPca, choice = "var", axes = 2)

#Biplot
fviz_pca_biplot(dryPca, 
                col.ind = gpMetrics$Site,
                palette =  RColorBrewer::brewer.pal(3, "Dark2"), 
                addEllipses = F,
                label = "var",
                col.var = "black",
                repel = TRUE,
                legend.title = "")+
  theme_few()
```

## USGS Dat

```{r}
usgsMet <- read_csv(here("Data/USGS_AllMetrics.csv")) %>% 
  rename(Site = siteName)

allMetrics <- gpMetrics %>% 
  full_join(usgsMet) %>% 
  select(-Q_Ls, -SW_Level_cm, -cvDO.pctsat, -event_id, -WaterYear)
```

### DO PCA

```{r}
#PCA
allDoPca <- allMetrics %>%
  ungroup() %>% 
  select(minDO.pctsat, maxDO.pctsat, avgDO.pctsat, DOpctsat.amp, probHypSat,  NHR_sat) %>% 
  #m dayProbSat, nightProbSat,
  select(where(is.numeric)) %>% # retain only numeric columns
  prcomp(center = T, scale = T) # do PCA on scaled data

#extract loadings
allDoPcaLoad <- doPca %>%
  tidy(matrix = "loadings")
allDoPcaLoad
#Variance explained
allDoPcaEig <- doPca %>%
  tidy(matrix = "eigenvalues")
allDoPcaEig
#Scree plot
fviz_eig(allDoPca, addlabels = TRUE)

#Qual of representation
fviz_cos2(allDoPca, choice = "var", axes = 1)
fviz_cos2(allDoPca, choice = "var", axes = 2)

#Biplot
fviz_pca_biplot(allDoPca, 
                col.ind = allMetrics$Site,
                palette =  RColorBrewer::brewer.pal(7, "Dark2"), 
                addEllipses = F,
                label = "var",
                col.var = "black",
                repel = TRUE,
                legend.title = "")+
  scale_shape_manual(values = c(0, 1, 2, 3, 4, 5, 6))+
  theme_few()
```

### Dry PCA

```{r}
#PCA
allDryPca <- allMetrics %>% 
  ungroup() %>% 
  select(peak2zero, drying_rate, dry_dur, peak_value, peak_quantile) %>% 
  select(where(is.numeric)) %>% # retain only numeric columns
  prcomp(center = T, scale = T) # do PCA on scaled data

#extract loadings
allDryPcaLoad <- allDryPca %>%
  tidy(matrix = "loadings")
allDryPcaLoad
#Variance explained
allDryPcaEig<- allDryPca %>%
  tidy(matrix = "eigenvalues")
allDryPcaEig
#Scree plot
fviz_eig(allDryPca, addlabels = TRUE)

#Qual of representation
fviz_cos2(allDryPca, choice = "var", axes = 1)
fviz_cos2(allDryPca, choice = "var", axes = 2)

#Biplot
fviz_pca_biplot(allDryPca, 
                col.ind = allMetrics$Site,
                palette =  RColorBrewer::brewer.pal(7, "Dark2"), 
                addEllipses = F,
                label = "var",
                col.var = "black",
                repel = TRUE,
                legend.title = "")+
  scale_shape_manual(values = c(0, 1, 2, 3, 4, 5, 6))+
  theme_few()
```

# Clustering

```{r}
allMetricsScaled <- allMetrics %>%
  select(minDO.pctsat, maxDO.pctsat, avgDO.pctsat, DOpctsat.amp, probHypSat, cvDO.pctsat, dayProbSat, nightProbSat, NHR_sat) %>% 
  ungroup() %>% 
  select(where(is.numeric)) %>% # retain only numeric columns
  scale_numeric()
```

```{r}
fviz_nbclust(allMetricsScaled, kmeans, method = "wss")
gapStat <- clusGap(allMetricsScaled, FUN = kmeans, K.max = 10, B = 100)
fviz_gap_stat(gapStat)

gpKmeans <- kmeans(allMetricsScaled, centers = 6, nstart = 25)
fviz_cluster(gpKmeans, allMetricsScaled)
```

```{r}
# Determine the optimal number of clusters using the Elbow method
fviz_nbclust(allMetricsScaled, kmeans, method = "wss") +
  labs(subtitle = "Elbow method")

# Set the number of clusters (k)
set.seed(123) # For reproducibility
k <- 6  # You can choose the optimal k based on the Elbow method

# Perform k-means clustering
kmeans_result <- kmeans(allMetricsScaled, centers = k, nstart = 25)

# Ungroup the tibble if it is grouped
allMetrics <- allMetrics %>% ungroup()

# Add the cluster assignment to the original data
allMetrics <- allMetrics %>%
  mutate(cluster = kmeans_result$cluster)

# Visualize the clusters using the first two principal components
fviz_cluster(kmeans_result, data = allMetricsScaled,
             geom = "point", ellipse.type = "norm") +
  labs(title = "K-means Clustering",
       x = "Principal Component 1",
       y = "Principal Component 2")

# If you want to visualize clusters based on the original features,
# you can use pairs plot or ggplot for specific feature pairs
ggplot(allMetrics, aes(x = minDO.pctsat, y = maxDO.pctsat, color = as.factor(cluster))) +
  geom_point() +
  labs(title = "Clusters based on minDO.pctsat and maxDO.pctsat",
       x = "Minimum DO % Saturation",
       y = "Maximum DO % Saturation",
       color = "Cluster")
```

```{r}
# Compute the distance matrix
distMatrix <- dist(allMetricsScaled)

# Perform hierarchical clustering using the complete linkage method
hiClust <- hclust(distMatrix, method = "ward.D2")

# Plot the dendrogram
fviz_dend(hiClust, rect = TRUE, show_labels = FALSE) +
  labs(title = "Hierarchical Clustering Dendrogram",
       x = "Observations",
       y = "Height")

# Cut the dendrogram to obtain the desired number of clusters
k <- 5  # You can choose the number of clusters based on the dendrogram
clusterAssignments <- cutree(hiClust, k = k)

# Add the cluster assignment to the original data
allMetrics <- allMetrics %>%
  ungroup() %>%
  mutate(cluster = clusterAssignments)

# Visualize the clusters using the first two principal components
fviz_cluster(list(data = allMetricsScaled, cluster = clusterAssignments),
             geom = "point", ellipse.type = "norm") +
  labs(title = "Hierarchical Clustering",
       x = "Principal Component 1",
       y = "Principal Component 2")

# Visualize clusters based on the original features and add site names
ggplot(allMetrics, aes(x = minDO.pctsat, y = maxDO.pctsat, color = as.factor(cluster))) +
  geom_point() +
  geom_text(aes(label = Site), hjust = 1.5, vjust = 1.5) +
  labs(title = "Clusters based on minDO.pctsat and maxDO.pctsat",
       x = "Minimum DO % Saturation",
       y = "Maximum DO % Saturation",
       color = "Cluster")
```

```{r}

```

```{r}

```

```{r}
fviz_nbclust(allMetricsScaled, kmeans, method = "wss")
gapStat <- clusGap(allMetricsScaled, FUN = kmeans, K.max = 10, B = 100)
fviz_gap_stat(gapStat)

gpKmeans <- kmeans(allMetricsScaled, centers = 6, nstart = 25)

# Compute the distance matrix
distMatrix <- dist(allSiteScaled)

# Perform hierarchical clustering using the complete linkage method
hiClust <- hclust(distMatrix, method = "ward.D2")

# Plot the dendrogram
fviz_dend(hiClust, rect = TRUE, show_labels = FALSE) +
  labs(title = "Hierarchical Clustering Dendrogram",
       x = "Observations",
       y = "Height")

# Cut the dendrogram to obtain the desired number of clusters
k <- 6  # You can choose the number of clusters based on the dendrogram
clusterAssignments <- cutree(hiClust, k = k)

# Add the cluster assignment to the original data
allSiteMet <- allSiteMet %>%
  ungroup() %>%
  mutate(cluster = clusterAssignments)

# Visualize the clusters using the first two principal components
fviz_cluster(list(data = allSiteScaled, cluster = clusterAssignments),
             geom = "point", ellipse.type = "norm") +
  labs(title = "Hierarchical Clustering",
       x = "Principal Component 1",
       y = "Principal Component 2")

# Visualize clusters based on the original features and add site names
ggplot(allSiteMet, aes(x = minDO.pctsat, y = maxDO.pctsat, color = as.factor(cluster))) +
  geom_point() +
  geom_text(aes(label = Site), hjust = 1.5, vjust = 1.5) +
  labs(title = "Clusters based on minDO.pctsat and maxDO.pctsat",
       x = "Minimum DO % Saturation",
       y = "Maximum DO % Saturation",
       color = "Cluster")
```

```{r}
# Aggregate data by site
data_agg <- allSiteMet %>%
  #select(Site, minDO.pctsat, maxDO.pctsat, avgDO.pctsat, DOpctsat.amp, probHypSat,  NHR_sat) %>% 
  #select(Site, peak2zero, drying_rate, dry_dur, peak_value, peak_quantile) %>% 
  select(-Date, -cluster) %>% 
  filter(Site %in% c ("KMZ", "SHN", "YMR")) %>% 
  group_by(Site) %>%
  summarise(across(everything(), mean))

# Transform data for heatmap
heatmap_data <- data_agg %>%
  column_to_rownames("Site")

# Create heatmap
pheatmap::pheatmap(heatmap_data, scale = "row", clustering_distance_rows = "euclidean",
         clustering_distance_cols = "euclidean", clustering_method = "ward.D2",
         color = colorRampPalette(c("blue", "white", "red"))(50))

# Distance matrix
dist_matrix <- dist(heatmap_data)

# Hierarchical clustering
hc <- hclust(dist_matrix, method = "ward.D2")

# Plot dendrogram
plot(hc, labels = rownames(heatmap_data), main = "Dendrogram of Sites")
```
