---
title: "Predict Stream and Benthic Light using StreamLight"
author: "CLB"
format: html
editor: visual
---

```{r}
#| include: false
#| echo: false
library(here)
source(here("Code/paths+packages.R"))
library(future)
library(furrr)
library(purrr)
library(googlesheets4)
#remotes::install_github("psavoy/StreamLightUtils")
#remotes::install_github("psavoy/StreamLight", build_vignettes = T, build_manual = T)
library(StreamLightUtils)
library(StreamLight)
okabeIto <- c("#0072B2", "#E69F00", "#CC79A7", "#009E73", "#56B4E9", "#D55E00", "#F0E442", "#999999", "#000000")

pathDir <- here("Data/streamLight")
```

# Get Data

## AIMS Sites

```{r}
siteInfo <- st_read("~/Dropbox/AIMS/GIS/Results/AIMS_GreatPlains.gpkg", layer = "Sites") |> 
  dplyr::filter(siteId %in% c("SFM01", "SHM01", "ENM01")) |> 
  dplyr::rename(Site_ID = siteId) |> 
  st_transform(crs = 4326) |>
  dplyr::mutate(
    Lon = st_coordinates(geom)[, 1],
    Lat = st_coordinates(geom)[, 2]
  ) |> 
  mutate(
    bank_height = (left_bank_height + right_bank_height) / 2,
    bank_slope  = (left_bank_slope + right_bank_slope) / 2
  )
```

## NLDAS

```{r}
#| include: false
#| echo: false
# #Download NLDAS data
NLDAS_DL_bulk(
    save_dir = file.path(pathDir, "NLDAS"),
    site_locs = siteInfo,
    startDate = "2021-01-01"
  )

#List of successfully downloaded sites
nldasList <- stringr::str_sub(list.files(file.path(pathDir, "NLDAS")), 1, -11)
  
#Processing the downloaded NLDAS data
NLDAS_processed <- StreamLightUtils::NLDAS_proc(read_dir = file.path(pathDir, "NLDAS"), nldasList)
```

## MODIS LAI

```{r}
#Make a table for the MODIS request 
site_locs <- siteInfo |> dplyr::select("Site_ID", "Lat", "Lon")
          
#Export your sites as a .csv for the AppEEARS request  
  write.table(
    site_locs, 
    file.path(pathDir, "sites.csv"), 
    sep = ",", 
    row.names = FALSE,
    quote = FALSE, 
    col.names = FALSE
  )
  
# Unzip MODIS Data
modDat  <- AppEEARS_unpack_QC(
  zip_file = "AIMS_MODIS.zip", 
  zip_dir = file.path(pathDir, "MODIS"), 
  request_sites = site_locs$Site_ID)

modDat <- modDat[names(modDat) != "" & !is.na(names(modDat))]
```

## Process MODIS

```{r}
# List of available fit methods from phenofit
fit_methods <- c("AG", "Beck", "Elmore", "Gu", "Klos", "Zhang")

# Function to calculate evaluation metrics
calculate_metrics <- function(observed, predicted) {
  # Remove NA values
  valid <- !is.na(observed) & !is.na(predicted)
  obs <- observed[valid]
  pred <- predicted[valid]
  
  if (length(obs) == 0) return(list(RMSE = NA, R2 = NA))
  
  rmse <- sqrt(mean((obs - pred)^2))
  r2 <- cor(obs, pred)^2
  
  return(list(RMSE = rmse, R2 = r2))
}

# Set up parallel plan
plan(multisession) 
# Alternative: plan() on Linux/Mac for potentially better performance

# Main parallel processing using furrr
results <- future_map(fit_methods, function(method) {
  # Process data with current method
  modProc <- suppressWarnings(AppEEARS_proc(unpacked_LAI = modDat, fit_method = method, plot = FALSE))
  
  # Calculate metrics for all sites
  site_metrics <- map(names(modProc), function(site_name) {
    site_data <- modProc[[site_name]]
    calculate_metrics(site_data$Lai, site_data$LAI_proc)
  })
  
  # Extract metrics
  all_rmse <- map_dbl(site_metrics, "RMSE") |> na.omit()
  all_r2 <- map_dbl(site_metrics, "R2") |> na.omit()
  
  # Return summary for this method
  list(
    Method = method,
    RMSE_mean = mean(all_rmse),
    R2_mean = mean(all_r2),
    RMSE_sd = sd(all_rmse),
    R2_sd = sd(all_r2)
  )
}, .options = furrr_options(seed = TRUE), .progress = T)  

# Convert results to a data frame
results_df <- map_dfr(results, as.data.frame)

# Find the best method based on lowest RMSE and highest R2
best_method <- results_df |> 
  arrange(RMSE_mean, desc(R2_mean)) |> 
  slice(1) |> 
  pull(Method)

# Clean up parallel workers
plan(sequential)

```

### 

```{r}
MOD_processed <- AppEEARS_proc(
  unpacked_LAI = modDat,  
  fit_method = best_method, 
  plot = T)
```

# Driver Files

```{r}
#Set coordinate reference system with EPSG code, "4326" is WGS84
siteInfo$epsg_crs <- 4326
siteInfo <- siteInfo |> 
  as.data.frame()
#Make driver file for StreamLight
driverFile <- StreamLightUtils::make_driver(site_locs = siteInfo,
            NLDAS_processed = NLDAS_processed,
            MOD_processed = MOD_processed,
            write_output = T,
            save_dir = file.path(pathDir, "driverFiles")
            )
```

# Parameter Files

We need to make a parameter file for streamLight, parameters include:

-   Lat: The site latitude
-   Lon: The site longitude
-   channel_azimuth: Channel azimuth
-   bottom_width: Bottom width (m)
-   BH: Bank height (m)
-   BS: Bank slope
-   WL: Water level (m)
-   TH: Tree height (m)
-   overhang: Maximum canopy overhang (m)
-   overhang_height: Height of the maximum canopy overhang (m). If overhang_height = NA, then the model defaults to a value of 75% of tree height.
-   x_LAD: Leaf angle distribution, default = 1

```{r}
extract_height <- function(Site_ID, Lat, Lon, site_crs){
  #Import the Simard et al. (2011) dataset 
    simard2011 <- raster::raster(paste0(.libPaths()[1], "/StreamLightUtils/data/simard2011.tif"))
  
  # #Import the Simard et al. (2011) dataset if it is not already loaded
  #   if(!exists("simard2011")){data("simard2011", package = "StreamLightUtils")}

  #Defining the min and max values (by default these are not associated with the raster)
    map_ranges <- raster::setMinMax(simard2011)
    
  #Create a simple features object from the site location
    site_location <- data.frame(Lat, Lon)
    
  #Check if the data is in WGS84, if not reproject to WGS84
    if(site_crs == 4326){
      xy_sf <- sf::st_as_sf(site_location, coords = c('Lon', 'Lat'), crs = site_crs)
    } else{
      xy_native <- sf::st_as_sf(site_location, coords = c('Lon', 'Lat'), crs = site_crs)
      xy_sf <- sf::st_transform(xy_native, crs = 4326)
    } #End if else statement    
 
  #Extracting the canopy height at our site
    suppressWarnings(TH <- raster::extract(map_ranges, xy_sf))

  #Bind together the final information
    bound <- setNames(data.frame(Site_ID, Lat, Lon, TH), c("Site_ID", "Lat", "Lon", "TH"))

  return(bound)
} #End extract_height function
```

```{r}
#Get tree height from streamLightUtils
treeHeight <- extract_height(
    Site_ID = siteInfo$Site_ID, 
    Lat = siteInfo$Lat,
    Lon = siteInfo$Lon,
    site_crs = 4326)
setDT(treeHeight)
treeHeight <- treeHeight[, .(Site_ID, TH)]

waterDepth <- aims_read(input_dir = "~/Dropbox/AIMS/GP_SW-GW_QAQC/Data/Processed", 
                        rType = "PRES",
                        sublocation = "SW") |> 
  rename(Site_ID = siteId) |> 
  group_by(Site_ID) |>
  summarise(avg_depth = mean(waterDepth_m, na.rm = TRUE)) |>
  ungroup()


siteInfo <- siteInfo |> 
  left_join(waterDepth) |> 
  left_join(treeHeight)
```

```{r}
batch_model_to_csv <- function(sites, read_dir, save_path) {
  # Initialize an empty list to store all results
  all_results <- list()
  
  # Process each site
  for (Site in sites) {
    tryCatch({
      # Construct the driver file path
      driver_path <- file.path(read_dir, paste0(Site, "_driver.rds"))
      
      # Check if driver file exists
      if (!file.exists(driver_path)) {
        warning(paste("Driver file not found for site:", Site))
        next
      }
      
      # Get the model driver
      driver_file <- readRDS(driver_path)
      
      # Get model parameters for the site
      site_p <- siteInfo[siteInfo$Site_ID == Site, ]
      
      # Run the model
      modeled <- stream_light(
        driver_file, 
        Lat = site_p$Lat, 
        Lon = site_p$Lon,
        channel_azimuth = site_p$azimuth, 
        bottom_width = site_p$bottom_width_m, 
        TH = site_p$TH, 
        WL = site_p$avg_depth,
        BH = site_p$bank_height,
        BS = site_p$bank_slope,
        overhang = site_p$TH * 0.1, #Assume overhang is 10% of Tree Height 
        overhang_height = NA, 
        x_LAD = 1
      )
      
      # Add site ID to results
      modeled$Site_ID <- Site
      
      # Convert all problematic columns to character
      modeled <- modeled |>
        mutate(across(where(~is.list(.) || is.matrix(.) || !is.atomic(.)), 
               ~ sapply(.x, function(x) {
                 if (is.null(x)) return(NA_character_)
                 paste(as.character(x), collapse = "|")
               })))
      
      # Store results
      all_results[[Site]] <- modeled
      
      message("Successfully processed ", Site)
      
    }, error = function(e) {
      warning(paste("Error processing site", Site, ":", e$message))
    })
  }
  
  # Combine all results into one data frame
  combined_results <- bind_rows(all_results)
  
  # Create directory if it doesn't exist
  dir.create(dirname(save_path), showWarnings = FALSE, recursive = TRUE)
  
  # Save to CSV - using data.table::fwrite which handles complex types better
  if (!requireNamespace("data.table", quietly = TRUE)) {
    install.packages("data.table")
  }
  data.table::fwrite(combined_results, save_path)
  
  message("All results saved to: ", save_path)
  
  return(combined_results)
}

# Run the model again
results <- batch_model_to_csv(
  sites = siteInfo$Site_ID,
  read_dir = file.path(pathDir, "driverFiles"),
  save_path = file.path(pathDir, "predLight/all_sites_results.csv")
)
```

```{r}

results |> 
  ggplot(aes(local_time, PAR_surface))+
  geom_line(color = "darkgoldenrod1")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  labs(y = expression(PAR~(mu*"mol m"^-2*"s"^-1)))+
  theme_bw()+
  theme(axis.text = element_text(size = 10), axis.title = element_text(size = 12),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.text = element_text(size = 8))+
  facet_wrap(~Site_ID, ncol = 3, scales = "fixed")
```

```{r}
waterDepth <- aims_read(input_dir = "~/Dropbox/AIMS/GP_SW-GW_QAQC/Data/Processed", 
                        rType = "PRES",
                        sublocation = "SW") |> 
  rename(Site_ID = siteId) |> 
  group_by(Site_ID) |>
  summarise(avg_depth = mean(waterDepth_m, na.rm = TRUE)) |>
  ungroup()
```

# Get Turb

```{r}
# Path to QAQC pipeline
qaqcPath <- here("Data/EXO_QAQC/01_data/04_postQaqc/")
# Site to load data
Sites <- c("SHN", "KMZ", "YMR")
# Read data for each site and combine
gpTurb <- aims_read(input_dir = qaqcPath, siteId = Sites, rType = "EXOS") |> 
  dplyr::filter(alphaQf == 0 & betaQfSd == 0, parameterName == "Turbidity_NTU") |>
  dplyr::select(Site, dateTime = timestamp, Turbidity_NTU = value) |>
  mutate(
    dateTime = ymd_hms(dateTime)
  ) |>
  mutate(date = as_date(dateTime)) |> 
  dplyr::filter(!(Site == "YMR" & dateTime < "2021-07-18"),
         !(Site == "SHN" & dateTime < "2023-03-24"),
         !(Site == "KMZ" & dateTime < "2021-06-10")
         ) |> 
  rename(Site_ID = Site) |> 
  group_by(Site_ID, date) |>
  summarise(turb = mean(Turbidity_NTU, na.rm = TRUE)) |>
  ungroup()

gpTurb <- gpTurb |> 
  mutate(kd_pred = 0.52*log(turb)-0.26 )


```
