---
title: "DO Regimes"
author: "CLB"
format: html
editor: visual
---

```{r}
#| include: false
library(here)
source(here("Code/paths+packages.R"))
library(preMetabolizer)
library(furrr)
library(PerformanceAnalytics)
library(corrr)
library(cluster)
library(factoextra)
library(beepr)
library(glue)
library(cowplot)
library(ggExtra)
library(caret)
library(randomForest)
library(gganimate)
library(gt)
set.seed(4) # For reproducibility
gpSites <- c("#1b9e77", "#d95f02", "#7570b3")
cluster_colors <- c("1" = "#0072B2","2" = "#009E73","3" = "#E69F00")
#cluster_colors <- c("1" = "#4A6B7C","2" = "#5A7A5A","3" = "#B8834A")

# site_colors <- tibble(Site = c("JDR", "GBJ", "KNZ", "SHN", "YMR", "WHR", "PRF", "TAL"),
#                       color = c("#0072B2", "#E69F00", "#CC79A7", "#009E73", "#56B4E9", "#D55E00", "#F0E442", "#999999"))
# site_colors <- tibble(Site = c("JDR", "GBJ", "KNZ", "SHN", "YMR", "WHR", "PRF", "TAL"),
#                       color = c("#006475", "#56CFD8", "#613921", "#F2C27B", "#A77652", "#345022", "#5A6C0A", "#83A102"))
site_colors <- tibble(
    Site = c("JDR", "GBJ", "KNZ", "SHN", "YMR", "WHR", "PRF", "TAL"),
    color = c("#006475", "#56CFD8", "#613921", "#F2C27B", "#A77652", "#345022", "#5A6C0A", "#83A102")
  )
```

```{r}
theme_set(
  theme_bw() +
  theme(text = element_text(size = 20),          # overall text size
    legend.text = element_text(size = 20),   # legend entries
    legend.title = element_text(size = 20),  # legend title
    axis.title.x = element_text(size = 20, face = "bold"),    # axis titles
    axis.title.y = element_text(size = 20, face = "bold"),    # axis titles
    axis.text = element_text(size = 20),     # axis tick labels
    plot.title = element_text(size = 20, hjust = 0.5, face = "bold"),
    legend.position = "bottom",
    plot.tag = element_text(size = 20, face = "bold"),
    strip.background = element_blank())
)
```

# Load Data

```{r}

gpSheds <- st_read("~/Dropbox/AIMS/GIS/Results/AIMS_GreatPlains.gpkg", layer = "Watersheds", quiet = T) |> 
  filter(watershed %in% c("KNZ", "SHN", "YMR")) |> 
  st_transform(crs = 4326)
seSheds <- st_read("~/Dropbox/AIMS/GIS/Results/AIMS_SouthEast.gpkg", layer = "Watersheds") |> 
  st_transform(4326) 
mwSheds <- st_read("~/Dropbox/AIMS/GIS/Results/AIMS_MountainWest.gpkg", layer = "Watersheds") |> 
  st_transform(4326) |> 
  filter(watershed != "DRC")

all_sheds <- gpSheds |> 
  bind_rows(seSheds) |> 
  bind_rows(mwSheds) |> 
  rename(Site = watershed) |>
  st_transform(5070) |> 
  st_point_on_surface() |>
  st_transform(4326) |> 
  mutate(lon = st_coordinates(geom)[, 1],
         lat = st_coordinates(geom)[, 2]) |>
  st_drop_geometry()

gpStreams <- st_read("~/Dropbox/AIMS/GIS/Results/AIMS_GreatPlains.gpkg", layer = "Streams", quiet = T) |> 
  filter(watershed %in% c("KNZ", "SHN", "YMR")) |> 
  st_transform(crs = 4326)
seStreams <- st_read("~/Dropbox/AIMS/GIS/Results/AIMS_SouthEast.gpkg", layer = "Streams", quiet = T) |> 
  mutate(length_m = units::drop_units(st_length(geom))) |> 
  st_transform(4326)
mwStreams <- st_read("~/Dropbox/AIMS/GIS/Results/AIMS_MountainWest.gpkg", layer = "Streams", quiet = T) |> 
  st_transform(4326) |> 
  filter(watershed != "DRC")

stream_lengths <- gpStreams |> 
  bind_rows(seStreams) |> 
  bind_rows(mwStreams) |> 
  st_drop_geometry() |> 
  select(Site = watershed, length_m) |> 
  group_by(Site) |>
  summarise(total_length_m = sum(length_m, na.rm = TRUE))

site_meta <- all_sheds |>
  full_join(stream_lengths) |>
  full_join(site_colors) |>
  as_tibble() |>
  select(Site, color, area_sqkm, total_length_m, lon, lat) |>
  mutate(Site = fct_reorder(Site, lon),
         shape = c(1, 0, 12, 2, 3, 8, 6, 11))

rm(gpSheds, gpStreams, seSheds, seStreams, mwSheds, mwStreams, stream_lengths, all_sheds, site_colors, okabeIto, gpSites, distColors)
```

```{r}
streamLight <- read_csv(here("Data/streamLight/predLight/all_sites_results.csv")) |> 
  rename(Site = Site_ID) |> 
  mutate(
    Date = as_date(local_time),
    Site = case_when(
    Site == "SFM01" ~ "KNZ",
    Site == "SHM01" ~ "SHN",
    Site == "ENM01" ~ "YMR",
    TRUE ~ Site  # keep other values unchanged
  )) |> 
    group_by(Site, Date) |>
  summarise(maxPar = max(PAR_surface, na.rm = TRUE),
            avgPar = mean(PAR_surface, na.rm = TRUE),
            .groups = "drop")

modisET <- read_csv(here("Data/ET/MODIS_Daily_ET.csv")) |> 
  filter(Site != "DRC")

gpDat <- read_csv(here("Data/GP_dailyData.csv")) |>
  select(-c(minDO.obs, maxDO.obs, avgDO.obs, DO.hyp, DO.amp, probHyp,
            dayProbConc, nightProbConc, NHR_obs, dPercCon, wetdry_SW,
            PercWet_1d, dPercWet_1d, ICSL, IntactStreamLength, AlphaCentrality,
            Harary, upstream_length, percentConnected, PercWet_7d, PercWet_14d,
            PercWet_30d, PercWet_60d, dPercWet_7d, dPercWet_14d, dPercWet_30d,
            dPercWet_60d, DOY, DOSat.hyp, cvDO.pctsat, sdDO.pctsat,
            DOSat.slope, dDOpctsat.amp, dMinDO.pctsat, dayProbSat,
            nightProbSat, NHR_sat)) |>
  mutate(Date = as_date(Date)) |>
  left_join(streamLight, by = c("Site", "Date"))

seDat <- read_csv(here("Data/SE_dailyData.csv")) |>
  select(-c(minDO.obs, maxDO.obs, avgDO.obs, DO.hyp, DO.amp, probHyp,
            dayProbConc, nightProbConc, NHR_obs, DOSat.hyp,
             cvDO.pctsat, sdDO.pctsat,
            DOSat.slope, dDOpctsat.amp, dMinDO.pctsat, dayProbSat,
            nightProbSat, NHR_sat)) |>
  mutate(Date = as_date(Date))

jdrDat <- read_csv(here("Data/JDR_dailyData.csv")) |> 
  select(-c(minDO.obs, maxDO.obs, avgDO.obs, DO.hyp, DO.amp, probHyp,
            dayProbConc, nightProbConc, NHR_obs, DOSat.hyp,
             cvDO.pctsat, sdDO.pctsat,
            DOSat.slope, dDOpctsat.amp, dMinDO.pctsat, dayProbSat,
            nightProbSat, NHR_sat)) |>
  mutate(Date = as_date(Date))

gbjDat <- read_csv(here("Data/GBJ_dailyData.csv")) |> 
  select(-c(minDO.obs, maxDO.obs, avgDO.obs, DO.hyp, DO.amp, probHyp, 
            dayProbConc, nightProbConc, NHR_obs, DOSat.hyp,
             dayProbSat, 
            nightProbSat, NHR_sat)) |> 
  mutate(Date = as_date(Date)) 


allDat <- gpDat |> 
  bind_rows(seDat) |> 
  bind_rows(jdrDat) |> 
  bind_rows(gbjDat) |> 
  full_join(modisET) |> 
  select(-c(excProb, DOSat.exc, minDepth, maxDepth, minQ, maxQ, cvDO.pctsat, DOSat.slope, dDOpctsat.amp, sdDO.pctsat, dMinDO.pctsat)) |> 
  mutate(waterYear = as.factor(calcWaterYear(Date)),
         Season = as.factor(get_season(Date)),
         Date = as.Date(Date),
         DOY = as.numeric(format(Date, "%j"))
         ) |> 
  # Fill Q with 0 if na or inf AND depth is 0
  mutate(Q_1d = case_when((is.na(Q_1d) | is.infinite(Q_1d)) & Depth_1d == 0 ~ 0,
                          TRUE ~ Q_1d
                          )) |> 
  select(-totalDO) |> 
    group_by(Site) |> 
    mutate(excProb = calc_exceedance_prob(Q_1d, rm.zero = T) * 100,
           DOSat.exc = calc_exceedance_prob(avgDO.pctsat, rm.zero = T)) |> 
    ungroup() |> 
  mutate(excProb = case_when(is.na(excProb) ~ 0, T ~ excProb))

## New Alpha centrality
sticAC <- read_csv(here("Data/STICs/streamDAG_metrics.csv")) |> 
  rename(Date = date, Site = watershed) |> 
  filter(Site != "DRC") |> 
  select(Site, Date, active_upstream_length, active_network_length_m,  percent_network_wet, a_cent, a_cent_weighted) |> 
  mutate(active_upstream_length = case_when(is.na(active_upstream_length) ~ 0, T~ active_upstream_length),
         percent_network_wet = case_when(is.na(percent_network_wet) ~ 0, T~ percent_network_wet),
         active_network_length_m = case_when(is.na(active_network_length_m) ~ 0, T~ active_network_length_m))

allDat <- allDat |> 
  full_join(sticAC)

# Fill NAs in ASDN with 0 only if between dry periods
allDat <- allDat |>
  arrange(Site, Date) |>
  group_by(Site) |>
  mutate(
    prev_non_na = zoo::na.locf(active_network_length_m, na.rm = FALSE),
    next_non_na = zoo::na.locf(active_network_length_m, fromLast = TRUE, na.rm = FALSE),
    active_network_length_m =
      ifelse(
        is.na(active_network_length_m) &
          prev_non_na == 0 &
          next_non_na == 0,
        0,
        active_network_length_m
      )
  ) |>
  select(-prev_non_na, -next_non_na) |>
  ungroup()

# Define the variables you want to calculate rolling averages for
vars_to_roll <- c("active_upstream_length", "active_network_length_m", 
                  "percent_network_wet", "a_cent", "a_cent_weighted", "avgPar", "water.temp")

# Define rolling window sizes
window_sizes <- c(7, 14, 30, 60)

# List of variables to compute rolling means for
vars <- list(
 # active_upstream_length = "active_upstream_length", 
#  active_network_length_m = "active_network_length_m",
#  percent_network_wet = "percent_network_wet",
#  a_cent = "a_cent",
#  a_cent_weighted = "a_cent_weighted",
  avgPar = "avgPar",
  water.temp = "water.temp"
)
allDat <- allDat |>
  group_by(Site) |>
  arrange(Date)
setDT(allDat)
# Generate rolling means for each variable and window size
for (var in names(vars)) {
  prefix <- vars[[var]]  # Get the appropriate prefix for the variable
  allDat[, (paste0(prefix, "_", window_sizes, "d")) := 
              lapply(window_sizes, function(w) frollmean(get(var), n = w, align = "right", na.rm = TRUE)),
           by = Site]
}

dailyDat <- allDat |>
  filter(Date >= "2021-03-26")
#rm(streamLight, gpDat, seDat, gbjDat, jdrDat, modisET, window_sizes, vars)

n_days <- dailyDat |>
  group_by(Site) |>
  summarise(
    n_days_avgDO = sum(!is.na(avgDO.pctsat)),
    n_days_upstream = sum(!is.na(active_upstream_length)),
    mean_perc_wet = mean(percent_network_wet, na.rm = TRUE),
    .groups = "drop"
  )

site_meta <- site_meta |>
  mutate(name = case_when(
    Site %in% c("KNZ", "SHN", "YMR") ~ c("GP-1", "GP-2", "GP-3")[match(Site, c("KNZ", "SHN", "YMR"))],
    Site %in% c("TAL", "PRF", "WHR") ~ c("EF-1", "EF-2", "EF-3")[match(Site, c("TAL", "PRF", "WHR"))],
    Site %in% c("GBJ", "JDR") ~ c("ND-1", "ND-2")[match(Site, c("GBJ", "JDR"))],
    TRUE ~ NA_character_
  )) |> 
  mutate(name = factor(name, levels = c("GP-1", "GP-2", "GP-3",
                                        "EF-1", "EF-2", "EF-3",
                                        "ND-1", "ND-2")))

site_info <- site_meta |> 
  left_join(n_days) |> 
  mutate(n_years = n_days_upstream/365.25,
         Region = case_when(
    Site %in% c("KNZ","SHN", "YMR") ~ "Great Plains",
    Site %in% c("TAL","WHR", "PRF") ~ "Eastern Temperate Forests" ,
    Site %in% c("GBJ", "JDR") ~ "North American Deserts"
    )) |> 
  arrange(name)

total <- site_info |> 
  summarise(n_days = sum(n_days_upstream))


n_stics <- read_csv(here("Data/n_stics.csv")) |> 
  rename(Site = watershed)

site_info <- site_info |> 
  left_join(n_stics)

avg_ppt <- dailyDat |> 
  group_by(Site) |> 
  summarise(mean_ppt_mm = mean(PPT_1d, na.rm = TRUE)) |> 
  ungroup()

site_info |> 
  left_join(avg_ppt, by = "Site") |> 
  mutate(area_sqkm = case_when(Site == "GBJ" ~ 16.20, T~ area_sqkm)) |> 
  select(Region, Watershed = name, km2 = area_sqkm, n_days_upstream, n_sites, mean_perc_wet) |>
  gt() |> 
  fmt_number(
    columns = c(km2, mean_perc_wet),
    decimals = 2
  ) |> 
  fmt_number(
    columns = c(n_days_upstream),
    decimals = 0
  ) |> 
  cols_label(
    km2 =  "Area ({{km^2}})",
    n_days_upstream = "# of Days",
    mean_perc_wet = "Average % Wet",
    n_sites = "# of STICs",
  ) |>
  cols_align(
    align = "center",
    columns = -Watershed 
  ) |> 
  gtsave(here("Figs/site_info.png"), zoom = 6, vwidth = 2000, vheight = 1200)

site_info |> 
  left_join(avg_ppt, by = "Site") |> 
  mutate(area_sqkm = case_when(Site == "GBJ" ~ 16.20, T~ area_sqkm)) |> 
  select(Region, Watershed = name, km2 = area_sqkm, n_days_upstream, n_sites, mean_perc_wet) |>
  gt() |> 
  fmt_number(
    columns = c(km2, mean_perc_wet),
    decimals = 2
  ) |> 
  fmt_number(
    columns = c(n_days_upstream),
    decimals = 0
  ) |> 
  cols_label(
    km2 =  "Area ({{km^2}})",
    n_days_upstream = "# of Days",
    mean_perc_wet = "Average % Wet",
    n_sites = "# of STICs",
  ) |>
  cols_align(
    align = "center",
    columns = -Watershed 
  ) |> 
  gtsave(here("Figs/site_info.docx"), zoom = 6, vwidth = 2000, vheight = 1200)

```

## Add Name to DailyDat

```{r}
dailyDat <- dailyDat |> 
  left_join(site_meta |> select(Site, name))
```

## Look at hydro between regions

```{r}

flow_prop <- dailyDat |> 
  group_by(Site, name) |> 
  filter(Date >= min(Date[!is.na(Q_1d)], na.rm = TRUE),
         Date <= max(Date[!is.na(Q_1d)], na.rm = TRUE)) |> 
  summarise(
    total_days = n(),
    days_with_flow = sum(Q_1d > 0, na.rm = TRUE),
    prop_flowing = days_with_flow / total_days
  )

flow_prop |> 
  ggplot(aes(x = name, y = prop_flowing, fill = name)) +
  geom_col() +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  labs(
    title = "Proportion of Days with Flow by Site",
    subtitle = "(Days with Q > 0)",
    x = "Site",
    y = "Proportion of Days Flowing"
  ) +
  theme(legend.position = "none")
ggsave(filename = "flowProp.png", path = here("Figs/"),device = "png",width = 20,height = 20,units = "cm", dpi = 600)
q50 <- dailyDat |>
  filter(Q_1d > 0) |> 
  group_by(Site, name) |>
  filter(Date >= min(Date[!is.na(Q_1d)], na.rm = TRUE),
         Date <= max(Date[!is.na(Q_1d)], na.rm = TRUE)) |> 
  summarize(
    days_below_50th = sum(Q_1d < median(Q_1d, na.rm = TRUE), na.rm = TRUE),
    total_days = sum(!is.na(Q_1d)),
    median_Q = median(Q_1d, na.rm = TRUE)
  )

q50 |> 
  ggplot(aes(x = name, y = days_below_50th, fill = name)) +
  geom_col() +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  labs(
    title = "Number of Days with Below-Median Discharge",
    x = "Site",
    y = "Days < 50th Percentile"
  )  +
  theme(legend.position = "none")
ggsave(filename = "q50.png", path = here("Figs/"),device = "png",width = 20,height = 20,units = "cm", dpi = 600)

dailyDat |>
  mutate(Region = case_when(
    Site %in% c("KNZ","SHN", "YMR") ~ "Great Plains",
    Site %in% c("TAL","WHR", "PRF") ~ "Eastern Temperate Forests" ,
    Site %in% c("GBJ", "JDR") ~ "North American Deserts"
    ),
    Region = factor(Region, levels = c("Great Plains", "Eastern Temperate Forests", "North American Deserts"))) |> 
  group_by(Site, name) |> 
  mutate(active_network_length_m = active_network_length_m/max(active_network_length_m, na.rm = T)*100) |>
  ungroup() |> 
  ggplot(aes(x = active_network_length_m, fill = name)) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  labs(
    title = "Distribution of Active Stream \nNetwork Extent by Region",
    x = "ASDN (%)",
    y = "Relative Frequency"
  ) +
  facet_wrap(~ Region, ncol = 1, scales = "free_y") +
  guides(color = guide_legend(title = NULL)) +
  theme(legend.title = element_blank())
ggsave(filename = "asdnPerc.png", path = here("Figs/"),device = "png",width = 20,height = 20,units = "cm", dpi = 600)

dailyDat |>
  mutate(Region = case_when(
    Site %in% c("KNZ","SHN", "YMR") ~ "Great Plains",
    Site %in% c("TAL","WHR", "PRF") ~ "Eastern Temperate Forests" ,
    Site %in% c("GBJ", "JDR") ~ "North American Deserts"
    ),
    Region = factor(Region, levels = c("Great Plains", "Eastern Temperate Forests", "North American Deserts"))) |> 
  group_by(Site, name) |> 
  mutate(active_upstream_length = active_upstream_length/max(active_upstream_length, na.rm = T)*100) |>
  ungroup() |> 
  ggplot(aes(x = active_upstream_length, fill = name)) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  labs(
    title = "Distribution of Active Upstream\nNetwork Extent by Region",
    x = "AUDN (%)",
    y = "Relative Frequency"
  ) +
  facet_wrap(~ Region, ncol = 1, scales = "free_y") +
  guides(color = guide_legend(title = NULL)) +
  theme(legend.title = element_blank())
ggsave(filename = "audnPerc.png", path = here("Figs/"),device = "png",width = 20,height = 20,units = "cm", dpi = 600)

dailyDat |>
  mutate(Region = case_when(
    Site %in% c("KNZ","SHN", "YMR") ~ "Great Plains",
    Site %in% c("TAL","WHR", "PRF") ~ "Eastern Temperate Forests" ,
    Site %in% c("GBJ", "JDR") ~ "North American Deserts"
    ),
    Region = factor(Region, levels = c("Great Plains", "Eastern Temperate Forests", "North American Deserts"))) |> 
  group_by(Site, name) |> 
  ungroup() |> 
  ggplot(aes(x = percent_network_wet, fill = name)) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  labs(
    title = "Distribution of Percent Network Wet by Region",
    x = "% Network Wet",
    y = "Relative Frequency"
  ) +
  facet_wrap(~ Region, ncol = 1, scales = "free_y") +
  guides(color = guide_legend(title = NULL)) +
  theme(legend.title = element_blank())
ggsave(filename = "PercNetWet.png", path = here("Figs/"),device = "png",width = 20,height = 20,units = "cm", dpi = 600)

dailyDat |>
  mutate(Region = case_when(
    Site %in% c("KNZ","SHN", "YMR") ~ "Great Plains",
    Site %in% c("TAL","WHR", "PRF") ~ "Eastern Temperate Forests" ,
    Site %in% c("GBJ", "JDR") ~ "North American Deserts"
    ),
    Region = factor(Region, levels = c("Great Plains", "Eastern Temperate Forests", "North American Deserts"))) |> 
  group_by(Site, name) |> 
  ungroup() |> 
  ggplot(aes(x = a_cent, fill = name)) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  labs(
    title = "Distribution of Alpha Centrality by Region",
    x = "Alpha Centrality",
    y = "Relative Frequency"
  ) +
  facet_wrap(~ Region, ncol = 1, scales = "free_y") +
  guides(color = guide_legend(title = NULL)) +
  theme(legend.title = element_blank())
ggsave(filename = "AlphaCentrality.png", path = here("Figs/"),device = "png",width = 20,height = 20,units = "cm", dpi = 600)

dailyDat |>
  mutate(Region = case_when(
    Site %in% c("KNZ","SHN", "YMR") ~ "Great Plains",
    Site %in% c("TAL","WHR", "PRF") ~ "Eastern Temperate Forests" ,
    Site %in% c("GBJ", "JDR") ~ "North American Deserts"
    ),
    Region = factor(Region, levels = c("Great Plains", "Eastern Temperate Forests", "North American Deserts"))) |> 
  group_by(Site, name) |> 
  ungroup() |> 
  ggplot(aes(x = a_cent_weighted, fill = name)) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  labs(
    title = "Distribution of Weighted Alpha Centrality by Region",
    x = "Weighted Alpha Centrality",
    y = "Relative Frequency"
  ) +
  facet_wrap(~ Region, ncol = 1, scales = "free_y") +
  guides(color = guide_legend(title = NULL)) +
  theme(legend.title = element_blank())
ggsave(filename = "AlphaCentralityWeighted.png", path = here("Figs/"),device = "png",width = 20,height = 20,units = "cm", dpi = 600)


runoff50 <- dailyDat |>
  filter(Q_1d > 0) |> 
  select(Site, name, Date, Q_1d, active_network_length_m) |>
  left_join(
    site_info |>
      select(Site, area_sqkm, total_length_m) |>
      mutate(area_sqkm = case_when(Site == "GBJ" ~ 16.20, T ~ area_sqkm))
  ) |>
  mutate(runoff = Q_1d * 0.0864 / area_sqkm) |> 
  group_by(Site, name) |>
  filter(Date >= min(Date[!is.na(runoff)], na.rm = TRUE),
         Date <= max(Date[!is.na(runoff)], na.rm = TRUE)) |> 
  summarize(
    days_below_50th = sum(runoff < median(runoff, na.rm = TRUE), na.rm = TRUE),
    total_days = sum(!is.na(runoff)),
    median_runoff = median(runoff, na.rm = TRUE)
  )

runoff50 |> 
  ggplot(aes(x = name, y = days_below_50th, fill = name)) +
  geom_col() +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  labs(
    title = "Number of Days with Below-Median Runoff",
    x = "Site",
    y = "Days < 50th Percentile"
  )  +
  theme(legend.position = "none")
ggsave(filename = "runoff50.png", path = here("Figs/"),device = "png",width = 20,height = 20,units = "cm", dpi = 600)

peakDry <- dailyDat |>
  select(Site, name, Date, percent_network_wet) |>
  mutate(
    Year = year(Date),
    water_year = calcWaterYear(Date)
  ) |>
  filter(!is.na(percent_network_wet)) |>
  group_by(Site, name, Year) |>
  summarise(
    peak_date = Date[which.max(percent_network_wet)],
    
    # minimum value AFTER the peak
    min_value = min(percent_network_wet[Date >= peak_date], na.rm = TRUE),
    
    # corresponding date
    min_date = Date[
      percent_network_wet == min_value &
      Date >= peak_date
    ][1],
    
    days_peak_to_min = as.numeric(min_date - peak_date),
    .groups = "drop"
  ) |> 
  filter(
    min_value < 60,      # require minimum < 50%
    days_peak_to_min >= 13,
    Year <= 2024
  )


# Prepare vline data with Region info
vline_dat <- peakDry |>
  mutate(
    Region = case_when(
      Site %in% c("KNZ","SHN","YMR") ~ "Great Plains",
      Site %in% c("TAL","WHR","PRF") ~ "Eastern Temperate Forests",
      Site %in% c("GBJ","JDR") ~ "North American Deserts",
      TRUE ~ NA_character_
    ),
    Region = factor(Region, levels = c("Great Plains", "Eastern Temperate Forests", "North American Deserts"))
  ) |>
  tidyr::pivot_longer(
    cols = c(peak_date, min_date),
    names_to = "type",
    values_to = "date"
  ) |> 
  mutate(type = factor(type, levels = c("peak_date", "min_date")))
# Then add to your plot:
dailyDat |>
  mutate(Region = case_when(
    Site %in% c("KNZ","SHN", "YMR") ~ "Great Plains",
    Site %in% c("TAL","WHR", "PRF") ~ "Eastern Temperate Forests",
    Site %in% c("GBJ", "JDR") ~ "North American Deserts"
  ),
  Region = factor(Region, levels = c("Great Plains", "Eastern Temperate Forests", "North American Deserts"))) |> 
  group_by(Site, name) |> 
  #mutate(active_network_length_m = active_network_length_m/max(active_network_length_m, na.rm = T)*100) |>
  ungroup() |> 
  ggplot(aes(x = Date, y = percent_network_wet, color = name)) +
  geom_line(alpha = 1) +
  # Add vertical lines
  geom_vline(data = vline_dat, 
             aes(xintercept = date, linetype = type), 
             alpha = 1) +
  scale_linetype_manual(
    values = c("min_date" = "dashed", "peak_date" = "solid"),
    labels = c("min_date" = "Minimum", "peak_date" = "Peak")
  ) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  labs(
    title = "Percent Wet Over Time",
    x = "Date",
    y = "Network Wet (%)"
  ) +
  facet_wrap(~ name, scales = "free") +
  guides(color = guide_legend(title = NULL),
         linetype = guide_legend(title = NULL)) +
  theme(legend.title = element_blank())
ggsave(filename = "peak_to_dry_overtime.png", path = here("Figs/"),device = "png",width = 40,height = 20,units = "cm", dpi = 600)
peak_to_dry <- peakDry |> 
  group_by(Site, name) |>
  summarise(
    mean_days_peak_to_min = mean(days_peak_to_min, na.rm = TRUE),
    n_years = n()
  )

peak_to_dry |> 
  ggplot(aes(x = name, y = mean_days_peak_to_min, fill = name)) +
  geom_col() +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  labs(
    title = "Average Time from Peak Wetness to Dry-Down",
    x = "Site",
    y = "Days"
  )  +
  theme(legend.position = "none")
ggsave(filename = "peak_to_dry.png", path = here("Figs/"),device = "png",width = 20,height = 20,units = "cm", dpi = 600)
```

# Exceedance Probability

```{r}
#| warning: false
#| message: false

runoff <- dailyDat |>
  select(Site, name, Date, Q_1d) |> 
  left_join(
    site_info |>
      select(Site, area_sqkm, total_length_m) |>
      mutate(area_sqkm = case_when(Site == "GBJ" ~ 16.20, T ~ area_sqkm))
  ) |>
  group_by(Site, name) |> 
  mutate(runoff_mm = (Q_1d * 86400) / (area_sqkm * 1e6) * 1000,
         runoff_exc = calc_exceedance_prob(runoff_mm, rm.zero = T)) |> 
  ungroup() 
# Runoff Exceedance Prob
runoff_expProb <- runoff |> 
  ggplot(aes(x = runoff_exc, y = runoff_mm, color = name)) +
  geom_line(na.rm = TRUE, linewidth = 2) +
  scale_y_continuous(
    trans = "log10",
    labels = label_number(big.mark = ",")
  ) +
  scale_x_continuous(labels = label_percent(accuracy = 1, scale = 100)) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  labs(
    x = "Runoff Exceedance Probability",
    y = bquote(bold("Runoff (mm day"^{-1}*")"))
  ) +
  guides(color = guide_legend(title = NULL))
ggsave(filename = "runoff_excProb.png",plot = runoff_expProb, path = here("Figs/"),device = "png",width = 20,height = 20,units = "cm", dpi = 600)
runoff_plt <- runoff |> 
  mutate(runoff_mm = if_else(runoff_mm <= 0, NA_real_, runoff_mm)) |> 
  ggplot(aes(x = Date, y = runoff_mm, color = name)) +
  geom_line(na.rm = TRUE) +
  facet_wrap(~ name, scales = "fixed") +
  scale_y_continuous(
    trans = "log10",
    labels = label_number(big.mark = ",")
  ) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  labs(
    x = "",
    y = bquote(bold("Runoff (mm day"^{-1}*")"))
  ) +
  guides(color = guide_legend(title = NULL))
ggsave(filename = "runoff.png",plot = runoff_plt, path = here("Figs/"),device = "png",width = 40,height = 20,units = "cm", dpi = 600)
# Discharge Exceedance Prob
qExcProb <- dailyDat |>
  ggplot(aes(x = excProb, y = Q_1d, color = name)) +
  geom_line(na.rm = TRUE, linewidth = 2) +
  scale_y_continuous(
    trans = "log10",
    labels = label_number(big.mark = ",")
  ) +
  scale_x_continuous(labels = label_percent(accuracy = 1, scale = 1)) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  labs(
    x = "Discharge Exceedance Probability",
    y = bquote(bold("Discharge ("*L~s^{-1}*")"))
  ) +
  guides(color = guide_legend(title = NULL))
qExcProb

# DO Exceedance Prob
doExcProb <- dailyDat |>
  ggplot(aes(x = DOSat.exc, y = avgDO.pctsat, color = name)) +
  geom_line(na.rm = TRUE, linewidth = 2) +
  geom_hline(yintercept = 50, linetype = "dashed", color = "grey60", linewidth = 1.5) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  scale_x_continuous(labels = label_percent(accuracy = 1, scale = 100)) +
  scale_y_continuous(
    breaks = c(0, 25, 50, 75, 100),
  ) +
  labs(
    x = expression(bold("DO"["% Sat"] ~ "Exceedance Probability")),
    y = expression(bold("\nDaily Mean DO"["% Sat"]))
  ) +
  guides(color = guide_legend(title = NULL))
doExcProb
ggsave(filename = "doExcProb.png",plot = doExcProb, path = here("Figs/"),device = "png",width = 20,height = 20,units = "cm", dpi = 600)


excProb_plt <- (runoff_expProb + doExcProb) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
ggsave(filename = "excProb.png",plot = excProb_plt, path = here("Figs/"),device = "png",width = 40,height = 20,units = "cm", dpi = 600)

#rm(doExcProb, qExcProb, combined_plot)
```

## Equation Table

```{r}
# Define the table data
eq_table <- tibble(
  Metric = c(
    "Mean DO",
    "Minimum DO",
    "Maximum DO",
    "Amplitude",
    "Exceedance Probability",
    "Probability of Hypoxia"#,
   # "Night Hypoxia Ratio"
  ),
  
  Equation = c(
    # Daily mean
    "$\\overline{DO}_{\\text{day}} = \\frac{1}{n_d} \\sum_{t=1}^{n_d} DO_{t,d}$",
    # Daily min
    "$DO_{\\min,d} = \\min_t(DO_{t,d})$",
    # Daily max
    "$DO_{\\max,d} = \\max_t(DO_{t,d})$",
    # Daily amplitude
    "$DO_{\\text{amp},d} = \\max_t(DO_{t,d}) - \\min_t(DO_{t,d})$",
    # Exceedance probability
    "$P_{\\text{exc},d} = \\frac{r_d}{N + 1}$",
    # Daily probability of hypoxia
    "$P_{\\text{hypoxia},d} = \\frac{1}{n_d} \\sum_{t=1}^{n_d}
      \\begin{cases}
        1, & \\text{if } DO_{t,d} < DO_{\\text{hyp}} \\\\
        0, & \\text{otherwise}
      \\end{cases}$"#,
  #  # Night hypoxia ratio
  #  "$R_{\\text{hypoxia},d} = \\frac{P_{\\text{hypoxia, night},d}}{P_{\\text{hypoxia, day},d}}$"
  ),
  Temporal_Scale = c(
    "Daily",
    "Daily",
    "Daily",
    "Daily",
    "Across Days",
    "Daily"
  )
)

# Build the gt table
footnote_text <- paste0(
  "Table 2. $DO_{t,d}$ is the dissolved oxygen observation at time $t$ on day $d$; $n_d$ is the number of valid observations per day; $\\overline{DO}_d$ is the daily mean DO; $r_d$ is the rank of $\\overline{DO}_d$ (1 = highest); $N$ is the total number of days; $DO_{\\text{hyp}}$ is the hypoxia threshold."
)

do_eq_table <- eq_table |>
  select(-Temporal_Scale) |> 
  gt() |>
  fmt_markdown(columns = "Equation") |>
  cols_label(
    Metric = "Metric",
    Equation = "Equation",
  ) |>
  #tab_header(title = "Equations for Daily Dissolved Oxygen Metrics") |>
  #tab_source_note(source_note = md(footnote_text)) |> 
  # tab_options(
  #   data_row.padding = px(1),             # tighten vertical spacing
  #   column_labels.padding = px(2),        # tighten header padding
  #   table_body.hlines.width = px(0.5),    # thin horizontal lines
  #   table_body.border.bottom.width = px(0.5),
  # ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "center", columns = Equation) |> 
  tab_options(
  data_row.padding = px(2),          # reduce from default ~8px
  column_labels.padding = px(4),     # tighten header row too
  table.font.names = "Times New Roman"
  )|> 
  #tab_source_note(source_note = md(footnote_text)) |> 
     tab_style(
     style = cell_text(weight = "bold"),    # 👈 bold font weight
     locations = cells_column_labels(columns = c(Metric, Equation))
   ) 
do_eq_table |> 
  gtsave(here("Figs/do_metrics_eqn.pdf"))
do_eq_table |> 
  gtsave(here("Figs/do_metrics_eqn.png"))
do_eq_table |> 
  gtsave(here("Figs/do_metrics_eqn.docx"))

```

# ASDN

```{r}
p <- dailyDat |>
  mutate(name = factor(name, levels = levels(site_meta$name))) |>
  ggplot(aes(
    x = Date,
    y = active_network_length_m / 1000,
    color = percent_network_wet
  )) +
  geom_point(size = 3, alpha = 0.8) +
  facet_wrap(~name, scales = "free_y", ncol = 2) +
  scale_x_date(labels = date_format("%b '%y")) +
  scale_color_viridis_c(option = "plasma", na.value = "grey70", direction = -1) +
  labs(
    x = "",
    y = "ASDN (km)",
    color = "% Network Wet"
  ) 
p
ggsave(filename = "ASDN.png", plot = p, path = here("Figs/Misc/"), device = "png", width = 40, height = 20, units = "cm", dpi = 600)

p <- dailyDat |>
  mutate(name = factor(name, levels = levels(site_meta$name))) |>
  group_by(Site) |> 
  mutate(asdn = (active_network_length_m / max(active_network_length_m, na.rm = TRUE)) * 100) |> 
  ungroup() |> 
  ggplot(aes(x = percent_network_wet, y = asdn)) +
  geom_point(aes(color = name), alpha = 0.7, size = 2.5, na.rm = TRUE) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  scale_x_continuous(labels = label_percent(accuracy = 1, scale = 1)) +
  scale_y_continuous(labels = label_percent(accuracy = 1, scale = 1)) +
  facet_wrap(~name, scales = "fixed", ncol = 2) +
  labs(x = "Network Wet", y = "ASDN", color = "Watershed")
p
ggsave(filename = "percWet_ASDN.png", plot = p, path = here("Figs/Misc/"), device = "png", width = 40, height = 20, units = "cm", dpi = 600)


dailyDat |>
  mutate(name = factor(name, levels = levels(site_meta$name))) |>
  group_by(Site) |> 
  mutate(asdn = (active_network_length_m / max(active_network_length_m, na.rm = TRUE)) * 100) |> 
  ungroup() |> 
  ggplot(aes(
    x = Date,
    y =  asdn,
    color = a_cent_weighted
  )) +
  geom_point(size = 3, alpha = 0.8) +
  facet_wrap(~name, scales = "free_y", ncol = 2) +
  scale_x_date(labels = date_format("%b '%y")) +
  scale_color_viridis_c(option = "plasma", na.value = "grey70", direction = -1) +
  labs(
    x = "",
    y = "ASDN (%)",
    color = "Alpha Centrality (Weighted)"
  ) -> p
p
ggsave(filename = "a_cent_asdn.png", plot = p, path = here("Figs/Misc/"), device = "png", width = 40, height = 20, units = "cm", dpi = 600)

dailyDat |>
  mutate(name = factor(name, levels = levels(site_meta$name))) |>
  group_by(Site) |> 
  mutate(asdn = (active_network_length_m / max(active_network_length_m, na.rm = TRUE)) * 100) |> 
  ungroup() |> 
  ggplot(aes(
    x = Date,
    y =  a_cent_weighted,
    color = percent_network_wet,
  )) +
  geom_point(size = 3, alpha = 0.8) +
  facet_wrap(~name, scales = "free_y", ncol = 2) +
  scale_x_date(labels = date_format("%b '%y")) +
  scale_color_viridis_c(option = "plasma", na.value = "grey70", direction = -1) +
  labs(
    x = "",
    y = "Alpha Centrality (Weighted)",
    color = "Network Wet"
  ) -> p
p
ggsave(filename = "a_cent_perc_wet.png", plot = p, path = here("Figs/Misc/"), device = "png", width = 40, height = 20, units = "cm", dpi = 600)

```

# DO Metrics PCA

```{r}
var_labels <- c(
  avgDO.pctsat   = "Mean",
  DOpctsat.amp   = "Amplitude",
  DOSat.exc      = "Exceedence Probability",
  probHypSat     = "Probability of Hypoxia",
  minDO.pctsat   = "Min",
  maxDO.pctsat   = "Max"
)

# Prepare PCA data
pca_data <- dailyDat |>
  select(name, Date, avgDO.pctsat, DOpctsat.amp, DOSat.exc, probHypSat, minDO.pctsat, maxDO.pctsat) |>
  mutate(probHypSat = case_when(is.nan(probHypSat) ~ 0, TRUE ~ probHypSat))

# Create and fit PCA recipe
pca_recipe <- recipe(~., data = pca_data) |>
  update_role(name, Date, new_role = "Id") |>
  step_naomit(all_predictors()) |>
  step_normalize(all_predictors()) |>
  step_pca(all_predictors(), id = "PCA")

do_pca <- prep(pca_recipe)

# Extract PCA loadings for biplot (PC1–PC3, wide form)
loadings <- do_pca |>
  tidy(id = "PCA") |>
  filter(component %in% c("PC1", "PC2", "PC3")) |>
  pivot_wider(names_from = component, values_from = value) |>
  mutate(var_label = recode(terms, !!!var_labels))

# Calculate arrow scale for biplot
pca_scores <- juice(do_pca)

arrow_scale <- min(
  (max(pca_scores$PC1) - min(pca_scores$PC1) / (max(pca_scores$PC1) - min(pca_scores$PC1))),
  (max(pca_scores$PC2) - min(pca_scores$PC2) / (max(pca_scores$PC2) - min(pca_scores$PC2)))
) * 2

loadings <- loadings |>
  mutate(across(where(is.numeric), ~ . * arrow_scale))

# Extract variance explained
var_explained <- do_pca |>
  tidy(id = "PCA", type = "variance") |>
  filter(terms == "percent variance") |>
  arrange(component)

eigvals <- do_pca |>
  tidy(id = "PCA", type = "variance") |>
  filter(terms == "variance") |>
  mutate(sig = value > 1)

pc1_label <- paste0(
  "PC1 (",
  round(var_explained |> filter(component == 1) |> pull(value), 1),
  "%)"
)

pc2_label <- paste0(
  "PC2 (",
  round(var_explained |> filter(component == 2) |> pull(value), 1),
  "%)"
)

# -------------------------------------------------------------------
# NEW: Table of eigenvalues, % explained, and loadings for PCs with
#      eigenvalues > 1
# -------------------------------------------------------------------

# Combine eigenvalues and percent variance
pc_table <- eigvals |>
  select(component, eigenvalue = value, sig) |>
  left_join(
    var_explained |>
      select(component, pct_explained = value),
    by = "component"
  ) |>
  filter(eigenvalue > 1)

# Long table: one row per variable per retained PC
pca_table <- do_pca |>
  tidy(id = "PCA") |>
  # component is "PC1", "PC2", ...; make numeric to join to pc_table
  mutate(component_num = as.integer(gsub("PC", "", component))) |>
  inner_join(pc_table, by = c("component_num" = "component")) |>
  mutate(
    var_label = recode(terms, !!!var_labels)
  ) |>
  select(
    component,                 # "PC1", "PC2", ...
    eigenvalue,
    pct_explained,
    term = terms,
    var_label,
    loading = value
  ) |>
  arrange(component, desc(abs(loading)))

# pca_table now has eigenvalues, % explained and loadings
# for all PCs with eigenvalues > 1.
# -------------------------------------------------------------------

# Scree plot
do_pca |>
  tidy(id = "PCA", type = "variance") |>
  filter(terms == "percent variance") |>
  ggplot(aes(x = factor(component), y = value)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = paste0(round(value, 1), "%")),
            vjust = -0.5, size = 3) +
  labs(
    title = "Scree Plot",
    x = "Principal Component",
    y = "Percent Variance Explained"
  )

# Variable contribution plot
do_pca |>
  tidy(id = "PCA") |>
  mutate(
    Direction = if_else(value > 0, "Positive", "Negative"),
    terms = tidytext::reorder_within(terms, abs(value), component)
  ) |>
  filter(component %in% c("PC1", "PC2", "PC3", "PC4")) |>
  ggplot(aes(abs(value), terms, fill = Direction)) +
  geom_col() +
  facet_wrap(~component, scales = "free_y", ncol = 2) +
  tidytext::scale_y_reordered() +
  scale_fill_manual(values = c("Negative" = "#D55E00", "Positive" = "#56B4E9")) +
  labs(
    x = "Absolute Value of Contribution",
    y = NULL,
    fill = "Direction"
  ) +
  theme_few() +
  theme(legend.position = "bottom")

# Create PCA biplot
do_pca_plt <- pca_scores |>
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(aes(fill = name, shape = name, color = name),
             size = 3, alpha = 1, stroke = 0.3) +
  stat_ellipse(aes(color = name),
               geom = "polygon", level = 0.95,
               alpha = 0.2, fill = NA) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  # Add loading arrows
  geom_segment(
    data = loadings,
    aes(x = 0, y = 0, xend = PC1, yend = PC2),
    arrow = arrow(length = unit(0.2, "cm")),
    color = "black"
  ) +
  geom_text_repel(
    data = loadings,
    aes(x = PC1 * 1.1, y = PC2 * 1.1, label = var_label),
    size = 5,
    direction = "both"
  ) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  scale_shape_manual(
    values = setNames(site_meta$shape, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  labs(
    x = pc1_label,
    y = pc2_label
  ) +
  theme(
    legend.title = element_blank()
  )

do_pca_plt
# Save main PCA plot
ggsave(filename = "doPCA.png", plot = do_pca_plt, path = here("Figs/"), device = "png", width = 20, height = 20, units = "cm", dpi = 600)

```

```{r}
## 1. Get eigenvalues and % variance for PC1–PC3
pca_var_info <- do_pca |>
  tidy(id = "PCA", type = "variance") |>
  filter(
    component %in% c("1", "2", "3"),
    terms %in% c("variance", "percent variance")
  ) |>
  select(component, terms, value) |>
  pivot_wider(names_from = terms, values_from = value) |>
  mutate(
    eigenvalue = variance,
    pct        = `percent variance`,
    component = case_when(component == 1~"PC1", component == 2~"PC2", component == 3~"PC3", TRUE~NA)
  )

## 2. Build pretty labels like
pc_labels <- pca_var_info |> 
  arrange(component) |> 
  mutate(label = sprintf( "%s\n(λ = %.2f, %.1f%%)", component, eigenvalue, pct )) |> 
  pull(label)
# pc_labels will be a character vector of length 3 # in order: PC1, PC2
## 3. Your loadings table (as before) 
pca_loadings_table <- loadings |> 
  select(Variable = var_label, PC1, PC2, PC3) |> 
  mutate( across(c(PC1, PC2), ~ round(.x, 2)) ) 
## 4. Replace the PC column names with the new labels 
names(pca_loadings_table)[2:4] <- pc_labels 

pca_table_png <- pca_loadings_table |> 
  gt() |>
  tab_header(title = "PCA Loadings",) |> 
  fmt_number(columns = everything(), decimals = 2) |> 
  cols_align(align = "center")



pca_table_png |> 
  gtsave(here("Figs/Supplement/pca_table.png"))
pca_table_png |> 
  gtsave(here("Figs/Supplement/pca_table.docx"))
pca_table_png
```

```{r}
library(plotly)


# Build color vector in Site level order
site_colors <- site_meta |>
  arrange(fct_relevel(name, levels(site_meta$name))) |>
  pull(color) |>
  unname()

# Rescale loadings for 3D arrows
arrow_scale_3d <- min(
  (max(pca_scores$PC1) - min(pca_scores$PC1)) / (max(loadings$PC1) - min(loadings$PC1)),
  (max(pca_scores$PC2) - min(pca_scores$PC2)) / (max(loadings$PC2) - min(loadings$PC2)),
  (max(pca_scores$PC3) - min(pca_scores$PC3)) / (max(loadings$PC3) - min(loadings$PC3))
) * 1.5

loadings_3d <- loadings |>
  mutate(across(starts_with("PC"), ~ . * arrow_scale_3d))

loading_lines <- loadings_3d |>
  rowwise() |>
  do({
    data.frame(
      var_label = .$var_label,
      x = c(0, .$PC1),
      y = c(0, .$PC2),
      z = c(0, .$PC3)
    )
  }) |>
  ungroup()

# 3D PCA plot with forcats-based factor order & colors
plot_ly() |>
  add_trace(
    data = pca_scores,
    x = ~PC1, y = ~PC2, z = ~PC3,
    type = "scatter3d", mode = "markers",
    color = ~fct_relevel(name, levels(site_meta$name), after = 0), colors = site_colors,
    marker = list(size = 4, opacity = 0.7)
  ) |>
  add_trace(
    data = loading_lines,
    x = ~x, y = ~y, z = ~z,
    type = "scatter3d", mode = "lines",
    line = list(color = "black", width = 3),
    split = ~var_label,
    showlegend = FALSE
  ) |>
  add_trace(
    data = loadings_3d,
    x = ~PC1, y = ~PC2, z = ~PC3,
    type = "scatter3d", mode = "text",
    text = ~var_label,
    textposition = "top center",
    showlegend = FALSE
  ) |>
  layout(
    legend = list(title = list(text = "")),
    scene = list(
      xaxis = list(title = paste0("PC1 (", round(var_explained |> filter(component == 1) |> pull(value), 1), "%)")),
      yaxis = list(title = paste0("PC2 (", round(var_explained |> filter(component == 2) |> pull(value), 1), "%)")),
      zaxis = list(title = paste0("PC3 (", round(var_explained |> filter(component == 3) |> pull(value), 1), "%)"))
    ),
    title = "3D PCA of Daily DO Metrics"
  )
```

# K-Means

## Optimal Number of Cluster Selection

```{r}
# nbclust_res <- NbClust::NbClust(data = pca_scores |> select(PC1:PC2) |> as.matrix(),
#                                   distance = "euclidean",
#                                   min.nc = 2, max.nc = 9,
#                                   method = "kmeans",
#                                   index = "all")
# 
# majority_rule <- nbclust_res |>
#   (\(x) {
#     tbl <- table(x$Best.nc["Number_clusters", ])
#     as.numeric(names(tbl)[which.max(tbl)])
#   })()
# 
# # elbow method
# res_ch <- NbClust::NbClust(
#   data = pca_scores |> select(PC1:PC2) |> as.matrix(),
#   distance = "euclidean",
#   min.nc = 3,
#   max.nc = 15,
#   method = "kmeans",
#   index = "ch"
# ) 
# # Silhouette index
# res_sil <- NbClust::NbClust(
#   data = pca_scores |> select(PC1:PC2) |> as.matrix(),
#   distance = "euclidean",
#   min.nc = 3,
#   max.nc = 15,
#   method = "kmeans",
#   index = "silhouette"
# )
# majority_rule <- tibble(
#   method = c("Elbow", "Silhouette"),
#   num_clust = c(res_ch$Best.nc[["Number_clusters"]], res_sil$Best.nc[["Number_clusters"]])) |>
#   (\(df){floor(mean(df$num_clust))})()
majority_rule <- 3
```

```{r}
#K-means clustering
kmeans_result <- pca_scores |> 
  select(PC1:PC2) |> 
  as.matrix() |>
  eclust(
    FUNcluster = "kmeans", # Use K-means clustering
    k = majority_rule, # Number of clusters
    nstart = 25, # Number of random starts for K-means
    graph = F,
    nboot = 500
  )
pca_scores <- pca_scores |>
  mutate(Cluster = as_factor(kmeans_result$cluster)) |> 
  left_join(dailyDat |> select(name, Date, avgDO.pctsat), by = c("name", "Date")) |>
  group_by(Cluster) |>
  mutate(cluster_mean_DO = mean(avgDO.pctsat, na.rm = TRUE)) |>
  ungroup() |>
  mutate(Cluster = as_factor(dense_rank(desc(cluster_mean_DO)))) |>
  select(-cluster_mean_DO)

#Add cluster information back to the original dataset
clustered_dat <- dailyDat |>
  left_join(pca_scores |> select(name, Date, Cluster), by = c("name", "Date")) |>
  mutate(name = fct_relevel(name, levels(site_meta$name))) |> 
  write_csv(here("Data/rf_data.csv"))
```

## Plot K-Means

```{r}
# Summary statistics by cluster
cluster_summary <- clustered_dat |>
  drop_na(Cluster) |> 
  group_by(Cluster) |>
  summarise(
    n_sites = n_distinct(Site),
    n_observations = n(),
    avg_DO_pctsat = mean(avgDO.pctsat, na.rm = TRUE),
    avg_DO_amp = mean(DOpctsat.amp, na.rm = TRUE),
    avg_DO_exc = mean(DOSat.exc, na.rm = TRUE),
    avg_ProbHyp = mean(probHypSat, na.rm = TRUE),
    .groups = "drop"
  )

cli::cli_inform("Cluster summary statistics:")
print(cluster_summary)

kmeans_plt <- ggplot(pca_scores, aes(x = PC1, y = PC2)) +
  geom_point(aes(fill = Cluster, shape = name, color = Cluster), size = 3, alpha = 0.6, stroke = 0.3) +
  stat_ellipse(aes(color = Cluster), geom = "polygon", level = 0.95, alpha = 0.2, fill = NA) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  # Add loading arrows
  geom_segment(data = loadings, 
               aes(x = 0, y = 0, xend = PC1, yend = PC2),
               arrow = arrow(length = unit(0.2, "cm")), 
               color = "black") +
  # Use ggrepel for smart label positioning
  geom_text_repel(data = loadings, 
                  aes(x = PC1 * 1.1, y = PC2 * 1.1, label = var_label),
                  size = 5, 
                  direction = "both") +
  scale_color_manual(values = cluster_colors) +
  scale_fill_manual(values = cluster_colors) +
  #scale_shape(solid = FALSE) +
  scale_shape_manual(
    values = setNames(site_meta$shape, site_meta$name),
    breaks = levels(site_meta$name)
    ) +
  labs(
    x = pc1_label, 
    y = pc2_label,
    shape = "Watershed",
    color = "State",
    fill = "State"
  ) +
  theme(legend.box = "vertical")
kmeans_plt
# Save the plot
ggsave(filename = "kmeansPCA.png", plot = kmeans_plt, path = here("Figs/"), device = "png", width = 20, height = 20, units = "cm", dpi = "retina")
```

```{r}
plot_ly() |>
  # PCA scores as points
  add_trace(
    data = pca_scores,
    x = ~PC1, y = ~PC2, z = ~PC3,
    type = "scatter3d", mode = "markers",
    color = ~Cluster, colors = cluster_colors,
    marker = list(size = 4, opacity = 0.7)
  ) |>
  add_trace(
    data = loading_lines,
    x = ~x, y = ~y, z = ~z,
    type = "scatter3d", mode = "lines",
    line = list(color = "black", width = 3),
    split = ~var_label,
    showlegend = FALSE
  ) |>
  # Add loading labels at arrow tips
  add_trace(
    data = loadings_3d,
    x = ~PC1, y = ~PC2, z = ~PC3,
    type = "scatter3d", mode = "text",
    text = ~var_label,
    textposition = "top center",
    showlegend = FALSE
  ) |>
  layout(
    scene = list(
      xaxis = list(title = paste0("PC1 (", round(var_explained |> filter(component == 1) |> pull(value), 1), "%)")),
      yaxis = list(title = paste0("PC2 (", round(var_explained |> filter(component == 2) |> pull(value), 1), "%)")),
      zaxis = list(title = paste0("PC3 (", round(var_explained |> filter(component == 3) |> pull(value), 1), "%)"))
    ),
    title = "3D PCA of Daily DO Metrics"
  )
rm(arrow_scale, arrow_scale_3d, pc1_label, pc2_label, loading_lines, loadings_3d, loadings)
```

```{r}
do_metrics <- c(
  "avgDO.pctsat" = "Avg.",
  "DOpctsat.amp" = "Amplitude",
  "DOSat.exc" = "Exceedence\nProbability",
  "probHypSat" = "Probability of\nHypoxia",
  "minDO.pctsat" = "Min",
  "maxDO.pctsat" = "Max" 
)
# Reshape data to long format for faceting
metrics <- c("avgDO.pctsat", "DOpctsat.amp", "DOSat.exc", "probHypSat", "minDO.pctsat", "maxDO.pctsat")

plot_data <- clustered_dat |>
  filter(!is.na(Cluster)) |> 
  select(name, Cluster, all_of(names(do_metrics))) |>
  pivot_longer(cols = all_of(names(do_metrics)), 
               names_to = "Metric", 
               values_to = "Value") |> 
  mutate(Metric = recode(Metric, !!!do_metrics))

# Create faceted plot
ggplot(plot_data, aes(x = Cluster, y = Value, fill = Cluster)) +
  geom_boxplot(outliers = FALSE) +
  scale_fill_manual(values = cluster_colors) +
  scale_color_manual(values = cluster_colors) +
  geom_jitter(alpha = 0.1, aes(color = Cluster),
              position = position_jitterdodge(jitter.width = 0.5, dodge.width = 0.75)) +
  facet_wrap(~ Metric, ncol = 3, scales = "free") +
  labs(title = "DO Metrics by Cluster",
       y = "Value",
       x = "Cluster") +
  theme_minimal() +
  theme(legend.position = "none",
        strip.text = element_text(size = 12, face = "bold"))

# Create faceted plot
site_cluster_plt <- plot_data |> 
  ggplot(aes(x = name, y = Value, fill = Cluster)) +
  geom_boxplot(outliers = FALSE) +
  scale_fill_manual(values = cluster_colors) +
  scale_color_manual(values = cluster_colors) +
  geom_jitter(alpha = 0.1, aes(color = Cluster),
              position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.75)) +
  facet_wrap(~ Metric, ncol = 3, scales = "free") +
  labs(title = "DO Metrics by Cluster",
       y = "Value",
       x = "") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
site_cluster_plt
ggsave(filename = "site_cluster_metric.png", plot = site_cluster_plt, path = here("Figs/Supplement"), device = "png", width = 40, height = 20, units = "cm", dpi = "retina")

```

# K-Means Validation

```{r}
library(vegan)
# MRPP Validation
mrpp_data <- pca_scores |> select(PC1:PC2) |> as.matrix()

# Run MRPP test
mrpp_result <- mrpp(
  dat = mrpp_data,
  grouping = pca_scores$Cluster,
  permutations = 1000, 
  distance = "euclidean"
)

# Group by Site instead of clusters
site_mrpp <- mrpp(
  dat = mrpp_data,
  grouping = pca_scores$name, 
  permutations = 1000,
  distance = "euclidean"
)

# Display results
cli::cli_inform(c(
  "v" = "MRPP Cluster Validation Results:",
  " " = "",
  "i" = "Number of clusters: {.val {majority_rule}}",
  "i" = "A-statistic: {.val {round(mrpp_result$A, 4)}}",
  "i" = "P-value: {.val {round(mrpp_result$Pvalue, 4)}}",
  "i" = "Delta (observed): {.val {round(mrpp_result$delta, 4)}}",
  "i" = "Expected delta: {.val {round(mrpp_result$E.delta, 4)}}"
))

# Interpretation
if (mrpp_result$Pvalue < 0.05) {
  if (mrpp_result$A > 0.3) {
    cli::cli_inform(c("v" = "Interpretation: {.strong STRONG} evidence of meaningful cluster separation"))
  } else if (mrpp_result$A > 0.1) {
    cli::cli_inform(c("v" = "Interpretation: {.strong MODERATE} evidence of meaningful cluster separation"))
  } else if (mrpp_result$A > 0.05) {
    cli::cli_inform(c("!" = "Interpretation: {.strong WEAK} but significant cluster separation"))
  } else {
    cli::cli_warn(c("!" = "Interpretation: {.strong SIGNIFICANT} but very weak cluster separation"))
  }
} else {
  cli::cli_warn(c("x" = "Interpretation: {.strong NO} significant evidence of cluster separation"))
}

cli::cli_inform(c(
  "i" = "Testing Site Effect:",
  "i" = "Site-based A-statistic: {.val {round(site_mrpp$A, 4)}}",
  "i" = "Site-based P-value: {.val {round(site_mrpp$Pvalue, 4)}}",
  " " = "",
  "i" = "Cluster-based A-statistic: {.val {round(mrpp_result$A, 4)}}",
  "i" = "Cluster-based P-value: {.val {round(mrpp_result$Pvalue, 4)}}"
))

# Compare effect sizes
if (mrpp_result$A > site_mrpp$A) {
  cli::cli_inform(c("v" = "Clusters explain MORE variance than sites alone"))
} else {
  cli::cli_inform(c("!" = "Sites may be driving the clustering patterns"))
}

```

### Cluster Time-series

```{r}
doDates <- clustered_dat |>
 # filter(Site %in% c("KNZ", "SHN", "YMR")) |> 
  mutate(Year = year(Date)) |> 
  filter(!is.na(avgDO.pctsat)) |>
  group_by(name, Year) |>
  summarise(
    first_date = min(Date),
    last_date = max(Date),
    .groups = "drop"
  ) |> 
  mutate(WinterStart_preY = as.Date(paste0(as.integer(as.character(Year))-1, "-12-01")),
         WinterStart = as.Date(paste0(as.integer(as.character(Year)), "-12-01"))) |> 
  mutate(WinterStart = case_when(name == "GP-3" & Year == "2021"~ as_date("2021-12-20"), T~WinterStart),
         WinterStart_preY = case_when(name == "GP-3" & Year == "2022"~ as_date("2021-12-20"), T~WinterStart_preY)) |> 
  mutate(WinterStart_preY = case_when(name == "GP-2" & Year == "2023" ~ NA, T~WinterStart_preY))


# Create separate data frame for the SHN pre-first-date grey box
preFirstDate_SHN <- doDates |>
  filter(name == "GP-2") |>
  mutate(xmin = as.Date("2021-06-01"),  # or your actual xlim min
         xmax = first_date,
         ymin = -Inf,
         ymax = Inf) |> 
  filter(Year == "2023")


# Create the plot
clustered_dat |> 
  drop_na(Cluster) |> 
  ggplot(aes(x = Date, y = avgDO.pctsat, color = Cluster)) +
  geom_point(size = 3, alpha = 0.8) +
  facet_wrap(~name, scales = "free_y", ncol = 1) +
  scale_color_manual(values = cluster_colors, name = "Cluster") +
  # geom_rect(data = doDates,
  #           aes(xmin = WinterStart_preY, xmax = first_date, ymin = -Inf, ymax = Inf),
  #           inherit.aes = FALSE,
  #           fill = "#42d4f4", alpha = 0.2) +
  # geom_rect(data = doDates,
  #           aes(xmin = last_date, xmax = WinterStart, ymin = -Inf, ymax = Inf),
  #           inherit.aes = FALSE,
  #           fill = "#9A6324", alpha = 0.2) +
  # geom_rect(data = preFirstDate_SHN,
  #         aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
  #         inherit.aes = FALSE,
  #         fill = "grey5", alpha = 0.2) +
  ylim(0,100)+
  scale_x_date(labels = date_format("%b '%y")) +
  xlim(as_date("2021-06-01"), as_date("2024-12-01")) +
  labs(
  x = "",
  y = expression("Daily Average DO"["% Sat"])
  ) +
  theme(axis.text = element_text(size = 15)) -> p 
p
ggsave(filename = "Average.png", plot = p, path = here("Figs/ClusterTimeseries"), device = "png", width = 25.4, height = 25.4, units = "cm", dpi = "retina")

doDates <- doDates |> 
  filter(Year == "2024") |> 
  mutate(WinterStart_preY = as_date("2024-01-01"))

clustered_dat |> 
  filter(year(Date) == 2024) |> 
  filter(!(name == "EF-1" & Date == as_date("2024-01-01"))) |> 
  ggplot(aes(x = Date)) +
  geom_point(aes(y = avgDO.pctsat, color = Cluster), size = 3, alpha = 0.8) +
  facet_wrap(~name, scales = "free_y", ncol = 1) +
  geom_rect(data = doDates,
            aes(xmin = WinterStart_preY, xmax = first_date, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE,
            fill = "#42d4f4", alpha = 0.2) +
  geom_rect(data = doDates,
            aes(xmin = last_date, xmax = WinterStart, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE,
            fill = "#9A6324", alpha = 0.2) +
  scale_color_manual(values = cluster_colors, name = "Cluster", na.translate = FALSE) +
  scale_x_date(labels = date_format("%b"), breaks = "2 months") +
  ylim(0,100)+
  xlim(as_date("2024-01-01"), as_date("2024-12-01")) +
  labs(x = "", y = expression("Daily Average DO"["% Sat"])) +
  theme(axis.text = element_text(size = 15)) -> p 
p
ggsave(filename = "Average_2024.png", plot = p, path = here("Figs/ClusterTimeseries"), device = "png", width = 25.4, height = 25.4, units = "cm", dpi = "retina")

clustered_dat |>
  filter(!(name == "EF-1" & Date == as_date("2024-01-01"))) |> 
  filter(year(Date) == 2024) |> 
  ggplot(aes(x = Date)) +
  geom_point(aes(y = avgDO.pctsat, color = Cluster), size = 3, alpha = 0.8) +
  geom_line(aes(y = percent_network_wet), color = "black", linewidth = 1) +
  facet_wrap(~name, scales = "free_y", ncol = 1, drop = FALSE) +
  geom_rect(data = doDates,
            aes(xmin = WinterStart_preY, xmax = first_date, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE,
            fill = "#42d4f4", alpha = 0.2) +
  geom_rect(data = doDates,
            aes(xmin = last_date, xmax = WinterStart, ymin = -Inf, ymax = Inf),
            inherit.aes = FALSE,
            fill = "#9A6324", alpha = 0.2) +
  scale_color_manual(values = cluster_colors, name = "Cluster", na.translate = FALSE) +
  scale_x_date(labels = date_format("%b"), breaks = "2 months") +
  scale_y_continuous(
    limits = c(0, 100),
    name = expression("Daily Average DO"["% Sat"]),
    sec.axis = sec_axis(~ ., name = "Network Wet (%)")
  ) +
  xlim(as_date("2024-01-01"), as_date("2024-12-01")) +
  labs(x = "") +
  theme(axis.text = element_text(size = 15)) -> p
p
ggsave(filename = "Average_2024_PercentWet.png", plot = p, path = here("Figs/ClusterTimeseries"), device = "png", width = 25.4, height = 25.4, units = "cm", dpi = "retina")
```

## Cluster Time Series for Selected Year

```{r}
clustered_dat |>
  filter(!(name == "EF-1" & Date == as_date("2024-01-01"))) |>
  filter(
    (name %in% c("GP-1", "GP-2", "GP-3", "EF-1") & year(Date) == 2024) |
    (name %in% c("ND-2", "EF-3", "EF-2") & year(Date) == 2023) |
    (name %in% c("ND-1") & year(Date) == 2022)
  ) |> 
  mutate(
    Year = year(Date),
    SiteLabel = paste(name, Year),
    fakeDate = as.Date(paste0("2000-", month(Date), "-", day(Date)))
  ) |> 
  mutate(SiteLabel = factor(SiteLabel, levels = c("GP-1 2024", "GP-2 2024", "GP-3 2024", "EF-1 2024", "EF-2 2023", "EF-3 2023", "ND-1 2022", "ND-2 2023"))) |> 
  ggplot(aes(x = fakeDate)) +
  geom_point(aes(y = avgDO.pctsat, color = Cluster), size = 3, alpha = 0.8) +
  geom_line(aes(y = percent_network_wet), color = "black", linewidth = 1) +
  facet_wrap(~SiteLabel, scales = "fixed", ncol = 2, drop = FALSE) +
  scale_color_manual(values = cluster_colors, name = "State", na.translate = FALSE) +
  xlim(as_date("2000-01-01"), as_date("2000-12-01")) +
  scale_x_date(labels = date_format("%b"), breaks = "1 months", expand = c(0,0)) +
  scale_y_continuous(
    #limits = c(0, 100),
    name = expression(bold("Daily Average DO"["% Sat"])),
    sec.axis = sec_axis(~ ., name = "Network Wet (%)")
  ) +
  labs(x = "") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.text = element_text(size = 12),
        text = element_text(size = 20),          # overall text size
    legend.text = element_text(size = 20),   # legend entries
    legend.title = element_text(size = 20),  # legend title
    axis.title = element_text(size = 20),    # axis titles
    axis.text = element_text(size = 20),     # axis tick labels
    plot.title = element_text(size = 20, hjust = 0.5),
    legend.position = "bottom") -> p

p
ggsave(filename = "Average_YEAR_PercentWet.png", plot = p, path = here("Figs/ClusterTimeseries"), device = "png", width = 40, height = 25.4, units = "cm", dpi = "retina")



year_map <- c(
  "GP-1" = 2024, "GP-2" = 2023, "GP-3" = 2024, 
  "EF-1" = 2024, "EF-2" = 2023, "EF-3" = 2023, 
  "ND-2" = 2023,"ND-1" = 2022
)

clustered_dat |>
  group_by(name) |>
  mutate(active_network_length_m = active_network_length_m / max(active_network_length_m, na.rm = T) * 100) |>
  ungroup() |>
  filter(!(name == "EF-1" & Date == as_date("2024-01-01"))) |>
  mutate(target_year = year_map[as.character(name)]) |>
  filter(year(Date) == target_year) |>
    mutate(
    Date = update(Date, year = 2000),
    name = paste0(name, " (", target_year, ")"),
    name = factor(name, levels = c(
      "GP-1 (2024)", "GP-2 (2023)", "GP-3 (2024)", "EF-1 (2024)",
      "EF-2 (2023)", "EF-3 (2023)", "ND-1 (2022)", "ND-2 (2023)"
    ))
  ) |> 
  ggplot(aes(x = Date)) +
  see::geom_point2(aes(y = avgDO.pctsat, color = Cluster), size = 4, alpha = 0.7) +
  geom_line(aes(y = active_network_length_m), color = "black", linewidth = 1, alpha = 0.7) +
  facet_wrap(~name, scales = "fixed", ncol = 2, drop = FALSE) +
  scale_color_manual(values = cluster_colors, name = "State", na.translate = FALSE) +
  scale_x_date(
    labels = date_format("%b"),
    date_breaks = "1 month",
    limits = c(as_date("2000-01-01"), as_date("2000-09-30")),
    expand = expansion(0, 0)
  ) +
  scale_y_continuous(
    limits = c(0, 100),
    name = expression(bold("Daily Average DO"["% Sat"])),
    sec.axis = sec_axis(~., name = "ASDN (%)")
  ) +
  labs(x = "") +
  guides(color = guide_legend(override.aes = list(size = 4))) +
  theme(
    strip.text = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) -> p
ggsave(filename = "Average_YEAR_ASDN.png", plot = p, path = here("Figs/ClusterTimeseries"), device = "png", width = 40, height = 25.4, units = "cm", dpi = "retina")
```

```{r}
clustered_dat |> 
  drop_na(Cluster) |> 
  ggplot(aes(x = Date, y = DOpctsat.amp, color = Cluster)) +
  geom_point(size = 3, alpha = 0.8) +
  facet_wrap(~name, scales = "free_y", ncol = 1) +
  scale_color_manual(values = cluster_colors, name = "Cluster") +
  ylim(0,100)+
  scale_x_date(labels = date_format("%b '%y")) +
  xlim(as_date("2021-06-01"), as_date("2024-12-01")) +
  labs(
  x = "",
  y = expression("Daily DO"["% Sat"]*" Amplitude")
  ) +
  theme(axis.text = element_text(size = 15)) -> p 
p
ggsave(filename = "Amplitude.png", plot = p, path = here("Figs/ClusterTimeseries"), device = "png", width = 25.4, height = 25.4, units = "cm", dpi = "retina")

clustered_dat |> 
  drop_na(Cluster) |> 
  ggplot(aes(x = Date, y = minDO.pctsat, color = Cluster)) +
  geom_point(size = 3, alpha = 0.8) +
  facet_wrap(~name, scales = "free_y", ncol = 1) +
  scale_color_manual(values = cluster_colors, name = "Cluster") +
  ylim(0,100)+
  scale_x_date(labels = date_format("%b '%y")) +
  xlim(as_date("2021-06-01"), as_date("2024-12-01")) +
  labs(
  x = "",
  y = expression("Daily DO"["% Sat"]*" Minimum")
  ) +
  theme(axis.text = element_text(size = 15)) -> p 
p
ggsave(filename = "Minimum.png", plot = p, path = here("Figs/ClusterTimeseries"), device = "png", width = 25.4, height = 25.4, units = "cm", dpi = "retina")

clustered_dat |> 
  drop_na(Cluster) |> 
  ggplot(aes(x = Date, y = maxDO.pctsat, color = Cluster)) +
  geom_point(size = 3, alpha = 0.8) +
  facet_wrap(~name, scales = "free_y", ncol = 1) +
  scale_color_manual(values = cluster_colors, name = "Cluster") +
  ylim(0,100)+
  scale_x_date(labels = date_format("%b '%y")) +
  xlim(as_date("2021-06-01"), as_date("2024-12-01")) +
  labs(
  x = "",
  y = expression("Daily DO"["% Sat"]*" Maximum")
  ) +
  theme(axis.text = element_text(size = 15)) -> p 
p
ggsave(filename = "Maximum.png", plot = p, path = here("Figs/ClusterTimeseries"), device = "png", width = 25.4, height = 25.4, units = "cm", dpi = "retina")


clustered_dat |> 
  drop_na(Cluster) |> 
  ggplot(aes(x = Date, y = DOSat.exc, color = Cluster)) +
  geom_point(size = 3, alpha = 0.8) +
  facet_wrap(~name, scales = "free_y", ncol = 1) +
  scale_color_manual(values = cluster_colors, name = "Cluster") +
  ylim(0,1)+
  scale_x_date(labels = date_format("%b '%y")) +
  xlim(as_date("2021-06-01"), as_date("2024-12-01")) +
  labs(
  x = "",
  y = expression("DO"["% Sat"]*" Exceedance Probability")
  ) +
  theme(axis.text = element_text(size = 15)) -> p 
p
ggsave(filename = "ExcProb.png", plot = p, path = here("Figs/ClusterTimeseries"), device = "png", width = 25.4, height = 25.4, units = "cm", dpi = "retina")

clustered_dat |> 
  drop_na(Cluster) |> 
  ggplot(aes(x = Date, y = probHypSat, color = Cluster)) +
  geom_point(size = 3, alpha = 0.8) +
  facet_wrap(~name, scales = "free_y", ncol = 1) +
  scale_color_manual(values = cluster_colors, name = "Cluster") +
  ylim(0,1)+
  scale_x_date(labels = date_format("%b '%y")) +
  xlim(as_date("2021-06-01"), as_date("2024-12-01")) +
  labs(
  x = "",
  y = "Probability of Hypoxia"
  ) +
  theme(axis.text = element_text(size = 15)) -> p 
p
ggsave(filename = "ProbHyp.png", plot = p, path = here("Figs/ClusterTimeseries"), device = "png", width = 25.4, height = 25.4, units = "cm", dpi = "retina")

```

```{r}
clustered_dat |>
  group_by(name) |> 
  mutate(headGradient_scaled = scales::rescale(headGradient_1d, to = c(0, 100))) |> 
  ungroup() |> 
  ggplot(aes(x = Date)) +
  geom_point(aes(y = avgDO.pctsat, color = Cluster), size = 1, alpha = 0.8) +
  geom_line(aes(y = headGradient_scaled), color = "black", linewidth = 1) +
  facet_wrap(~name, scales = "free_y", ncol = 1, drop = FALSE) +
  scale_color_manual(values = cluster_colors, name = "Cluster", na.translate = FALSE) +
  scale_x_date(labels = date_format("%b '%y")) +
  scale_y_continuous(
    limits = c(0, 100),
    name = expression("Daily Average DO"["% Sat"]),
    sec.axis = sec_axis(~ ., name = "Head Greadient (Scaled)")
  ) +
  labs(x = "") +
  theme(axis.text = element_text(size = 15)) -> p 
ggsave(filename = "HeadGradient.png", plot = p, path = here("Figs/ClusterTimeseries"), device = "png", width = 25.4, height = 25.4, units = "cm", dpi = "retina")

clustered_dat |>
  group_by(name) |> 
  mutate(active_upstream_length = active_upstream_length/max(active_upstream_length, na.rm = T)*100) |> 
  ungroup() |> 
  ggplot(aes(x = Date)) +
  geom_point(aes(y = avgDO.pctsat, color = Cluster), size = 1, alpha = 0.8) +
  geom_line(aes(y = active_upstream_length), color = "black", linewidth = 1) +
  facet_wrap(~name, scales = "free_y", ncol = 1, drop = FALSE) +
  scale_color_manual(values = cluster_colors, name = "Cluster", na.translate = FALSE) +
  scale_x_date(labels = date_format("%b '%y")) +
  scale_y_continuous(
    limits = c(0, 100),
    name = expression("Daily Average DO"["% Sat"]),
    sec.axis = sec_axis(~ ., name = "Active Upstream Length")
  ) +
  labs(x = "") +
  theme(axis.text = element_text(size = 15)) -> p 
ggsave(filename = "ActiveUpstreamLength.png", plot = p, path = here("Figs/ClusterTimeseries"), device = "png", width = 25.4, height = 25.4, units = "cm", dpi = "retina")


clustered_dat |>
  group_by(name) |> 
  mutate(active_network_length_m = active_network_length_m/max(active_network_length_m, na.rm = T)*100) |> 
  ungroup() |> 
  ggplot(aes(x = Date)) +
  geom_point(aes(y = avgDO.pctsat, color = Cluster), size = 1, alpha = 0.8) +
  geom_line(aes(y = active_network_length_m), color = "black", linewidth = 1, alpha = 0.7) +
  facet_wrap(~name, scales = "fixed", ncol = 1, drop = FALSE) +
  scale_color_manual(values = cluster_colors, name = "State", na.translate = FALSE) +
  scale_x_date(labels = date_format("%b '%y"), 
               date_breaks = "3 months", 
               limits = c(as_date(NA), as_date("2024-12-31")),
               expand = expansion(0, 0)) +
  scale_y_continuous(
    limits = c(0, 100),
    name = expression(bold("Daily Average DO"["% Sat"])),
    sec.axis = sec_axis(~ ., name = "ASDN (%)")
  ) +
  labs(x = "") +
  guides(color = guide_legend(override.aes = list(size = 4))) +
  theme(axis.text = element_text(size = 15),
        strip.text = element_text(face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1)) -> p 
ggsave(filename = "ASDN_Timeseries.png", plot = p, path = here("Figs/ClusterTimeseries"), device = "png", width = 25.4, height = 25.4, units = "cm", dpi = "retina")

clustered_dat |>
  ggplot(aes(x = Date)) +
  geom_point(aes(y = avgDO.pctsat, color = Cluster), size = 1, alpha = 0.8) +
  geom_line(aes(y = percent_network_wet), color = "black", linewidth = 1) +
  facet_wrap(~name, scales = "free_y", ncol = 1, drop = FALSE) +
  scale_color_manual(values = cluster_colors, name = "Cluster", na.translate = FALSE) +
  scale_x_date(labels = date_format("%b '%y")) +
  scale_y_continuous(
    limits = c(0, 100),
    name = expression("Daily Average DO"["% Sat"]),
    sec.axis = sec_axis(~ ., name = "Network Wet (%)")
  ) +
  labs(x = "") +
  theme(axis.text = element_text(size = 15)) -> p 
ggsave(filename = "PercentNetworkWet.png", plot = p, path = here("Figs/ClusterTimeseries"), device = "png", width = 25.4, height = 25.4, units = "cm", dpi = "retina")

clustered_dat |>
  ggplot(aes(x = Date)) +
  geom_point(aes(y = avgDO.pctsat, color = name), size = 1, alpha = 0.8) +
  geom_line(aes(y = percent_network_wet), color = "black", linewidth = 1) +
  facet_wrap(~name, scales = "free_y", ncol = 1, drop = FALSE) +
  scale_color_manual(values = site_colors, name = "Cluster", na.translate = FALSE) +
  scale_x_date(labels = date_format("%b '%y")) +
  scale_y_continuous(
    limits = c(0, 100),
    name = expression("Daily Average DO"["% Sat"]),
    sec.axis = sec_axis(~ ., name = "Network Wet (%)")
  ) +
  labs(x = "") +
  theme(axis.text = element_text(size = 15), legend.position = "none") -> p 
ggsave(filename = "PercentNetworkWet_FinalReport.png", plot = p, path = here("Figs/ClusterTimeseries"), device = "png", width = 25.4, height = 25.4, units = "cm", dpi = "retina")

clustered_dat |>
  group_by(name) |> 
  mutate(a_cent = scales::rescale(a_cent, to = c(0, 100))) |> 
  ungroup() |> 
  ggplot(aes(x = Date)) +
  geom_point(aes(y = avgDO.pctsat, color = Cluster), size = 1, alpha = 0.8) +
  geom_line(aes(y = a_cent), color = "black", linewidth = 1) +
  facet_wrap(~name, scales = "free_y", ncol = 1, drop = FALSE) +
  scale_color_manual(values = cluster_colors, name = "Cluster", na.translate = FALSE) +
  scale_x_date(labels = date_format("%b '%y")) +
  scale_y_continuous(
    limits = c(0, 100),
    name = expression("Daily Average DO"["% Sat"]),
    sec.axis = sec_axis(~ ., name = "Alpha Centrality")
  ) +
  labs(x = "") +
  theme(axis.text = element_text(size = 15)) -> p 
ggsave(filename = "AlphaCentrality.png", plot = p, path = here("Figs/ClusterTimeseries"), device = "png", width = 25.4, height = 25.4, units = "cm", dpi = "retina")
```

## Head Gradient

```{r}
site_plot_data <- clustered_dat |>
  filter(!name %in% c("ND-1", "ND-2")) |> 
  select(name, Date, avgDO.pctsat, headGradient_1d, Cluster) |>
  pivot_longer(cols = c(avgDO.pctsat, headGradient_1d),
               names_to = "Variable",
               values_to = "Value") |>
  mutate(Variable = recode(Variable,
    "avgDO.pctsat" = "Daily Avg DO (%Sat)",
    "headGradient_1d" = "Head Gradient"
  ))

custom_labeller <- labeller(
  Variable = as_labeller(c(
    "Daily Avg DO (%Sat)" = "Daily~Avg.~DO['% Sat']",
    "Head Gradient"       = "Head~Gradient~(m)"
  ), default = label_parsed),
  name = label_value  # Leave Site labels unchanged
)

site_faceted_plot <- site_plot_data |> 
  ggplot(aes(x = Date, y = Value)) +
  geom_point(
    data = filter(site_plot_data, Variable == "Daily Avg DO (%Sat)"),
    aes(color = Cluster), size = 1, alpha = 0.8
  ) +
  geom_line(
    data = filter(site_plot_data, Variable == "Head Gradient"),
    color = "black", linewidth = 1
  ) +
  geom_hline(
    data = filter(site_plot_data, Variable == "Head Gradient") |> distinct(Variable, name),
    aes(yintercept = 0),
    linetype = "dashed", color = "grey50"
  ) +
  facet_grid(Variable ~ name, scales = "free_y", labeller = custom_labeller) +
  scale_color_manual(values = cluster_colors, name = "Cluster", na.translate = FALSE) +
  scale_x_date() +
  labs(x = "", y = "") +
  theme(
    strip.placement = "outside",
    strip.text.y.left = element_text(angle = 0),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

site_faceted_plot

ggsave(
    filename = "HeadGradient.png",
    plot = site_faceted_plot,
    path = here::here("Figs/Supplement/"),
    width = 40, height = 20, units = "cm", dpi = "retina"
  )
```

## Stacked Bars

```{r}
# Create month column and calculate frequencies
cluster_freq <- clustered_dat |>
  mutate(
         Month = factor(lubridate::month(Date, label = T))) |>
  drop_na(Cluster) |> 
  group_by(name, Month, Cluster) |>
  summarise(count = n(), .groups = 'drop') 

monthlyPercWet <- allDat |>
  mutate(name = case_when(
    Site %in% c("KNZ", "SHN", "YMR") ~ c("GP-1", "GP-2", "GP-3")[match(Site, c("KNZ", "SHN", "YMR"))],
    Site %in% c("TAL", "PRF", "WHR") ~ c("EF-1", "EF-2", "EF-3")[match(Site, c("TAL", "PRF", "WHR"))],
    Site %in% c("GBJ", "JDR") ~ c("ND-1", "ND-2")[match(Site, c("GBJ", "JDR"))],
    TRUE ~ NA_character_
  )) |> 
    mutate(name = factor(name, levels = c("GP-1", "GP-2", "GP-3", "EF-1", "EF-2", "EF-3", "ND-1", "ND-2")),
           Month = factor(lubridate::month(Date, label = TRUE))) |>
  group_by(name, Month) |>
  summarise(meanPercWet = mean(percent_network_wet, na.rm = TRUE) , .groups = "drop")

cluster_freq |> 
  filter(!Month %in% c("Oct", "Nov", "Dec", "Jan")) |> 
  ggplot(aes(x = Month, y = count, fill = Cluster)) +
  geom_bar(stat = "identity", position = "fill") +
  geom_point(
    data = monthlyPercWet |> filter(!Month %in% c("Oct", "Nov", "Dec", "Jan")),
    aes(x = Month, y = meanPercWet / 100),  # scale to proportion
    inherit.aes = FALSE,
    size = 3,
    shape = 21,
    fill = "black"
  ) +
  facet_wrap(~name, ncol = 3) +
  scale_fill_manual(values = cluster_colors, name = "Cluster") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  labs(
    title = "",
    x = "",
    y = "Proportion"
  ) +
  theme_few() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    text = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 16),
    plot.title = element_text(size = 20, hjust = 0.5),
    plot.margin = grid::unit(c(0, 0, 0, 0), "lines"),
    legend.position = "bottom"
  ) -> p
p
ggsave(filename = "clusterDist_PercWet.png", plot = p, path = here("Figs/"), device = "png", width = 25.4, height = 25.4, units = "cm", dpi = "retina")


cluster_freq |> 
  #filter(!Month %in% c("Oct", "Nov", "Dec")) |> 
  ggplot(aes(x = Month, y = count, fill = Cluster)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~name, nrow = 2) +
  scale_fill_manual(values = cluster_colors, name = "Cluster") +
  scale_y_continuous(labels = scales::percent, expand = c(0, 0)) +
  theme_minimal() +
  labs(
    title = "",
    x = "",
    y = "Proportion"
  ) +
  theme_few() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    text = element_text(size = 20),          # overall text size
    legend.text = element_text(size = 20),   # legend entries
    legend.title = element_text(size = 20),  # legend title
    axis.title = element_text(size = 20, face = "bold"),    # axis titles
    axis.text = element_text(size = 20),     # axis tick labels
    plot.title = element_text(size = 20, hjust = 0.5),
    plot.margin = grid::unit(c(0, 0, 0, 0), "lines"),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  ) -> p 
p

# Find month/name combos that have no data and add a "No Data" row
all_months <- unique(cluster_freq$Month)  # or use factor levels to preserve order

cluster_freq_plot <- cluster_freq |>
  complete(Month = all_months, name) |>
  group_by(Month, name) |>
  mutate(has_data = any(!is.na(count) & count > 0)) |>
  ungroup() |>
  filter(!is.na(Cluster)) |>  # drop NAs introduced by complete()
  bind_rows(
    cluster_freq |>
      complete(Month = all_months, name) |>
      group_by(Month, name) |>
      filter(all(is.na(count) | count == 0)) |>
      ungroup() |>
      distinct(Month, name) |>
      mutate(Cluster = "No Data", count = 1)
  )

# Add grey to color scale
cluster_colors_plot <- c(cluster_colors, "No Data" = "grey80")
cluster_freq_plot |>
  ggplot(aes(x = Month, y = count, fill = Cluster)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~name, nrow = 2) +
  scale_fill_manual(values = cluster_colors_plot, name = "State") +
  scale_x_discrete(drop = FALSE) +  # keep all months even if still missing
  scale_y_continuous(
    expand = c(0, 0),
    labels = function(x) x * 100,
    breaks = seq(0, 1, by = 0.25)
  ) +
  labs(x = "", y = "Proportion (%)") +
  theme_few() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    text = element_text(size = 20),
    legend.text = element_text(size = 20),
    legend.title = element_text(size = 20),
    axis.title = element_text(size = 20, face = "bold"),
    axis.text = element_text(size = 20),
    plot.title = element_text(size = 20, hjust = 0.5),
    plot.margin = grid::unit(c(0, 0, 0, 0), "lines"),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  ) -> p
p
ggsave(filename = "clusterDist.png", plot = p, path = here("Figs/"), device = "png", width = 40, height = 20, units = "cm", dpi = "retina")
```

## Correlation

```{r}
rf_dat <- clustered_dat |>
  group_by(Site) |> 
  mutate(active_network_length_m = active_network_length_m/max(active_network_length_m, na.rm = T)*100,
         active_upstream_length = active_upstream_length/max(active_upstream_length, na.rm = T)*100) |>  
  ungroup() |> 
  select(-c(avgDO.pctsat, DOSat.exc, DOpctsat.amp, maxDO.pctsat, probHypSat, minDO.pctsat, DOY, waterYear, maxPar)) |> 
  arrange(Date) |>
  mutate(across(
    where(is.numeric),
    ~ na.approx(.x, maxgap = 7, na.rm = FALSE),
    .names = "{.col}"
    ),
    across(.cols = matches("^(Q_)"), ~ scale(.)[, 1]),
    .by = "Site"
  ) |>
  #mutate(across(.cols = matches("^(Depth_)"), ~ scale(.)[, 1])) |> 
  drop_na(Cluster)
  


png(here("Figs/Corr_00.png"), width = 2160, height = 2160)
rf_dat |>
  group_by(Site) |> 
  mutate(across(where(is.numeric), ~ scale(.)[, 1])) |>
  ungroup() |>
  select(where(is.numeric)) |>
  cor(use = "complete.obs") |>
  (\(cor_mat) {
    corrplot(cor_mat, method = "circle", type = "upper", order = "alphabet", tl.col = "black", tl.pos = "lt", tl.cex = 2,
             cl.pos = "n", na.label = " ", diag = TRUE)
    corrplot(cor_mat, method = "number", type = "lower", order = "alphabet", tl.pos = "n", add = TRUE, number.cex = 1.5, na.label = " ", diag = FALSE)
  })()
dev.off()

png(here("Figs/Corr_50.png"), width = 2160, height = 2160)
rf_dat |>
  group_by(Site) |> 
  mutate(across(where(is.numeric), ~ scale(.)[, 1])) |>
  ungroup() |>
  select(where(is.numeric)) |>
  cor(use = "complete.obs") |>
  (\(cor_mat) {
    cor_mat[abs(cor_mat) < 0.5] <- NA
    corrplot(cor_mat, method = "circle", type = "upper", order = "alphabet", tl.col = "black", tl.pos = "lt", tl.cex = 2,
             cl.pos = "n", na.label = " ", diag = TRUE)
    corrplot(cor_mat, method = "number", type = "lower", order = "alphabet", tl.pos = "n", add = TRUE, number.cex = 1.5, na.label = " ", diag = FALSE)
  })()
dev.off()

png(here("Figs/Corr_75.png"), width = 2160, height = 2160)
rf_dat |>
  group_by(Site) |> 
  mutate(across(where(is.numeric), ~ scale(.)[, 1])) |>
  ungroup() |> 
  select(where(is.numeric)) |>
  cor(use = "complete.obs") |>
  (\(cor_mat) {
    cor_mat[abs(cor_mat) < 0.75] <- NA
    corrplot(cor_mat, method = "circle", type = "upper", order = "alphabet", tl.col = "black", tl.pos = "lt", tl.cex = 2,
             cl.pos = "n", na.label = " ", diag = TRUE)
    corrplot(cor_mat, method = "number", type = "lower", order = "alphabet", tl.pos = "n", add = TRUE, number.cex = 1.5, na.label = " ", diag = FALSE)
  })()
dev.off()


png(here("Figs/Corr_70.png"), width = 2160, height = 2160)
rf_dat |>
  group_by(Site) |> 
  mutate(across(where(is.numeric), ~ scale(.)[, 1])) |>
  ungroup() |> 
  select(where(is.numeric)) |>
  cor(use = "complete.obs") |>
  (\(cor_mat) {
    cor_mat[abs(cor_mat) < 0.70] <- NA
    corrplot(cor_mat, method = "circle", type = "upper", order = "alphabet", tl.col = "black", tl.pos = "lt", tl.cex = 2,
             cl.pos = "n", na.label = " ", diag = TRUE)
    corrplot(cor_mat, method = "number", type = "lower", order = "alphabet", tl.pos = "n", add = TRUE, number.cex = 1.5, na.label = " ", diag = FALSE)
  })()
dev.off()
```

## Look at vars to standardize

```{r}
# density plots with means
dat_dens <- clustered_dat |>
  filter(PPT_1d < 1, Q_1d < 5) |>
  select(name, Date, active_network_length_m, PPT_1d, Q_1d, Depth_1d, ET_1d, avgPar, headGradient_1d) |>
  pivot_longer(
    cols = -c(name, Date),
    names_to = "variable",
    values_to = "value"
  )

means_dens <- dat_dens |>
  group_by(name, variable) |>
  summarise(mean_val = mean(value, na.rm = TRUE), .groups = "drop")

ggplot(dat_dens, aes(x = value, fill = name)) +
  geom_density(position = "identity", alpha = 0.5) +
  geom_vline(
    data = means_dens,
    aes(xintercept = mean_val, color = name),
    linetype = "dashed",
    linewidth = 0.6,
    show.legend = FALSE
  ) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  labs(title = "RF Variables") +
  facet_wrap(~variable, scales = "free") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) ->p
ggsave(
    filename = "vars_dens.png",
    plot = p,
    path = here::here("Figs/Supplement/"),
    width = 40, height = 20, units = "cm", dpi = "retina"
  )


# Z-Score transformed density plots 
dat_dens_z <- clustered_dat |> 
  group_by(name) |> 
  mutate(across(where(is.numeric), ~ scale(.)[, 1])) |>
  ungroup() |> 
  select(name, Date, active_network_length_m, PPT_1d, Q_1d, Depth_1d, ET_1d, avgPar, headGradient_1d) |> 
  pivot_longer(
    cols = -c(name, Date),  # keep Site and Date, melt everything else
    names_to = "variable",
    values_to = "value"
  )
means_dens_z <- dat_dens_z |>
  group_by(name, variable) |>
  summarise(mean_val = mean(value, na.rm = TRUE), .groups = "drop")

ggplot(dat_dens_z |> 
         filter(
      (variable == "Q_1d"   & value < 4) |
      (variable == "PPT_1d" & value < 2.5) |
      !(variable %in% c("Q_1d", "PPT_1d"))
    ), aes(x = value, fill = name)) +
  geom_density(position = "identity", alpha = 0.5) +
  geom_vline(
    data = means_dens_z,
    aes(xintercept = mean_val, color = name),
    linetype = "dashed",
    linewidth = 0.6,
    show.legend = FALSE
  ) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  labs(title = "RF Variables Z-Score Transformed") +
  facet_wrap(~variable, scales = "free") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) -> p
ggsave(
    filename = "vars_dens_zscore.png",
    plot = p,
    path = here::here("Figs/Supplement/"),
    width = 40, height = 20, units = "cm", dpi = "retina"
  )

# 0-100 transformed density plots 
dat_dens_scaled <- clustered_dat |>
  select(name, Date, active_network_length_m, PPT_1d, Q_1d, Depth_1d, ET_1d, avgPar, headGradient_1d) |>
  pivot_longer(
    cols = -c(name, Date),
    names_to = "variable",
    values_to = "value"
  ) |>
  group_by(name, variable) |>
  mutate(
    scaled_value = (value - min(value, na.rm = TRUE)) / 
                   (max(value, na.rm = TRUE) - min(value, na.rm = TRUE)) * 100
  ) |>
  ungroup()

means_dens_scaled <- dat_dens_scaled |>
  group_by(name, variable) |>
  summarise(mean_val = mean(scaled_value, na.rm = TRUE), .groups = "drop")

ggplot(dat_dens_scaled, aes(x = scaled_value, fill = name)) +
  geom_density(position = "identity", alpha = 0.5) +
  geom_vline(
    data = means_dens_scaled,
    aes(xintercept = mean_val, color = name),
    linetype = "dashed",
    linewidth = 0.6,
    show.legend = FALSE
  ) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  labs(title = "RF Variables 0-100 Scale") +
  facet_wrap(~variable, scales = "free") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) ->p
ggsave(
    filename = "vars_dens_0-100.png",
    plot = p,
    path = here("Figs/Supplement/"),
    width = 40, height = 20, units = "cm", dpi = "retina"
  )

# Site centered density plots 
dat_dens_cent <- clustered_dat |> 
  select(name, Date, active_network_length_m, PPT_1d, Q_1d, Depth_1d, ET_1d, avgPar, headGradient_1d) |>
  pivot_longer(
    cols = -c(name, Date),
    names_to = "variable",
    values_to = "value"
  ) |>
  group_by(name, variable) |> 
    mutate(
    value = value - mean(value, na.rm = TRUE)
  ) |>
  ungroup()
means_dens_cent <- dat_dens_z |>
  group_by(name, variable) |>
  summarise(mean_val = mean(value, na.rm = TRUE), .groups = "drop")

ggplot(dat_dens_cent |> 
         filter(
      (variable == "Q_1d"   & value < 4) |
      (variable == "PPT_1d" & value < 2.5) |
      !(variable %in% c("Q_1d", "PPT_1d"))
    ), aes(x = value, fill = name)) +
  geom_density(position = "identity", alpha = 0.5) +
  geom_vline(
    data = means_dens_cent,
    aes(xintercept = mean_val, color = name),
    linetype = "dashed",
    linewidth = 0.6,
    show.legend = FALSE
  ) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$name),
    breaks = levels(site_meta$name)
  ) +
  labs(title = "RF Variables Centered") +
  facet_wrap(~variable, scales = "free") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) -> p
ggsave(
    filename = "vars_dens_mean_centered.png",
    plot = p,
    path = here::here("Figs/Supplement/"),
    width = 40, height = 20, units = "cm", dpi = "retina"
  )
```

```{r}


metrics <- c(
  "PPT_7d",
  "Q_1d", "Q_7d", "Q_60d",
  "Depth_1d", "Depth_60d",
  "headGradient_1d", "headGradient_30d",
  "ET_7d",
  "avgPar",
  "active_network_length_m",
  "water.temp"
)

metric_names <- c(
  PPT_7d = "Precip (7 Day Avg.)",
  Q_1d = "Discharge", Q_7d = "Discharge (7 Day Avg.)", Q_60d = "Discharge (60 Day Avg.)",
  Depth_1d = "Depth", Depth_60d = "Depth (60 Day Avg.)",
  headGradient_1d = "Head Gradient", headGradient_30d = "Head Gradient (30 Day Avg.)",
   ET_7d = "ET (7 Day Avg.)",
  avgPar = "Light",
  active_network_length_m = "ASDN",
  water.temp = "Water Temp"
)

# Pivot data long
plot_data <- rf_dat |>
  filter(!is.na(Cluster)) |> 
  select(name, Cluster, any_of(metrics)) |>
  pivot_longer(cols = any_of(metrics), names_to = "Metric", values_to = "Value") |>
  mutate(Metric_label = metric_names[Metric],
         name = factor(name, levels = c("GP-1", "GP-2", "GP-3", "EF-1", "EF-2", "EF-3", "ND-1", "ND-2")))



p <- ggplot(plot_data, aes(x = name, y = Value, fill = Cluster)) +
  geom_boxplot(outlier.shape = NA, position = position_dodge(width = 0.75)) +
  geom_jitter(
    alpha = 0.1,
    aes(color = Cluster),
    position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.75)
    ) +
  facet_wrap(~Metric_label, scales = "free_y") +  # facet by metric
  scale_fill_manual(values = cluster_colors) +
  scale_color_manual(values = cluster_colors) +
  labs(title = "Random Forest Data by Cluster",
    x = "", y = "Value") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

ggsave(
    filename = "rf_data.png",
    plot = p,
    path = here::here("Figs/Supplement/"),
    width = 40, height = 20, units = "cm", dpi = "retina"
  )

```

# Find Best Lag

```{r}
source(here("Code/functions/rf_model.R"))

plan(multisession, workers = parallel::detectCores() - 1)

evaluate_lag <- function(dat, var, lag_k) {
  # Case weights
  cluster_counts <- table(dat$Cluster)
  #cluster_weights <- sqrt(max(cluster_counts) / cluster_counts)
  cluster_weights <- c("1" = .2, "2" = .3, "3" = .5)
  roll_var <- paste0(var, "_roll", lag_k)

  dat_roll <- dat |>
    group_by(Site, name) |>
    arrange(Date) |>
    mutate(
      "{roll_var}" := slider::slide_dbl(.data[[var]],
                                .f = mean,
                                .before = lag_k,
                                .after = 0,
                                .complete = FALSE)
    ) |>
    ungroup() |>
    filter(!is.na(.data[[roll_var]]),
           !is.na(Cluster)) |>
    mutate(
      case_wt = case_when(
        Cluster == "1" ~ cluster_weights["1"],
        Cluster == "2" ~ cluster_weights["2"],
        Cluster == "3" ~ cluster_weights["3"]
      ),
      case_wt = importance_weights(case_wt)
    ) |> 
    select(Cluster, all_of(roll_var), case_wt) 


  folds <- vfold_cv(dat_roll, v = 5, strata = "Cluster")

  # Recipe
  rec <- recipe(Cluster ~ ., data = dat_roll) |>
    step_zv()
    
  # RF Spec
  model_spec <- rand_forest(trees = 500) |>
    set_engine("ranger", seed = 42, num.threads = 1) |>
    set_mode("classification")
  # Workflow
  wf <- workflow() |>
    add_model(model_spec) |>
    add_recipe(rec) |> 
    add_case_weights(case_wt)

  # Results
  res <- wf |>
    fit_resamples(folds, metrics = metric_set(kap, f_meas, accuracy, roc_auc),
                  control = control_resamples(save_pred = TRUE))

  res |>
    collect_metrics() |>
    mutate(variable = var, lag = lag_k)
}


vars <- c("water.temp", "Q", "Depth", "headGradient",
          "PPT", "avgPar", "ET", "active_network_length_m")

max_lag <- 60

grid <- tidyr::expand_grid(
  variable = vars,
  lag = 0:max_lag
)

lagged_data <- dailyDat |>
  as_tibble() |>
  # group_by(Site) |>
  # mutate(active_network_length_m = active_network_length_m/max(active_network_length_m, na.rm = T)*100,
  #        active_upstream_length = active_upstream_length/max(active_upstream_length, na.rm = T)*100) |>
  # ungroup() |>
  select(-ends_with(c("_7d", "_14d", "_30d", "_60d"))) |>
  rename_with(~ gsub("_1d$", "", .x), ends_with("_1d")) |>
  select(-c("active_upstream_length", "a_cent", "a_cent_weighted", "Season", "DOY", "excProb",
            "DOSat.exc", "air.temp", "minDO.pctsat", "maxDO.pctsat", "avgDO.pctsat", "DOpctsat.amp", "probHypSat",
            "wetQf", "maxPar", "waterYear", "percent_network_wet")) |>
  left_join(pca_scores |> select(name, Date, Cluster), by = c("name", "Date")) |>
  relocate(c("name", "Cluster"), .after = Site) |>
  arrange(Site, Date)


# scale data
lagged_data_scaled <- lagged_data |>
  group_by(Site) |>
  # # mutate(across(
  # #   all_of(c("active_network_length_m", "avgPar", "Depth", "Q", "ET", "PPT")),
  # #   ~ ( (.x - min(.x, na.rm = TRUE)) /
  # #       (max(.x, na.rm = TRUE) - min(.x, na.rm = TRUE)) ) * 100
  # # )) |>
  mutate(active_network_length_m = active_network_length_m/max(active_network_length_m, na.rm = T)*100)# |>
  # mutate(
  #   across(all_of(c("avgPar", "Depth", "Q", "ET", "PPT")),
  #          ~ .x - mean(.x, na.rm = TRUE))
  # ) |>
  #  # mutate(across(all_of(c("active_network_length_m", "avgPar", "Depth", "Q", "ET", "PPT")), ~ scale(.)[, 1])) |>
  # ungroup() |>
  # group_by(Site) |>
  # mutate(across(where(is.numeric), ~ as.numeric(scale(.x)))) |>
  # ungroup()

lag_results <- future_map_dfr(
  seq_len(nrow(grid)),
  function(i) {
    evaluate_lag(
      dat = lagged_data_scaled,
      var = grid$variable[i],
      lag_k = grid$lag[i]
    )
  },
  .options = furrr_options(seed = TRUE)
)
plan(sequential)
```

```{r}
best_points <- lag_results |>
  group_by(variable, .metric) |>
  #mutate(mean = round(mean, 2)) |>
  mutate(
         best_mean = max(mean, na.rm = TRUE)) |>
  filter(mean >= best_mean *.97) |>
  slice_min(lag, n = 1) |>
  ungroup() |>
  select(-best_mean) |>
  mutate(variable = case_when(variable %in% names(var_names) ~ var_names[variable],TRUE ~ variable),
         .metric = case_when(.metric == "accuracy" ~ "Accuracy",
                             .metric == "f_meas" ~ "F1",
                             .metric == "kap" ~ "Kappa",
                             .metric ==  "roc_auc" ~ "ROC AUC"))


lag_results |>
  mutate(variable = case_when(variable %in% names(var_names) ~ var_names[variable],TRUE ~ variable),
         .metric = case_when(.metric == "accuracy" ~ "Accuracy",
                             .metric == "f_meas" ~ "F1",
                             .metric == "kap" ~ "Kappa",
                             .metric ==  "roc_auc" ~ "ROC AUC")) |>
  #filter(.metric %in% c("F1")) |>
  ggplot(aes(x = lag, y = mean, color = variable)) +
  geom_line() +
  see::scale_color_okabeito() +
  # geom_point(
  #   data = best_points,#|> filter(.metric %in% c("F1")),
  #   aes(x = lag, y = mean),
  #   color = "red",
  #   alpha = 0.5,
  #   show.legend = FALSE
  # ) +
  #   geom_text_repel(
  #   data = best_points,#|> filter(.metric %in% c("F1")),
  #   aes(x = lag, y = mean, label = lag),
  #   vjust = -0.6,
  #   color = "red",
  #   size = 4,
  #   show.legend = FALSE
  # ) +
  facet_grid(.metric ~ variable, scales = "free_y") +
  labs(
    title = "Performance Across Metrics",
    subtitle = "Higher is better",
    x = "Days Prior",
    y = "Performance",
    color = ""
  ) +
  theme_bw() +
  theme(    text = element_text(size = 18),          # overall text size
    legend.text = element_text(size = 16),   # legend entries
    legend.title = element_text(size = 18),  # legend title
    axis.title = element_text(size = 18),    # axis titles
    axis.text = element_text(size = 16),     # axis tick labels
    plot.title = element_text(size = 20),
    legend.position = "none") -> p
p


ggsave(
  filename = "best_lags.png",
  plot = p,
  path = here("Figs/Supplement/"),
  width = 40,
  height = 20,
  units = "cm",
  dpi = "retina"
)
```

# Random Forest

```{r}
# Find best lag
best_lag <- lag_results |>
  filter(.metric == c("f_meas")) |>
  group_by(variable) |>
  mutate(
         best_mean = max(mean, na.rm = TRUE)) |>
  filter(mean >= best_mean * 0.97) |>
  slice_min(lag, n = 1) |>
  ungroup() |>
  select(-best_mean)

# - Day 0: ASDN, depth, discharge
# - Day 14: PPT, water temp
# - Day 30: ET, head gradient (?), light
# - Day 60: head gradient (?)
day0_lag <- best_lag |> 
  mutate(lag = 0)
best_lag <- best_lag |> 
  mutate(lag = c(0, 30, 14, 0, 0, 30, 30, 14))
  #mutate(lag = 0)

# best_lag <- rbind(best_lag, day0_lag) |>
#  distinct(variable, lag)

best_lags_data <- lagged_data_scaled

for (i in seq_len(nrow(best_lag))) {
  var  <- best_lag$variable[i]
  lagn <- best_lag$lag[i]
  col_name <- paste0(var, "_", lagn, "d")
  
  best_lags_data <- best_lags_data |>
    group_by(Site, name) |>
    arrange(Date) |>
    mutate(
      !!col_name := slider::slide_dbl(
        .data[[var]],
        mean,
        .before = lagn,
        .after = 0,
        .complete = FALSE
      )
    ) |>
    ungroup()
}

# Drop the original (un-lagged) variable columns
best_lags_data <- best_lags_data |>
  select(-any_of(best_lag$variable)) |>
  drop_na(Cluster) |> 
  group_by(Site) |> 
  mutate(across(
      where(is.numeric) & !active_network_length_m_0d,
      ~ scale(.) |> as.numeric()
    )) |> 
   ungroup()

png(here("Figs/best_lag_corr.png"), width = 2160, height = 2160)
best_lags_data |>
  select(where(is.numeric)) |>
  cor(use = "complete.obs") |>
  (\(cor_mat) {
    pretty_names <- var_names[colnames(cor_mat)] |>
      stringr::str_wrap(width = 10)
    dimnames(cor_mat) <- list(pretty_names, pretty_names)
    corrplot(
      cor_mat,
      method = "number",
      type = "lower",
      order = "hclust",
      tl.pos = "d",
      tl.col = "black",
      diag = TRUE,
      tl.cex = 3,
      number.cex = 4,
      cl.cex = 3,
      col = colorRampPalette(c("#e8000d", "#dde5ed", "#0051ba"))(200),
      cl.lim = c(-1, 1),
    )
  })()

dev.off()

rf_lagged_mod <- fit_full_model(df = best_lags_data, response = "Cluster", label = "best_lag_rounded_scaled")
final_wf <- rf_lagged_mod$final_wf
final_fit <- rf_lagged_mod$final_fit
fitted_model <- extract_fit_parsnip(final_fit)$fit
train_data <- rf_lagged_mod$train_data
test_data <- rf_lagged_mod$test_data

selected_vars <- rf_lagged_mod$selected_vars
tune_res <- rf_lagged_mod$tune_results
best_params <- rf_lagged_mod$best_params
autoplot(tune_res)
conf_matrix <- rf_lagged_mod$confusion_matrix
print(conf_matrix)
save(fitted_model, selected_vars, test_data, file = here("Results/rf/model_objects_lag_rounded_scaled.RData"))
beep(2)
```

```{r}
cor_mtest <- function(mat, conf.level = 0.95) {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA_real_, n, n)
  lowCI <- matrix(NA_real_, n, n)
  uppCI <- matrix(NA_real_, n, n)

  diag(p.mat) <- 0
  diag(lowCI) <- 1
  diag(uppCI) <- 1

  for (i in seq_len(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(mat[, i], mat[, j], use = "complete.obs", conf.level = conf.level)

      p.mat[i, j] <- tmp$p.value
      p.mat[j, i] <- tmp$p.value

      lowCI[i, j] <- tmp$conf.int[1]
      lowCI[j, i] <- tmp$conf.int[1]

      uppCI[i, j] <- tmp$conf.int[2]
      uppCI[j, i] <- tmp$conf.int[2]
    }
  }

  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  colnames(lowCI) <- rownames(lowCI) <- colnames(mat)
  colnames(uppCI) <- rownames(uppCI) <- colnames(mat)

  list(p = p.mat, lowCI = lowCI, uppCI = uppCI)
}

num_data <- best_lags_data |>
  dplyr::select(where(is.numeric))

cor_mat <- cor(num_data, use = "complete.obs")
test_res <- cor_mtest(num_data)

pretty_names <- var_names[colnames(cor_mat)] |>
  stringr::str_wrap(width = 10)

dimnames(cor_mat) <- list(pretty_names, pretty_names)
dimnames(test_res$p) <- list(pretty_names, pretty_names)

png(here("Figs/best_lag_corr.png"), width = 2160, height = 2160, res = 300)

corrplot(
  cor_mat,
  method = "number",
  type = "lower",
  order = "hclust",
  number.cex = 0.9,
  tl.pos = "d",
  tl.col = "black",
  tl.cex = 0.75,
  diag = T,
  p.mat = test_res$p,
  sig.level = 0.05,
  insig = "blank",
  col = colorRampPalette(c("#e8000d", "#dde5ed", "#0051ba"))(200),
  cl.lim = c(-1, 1)
)

dev.off()
```

### SHAP w/ weighted model

### 05_calc_bootstrapped_shap.R

```{r}
# Read all saved kernelshap objects
files <- list.files(path = here("Results/SHAP_Rounded_Lags_Scaled"), pattern = "kernelshap_iter_\\d+\\.rds", full.names = TRUE)
shap_list <- lapply(files, readRDS)

# Overall Mean |SHAP| with 95% CI
overall_mean_abs_shap <- shap_list |>
  seq_along() |>
  lapply(\(b) {
    ks <- shap_list[[b]]
    ks$S |>
      lapply(abs) |>
      (\(lst) Reduce("+", lst) / length(lst))() |>
      colMeans() |>
      (\(x) tibble(predictor = names(x), mean_abs_shap = as.numeric(x), Boot = b))()
  }) |>
  do.call(what = rbind) |>
  group_by(predictor) |>
  summarise(
    lo = quantile(mean_abs_shap, 0.025),
    hi = quantile(mean_abs_shap, 0.975),
    mean_abs_shap = mean(mean_abs_shap),
    .groups = "drop"
  ) |>
  mutate(
    predictor = case_when(
      predictor %in% names(var_names) ~ var_names[predictor],
      TRUE ~ predictor
    )
  )

shap_imp <- overall_mean_abs_shap |>
  ggplot(aes(x = mean_abs_shap, y = reorder(predictor, mean_abs_shap))) +
  geom_col(fill = "#2C3E50", width = 0.7) +
  geom_errorbarh(aes(xmin = lo, xmax = hi), height = 0.25) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
       title = "SHAP Variable Importance",
       x = "Mean |SHAP|",
       y = "Predictor")
shap_imp

# Cluster-level SHAP with 95% CI
clust_shap <- seq_along(shap_list) |>
  lapply(\(b) {
    ks <- shap_list[[b]]
    names(ks$S) |>
      lapply(\(clust_name) {
        shap_mat <- ks$S[[clust_name]]
        as_tibble(shap_mat) |>
          cbind(ks$X |> select(Site, Date)) |>
          transform(Cluster = clust_name, Boot = b)
      }) |>
      do.call(what = rbind)
  }) |>
  do.call(what = rbind) |>
  as_tibble() |>
  mutate(across(where(is.numeric) & !matches("Boot"), abs)) |>
  # summarise per bootstrap × cluster first

  group_by(Boot, Cluster) |>
  summarise(across(where(is.numeric) & !matches("Boot"), mean), .groups = "drop") |>
  pivot_longer(-c(Cluster, Boot), names_to = "predictor", values_to = "shap") |>
  # normalise within each bootstrap × cluster

  group_by(Boot, Cluster) |>
  mutate(norm_shap = shap / sum(shap)) |>
  ungroup() |>
  # now summarise across bootstraps
  group_by(Cluster, predictor) |>
  summarise(
    lo = quantile(norm_shap, 0.025),
    hi = quantile(norm_shap, 0.975),
    norm_shap = mean(norm_shap),
    .groups = "drop"
  ) |>
  mutate(
    ClusterLabel = paste("State", Cluster),
    predictor = case_when(
      predictor %in% names(var_names) ~ var_names[predictor],
      TRUE ~ predictor
    )
  )

cluster_var_prop <- clust_shap |>
  ggplot(aes(x = reorder(predictor, -norm_shap), y = norm_shap, fill = Cluster)) +
  geom_col(position = position_dodge(width = 0.8), alpha = 0.8, width = 0.75) +
  geom_errorbar(
    aes(ymin = lo, ymax = hi),
    position = position_dodge(width = 0.8), width = 0.25
  ) +
  scale_x_discrete(labels = label_wrap(15)) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Relative SHAP Importance by Regime",
       x = "Predictor",
       y = "Proportion of Total |SHAP|",
       fill = "State")
cluster_var_prop


ggsave(
  filename = "rounded_lags_cluster_var_prop.png",
  plot = cluster_var_prop,
  path = here("Figs/RF_Models/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = "retina"
  )

# Region-level SHAP with 95% CI
region_shap <- seq_along(shap_list) |>
  lapply(\(b) {
    ks <- shap_list[[b]]
    names(ks$S) |>
      lapply(\(clust_name) {
        shap_mat <- ks$S[[clust_name]]
        as_tibble(shap_mat) |>
          cbind(ks$X |> select(Site, Date)) |>
          transform(Cluster = clust_name, Boot = b)
      }) |>
      do.call(what = rbind)
  }) |>
  do.call(what = rbind) |>
  as_tibble() |>
  left_join(site_info |> select(Site, Region)) |>
  mutate(across(where(is.numeric) & !matches("Boot"), abs)) |>
  # summarise per bootstrap × region × cluster
  group_by(Boot, Region, Cluster) |>
  summarise(across(where(is.numeric) & !matches("Boot"), mean), .groups = "drop") |>
  pivot_longer(-c(Cluster, Region, Boot), names_to = "predictor", values_to = "shap") |>
  # normalise within each bootstrap × region × cluster
  group_by(Boot, Cluster, Region) |>
  mutate(norm_shap = shap / sum(shap)) |>
  ungroup() |>
  # summarise across bootstraps
  group_by(Cluster, Region, predictor) |>
  summarise(
    lo = quantile(norm_shap, 0.025),
    hi = quantile(norm_shap, 0.975),
    norm_shap = mean(norm_shap),
    .groups = "drop"
  ) |>
  mutate(
    ClusterLabel = paste("State", Cluster),
    predictor = case_when(
      predictor %in% names(var_names) ~ var_names[predictor],
      TRUE ~ predictor
    )
  ) |> 
  group_by(Region, predictor) |>
  mutate(pred_mean_region = mean(norm_shap, na.rm = TRUE)) |>
  ungroup() |>
  mutate(
    predictor = reorder(predictor, pred_mean_region),
    Cluster = factor(Cluster, levels = c("1", "2", "3"))
  )


zero_trim <- function(x) {
  ifelse(x == 0, "0", label_number(accuracy = 0.01)(x))
}

region_cluster_var_imp <- ggplot(region_shap,
                               aes(x = predictor, y = norm_shap, fill = Cluster)) +
  geom_col(position = position_dodge(width = 0.8, reverse = TRUE),
           width = 0.75, alpha = 0.9) +
  geom_errorbar(
    aes(ymin = lo, ymax = hi),
    position = position_dodge(width = 0.8, reverse = TRUE), width = 0.25
  ) +
  facet_wrap(~ Region, scales = "fixed") +
  coord_flip() +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(
    labels = zero_trim,
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(x = NULL, y = "Proportion of Total |SHAP|", fill = "State") +
  theme(
    panel.spacing.x = unit(2.0, "lines"),
    strip.text = element_text(face = "bold", size = 20)
  )

region_cluster_var_imp

ggsave(
  filename = "rounded_lags_region_shap.png",
  plot = region_cluster_var_imp,
  path = here("Figs/RF_Models/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = "retina"
)
```

```{r}
# Read all saved kernelshap objects
# files <- list.files(path = "~/Desktop/SHAP_Rounded_Lags_Scaled/", pattern = "kernelshap_iter_\\d+\\.rds", full.names = TRUE)
# shap_list <- lapply(files, readRDS)

# Build long-format individual observation SHAP values with feature values
shap_beeswarm <- seq_along(shap_list) |>
  lapply(\(b) {
    ks <- shap_list[[b]]
    # identify predictor columns from the SHAP matrix
    pred_cols <- colnames(ks$S[[1]])
    names(ks$S) |>
      lapply(\(clust_name) {
        shap_mat  <- ks$S[[clust_name]]
        feat_mat  <- ks$X |> select(all_of(pred_cols))   # predictors only
        as_tibble(shap_mat) |>
          rename_with(\(nm) paste0("shap__", nm)) |>
          bind_cols(
            feat_mat |>
              rename_with(\(nm) paste0("feat__", nm))
          ) |>
          bind_cols(ks$X |> select(Site, Date)) |>        # keep identifiers separate
          mutate(Cluster = clust_name, Boot = b)
      }) |>
      do.call(what = rbind)
  }) |>
  do.call(what = rbind) |>
  as_tibble() |>
  # average over bootstraps per observation (Site × Date × Cluster)
  group_by(across(c(starts_with("feat__"), Site, Date, Cluster))) |>
  summarise(across(starts_with("shap__"), mean), .groups = "drop") |>
  # pivot shap values long
  pivot_longer(
    cols         = starts_with("shap__"),
    names_to     = "predictor",
    names_prefix = "shap__",
    values_to    = "shap"
  ) |>
  # bring in matching feature value for colour scale
  pivot_longer(
    cols         = starts_with("feat__"),
    names_to     = "feat_name",
    names_prefix = "feat__",
    values_to    = "feat_value"
  ) |>
  filter(predictor == feat_name) |>
  drop_na(feat_value) |> 
  select(-feat_name) |>
  # rescale feature value 0–1 within each predictor
  group_by(predictor) |>
  mutate(feat_scaled = (feat_value - min(feat_value, na.rm = TRUE)) /
           (max(feat_value, na.rm = TRUE) - min(feat_value, na.rm = TRUE))) |>
  ungroup() |>
  mutate(
    predictor    = case_when(
      predictor %in% names(var_names) ~ var_names[predictor],
      TRUE ~ predictor
    ),
    ClusterLabel = paste("State", Cluster),
    Cluster      = factor(Cluster, levels = c("1", "2", "3"))
  ) |>
  group_by(predictor) |>
  mutate(pred_order = mean(abs(shap), na.rm = TRUE)) |>
  ungroup() |>
  mutate(predictor = reorder(predictor, pred_order))


shap_bee <- ggplot(shap_beeswarm,
                   aes(x = shap, y = predictor, colour = feat_scaled)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
  ggbeeswarm::geom_quasirandom(
    groupOnX  = FALSE,
    alpha     = 0.6,
  ) +
  facet_wrap(~ ClusterLabel) +
  scale_color_viridis_c(
    option = "plasma",
    breaks = c(0, 1),
    labels = c("Low", "High")
  ) +
  #scale_x_continuous(labels = scales::label_number(accuracy = 0.001)) +
  labs(
    title   = "",
    x       = "SHAP Value",
    y       = NULL,
    colour  = "Feature\nValue"
  ) +
  coord_cartesian(clip = "off") +
  theme(
    plot.margin = ggplot2::margin(5, 20, 5, 5, "mm"),
    panel.spacing = unit(1.5, "cm"),
    legend.key.height = unit(1.5, "cm"),
    strip.text = element_text(face = "bold", size = 14)
  )

shap_bee

ggsave(
  filename = "shap_beeswarm.png",
  plot     = shap_bee,
  path     = here("Figs/"),
  device   = "png",
  width    = 45,
  height   = 25,
  units    = "cm",
  dpi      = "retina"
)

shap_beeswarm_region <- shap_beeswarm |>
  left_join(site_info |> select(Site, Region), by = "Site")

shap_bee_region <- ggplot(shap_beeswarm_region,
                          aes(x = shap, y = predictor, colour = feat_scaled)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
  ggbeeswarm::geom_quasirandom(
    groupOnX = FALSE,
    alpha    = 0.4,
    size     = 0.6
  ) +
  facet_grid(Region ~ Cluster,
             labeller = labeller(Cluster = label_both)) +
  scale_color_viridis_c(
    option = "plasma",
    breaks = c(0, 1),
    labels = c("Low", "High")
  ) +
 # scale_x_continuous(labels = scales::label_number(accuracy = 0.001)) +
  labs(
    title  = "SHAP Summary Plot by Region and State",
    x      = "SHAP Value",
    y      = NULL,
    colour = "Feature\nValue"
  ) +
  theme(
    legend.key.height = unit(1.5, "cm"),
    strip.text        = element_text(face = "bold", size = 14),
    strip.text.y      = element_text(angle = 0),          # readable region labels
    panel.spacing     = unit(1.2, "lines"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

shap_bee_region

ggsave(
  filename = "shap_beeswarm_region.png",
  plot     = shap_bee_region,
  path     = here("Figs/"),
  device   = "png",
  width    = 45,
  height   = 40,                                          # taller to fit region rows
  units    = "cm",
  dpi      = "retina"
)
```

### Conf Mat

```{r}
# List of cluster labels
labels <- c("all_clusters", "cluster_1", "cluster_2", "cluster_3")
cluster_colors <- c("1" = "#0072B2","2" = "#009E73","3" = "#E69F00")

# Read all importance files into one long dataframe
importance_df <-read_csv(here("Results/rf/best_lag_rounded_scaled_model_importance.csv")) |>
  #mutate(Cluster = label) |>
  mutate(
    predictor = case_when(predictor %in% names(var_names) ~ var_names[predictor],
                         TRUE ~ predictor  # Keep original if not found in var_names
                         )
    ) |> 
  mutate(norm_importance = importance / max(importance, na.rm = TRUE)
  ) |>
  ungroup()
  

impVars <- importance_df |>
  mutate(predictor = str_wrap(predictor, width = 15)) |> 
  ggplot(aes(x = reorder(predictor, importance), y = importance)) +
  geom_col(position = position_dodge(width = 0.8), fill = "#2C3E50") +
  coord_flip() +
  labs(x = "Predictor", y = "Scaled Variable Importance", title = "Random Forest Variable Importance") +
  theme_few() + 
  theme(
    text = element_text(size = 18),          # overall text size
    legend.text = element_text(size = 16),   # legend entries
    legend.title = element_text(size = 18),  # legend title
    axis.title = element_text(size = 18),    # axis titles
    axis.text = element_text(size = 16),     # axis tick labels
    plot.title = element_text(size = 20, hjust = 0.5),
    legend.position = "none"
    )
impVars


conf_mat <- read_csv(here("Results/rf/best_lag_rounded_scaled_model_confusion_matrix.csv")) |> 
 group_by(Prediction) |>
  mutate(precision = n / sum(n),
         freq.Perc = round(n / sum(n), 3)) |>
  ungroup() |>
  group_by(Reference) |>
  mutate(recall = n / sum(n)) |>
  ungroup() |>
  mutate(label = paste0("(", round(precision, 2), " / ", round(recall, 2), ")")) |> 
    mutate(label = if_else(Prediction == Reference, label, NA_character_))

conf_mat_plt <- ggplot(conf_mat) +
  geom_tile(aes(x = Reference, y = Prediction, fill = freq.Perc)) +
  scale_fill_viridis_c(option = "viridis", direction = 1) +
  geom_text(
  aes(x = Reference, y = Prediction, label = n, color = Prediction == Reference), size = 8) +
  scale_color_manual(values = c("TRUE" = "black", "FALSE" = "white"), guide = "none") +
  geom_text(aes(x = Reference, y = Prediction + 0.2, label = label), color = "black", size = 6) +
  scale_y_continuous(
    trans = "reverse",
    expand = c(0, 0),
    breaks = 1:3,
    labels = paste("State", 1:3)
  ) +
  scale_x_continuous(
    position = "top",
    expand = c(0, 0),
    breaks = 1:3,
    labels = paste("State", 1:3)
  ) +
  labs(
    fill = "Fraction Predicted",
    x = "Actual",
    y = "Predicted"
  ) +
  theme(
    panel.background = element_rect(fill = NA, colour = "black", linewidth = 1),
    axis.line = element_line(colour = "black"),
    legend.position = "bottom",
    legend.key.width = unit(1.5, "cm"),
    legend.key.height = unit(0.3, "cm"),
    legend.title = element_text(vjust = 1),
    axis.text.y = element_text(angle = 90, hjust = .5)
  )
conf_mat_plt
shap_imp

shap_conf <- plot_grid(conf_mat_plt, shap_imp, label_size = 0) + theme(plot.tag = element_text(face = "bold", size = 20))
shap_conf
ggsave(
  filename = "rounded_lags_shap_confmat.png",
  plot = shap_conf,
  path = here("Figs/RF_Models/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = 600
)
```

## ALE

```{r}
library(ale)
rf_pred_fun <- function(object, newdata, type = "response") {
  predict(object, data = newdata)$predictions
}
plan(sequential)
plan(multisession, workers = parallel::detectCores() - 1)
# Create ALE plots with the custom prediction function
if (file.exists(here("Results/rf/ale_env.rds"))) {
  ale_rf <- read_rds(here("Results/rf/ale_env.rds"))
} else {
  ale_rf <- ale::ALE(
    model = fitted_model,
    data = train_data |> select(selected_vars, Cluster) |> drop_na(),
    y_col = "Cluster",
   # y_type = "categorical",
    pred_fun = rf_pred_fun,
    x_cols = list(d1 = TRUE), # All 1D effects
    boot_it = 1000,
    parallel = "all but one",
   seed = 42
  )
  write_rds(ale_rf, here("Results/rf/ale_env.rds"))
}
plan(sequential)

ale_plts <- suppressWarnings(plot(ale_rf))
ale_plts
for (i in selected_vars) {
  get(ale_plts, i) |> print(ncol = 2)
}


# Extract the data
ale_plot_data <- get_ale_data(ale_rf) |> 
  mutate(
    variable = case_when(variable %in% names(var_names) ~ var_names[variable],
                         TRUE ~ variable  # Keep original if not found in var_names
                         )
    ) |> 
  # Normalize to 0-1 scale for each cluster-variable combination
  group_by(Cluster, variable) |>
  mutate(
    # Scale to 0-1 range
    ale_scaled = (ale_values - min(ale_values, na.rm = TRUE)) / 
                       (max(ale_values, na.rm = TRUE) - min(ale_values, na.rm = TRUE)),
    ale_lower_scaled = (ale_lower - min(ale_values, na.rm = TRUE)) /
                       (max(ale_values, na.rm = TRUE) - min(ale_values, na.rm = TRUE)),
    ale_upper_scaled = (ale_upper - min(ale_values, na.rm = TRUE)) /
                       (max(ale_values, na.rm = TRUE) - min(ale_values, na.rm = TRUE))
  ) |>
  # mutate(
  #   ale_lower_scaled = pmax(ale_lower_scaled, -0.01),
  #   ale_upper_scaled = pmin(ale_upper_scaled, 1.01)
  # ) |>
  ungroup()

# Reorder variables by SHAP importance
var_order <- overall_mean_abs_shap |>
  arrange(desc(mean_abs_shap)) |>
  pull(predictor)

ale_plot_data <- ale_plot_data |>
  mutate(variable = factor(variable, levels = var_order))

# ale_plot_data |> 
#   ggplot(aes(x = x_values, y = ale_values, color = Cluster)) +
#   scale_color_manual(values = cluster_colors) +
#   scale_fill_manual(values = cluster_colors) +
#   # Add confidence intervals as ribbons (behind the lines)
#   geom_ribbon(aes(ymin = ale_lower, ymax = ale_upper, fill = Cluster), alpha = 0.4) +
#   # Add the main ALE lines
#   geom_line(linewidth = 1.2, alpha = 0.9) +
#   geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
#   facet_wrap(~variable, scales = "free", ncol = 3) +
#   labs(
#     title = "ALE Effects Across All Clusters and Variables",
#     x = "Feature Values",
#     y = "ALE Values",
#     subtitle = paste("Based on", nrow(train_data), "observations with", 
#                     length(unique(train_data$Cluster)), "clusters")
#   ) +
#   theme_minimal() +
#   theme(
#     strip.text = element_text(size = 10, face = "bold"),
#     legend.position = "bottom",
#     panel.grid.minor = element_blank(),
#     axis.text.x = element_text(angle = 45, hjust = 1)
#   )

# Plot scaled ALE effects
ale_scaled <- ale_plot_data |>
  filter(!is.na(x_values), !is.na(ale_scaled)) |>
  ggplot(aes(x = x_values, y = ale_scaled, color = Cluster, fill = Cluster)) +
  geom_ribbon(
    aes(ymin = ale_lower_scaled, ymax = ale_upper_scaled),
    alpha = 0.2, color = NA, na.rm = FALSE
  ) +
  geom_line(linewidth = 1.3, alpha = 0.9) +
  geom_hline(yintercept = 0.5, linetype = "dashed", alpha = 0.5) +
  scale_color_manual(values = cluster_colors) +
  scale_fill_manual(values = cluster_colors) +
  facet_wrap(~variable, scales = "free_x", nrow = 2) +
  labs(
    x = "Feature Values",
    y = "Normalized ALE Values (0–1)",
    fill = "Regime",
    color = "Regime"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold")
  ) +
  coord_cartesian(ylim = c(0, 1))
ale_scaled
ggsave(
  filename = "scaled_ale_plt.png",
  plot = ale_scaled,
  path = here("Figs/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = "retina"
)
beepr::beep(2)
```

```{r}
# Extract ALE data for all clusters and variables
vars <- c("Depth_0d", "ET_30d", "PPT_14d", "Q_0d",
          "active_network_length_m_0d", "avgPar_30d",
          "headGradient_30d", "water.temp_14d")

ale_data <- map_dfr(1:3, function(cat) {
  map_dfr(vars, function(var) {
    df <- ale_rf@effect[[cat]]$ale$d1[[var]]
    x_col <- names(df)[1]
    df |>
      rename(x_ceil = all_of(x_col)) |>
      select(x_ceil, .y, .y_lo, .y_hi, .n) |>
      mutate(variable = var, cluster = factor(cat))
  })
}) |>
  mutate(
    variable = case_when(
      variable %in% names(var_names) ~ var_names[variable],
      TRUE ~ variable
    )
  )

vars_ordered <- var_order[var_order %in% unique(ale_data$variable)]

# Apply factor ordering
ale_data <- ale_data |>
  mutate(variable = factor(variable, levels = vars_ordered))
# Get y reference lines from y_summary
y_refs <- as.data.frame(ale_rf@params$y_summary) |>
  rownames_to_column("quantile") |>
  filter(quantile %in% c("25%", "50%", "75%")) |>
  pivot_longer(-quantile, names_to = "cluster", values_to = "value") |>
  filter(cluster != "Cluster") |>
  mutate(cluster = factor(cluster),
         label = case_when(
           quantile == "25%" ~ "25%",
           quantile == "50%" ~ "median",
           quantile == "75%" ~ "75%"
         ))

# Extract raw data for rugs
raw_data <- ale_rf@params$data$data_sample
raw_data_cols <- names(raw_data)[names(raw_data) != "Cluster"]
vars_rename <- var_names[names(var_names) %in% raw_data_cols]
reverse_var_names <- setNames(names(vars_rename), vars_rename)

# Build rug data for all variables at once
rug_data <- map_dfr(vars_ordered, function(var) {
  orig_var <- unname(reverse_var_names[var])
  if (is.na(orig_var)) orig_var <- var
  raw_data |>
    select(x = all_of(orig_var)) |>
    mutate(variable = factor(var, levels = vars_ordered))
})

# Plot with facet_wrap
test_ale <- ale_data |>
  filter(!is.na(x_ceil), !is.na(.y)) |>
  ggplot(aes(x = x_ceil, color = cluster, fill = cluster)) +
  geom_ribbon(aes(ymin = .y_lo, ymax = .y_hi), alpha = 0.15, color = NA) +
  geom_line(aes(y = .y), linewidth = 0.8) +
  geom_rug(data = rug_data, aes(x = x, y = NULL),
           inherit.aes = FALSE,
           sides = "b", alpha = 0.2, linewidth = 0.3, color = "grey30") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40", linewidth = 0.5) +
  scale_color_manual(values = cluster_colors, name = "State") +
  scale_fill_manual(values = cluster_colors, name = "State") +
  facet_wrap(~variable, scales = "free_x", ncol = 4) +
  labs(x = "Feature Values", y = "ALE Effect") +
 # coord_cartesian(clip = "off") +
  theme(
    #axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold"),
    plot.margin = ggplot2::margin(5, 6, 5, 5, "mm")  # extra right margin
  )
test_ale
ggsave(
  filename = "ale_plt.png",
  plot = test_ale,
  here("Figs/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = "retina"
)
```

```{r}
site_shap <-  names(all_shap$S) |>
  lapply(\(clust_name) {
    shap_mat <- all_shap$S[[clust_name]]
    as_tibble(shap_mat) |>
      cbind(all_shap$X |> select(Site, Date)) |> 
      transform(Cluster = clust_name)
  }) |>
  do.call(what = rbind) |> 
  as_tibble() |>
  left_join(site_info |> select(Site, Region)) |> 
  mutate(across(where(is.numeric), abs)) |> 
  group_by(Region, Cluster) |> 
  summarise(across(where(is.numeric), mean)) |> 
  pivot_longer(-c(Cluster, Region), names_to = "predictor", values_to = "shap") |> 
  group_by(Cluster, Region) |> 
  #mutate(norm_shap = shap / max(shap)) |> 
  mutate(norm_shap = shap / sum(shap)) |> 
  ungroup() |> 
    mutate(
    ClusterLabel = paste("Cluster", Cluster),
     predictor = case_when(predictor %in% names(var_names) ~ var_names[predictor],
                         TRUE ~ predictor  # Keep original if not found in var_names
                         )
    ) 

site_cluster_var_imp <- site_shap |>
  ggplot(aes(x = reorder(predictor, -shap), y = norm_shap, fill = Region)) +
  geom_col(position = "dodge", alpha = 0.8) +
  facet_grid(Region~Cluster, scales = "free_x") +
  #coord_flip() +
  scale_fill_manual(
    values = c("#613921", "#83A102", "#006475"),
    breaks = c("Great Plains", "Eastern Temperate Forests", "North American Deserts")
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = "Predictor",
    y = "Proportion of Total |SHAP|"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

site_cluster_var_imp


# Check if the sum of the normalized shap is 1
norm_check <- site_shap |> group_by(Cluster, Region) |>
    summarise(
        sum_shap = sum(shap, na.rm = TRUE),
        sum_norm_shap = sum(norm_shap, na.rm = TRUE)
    )

site_cluster_var_imp <- site_shap |>
  ggplot(aes(x = reorder(predictor, -shap), y = norm_shap, fill = Region)) +
  geom_col(position = "dodge", alpha = 0.8) +
  facet_wrap(~Cluster, scales = "free_x") +
  #coord_flip() +
  scale_fill_manual(
    values = c("#613921", "#83A102", "#006475"),
    breaks = c("Great Plains", "Eastern Temperate Forests", "North American Deserts")
  ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = "Predictor",
    y = "Proportion of Total |SHAP|"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

site_cluster_var_imp


site_cluster_var_imp <- site_shap |>
  ggplot(aes(x = reorder(predictor, -shap), y = norm_shap, fill = Cluster)) +
  geom_col(position = "dodge", alpha = 0.8) +
  facet_wrap(~Region, scales = "free_x") +
  #coord_flip() +
  # scale_fill_manual(
  #   values = c("#613921", "#83A102", "#006475"),
  #   breaks = c("Great Plains", "Eastern Temperate Forests", "North American Deserts")
  # ) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = "Predictor",
    y = "Proportion of Total |SHAP|",
    fill = "Regime"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
  )
site_cluster_var_imp

site_shap |> group_by(Region) |>
    summarise(
        sum_shap = sum(shap, na.rm = TRUE),
        sum_norm_shap = sum(norm_shap, na.rm = TRUE)
    )

ggsave(
  filename = "site_shap.png",
  plot = site_cluster_var_imp,
  path = here("Figs/Supplement/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = "retina"
)

# order predictors by overall mean importance (across all region-regime combos)
pred_order <- site_shap |>
  group_by(predictor) |>
  summarise(m = mean(norm_shap, na.rm = TRUE), .groups = "drop") |>
  arrange(desc(m)) |>
  pull(predictor)

site_shap |>
  mutate(
    predictor = factor(predictor, levels = rev(pred_order)),
    Cluster = factor(Cluster, levels = c("1","2","3"))
  ) |>
  ggplot(aes(x = Cluster, y = predictor, fill = norm_shap)) +
  geom_tile(color = "white", linewidth = 0.2) +
  facet_wrap(~ Region, nrow = 1) +
  scale_y_discrete(drop = FALSE) +
  scale_fill_viridis_c(option = "viridis", direction = 1) +
  labs(x = "Regime", y = NULL, fill = "Proportion of\nTotal |SHAP|") +
  theme_minimal(base_size = 11) +
  theme(panel.grid = element_blank())

site_shap_ranked <- site_shap |>
  group_by(Region, predictor) |>
  mutate(pred_mean_region = mean(norm_shap, na.rm = TRUE)) |>
  ungroup() |>
  mutate(
    predictor = reorder(predictor, pred_mean_region),
    Cluster = factor(Cluster, levels = c("1","2","3"))
  )


zero_trim <- function(x) {
  ifelse(x == 0, "0", label_number(accuracy = 0.01)(x))
}

site_cluster_var_imp <- ggplot(site_shap_ranked,
                               aes(x = predictor, y = norm_shap, fill = Cluster)) +
  geom_col(position = position_dodge(width = 0.8, reverse = TRUE), width = 0.75, alpha = 0.9) +
  facet_wrap(~ Region, scales = "fixed") +
  coord_flip() +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(
    labels = zero_trim,
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(x = NULL, y = "Proportion of Total |SHAP|", fill = "Regime") +
  theme(
    panel.spacing.x = unit(2.0, "lines"),
    # plot.margin = ggplot2::margin(5.5, 20, 5.5, 5.5),
    # axis.text.y = element_text(margin = ggplot2::margin(r = 6)),
    strip.text = element_text(face = "bold", size = 20)
  )

site_cluster_var_imp
ggsave(
  filename = "site_shap.png",
  plot = site_cluster_var_imp,
  path = here("Figs/Supplement/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = "retina"
)

```

### PDP

```{r}
# classes <- levels(train_data$Cluster)
# 
# pdp_out <- list()
# 
# sites <- train_data |> 
#   select(all_of(selected_vars), Site) |> 
#   drop_na() |> 
#   pull(Site) |> 
#   unique()
# 
# unscalled_lag <- lagged_data |>
#   group_by(Site, name) |>
#   arrange(Date) |>
#   mutate(
#     across(
#       all_of(best_lag$variable),
#       ~ {
#         this_lag <- best_lag$lag[best_lag$variable == cur_column()]
#         slider::slide_dbl(
#           .x,
#           mean,
#           .before = this_lag,
#           .after = 0,
#           .complete = TRUE
#         )
#       },
#       .names = "{.col}_lag"
#     )
#   ) |>
#   ungroup()
# unscalled_lag <- rename_lagged(unscalled_lag, best_lag)
# 
# var_ranges <- unscalled_lag |>
#   group_by(Site) |>
#   summarise(across(all_of(selected_vars),
#                    \(x) list(c(min(x, na.rm = TRUE), max(x, na.rm = TRUE)))))
# 
# 
# for (var in selected_vars) {
#   for (cl in classes) {
#     for (s in sites) {
#       # site training data
#       train_s <- train_data |> 
#         filter(Site == s) |> 
#         select(all_of(selected_vars)) |> 
#         drop_na()
#       # if a site has zero or 1 rows, skip it
#       if (nrow(train_s) < 2) next
# 
#       pd <- pdp::partial(
#         object = fitted_model,
#         pred.var = var,
#         train = train_s,
#         prob = TRUE,
#         which.class = cl
#       ) |>
#         mutate(
#           variable = var,
#           cluster = cl,
#           Site = s
#         ) |>
#         rename(predictor_value = !!sym(var))
# 
#       pdp_out[[paste(var, cl, s, sep = "_")]] <- pd
#     }
#   }
# }
# pdp_df <- bind_rows(pdp_out)
# 
# p <- list()
# 
# for (var in selected_vars) {
# 
#   df_var <- pdp_df |> 
#     filter(variable == var) |> 
#     group_by(Site, cluster, variable) |> 
#     mutate(prob_scale = rescale(yhat, to = c(0, 1))) |> 
#     ungroup()
#   
#   
#     df_var <- df_var |>
#       left_join(var_ranges, by = "Site") |>
#       rowwise() |>
#       mutate(
#         predictor_orig =
#           (predictor_value / 100) *
#           (get(variable)[[2]] - get(variable)[[1]]) +
#           get(variable)[[1]]
#       ) |>
#       ungroup() |> 
#       mutate(
#       variable = case_when(
#         variable %in% names(var_names) ~ var_names[variable],
#         TRUE ~ variable
#       )
#     ) 
# 
#   p[[var]] <- df_var |> 
#     ggplot(aes(x = predictor_orig, y = prob_scale, color = cluster)) +
#     geom_line(linewidth = 0.8) +
#     scale_color_manual(values = cluster_colors) +
#     facet_wrap(~ Site, scales = "free_x") +
#     labs(
#       x = var_names[var],
#       y = "Scaled probability (0–1 within Regime)",
#       color = "Regime",
#       title = "PDP by Regime × Site",
#       subtitle = paste("Predictor:", var_names[var])
#     ) +
#     theme_few()
# }
# 
# p
# 
# 
# ggsave(
#   filename = "site_pdp_asdn.png",
#   plot = p$active_network_length_m_0d,
#   path = here("Figs/Supplement/"),
#   device = "png",
#   width = 40,
#   height = 20,
#   units = "cm",
#   dpi = "retina"
# )
```

```{r}
# 
# classes <- levels(train_data$Cluster)
# 
# pdp_out <- list()
# 
# regions <- train_data |> 
#   select(all_of(selected_vars), Site) |> 
#   left_join(site_info |> select(Site, Region)) |> 
#   pull(Region) |> 
#   unique()
# 
# train_data <- train_data |> 
#   left_join(site_info |> select(Site, Region))
# 
# for (var in selected_vars) {
#   for (cl in classes) {
#     for (region in regions) {
#       # region training data
#       train_region <- train_data |> 
#         filter(Region == region) |> 
#         select(all_of(selected_vars))
#       
#       # must have enough rows
#       if (nrow(train_region) < 2) next
# 
#       # drop rows where this var is NA (prevents min/max Inf/-Inf)
#       train_region_nonmiss <- train_region |>
#         filter(!is.na(.data[[var]]))
# 
#       if (nrow(train_region_nonmiss) < 2) next
# 
#       # need some variation; constant predictor can also cause trouble/useless PD
#       if (length(unique(train_region_nonmiss[[var]])) < 2) next
#       
#       pd <- pdp::partial(
#         object = fitted_model,
#         pred.var = var,
#         train = train_region,
#         prob = TRUE,
#         which.class = cl
#       ) |>
#         mutate(
#           variable = var,
#           cluster = cl,
#           Region = region
#         ) |>
#         rename(predictor_value = !!sym(var))
# 
#       pdp_out[[paste(var, cl, region, sep = "_")]] <- pd
#     }
#   }
# }
# pdp_df <- bind_rows(pdp_out)
# 
# p <- list()
# 
# for (var in selected_vars) {
#   
#   df_var <- pdp_df |> 
#     filter(variable == var) |> 
#     group_by(Region, cluster, variable) |> 
#     mutate(prob_scale = rescale(yhat, to = c(0, 1))) |> 
#     ungroup() |> 
#     mutate(
#       variable = case_when(
#         variable %in% names(var_names) ~ var_names[variable],
#         TRUE ~ variable
#       )
#     ) 
#   
#   p[[var]] <- df_var |> 
#     ggplot(aes(x = predictor_value, y = prob_scale, color = cluster)) +
#     geom_line(linewidth = 0.8) +
#     scale_color_manual(values = cluster_colors) +
#     facet_wrap(~ Region, scales = "free_x") +
#     labs(
#       x = var_names[var],
#       y = "Scaled probability (0–1 within Regime)",
#       color = "Regime",
#       title = "PDP by Regime × Region",
#       subtitle = paste("Predictor:", var_names[var])
#     ) +
#     theme_few()
# }
# 
# p
# 
# 
# ggsave(
#   filename = "site_pdp_asdn.png",
#   plot = p$active_network_length_m_0d,
#   path = here("Figs/Supplement/"),
#   device = "png",
#   width = 40,
#   height = 20,
#   units = "cm",
#   dpi = "retina"
# )

```

```{r}
classes <- levels(train_data$Cluster)

pdp_out <- list()

sites <- train_data |>
  select(all_of(selected_vars), Site) |>
  left_join(site_info |> select(Site, Region)) |>
  pull(Site) |>
  unique()

for (var in selected_vars) {
  for (cl in classes) {
    for (s in sites) {
      # region training data
      train_site <- train_data |>
        filter(Site == s) |>
        select(all_of(selected_vars))

      # must have enough rows
      if (nrow(train_site) < 2) next

      # drop rows where this var is NA (prevents min/max Inf/-Inf)
      train_site_nonmiss <- train_site |>
        filter(!is.na(.data[[var]]))

      if (nrow(train_site_nonmiss) < 2) next

      # need some variation; constant predictor can also cause trouble/useless PD
      if (length(unique(train_site_nonmiss[[var]])) < 2) next

      pd <- pdp::partial(
        object = fitted_model,
        pred.var = var,
        train = train_site,
        prob = TRUE,
        which.class = cl
      ) |>
        mutate(
          variable = var,
          cluster = cl,
          Site = s
        ) |>
        rename(predictor_value = !!sym(var))

      pdp_out[[paste(var, cl, s, sep = "_")]] <- pd
    }
  }
}
pdp_df <- bind_rows(pdp_out)

p <- list()

for (var in selected_vars) {

  df_var <- pdp_df |>
    filter(variable == var) |>
    group_by(Site, cluster, variable) |>
    mutate(prob_scale = rescale(yhat, to = c(0, 1))) |>
    ungroup() |>
    mutate(
      variable = case_when(
        variable %in% names(var_names) ~ var_names[variable],
        TRUE ~ variable
      )
    )

  p[[var]] <- df_var |>
    ggplot(aes(x = predictor_value, y = prob_scale, color = cluster)) +
    geom_line(linewidth = 0.8) +
    scale_color_manual(values = cluster_colors) +
    facet_wrap(~ Site, scales = "free_x") +
    labs(
      x = var_names[var],
      y = "Scaled probability (0–1 within Regime)",
      color = "Regime",
      title = "PDP by Regime × Site",
      subtitle = paste("Predictor:", var_names[var])
    ) +
    theme_few()
}

p

```

```{r}
beepr::beep(2)
# unlink(here("Results/rf/lagged_rf_exact_shap.rds"))
# unlink(here("Results/rf/ale_env.rds"))
```
