---
title: "DO Regimes"
author: "CLB"
format: html
---

```{r}
#| echo: false
#| label: Libraries
#| output: false
#| include: false
library(here)
source(here("Code/paths+packages.R"))
library(preMetabolizer)
library(furrr)
library(PerformanceAnalytics)
library(corrr)
library(cluster)
library(factoextra)
library(beepr)
library(glue)
library(cowplot)
library(ggExtra)
library(caret)
library(randomForest)
library(gganimate)
library(gt)
set.seed(4) # For reproducibility
gpSites <- c("#1b9e77", "#d95f02", "#7570b3")
cluster_colors <- c("1" = "#0072B2", "2" = "#009E73", "3" = "#E69F00")
site_colors <- tibble(
  Site = c("JDR", "GBJ", "KNZ", "SHN", "YMR", "WHR", "PRF", "TAL"),
  color = c(
    "#006475",
    "#56CFD8",
    "#613921",
    "#F2C27B",
    "#A77652",
    "#345022",
    "#5A6C0A",
    "#83A102"
  )
)
```

```{r}
#| label: Set Theme
theme_set(
  theme_bw() +
    theme(
      text = element_text(size = 20), # overall text size
      legend.text = element_text(size = 20), # legend entries
      legend.title = element_text(size = 20), # legend title
      axis.title.x = element_text(size = 20, face = "bold"), # axis titles
      axis.title.y = element_text(size = 20, face = "bold"), # axis titles
      axis.text = element_text(size = 20), # axis tick labels
      plot.title = element_text(size = 20, hjust = 0.5, face = "bold"),
      legend.position = "bottom",
      plot.tag = element_text(size = 20, face = "bold"),
      strip.background = element_blank()
    )
)
```

# Load Data

```{r}
gpSheds <- st_read(
  "~/Dropbox/AIMS/GIS/Results/AIMS_GreatPlains.gpkg",
  layer = "Watersheds",
  quiet = T
) |>
  filter(watershed %in% c("KNZ", "SHN", "YMR")) |>
  st_transform(crs = 4326)
seSheds <- st_read(
  "~/Dropbox/AIMS/GIS/Results/AIMS_SouthEast.gpkg",
  layer = "Watersheds"
) |>
  st_transform(4326)
mwSheds <- st_read(
  "~/Dropbox/AIMS/GIS/Results/AIMS_MountainWest.gpkg",
  layer = "Watersheds"
) |>
  st_transform(4326) |>
  filter(watershed != "DRC")

all_sheds <- gpSheds |>
  bind_rows(seSheds) |>
  bind_rows(mwSheds) |>
  rename(Site = watershed) |>
  st_transform(5070) |>
  st_point_on_surface() |>
  st_transform(4326) |>
  mutate(lon = st_coordinates(geom)[, 1], lat = st_coordinates(geom)[, 2]) |>
  st_drop_geometry()

gpStreams <- st_read(
  "~/Dropbox/AIMS/GIS/Results/AIMS_GreatPlains.gpkg",
  layer = "Streams",
  quiet = T
) |>
  filter(watershed %in% c("KNZ", "SHN", "YMR")) |>
  st_transform(crs = 4326)
seStreams <- st_read(
  "~/Dropbox/AIMS/GIS/Results/AIMS_SouthEast.gpkg",
  layer = "Streams",
  quiet = T
) |>
  mutate(length_m = units::drop_units(st_length(geom))) |>
  st_transform(4326)
mwStreams <- st_read(
  "~/Dropbox/AIMS/GIS/Results/AIMS_MountainWest.gpkg",
  layer = "Streams",
  quiet = T
) |>
  st_transform(4326) |>
  filter(watershed != "DRC")

stream_lengths <- gpStreams |>
  bind_rows(seStreams) |>
  bind_rows(mwStreams) |>
  st_drop_geometry() |>
  select(Site = watershed, length_m) |>
  group_by(Site) |>
  summarise(total_length_m = sum(length_m, na.rm = TRUE))

site_meta <- all_sheds |>
  full_join(stream_lengths) |>
  full_join(site_colors) |>
  as_tibble() |>
  select(Site, color, area_sqkm, total_length_m, lon, lat) |>
  mutate(Site = fct_reorder(Site, lon), shape = c(1, 0, 12, 2, 3, 8, 6, 11))

rm(
  gpSheds,
  gpStreams,
  seSheds,
  seStreams,
  mwSheds,
  mwStreams,
  stream_lengths,
  all_sheds,
  okabeIto,
  gpSites,
  distColors
)
```

```{r}
streamLight <- read_csv(here(
  "Data/streamLight/predLight/all_sites_results.csv"
)) |>
  rename(Site = Site_ID) |>
  mutate(
    Date = as_date(local_time),
    Site = case_when(
      Site == "SFM01" ~ "KNZ",
      Site == "SHM01" ~ "SHN",
      Site == "ENM01" ~ "YMR",
      TRUE ~ Site # keep other values unchanged
    )
  ) |>
  group_by(Site, Date) |>
  summarise(
    maxPar = max(PAR_surface, na.rm = TRUE),
    avgPar = mean(PAR_surface, na.rm = TRUE),
    .groups = "drop"
  )

modisET <- read_csv(here("Data/ET/MODIS_Daily_ET.csv")) |>
  filter(Site != "DRC")

gpDat <- read_csv(here("Data/GP_dailyData.csv")) |>
  select(
    -c(
      minDO.obs,
      maxDO.obs,
      avgDO.obs,
      DO.hyp,
      DO.amp,
      probHyp,
      dayProbConc,
      nightProbConc,
      NHR_obs,
      dPercCon,
      wetdry_SW,
      PercWet_1d,
      dPercWet_1d,
      ICSL,
      IntactStreamLength,
      AlphaCentrality,
      Harary,
      upstream_length,
      percentConnected,
      PercWet_7d,
      PercWet_14d,
      PercWet_30d,
      PercWet_60d,
      dPercWet_7d,
      dPercWet_14d,
      dPercWet_30d,
      dPercWet_60d,
      DOY,
      DOSat.hyp,
      cvDO.pctsat,
      sdDO.pctsat,
      DOSat.slope,
      dDOpctsat.amp,
      dMinDO.pctsat,
      dayProbSat,
      nightProbSat,
      NHR_sat
    )
  ) |>
  mutate(Date = as_date(Date)) |>
  left_join(streamLight, by = c("Site", "Date"))

seDat <- read_csv(here("Data/SE_dailyData.csv")) |>
  select(
    -c(
      minDO.obs,
      maxDO.obs,
      avgDO.obs,
      DO.hyp,
      DO.amp,
      probHyp,
      dayProbConc,
      nightProbConc,
      NHR_obs,
      DOSat.hyp,
      cvDO.pctsat,
      sdDO.pctsat,
      DOSat.slope,
      dDOpctsat.amp,
      dMinDO.pctsat,
      dayProbSat,
      nightProbSat,
      NHR_sat
    )
  ) |>
  mutate(Date = as_date(Date))

jdrDat <- read_csv(here("Data/JDR_dailyData.csv")) |>
  select(
    -c(
      minDO.obs,
      maxDO.obs,
      avgDO.obs,
      DO.hyp,
      DO.amp,
      probHyp,
      dayProbConc,
      nightProbConc,
      NHR_obs,
      DOSat.hyp,
      cvDO.pctsat,
      sdDO.pctsat,
      DOSat.slope,
      dDOpctsat.amp,
      dMinDO.pctsat,
      dayProbSat,
      nightProbSat,
      NHR_sat
    )
  ) |>
  mutate(Date = as_date(Date))

gbjDat <- read_csv(here("Data/GBJ_dailyData.csv")) |>
  select(
    -c(
      minDO.obs,
      maxDO.obs,
      avgDO.obs,
      DO.hyp,
      DO.amp,
      probHyp,
      dayProbConc,
      nightProbConc,
      NHR_obs,
      DOSat.hyp,
      dayProbSat,
      nightProbSat,
      NHR_sat
    )
  ) |>
  mutate(Date = as_date(Date))


allDat <- gpDat |>
  bind_rows(seDat) |>
  bind_rows(jdrDat) |>
  bind_rows(gbjDat) |>
  full_join(modisET) |>
  select(
    -c(
      excProb,
      DOSat.exc,
      minDepth,
      maxDepth,
      minQ,
      maxQ,
      cvDO.pctsat,
      DOSat.slope,
      dDOpctsat.amp,
      sdDO.pctsat,
      dMinDO.pctsat
    )
  ) |>
  mutate(
    waterYear = as.factor(calcWaterYear(Date)),
    Season = as.factor(get_season(Date)),
    Date = as.Date(Date),
    DOY = as.numeric(format(Date, "%j"))
  ) |>
  # Fill Q with 0 if na or inf AND depth is 0
  mutate(
    Q_1d = case_when(
      (is.na(Q_1d) | is.infinite(Q_1d)) & Depth_1d == 0 ~ 0,
      TRUE ~ Q_1d
    )
  ) |>
  select(-totalDO) |>
  group_by(Site) |>
  mutate(
    excProb = calc_exceedance_prob(Q_1d, remove_zeros = T) * 100,
    DOSat.exc = calc_exceedance_prob(avgDO.pctsat, remove_zeros = T)
  ) |>
  ungroup() |>
  mutate(excProb = case_when(is.na(excProb) ~ 0, T ~ excProb))

## New Alpha centrality
sticAC <- read_csv(here("Data/STICs/streamDAG_metrics.csv")) |>
  rename(Date = date, Site = watershed) |>
  filter(Site != "DRC") |>
  select(
    Site,
    Date,
    active_upstream_length,
    active_network_length_m,
    percent_network_wet,
    a_cent,
    a_cent_weighted
  ) |>
  mutate(
    active_upstream_length = case_when(
      is.na(active_upstream_length) ~ 0,
      T ~ active_upstream_length
    ),
    percent_network_wet = case_when(
      is.na(percent_network_wet) ~ 0,
      T ~ percent_network_wet
    ),
    active_network_length_m = case_when(
      is.na(active_network_length_m) ~ 0,
      T ~ active_network_length_m
    )
  )

allDat <- allDat |>
  full_join(sticAC)

# Fill NAs in ASDN with 0 only if between dry periods
allDat <- allDat |>
  arrange(Site, Date) |>
  group_by(Site) |>
  mutate(
    prev_non_na = zoo::na.locf(active_network_length_m, na.rm = FALSE),
    next_non_na = zoo::na.locf(
      active_network_length_m,
      fromLast = TRUE,
      na.rm = FALSE
    ),
    active_network_length_m = ifelse(
      is.na(active_network_length_m) &
        prev_non_na == 0 &
        next_non_na == 0,
      0,
      active_network_length_m
    )
  ) |>
  select(-prev_non_na, -next_non_na) |>
  ungroup()

# Define the variables you want to calculate rolling averages for
vars_to_roll <- c(
  "active_upstream_length",
  "active_network_length_m",
  "percent_network_wet",
  "a_cent",
  "a_cent_weighted",
  "avgPar",
  "water.temp"
)

# Define rolling window sizes
window_sizes <- c(7, 14, 30, 60)

# List of variables to compute rolling means for
vars <- list(
  # active_upstream_length = "active_upstream_length",
  #  active_network_length_m = "active_network_length_m",
  #  percent_network_wet = "percent_network_wet",
  #  a_cent = "a_cent",
  #  a_cent_weighted = "a_cent_weighted",
  avgPar = "avgPar",
  water.temp = "water.temp"
)
allDat <- allDat |>
  group_by(Site) |>
  arrange(Date)
setDT(allDat)
# Generate rolling means for each variable and window size
for (var in names(vars)) {
  prefix <- vars[[var]] # Get the appropriate prefix for the variable
  allDat[,
    (paste0(prefix, "_", window_sizes, "d")) := lapply(
      window_sizes,
      function(w) frollmean(get(var), n = w, align = "right", na.rm = TRUE)
    ),
    by = Site
  ]
}

dailyDat <- allDat |>
  filter(Date >= "2021-03-26")
#rm(streamLight, gpDat, seDat, gbjDat, jdrDat, modisET, window_sizes, vars)

n_days <- dailyDat |>
  group_by(Site) |>
  summarise(
    n_days_avgDO = sum(!is.na(avgDO.pctsat)),
    n_days_upstream = sum(!is.na(active_upstream_length)),
    mean_perc_wet = mean(percent_network_wet, na.rm = TRUE),
    .groups = "drop"
  )
site_koppen <- read_csv(here("Metadata/aims_koppen_geiger.csv")) |>
  mutate(
    hydro_clim = case_when(
      Site %in% c("KNZ", "SHN", "YMR") ~ "Cold Humid",
      Site %in% c("TAL", "WHR", "PRF") ~ "Warm Humid",
      Site %in% c("GBJ", "JDR") ~ "Cold Dry"
    )
  ) |>
  group_by(kg_zone) |>
  mutate(
    kg_name = paste0(toupper(kg_zone), "-", row_number())
  ) |>
  ungroup()

site_meta <- site_meta |>
  mutate(
    name = case_when(
      Site %in% c("KNZ", "SHN", "YMR") ~ c("GP-1", "GP-2", "GP-3")[match(
        Site,
        c("KNZ", "SHN", "YMR")
      )],
      Site %in% c("TAL", "PRF", "WHR") ~ c("EF-1", "EF-2", "EF-3")[match(
        Site,
        c("TAL", "PRF", "WHR")
      )],
      Site %in% c("GBJ", "JDR") ~ c("ND-1", "ND-2")[match(
        Site,
        c("GBJ", "JDR")
      )],
      TRUE ~ NA_character_
    )
  ) |>
  left_join(
    site_koppen |>
      select(Site, kg_zone = kg_zone_name, kg_name),
    by = "Site"
  ) |>
  arrange(lon) |>
  mutate(
    Site = fct_reorder(Site, lon),
    name = fct_reorder(name, lon),
    kg_name = fct_reorder(kg_name, lon)
  )

site_info <- site_meta |>
  left_join(n_days) |>
  mutate(
    n_years = n_days_upstream / 365.25,
    Region = case_when(
      Site %in% c("KNZ", "SHN", "YMR") ~ "Great Plains",
      Site %in% c("TAL", "WHR", "PRF") ~ "Eastern Temperate Forests",
      Site %in% c("GBJ", "JDR") ~ "North American Deserts"
    )
  )

total <- site_info |>
  summarise(n_days = sum(n_days_upstream))


n_stics <- read_csv(here("Data/n_stics.csv")) |>
  rename(Site = watershed)

site_info <- site_info |>
  left_join(n_stics)

avg_ppt <- dailyDat |>
  group_by(Site) |>
  summarise(mean_ppt_mm = mean(PPT_1d, na.rm = TRUE)) |>
  ungroup()

site_info |>
  left_join(avg_ppt, by = "Site") |>
  mutate(area_sqkm = case_when(Site == "GBJ" ~ 16.20, T ~ area_sqkm)) |>
  select(
    Region,
    Watershed = kg_name,
    km2 = area_sqkm,
    n_days_upstream,
    n_sites,
    mean_perc_wet
  ) |>
  gt() |>
  fmt_number(
    columns = c(km2, mean_perc_wet),
    decimals = 2
  ) |>
  fmt_number(
    columns = c(n_days_upstream),
    decimals = 0
  ) |>
  cols_label(
    km2 = "Area ({{km^2}})",
    n_days_upstream = "# of Days",
    mean_perc_wet = "Average % Wet",
    n_sites = "# of STICs",
  ) |>
  cols_align(
    align = "center",
    columns = -Watershed
  ) |>
  gtsave(here("Figs/site_info.png"), zoom = 6, vwidth = 2000, vheight = 1200)

site_info |>
  left_join(avg_ppt, by = "Site") |>
  mutate(area_sqkm = case_when(Site == "GBJ" ~ 16.20, T ~ area_sqkm)) |>
  select(
    Region,
    Watershed = kg_name,
    km2 = area_sqkm,
    n_days_upstream,
    n_sites,
    mean_perc_wet
  ) |>
  gt() |>
  fmt_number(
    columns = c(km2, mean_perc_wet),
    decimals = 2
  ) |>
  fmt_number(
    columns = c(n_days_upstream),
    decimals = 0
  ) |>
  cols_label(
    km2 = "Area ({{km^2}})",
    n_days_upstream = "# of Days",
    mean_perc_wet = "Average % Wet",
    n_sites = "# of STICs",
  ) |>
  cols_align(
    align = "center",
    columns = -Watershed
  ) |>
  gtsave(here("Figs/site_info.docx"), zoom = 6, vwidth = 2000, vheight = 1200)
```

## Add Name to DailyDat

```{r}
dailyDat <- dailyDat |>
  left_join(site_meta |> select(Site, name, kg_name))
```

## Get % Missingness for Sites

```{r}
daily_tibble <- dailyDat |>
  as_tibble() |>
  select(
    any_of(c(
      "kg_name",
      "Date",
      "water.temp",
      "avgDO.pctsat",
      "active_network_length_m"
    )),
    ends_with("_1d")
  ) |>
  select(-c(PPT_1d, ET_1d)) |>
  rename(Watershed = kg_name)


site_missingness <- daily_tibble |>
  rename_with(~ str_remove(.x, "_1d$"), ends_with("_1d")) |>
  group_by(Watershed) |>
  filter(
    row_number() >= min(which(!is.na(avgDO.pctsat))),
    row_number() <= max(which(!is.na(avgDO.pctsat)))
  ) |>
  complete(Date = seq.Date(min(Date), max(Date), by = "day")) |>
  summarise(
    total_possible_days = n_distinct(Date),
    Q_NA_days = sum(is.na(Q)),
    Depth_NA_days = sum(is.na(Depth)),
    active_network_length_NA_days = sum(is.na(active_network_length_m)),
    avgDO_pctsat_NA_days = sum(is.na(avgDO.pctsat)),
    .groups = "drop"
  )

site_missingness
```

## Look at hydro between regions

```{r}
flow_prop <- dailyDat |>
  group_by(Site, kg_name) |>
  filter(
    Date >= min(Date[!is.na(Q_1d)], na.rm = TRUE),
    Date <= max(Date[!is.na(Q_1d)], na.rm = TRUE)
  ) |>
  summarise(
    total_days = n(),
    days_with_flow = sum(Q_1d > 0, na.rm = TRUE),
    prop_flowing = days_with_flow / total_days
  )

flow_prop |>
  ggplot(aes(x = kg_name, y = prop_flowing, fill = kg_name)) +
  geom_col(na.rm = TRUE) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  labs(
    title = "Proportion of Days with Flow by Site",
    subtitle = "(Days with Q > 0)",
    x = "Site",
    y = "Proportion of Days Flowing"
  ) +
  theme(legend.position = "none")
ggsave(
  filename = "flowProp.png",
  path = here("Figs/"),
  device = "png",
  width = 20,
  height = 20,
  units = "cm",
  dpi = 600
)
q50 <- dailyDat |>
  filter(Q_1d > 0) |>
  group_by(Site, kg_name) |>
  filter(
    Date >= min(Date[!is.na(Q_1d)], na.rm = TRUE),
    Date <= max(Date[!is.na(Q_1d)], na.rm = TRUE)
  ) |>
  summarize(
    days_below_50th = sum(Q_1d < median(Q_1d, na.rm = TRUE), na.rm = TRUE),
    total_days = sum(!is.na(Q_1d)),
    median_Q = median(Q_1d, na.rm = TRUE)
  )

q50 |>
  ggplot(aes(x = kg_name, y = days_below_50th, fill = kg_name)) +
  geom_col(na.rm = TRUE) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  labs(
    title = "Number of Days with Below-Median Discharge",
    x = "Site",
    y = "Days < 50th Percentile"
  ) +
  theme(legend.position = "none")
ggsave(
  filename = "q50.png",
  path = here("Figs/"),
  device = "png",
  width = 20,
  height = 20,
  units = "cm",
  dpi = 600
)

dailyDat |>
  mutate(
    Region = case_when(
      Site %in% c("KNZ", "SHN", "YMR") ~ "Great Plains",
      Site %in% c("TAL", "WHR", "PRF") ~ "Eastern Temperate Forests",
      Site %in% c("GBJ", "JDR") ~ "North American Deserts"
    ),
    Region = factor(
      Region,
      levels = c(
        "Great Plains",
        "Eastern Temperate Forests",
        "North American Deserts"
      )
    )
  ) |>
  group_by(Site, kg_name) |>
  mutate(
    active_network_length_m = active_network_length_m /
      max(active_network_length_m, na.rm = T) *
      100
  ) |>
  ungroup() |>
  ggplot(aes(x = active_network_length_m, fill = kg_name)) +
  geom_density(alpha = 0.5, na.rm = TRUE) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  labs(
    title = "Distribution of Active Stream \nNetwork Extent by Region",
    x = "ASDN (%)",
    y = "Relative Frequency"
  ) +
  facet_wrap(~Region, ncol = 1, scales = "free_y") +
  guides(color = guide_legend(title = NULL)) +
  theme(legend.title = element_blank())
ggsave(
  filename = "asdnPerc.png",
  path = here("Figs/"),
  device = "png",
  width = 20,
  height = 20,
  units = "cm",
  dpi = 600
)

dailyDat |>
  mutate(
    Region = case_when(
      Site %in% c("KNZ", "SHN", "YMR") ~ "Great Plains",
      Site %in% c("TAL", "WHR", "PRF") ~ "Eastern Temperate Forests",
      Site %in% c("GBJ", "JDR") ~ "North American Deserts"
    ),
    Region = factor(
      Region,
      levels = c(
        "Great Plains",
        "Eastern Temperate Forests",
        "North American Deserts"
      )
    )
  ) |>
  group_by(Site, kg_name) |>
  mutate(
    active_upstream_length = active_upstream_length /
      max(active_upstream_length, na.rm = T) *
      100
  ) |>
  ungroup() |>
  ggplot(aes(x = active_upstream_length, fill = kg_name)) +
  geom_density(alpha = 0.5, na.rm = TRUE) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  labs(
    title = "Distribution of Active Upstream\nNetwork Extent by Region",
    x = "AUDN (%)",
    y = "Relative Frequency"
  ) +
  facet_wrap(~Region, ncol = 1, scales = "free_y") +
  guides(color = guide_legend(title = NULL)) +
  theme(legend.title = element_blank())
ggsave(
  filename = "audnPerc.png",
  path = here("Figs/"),
  device = "png",
  width = 20,
  height = 20,
  units = "cm",
  dpi = 600
)

dailyDat |>
  mutate(
    Region = case_when(
      Site %in% c("KNZ", "SHN", "YMR") ~ "Great Plains",
      Site %in% c("TAL", "WHR", "PRF") ~ "Eastern Temperate Forests",
      Site %in% c("GBJ", "JDR") ~ "North American Deserts"
    ),
    Region = factor(
      Region,
      levels = c(
        "Great Plains",
        "Eastern Temperate Forests",
        "North American Deserts"
      )
    )
  ) |>
  group_by(Site, kg_name) |>
  ungroup() |>
  ggplot(aes(x = percent_network_wet, fill = kg_name)) +
  geom_density(alpha = 0.5, na.rm = TRUE) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  labs(
    title = "Distribution of Percent Network Wet by Region",
    x = "% Network Wet",
    y = "Relative Frequency"
  ) +
  facet_wrap(~Region, ncol = 1, scales = "free_y") +
  guides(color = guide_legend(title = NULL)) +
  theme(legend.title = element_blank())
ggsave(
  filename = "PercNetWet.png",
  path = here("Figs/"),
  device = "png",
  width = 20,
  height = 20,
  units = "cm",
  dpi = 600
)

dailyDat |>
  mutate(
    Region = case_when(
      Site %in% c("KNZ", "SHN", "YMR") ~ "Great Plains",
      Site %in% c("TAL", "WHR", "PRF") ~ "Eastern Temperate Forests",
      Site %in% c("GBJ", "JDR") ~ "North American Deserts"
    ),
    Region = factor(
      Region,
      levels = c(
        "Great Plains",
        "Eastern Temperate Forests",
        "North American Deserts"
      )
    )
  ) |>
  group_by(Site, kg_name) |>
  ungroup() |>
  ggplot(aes(x = a_cent, fill = kg_name)) +
  geom_density(alpha = 0.5, na.rm = TRUE) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  labs(
    title = "Distribution of Alpha Centrality by Region",
    x = "Alpha Centrality",
    y = "Relative Frequency"
  ) +
  facet_wrap(~Region, ncol = 1, scales = "free_y") +
  guides(color = guide_legend(title = NULL)) +
  theme(legend.title = element_blank())
ggsave(
  filename = "AlphaCentrality.png",
  path = here("Figs/"),
  device = "png",
  width = 20,
  height = 20,
  units = "cm",
  dpi = 600
)

dailyDat |>
  mutate(
    Region = case_when(
      Site %in% c("KNZ", "SHN", "YMR") ~ "Great Plains",
      Site %in% c("TAL", "WHR", "PRF") ~ "Eastern Temperate Forests",
      Site %in% c("GBJ", "JDR") ~ "North American Deserts"
    ),
    Region = factor(
      Region,
      levels = c(
        "Great Plains",
        "Eastern Temperate Forests",
        "North American Deserts"
      )
    )
  ) |>
  group_by(Site, kg_name) |>
  ungroup() |>
  ggplot(aes(x = a_cent_weighted, fill = kg_name)) +
  geom_density(alpha = 0.5, na.rm = TRUE) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  labs(
    title = "Distribution of Weighted Alpha Centrality by Region",
    x = "Weighted Alpha Centrality",
    y = "Relative Frequency"
  ) +
  facet_wrap(~Region, ncol = 1, scales = "free_y") +
  guides(color = guide_legend(title = NULL)) +
  theme(legend.title = element_blank())
ggsave(
  filename = "AlphaCentralityWeighted.png",
  path = here("Figs/"),
  device = "png",
  width = 20,
  height = 20,
  units = "cm",
  dpi = 600
)


runoff50 <- dailyDat |>
  filter(Q_1d > 0) |>
  select(Site, kg_name, Date, Q_1d, active_network_length_m) |>
  left_join(
    site_info |>
      select(Site, area_sqkm, total_length_m) |>
      mutate(area_sqkm = case_when(Site == "GBJ" ~ 16.20, T ~ area_sqkm))
  ) |>
  mutate(runoff = Q_1d * 0.0864 / area_sqkm) |>
  group_by(Site, kg_name) |>
  filter(
    Date >= min(Date[!is.na(runoff)], na.rm = TRUE),
    Date <= max(Date[!is.na(runoff)], na.rm = TRUE)
  ) |>
  summarize(
    days_below_50th = sum(runoff < median(runoff, na.rm = TRUE), na.rm = TRUE),
    total_days = sum(!is.na(runoff)),
    median_runoff = median(runoff, na.rm = TRUE)
  )

runoff50 |>
  ggplot(aes(x = kg_name, y = days_below_50th, fill = kg_name)) +
  geom_col(na.rm = TRUE) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  labs(
    title = "Number of Days with Below-Median Runoff",
    x = "Site",
    y = "Days < 50th Percentile"
  ) +
  theme(legend.position = "none")
ggsave(
  filename = "runoff50.png",
  path = here("Figs/"),
  device = "png",
  width = 20,
  height = 20,
  units = "cm",
  dpi = 600
)

peakDry <- dailyDat |>
  select(Site, kg_name, Date, percent_network_wet) |>
  mutate(
    Year = year(Date),
    water_year = calcWaterYear(Date)
  ) |>
  filter(!is.na(percent_network_wet)) |>
  group_by(Site, kg_name, Year) |>
  summarise(
    peak_date = Date[which.max(percent_network_wet)],

    # minimum value AFTER the peak
    min_value = min(percent_network_wet[Date >= peak_date], na.rm = TRUE),

    # corresponding date
    min_date = Date[
      percent_network_wet == min_value &
        Date >= peak_date
    ][1],

    days_peak_to_min = as.numeric(min_date - peak_date),
    .groups = "drop"
  ) |>
  filter(
    min_value < 60, # require minimum < 50%
    days_peak_to_min >= 13,
    Year <= 2024
  )


# Prepare vline data with Region info
vline_dat <- peakDry |>
  mutate(
    Region = case_when(
      Site %in% c("KNZ", "SHN", "YMR") ~ "Great Plains",
      Site %in% c("TAL", "WHR", "PRF") ~ "Eastern Temperate Forests",
      Site %in% c("GBJ", "JDR") ~ "North American Deserts",
      TRUE ~ NA_character_
    ),
    Region = factor(
      Region,
      levels = c(
        "Great Plains",
        "Eastern Temperate Forests",
        "North American Deserts"
      )
    )
  ) |>
  tidyr::pivot_longer(
    cols = c(peak_date, min_date),
    names_to = "type",
    values_to = "date"
  ) |>
  mutate(type = factor(type, levels = c("peak_date", "min_date")))
# Then add to your plot:
dailyDat |>
  mutate(
    Region = case_when(
      Site %in% c("KNZ", "SHN", "YMR") ~ "Great Plains",
      Site %in% c("TAL", "WHR", "PRF") ~ "Eastern Temperate Forests",
      Site %in% c("GBJ", "JDR") ~ "North American Deserts"
    ),
    Region = factor(
      Region,
      levels = c(
        "Great Plains",
        "Eastern Temperate Forests",
        "North American Deserts"
      )
    )
  ) |>
  group_by(Site, kg_name) |>
  #mutate(active_network_length_m = active_network_length_m/max(active_network_length_m, na.rm = T)*100) |>
  ungroup() |>
  ggplot(aes(x = Date, y = percent_network_wet, color = kg_name)) +
  geom_line(alpha = 1, na.rm = TRUE) +
  # Add vertical lines
  geom_vline(
    data = vline_dat,
    aes(xintercept = date, linetype = type),
    alpha = 1
  ) +
  scale_linetype_manual(
    values = c("min_date" = "dashed", "peak_date" = "solid"),
    labels = c("min_date" = "Minimum", "peak_date" = "Peak")
  ) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  labs(
    title = "Percent Wet Over Time",
    x = "Date",
    y = "Network Wet (%)"
  ) +
  facet_wrap(~kg_name, scales = "free") +
  guides(
    color = guide_legend(title = NULL),
    linetype = guide_legend(title = NULL)
  ) +
  theme(legend.title = element_blank())

ggsave(
  filename = "peak_to_dry_overtime.png",
  path = here("Figs/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = 600
)
peak_to_dry <- peakDry |>
  group_by(Site, kg_name) |>
  summarise(
    mean_days_peak_to_min = mean(days_peak_to_min, na.rm = TRUE),
    n_years = n()
  )

peak_to_dry |>
  ggplot(aes(x = kg_name, y = mean_days_peak_to_min, fill = kg_name)) +
  geom_col(na.rm = TRUE) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  labs(
    title = "Average Time from Peak Wetness to Dry-Down",
    x = "Site",
    y = "Days"
  ) +
  theme(legend.position = "none")
ggsave(
  filename = "peak_to_dry.png",
  path = here("Figs/"),
  device = "png",
  width = 20,
  height = 20,
  units = "cm",
  dpi = 600
)
```

# MW Daily Q
```{r}
mw_q_plt <- dailyDat |>
  mutate(
    Year = lubridate::year(Date),
    Region = case_when(
      Site %in% c("KNZ", "SHN", "YMR") ~ "Great Plains",
      Site %in% c("TAL", "WHR", "PRF") ~ "Eastern Temperate Forests",
      Site %in% c("GBJ", "JDR") ~ "North American Deserts"
    )
  ) |>
  filter(Site %in% c("GBJ", "JDR")) |>
  group_by(Year) |>
  filter(any(!is.na(Q_1d))) |>
  ungroup() |>
  ggplot(aes(x = Date, y = Q_1d)) +
  geom_line(
    color = "dodgerblue4",
    linewidth = 0.6,
    na.rm = TRUE
  ) +
  scale_x_date(
    expand = c(0, 0),
    breaks = \(x) {
      lubridate::pretty_dates(x, n = 12) |>
        lubridate::as_date()
    },
    date_labels = "%b '%y"
  ) +
  scale_y_log10(
    #expand = c(0, 0),
    breaks = log_breaks(n = 6, base = 10),
    labels = scales::label_number(
      scale_cut = append(cut_short_scale(), 1, 1),
      accuracy = 0.001
    )
  ) +
  annotation_logticks(sides = "l") +
  labs(
    x = "",
    y = bquote(bold(
      "Discharge (" * L ~ s^{
        -1
      } *
        ")"
    ))
  ) +
  facet_wrap(~Site, scales = "free", ncol = 1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
mw_q_plt

ggsave(
  filename = "MW_Q.png",
  plot = mw_q_plt,
  path = here("Figs/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = 600
)
rm(mw_q_plt)
```


# Exceedance Probability

```{r}
#| warning: false
#| message: false
runoff <- dailyDat |>
  select(Site, kg_name, name, Date, Q_1d) |>
  left_join(
    site_info |>
      select(Site, area_sqkm, total_length_m) |>
      mutate(area_sqkm = case_when(Site == "GBJ" ~ 16.20, T ~ area_sqkm))
  ) |>
  group_by(Site, name) |>
  mutate(
    runoff_mm = (Q_1d * 86400) / (area_sqkm * 1e6) * 1000,
    runoff_exc = calc_exceedance_prob(runoff_mm, remove_zeros = T)
  ) |>
  ungroup()
# Runoff Exceedance Prob
runoff_expProb <- runoff |>
  ggplot(aes(x = runoff_exc, y = runoff_mm, color = name)) +
  geom_line(na.rm = TRUE, linewidth = 2) +
  scale_y_continuous(
    trans = "log10",
    labels = label_number(big.mark = ",")
  ) +
  scale_x_continuous(labels = label_percent(accuracy = 1, scale = 100)) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  labs(
    x = "Runoff Exceedance Probability",
    y = bquote(bold(
      "Runoff (mm day"^{
        -1
      } *
        ")"
    ))
  ) +
  guides(color = guide_legend(title = NULL))
ggsave(
  filename = "runoff_excProb.png",
  plot = runoff_expProb,
  path = here("Figs/"),
  device = "png",
  width = 20,
  height = 20,
  units = "cm",
  dpi = 600
)
runoff_plt <- runoff |>
  mutate(runoff_mm = if_else(runoff_mm <= 0, NA_real_, runoff_mm)) |>
  ggplot(aes(x = Date, y = runoff_mm, color = name)) +
  geom_line(na.rm = TRUE) +
  facet_wrap(~name, scales = "fixed") +
  scale_y_continuous(
    trans = "log10",
    labels = label_number(big.mark = ",")
  ) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  labs(
    x = "",
    y = bquote(bold(
      "Runoff (mm day"^{
        -1
      } *
        ")"
    ))
  ) +
  guides(color = guide_legend(title = NULL))
ggsave(
  filename = "runoff.png",
  plot = runoff_plt,
  path = here("Figs/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = 600
)
# Discharge Exceedance Prob
qExcProb <- dailyDat |>
  ggplot(aes(x = excProb, y = Q_1d, color = kg_name)) +
  geom_line(na.rm = TRUE, linewidth = 2) +
  scale_y_continuous(
    trans = "log10",
    labels = label_number(big.mark = ",")
  ) +
  scale_x_continuous(labels = label_percent(accuracy = 1, scale = 1)) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  labs(
    x = "Discharge Exceedance Probability",
    y = bquote(bold(
      "Discharge (" * L ~ s^{
        -1
      } *
        ")"
    ))
  ) +
  guides(color = guide_legend(title = NULL))
qExcProb

# DO Exceedance Prob
doExcProb <- dailyDat |>
  ggplot(aes(x = DOSat.exc, y = avgDO.pctsat, color = kg_name)) +
  geom_line(na.rm = TRUE, linewidth = 2) +
  geom_hline(
    yintercept = 50,
    linetype = "dashed",
    color = "grey60",
    linewidth = 1.5
  ) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  scale_x_continuous(labels = label_percent(accuracy = 1, scale = 100)) +
  scale_y_continuous(
    breaks = c(0, 25, 50, 75, 100),
  ) +
  labs(
    x = expression(bold("DO"["% Sat"] ~ "Exceedance Probability")),
    y = expression(bold("\nDaily Mean DO"["% Sat"]))
  ) +
  guides(color = guide_legend(title = NULL))
doExcProb
ggsave(
  filename = "doExcProb.png",
  plot = doExcProb,
  path = here("Figs/"),
  device = "png",
  width = 20,
  height = 20,
  units = "cm",
  dpi = 600
)


excProb_plt <- (runoff_expProb + doExcProb) +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
ggsave(
  filename = "excProb.png",
  plot = excProb_plt,
  path = here("Figs/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = 600
)

#rm(doExcProb, qExcProb, combined_plot)
```

## Equation Table

```{r}
# Define the table data
eq_table <- tibble(
  Metric = c(
    "Mean DO",
    "Minimum DO",
    "Maximum DO",
    "Amplitude",
    "Exceedance Probability",
    "Probability of Hypoxia" #,
    # "Night Hypoxia Ratio"
  ),

  Equation = c(
    # Daily mean
    "$\\overline{DO}_{\\text{day}} = \\frac{1}{n_d} \\sum_{t=1}^{n_d} DO_{t,d}$",
    # Daily min
    "$DO_{\\min,d} = \\min_t(DO_{t,d})$",
    # Daily max
    "$DO_{\\max,d} = \\max_t(DO_{t,d})$",
    # Daily amplitude
    "$DO_{\\text{amp},d} = \\max_t(DO_{t,d}) - \\min_t(DO_{t,d})$",
    # Exceedance probability
    "$P_{\\text{exc},d} = \\frac{r_d}{N + 1}$",
    # Daily probability of hypoxia
    "$P_{\\text{hypoxia},d} = \\frac{1}{n_d} \\sum_{t=1}^{n_d}
      \\begin{cases}
        1, & \\text{if } DO_{t,d} < DO_{\\text{hyp}} \\\\
        0, & \\text{otherwise}
      \\end{cases}$" #,
    #  # Night hypoxia ratio
    #  "$R_{\\text{hypoxia},d} = \\frac{P_{\\text{hypoxia, night},d}}{P_{\\text{hypoxia, day},d}}$"
  ),
  Temporal_Scale = c(
    "Daily",
    "Daily",
    "Daily",
    "Daily",
    "Across Days",
    "Daily"
  )
)

# Build the gt table
footnote_text <- paste0(
  "Table 2. $DO_{t,d}$ is the dissolved oxygen observation at time $t$ on day $d$; $n_d$ is the number of valid observations per day; $\\overline{DO}_d$ is the daily mean DO; $r_d$ is the rank of $\\overline{DO}_d$ (1 = highest); $N$ is the total number of days; $DO_{\\text{hyp}}$ is the hypoxia threshold."
)

do_eq_table <- eq_table |>
  select(-Temporal_Scale) |>
  gt() |>
  fmt_markdown(columns = "Equation") |>
  cols_label(
    Metric = "Metric",
    Equation = "Equation",
  ) |>
  #tab_header(title = "Equations for Daily Dissolved Oxygen Metrics") |>
  #tab_source_note(source_note = md(footnote_text)) |>
  # tab_options(
  #   data_row.padding = px(1),             # tighten vertical spacing
  #   column_labels.padding = px(2),        # tighten header padding
  #   table_body.hlines.width = px(0.5),    # thin horizontal lines
  #   table_body.border.bottom.width = px(0.5),
  # ) |>
  cols_align(align = "left", columns = Metric) |>
  cols_align(align = "center", columns = Equation) |>
  tab_options(
    data_row.padding = px(2), # reduce from default ~8px
    column_labels.padding = px(4), # tighten header row too
    table.font.names = "Times New Roman"
  ) |>
  #tab_source_note(source_note = md(footnote_text)) |>
  tab_style(
    style = cell_text(weight = "bold"), # 👈 bold font weight
    locations = cells_column_labels(columns = c(Metric, Equation))
  )
do_eq_table |>
  gtsave(here("Figs/do_metrics_eqn.pdf"))
do_eq_table |>
  gtsave(here("Figs/do_metrics_eqn.png"))
do_eq_table |>
  gtsave(here("Figs/do_metrics_eqn.docx"))
```

# ASDN

```{r}
p <- dailyDat |>
  mutate(name = factor(name, levels = levels(site_meta$kg_name))) |>
  ggplot(aes(
    x = Date,
    y = active_network_length_m / 1000,
    color = percent_network_wet
  )) +
  geom_point(size = 3, alpha = 0.8) +
  facet_wrap(~name, scales = "free_y", ncol = 2) +
  scale_x_date(labels = date_format("%b '%y")) +
  scale_color_viridis_c(
    option = "plasma",
    na.value = "grey70",
    direction = -1
  ) +
  labs(
    x = "",
    y = "ASDN (km)",
    color = "% Network Wet"
  )
p
ggsave(
  filename = "ASDN.png",
  plot = p,
  path = here("Figs/Misc/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = 600
)

p <- dailyDat |>
  mutate(name = factor(name, levels = levels(site_meta$kg_name))) |>
  group_by(Site) |>
  mutate(
    asdn = (active_network_length_m /
      max(active_network_length_m, na.rm = TRUE)) *
      100
  ) |>
  ungroup() |>
  ggplot(aes(x = percent_network_wet, y = asdn)) +
  geom_point(aes(color = name), alpha = 0.7, size = 2.5, na.rm = TRUE) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  scale_x_continuous(labels = label_percent(accuracy = 1, scale = 1)) +
  scale_y_continuous(labels = label_percent(accuracy = 1, scale = 1)) +
  facet_wrap(~name, scales = "fixed", ncol = 2) +
  labs(x = "Network Wet", y = "ASDN", color = "Watershed")
p
ggsave(
  filename = "percWet_ASDN.png",
  plot = p,
  path = here("Figs/Misc/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = 600
)


dailyDat |>
  mutate(name = factor(name, levels = levels(site_meta$kg_name))) |>
  group_by(Site) |>
  mutate(
    asdn = (active_network_length_m /
      max(active_network_length_m, na.rm = TRUE)) *
      100
  ) |>
  ungroup() |>
  ggplot(aes(
    x = Date,
    y = asdn,
    color = a_cent_weighted
  )) +
  geom_point(size = 3, alpha = 0.8) +
  facet_wrap(~name, scales = "free_y", ncol = 2) +
  scale_x_date(labels = date_format("%b '%y")) +
  scale_color_viridis_c(
    option = "plasma",
    na.value = "grey70",
    direction = -1
  ) +
  labs(
    x = "",
    y = "ASDN (%)",
    color = "Alpha Centrality (Weighted)"
  ) -> p
p
ggsave(
  filename = "a_cent_asdn.png",
  plot = p,
  path = here("Figs/Misc/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = 600
)

dailyDat |>
  mutate(name = factor(name, levels = levels(site_meta$kg_name))) |>
  group_by(Site) |>
  mutate(
    asdn = (active_network_length_m /
      max(active_network_length_m, na.rm = TRUE)) *
      100
  ) |>
  ungroup() |>
  ggplot(aes(
    x = Date,
    y = a_cent_weighted,
    color = percent_network_wet,
  )) +
  geom_point(size = 3, alpha = 0.8) +
  facet_wrap(~name, scales = "free_y", ncol = 2) +
  scale_x_date(labels = date_format("%b '%y")) +
  scale_color_viridis_c(
    option = "plasma",
    na.value = "grey70",
    direction = -1
  ) +
  labs(
    x = "",
    y = "Alpha Centrality (Weighted)",
    color = "Network Wet"
  ) -> p
p
ggsave(
  filename = "a_cent_perc_wet.png",
  plot = p,
  path = here("Figs/Misc/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = 600
)
```

# DO Metrics PCA

```{r}
var_labels <- c(
  DOpctsat.amp = "Amplitude",
  DOSat.exc = "Exceedence Probability",
  probHypSat = "Probability of Hypoxia",
  minDO.pctsat = "Min",
  # avgDO.pctsat = "Mean",
  maxDO.pctsat = "Max"
)

# Prepare PCA data
pca_data <- dailyDat |>
  select(
    kg_name,
    Date,
    DOpctsat.amp,
    DOSat.exc,
    probHypSat,
    minDO.pctsat,
    # avgDO.pctsat,
    maxDO.pctsat
  ) |>
  mutate(probHypSat = case_when(is.nan(probHypSat) ~ 0, TRUE ~ probHypSat))

# Create and fit PCA recipe
pca_recipe <- recipe(~., data = pca_data) |>
  update_role(kg_name, Date, new_role = "Id") |>
  step_naomit(all_predictors()) |>
  step_normalize(all_predictors()) |>
  step_pca(all_predictors(), id = "PCA")

do_pca <- prep(pca_recipe)

# Extract PCA loadings for biplot (PC1–PC3, wide form)
loadings <- do_pca |>
  tidy(id = "PCA") |>
  filter(component %in% c("PC1", "PC2", "PC3")) |>
  pivot_wider(names_from = component, values_from = value) |>
  mutate(var_label = recode(terms, !!!var_labels))

# Calculate arrow scale for biplot
pca_scores <- juice(do_pca)

arrow_scale <- min(
  (max(pca_scores$PC1) -
    min(pca_scores$PC1) / (max(pca_scores$PC1) - min(pca_scores$PC1))),
  (max(pca_scores$PC2) -
    min(pca_scores$PC2) / (max(pca_scores$PC2) - min(pca_scores$PC2)))
) *
  2

loadings <- loadings |>
  mutate(across(where(is.numeric), ~ . * arrow_scale))

# Extract variance explained
var_explained <- do_pca |>
  tidy(id = "PCA", type = "variance") |>
  filter(terms == "percent variance") |>
  arrange(component)

eigvals <- do_pca |>
  tidy(id = "PCA", type = "variance") |>
  filter(terms == "variance") |>
  mutate(sig = value > 1)

pc1_label <- paste0(
  "PC1 (",
  round(var_explained |> filter(component == 1) |> pull(value), 1),
  "%)"
)

pc2_label <- paste0(
  "PC2 (",
  round(var_explained |> filter(component == 2) |> pull(value), 1),
  "%)"
)

# -------------------------------------------------------------------
# NEW: Table of eigenvalues, % explained, and loadings for PCs with
#      eigenvalues > 1
# -------------------------------------------------------------------

# Combine eigenvalues and percent variance
pc_table <- eigvals |>
  select(component, eigenvalue = value, sig) |>
  left_join(
    var_explained |>
      select(component, pct_explained = value),
    by = "component"
  ) |>
  filter(eigenvalue > 1)

# Long table: one row per variable per retained PC
pca_table <- do_pca |>
  tidy(id = "PCA") |>
  # component is "PC1", "PC2", ...; make numeric to join to pc_table
  mutate(component_num = as.integer(gsub("PC", "", component))) |>
  inner_join(pc_table, by = c("component_num" = "component")) |>
  mutate(
    var_label = recode(terms, !!!var_labels)
  ) |>
  select(
    component, # "PC1", "PC2", ...
    eigenvalue,
    pct_explained,
    term = terms,
    var_label,
    loading = value
  ) |>
  arrange(component, desc(abs(loading)))
pca_table
# pca_table now has eigenvalues, % explained and loadings
# for all PCs with eigenvalues > 1.
# -------------------------------------------------------------------

# Scree plot
do_pca |>
  tidy(id = "PCA", type = "variance") |>
  filter(terms == "percent variance") |>
  ggplot(aes(x = factor(component), y = value)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = paste0(round(value, 1), "%")), vjust = -0.5, size = 3) +
  labs(
    title = "Scree Plot",
    x = "Principal Component",
    y = "Percent Variance Explained"
  )

# Variable contribution plot
do_pca |>
  tidy(id = "PCA") |>
  mutate(
    Direction = if_else(value > 0, "Positive", "Negative"),
    terms = tidytext::reorder_within(terms, abs(value), component)
  ) |>
  filter(component %in% c("PC1", "PC2", "PC3", "PC4")) |>
  ggplot(aes(abs(value), terms, fill = Direction)) +
  geom_col() +
  facet_wrap(~component, scales = "free_y", ncol = 2) +
  tidytext::scale_y_reordered() +
  scale_fill_manual(
    values = c("Negative" = "#D55E00", "Positive" = "#56B4E9")
  ) +
  labs(
    x = "Absolute Value of Contribution",
    y = NULL,
    fill = "Direction"
  ) +
  theme_few() +
  theme(legend.position = "bottom")

# Create PCA biplot
do_pca_plt <- pca_scores |>
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point(
    aes(fill = kg_name, shape = kg_name, color = kg_name),
    size = 3,
    alpha = 1,
    stroke = 0.3
  ) +
  stat_ellipse(
    aes(color = kg_name),
    geom = "polygon",
    level = 0.95,
    alpha = 0.2,
    fill = NA
  ) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  # Add loading arrows
  geom_segment(
    data = loadings,
    aes(x = 0, y = 0, xend = PC1, yend = PC2),
    arrow = arrow(length = unit(0.2, "cm")),
    color = "black"
  ) +
  geom_text_repel(
    data = loadings,
    aes(x = PC1 * 1.1, y = PC2 * 1.1, label = var_label),
    size = 5,
    direction = "both"
  ) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  scale_shape_manual(
    values = setNames(site_meta$shape, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  labs(
    x = pc1_label,
    y = pc2_label
  ) +
  theme(
    legend.title = element_blank()
  )

do_pca_plt
# Save main PCA plot
ggsave(
  filename = "doPCA.png",
  plot = do_pca_plt,
  path = here("Figs/"),
  device = "png",
  width = 20,
  height = 20,
  units = "cm",
  dpi = 600
)
```

```{r}
## 1. Get eigenvalues and % variance for PC1–PC3
pca_var_info <- do_pca |>
  tidy(id = "PCA", type = "variance") |>
  filter(
    component %in% c("1", "2", "3"),
    terms %in% c("variance", "percent variance")
  ) |>
  select(component, terms, value) |>
  pivot_wider(names_from = terms, values_from = value) |>
  mutate(
    eigenvalue = variance,
    pct = `percent variance`,
    component = case_when(
      component == 1 ~ "PC1",
      component == 2 ~ "PC2",
      component == 3 ~ "PC3",
      TRUE ~ NA
    )
  )

## 2. Build pretty labels like
pc_labels <- pca_var_info |>
  arrange(component) |>
  mutate(
    label = sprintf("%s\n(λ = %.2f, %.1f%%)", component, eigenvalue, pct)
  ) |>
  pull(label)
# pc_labels will be a character vector of length 3 # in order: PC1, PC2
## 3. Your loadings table (as before)
pca_loadings_table <- loadings |>
  select(Variable = var_label, PC1, PC2, PC3) |>
  mutate(across(c(PC1, PC2), ~ round(.x, 2)))
## 4. Replace the PC column names with the new labels
names(pca_loadings_table)[2:4] <- pc_labels

pca_table_png <- pca_loadings_table |>
  gt() |>
  tab_header(title = "PCA Loadings", ) |>
  fmt_number(columns = everything(), decimals = 2) |>
  cols_align(align = "center")


pca_table_png |>
  gtsave(here("Figs/Supplement/pca_table.png"))
pca_table_png |>
  gtsave(here("Figs/Supplement/pca_table.docx"))
pca_table_png
```

# Other Clustering on DO Metrics

```{r}
library(ggdendro)
library(cluster)

# Prepare daily dissolved oxygen metrics
do_metrics_daily <- dailyDat |>
  select(
    kg_name,
    Date,
    DOpctsat.amp,
    DOSat.exc,
    probHypSat,
    avgDO.pctsat,
    minDO.pctsat,
    maxDO.pctsat
  ) |>
  mutate(
    # Treat undefined hypoxia probabilities as zero.
    probHypSat = replace(probHypSat, is.nan(probHypSat), 0)
  ) |>
  drop_na() |>
  as_tibble() # |>
#mutate(across(where(is.numeric), \(x) (x - mean(x)) / sd(x)))

# Build and standardize the clustering feature matrix
metric_names <- c(
  "DOpctsat.amp",
  "DOSat.exc",
  "probHypSat",
  "avgDO.pctsat"
)

do_features <- do_metrics_daily |>
  mutate(obs_id = paste(kg_name, Date, sep = "_")) |>
  select(obs_id, all_of(metric_names))

# Convert the metric columns to a numeric matrix and preserve the observation
# identifier as the row name.
feature_matrix <- do_features |>
  select(all_of(metric_names)) |>
  as.matrix()

rownames(feature_matrix) <- do_features$obs_id

# Standardization gives each metric a mean of zero and standard deviation of
# one, preventing variables with larger numeric ranges from dominating.
feature_matrix_scaled <- scale(feature_matrix)
cor(feature_matrix_scaled)
# Euclidean distances among days based on the standardized DO metrics.
do_distance <- dist(feature_matrix_scaled, method = "euclidean")
```

```{r}
# Helper functions
calculate_silhouette_widths <- function(
  hc,
  distance_matrix,
  candidate_k = 2:8
) {
  tibble(k = candidate_k) |>
    mutate(
      avg_silhouette = map_dbl(
        k,
        function(current_k) {
          cluster_membership <- cutree(hc, k = current_k)

          silhouette_values <- cluster::silhouette(
            cluster_membership,
            distance_matrix
          )

          mean(silhouette_values[, "sil_width"])
        }
      )
    )
}


select_best_k <- function(silhouette_results) {
  silhouette_results |>
    slice_max(
      order_by = avg_silhouette,
      n = 1,
      with_ties = FALSE
    ) |>
    pull(k)
}


plot_silhouette_widths <- function(
  silhouette_results,
  title
) {
  best_result <- silhouette_results |>
    slice_max(
      order_by = avg_silhouette,
      n = 1,
      with_ties = FALSE
    )

  ggplot(
    silhouette_results,
    aes(x = k, y = avg_silhouette)
  ) +
    geom_line() +
    geom_point() +
    geom_point(
      data = best_result,
      color = "red",
      size = 3
    ) +
    scale_x_continuous(
      breaks = silhouette_results$k
    ) +
    labs(
      title = title,
      x = "Number of clusters",
      y = "Average silhouette width"
    ) +
    theme_minimal()
}


plot_pca_clusters <- function(
  pca_result,
  cluster_membership,
  title
) {
  variance_explained <- summary(pca_result)$importance[2, 1:2] * 100

  pca_scores <- pca_result$x[, 1:2, drop = FALSE] |>
    as_tibble() |>
    mutate(cluster = factor(cluster_membership))

  ggplot(
    pca_scores,
    aes(x = PC1, y = PC2, color = cluster)
  ) +
    geom_point(
      alpha = 0.4,
      size = 1
    ) +
    stat_ellipse(
      level = 0.95,
      linewidth = 0.6
    ) +
    labs(
      title = title,
      x = sprintf("PC1 (%.1f%%)", variance_explained[1]),
      y = sprintf("PC2 (%.1f%%)", variance_explained[2]),
      color = "Cluster"
    ) +
    theme_minimal()
}


plot_dendrogram <- function(
  hc,
  title
) {
  dendrogram_segments <- dendro_data(hc) |>
    segment()

  ggplot(
    dendrogram_segments,
    aes(
      x = x,
      y = y,
      xend = xend,
      yend = yend
    )
  ) +
    geom_segment(linewidth = 0.2) +
    labs(
      title = title,
      x = NULL,
      y = "Height"
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank()
    )
}


plot_colored_dendrogram <- function(
  hc,
  cluster_membership,
  title
) {
  dendrogram_data <- dendro_data(hc)

  # Associate each dendrogram leaf with its assigned cluster.
  leaf_ranges <- dendrogram_data$labels |>
    mutate(
      cluster = factor(
        cluster_membership[as.character(label)]
      )
    ) |>
    group_by(cluster) |>
    summarise(
      xmin = min(x),
      xmax = max(x),
      .groups = "drop"
    )

  # Color a branch only when both endpoints fall within the horizontal range
  # occupied by one cluster. Branches joining multiple clusters remain gray.
  assign_segment_cluster <- function(x, xend) {
    matching_ranges <- leaf_ranges |>
      filter(
        x >= xmin,
        x <= xmax,
        xend >= xmin,
        xend <= xmax
      )

    if (nrow(matching_ranges) == 1) {
      as.character(matching_ranges$cluster)
    } else {
      NA_character_
    }
  }

  colored_segments <- dendrogram_data$segments |>
    mutate(
      cluster = map2_chr(
        x,
        xend,
        assign_segment_cluster
      )
    )

  ggplot(
    colored_segments,
    aes(
      x = x,
      y = y,
      xend = xend,
      yend = yend,
      color = cluster
    )
  ) +
    geom_segment(linewidth = 0.3) +
    scale_color_discrete(na.value = "grey60") +
    labs(
      title = title,
      x = NULL,
      y = "Height",
      color = "Cluster"
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank()
    )
}
```

```{r}
# Principal component analysis
# PCA is used only for two-dimensional visualization of the clusters.
pca_result <- prcomp(feature_matrix_scaled)

# Agglomerative hierarchical clustering: AGNES with Ward linkage
# AGNES begins with each day in its own cluster and repeatedly merges the
# closest clusters. Ward linkage favors merges that minimize the increase in
# within-cluster variance.
agnes_result <- agnes(
  do_distance,
  method = "ward"
)

agnes_hc <- as.hclust(agnes_result)

# Values closer to one indicate a stronger hierarchical clustering structure.
agnes_coefficient <- agnes_result$ac
agnes_coefficient

plot_dendrogram(
  hc = agnes_hc,
  title = "Agglomerative clustering (AGNES, Ward)"
)

# Estimate the number of AGNES clusters with NbClust
# NbClust evaluates several cluster-validity indices. The selected value is
# the number of clusters receiving the most recommendations across indices.
nbclust_agnes <- NbClust::NbClust(
  data = feature_matrix_scaled,
  distance = "euclidean",
  min.nc = 2,
  max.nc = 8,
  method = "ward.D2",
  index = "all"
)

agnes_k_nbclust <- nbclust_agnes$Best.nc["Number_clusters", ] |>
  table() |>
  (\(counts) {
    as.numeric(names(counts)[which.max(counts)])
  })()

agnes_clusters_nbclust <- cutree(
  agnes_hc,
  k = agnes_k_nbclust
)

plot_pca_clusters(
  pca_result = pca_result,
  cluster_membership = agnes_clusters_nbclust,
  title = "AGNES day clusters in PCA space"
)

plot_colored_dendrogram(
  hc = agnes_hc,
  cluster_membership = agnes_clusters_nbclust,
  title = "Agglomerative clustering (AGNES, Ward)"
)

# Estimate the number of AGNES clusters with silhouette width
agnes_silhouettes <- calculate_silhouette_widths(
  hc = agnes_hc,
  distance_matrix = do_distance,
  candidate_k = 2:8
)

agnes_k_silhouette <- select_best_k(agnes_silhouettes)
agnes_k_silhouette

plot_silhouette_widths(
  silhouette_results = agnes_silhouettes,
  title = "Silhouette width by cluster count (AGNES, Ward)"
)
# Divisive hierarchical clustering: DIANA

# DIANA begins with all days in one cluster and repeatedly splits clusters
# until each observation is separate.
diana_result <- diana(do_distance)
diana_hc <- as.hclust(diana_result)

# The divisive coefficient is analogous to the AGNES agglomerative
# coefficient. Values closer to one indicate stronger clustering structure.
diana_coefficient <- diana_result$dc
diana_coefficient

plot_dendrogram(
  hc = diana_hc,
  title = "Divisive clustering (DIANA)"
)


# Estimate the number of DIANA clusters with silhouette width
diana_silhouettes <- calculate_silhouette_widths(
  hc = diana_hc,
  distance_matrix = do_distance,
  candidate_k = 2:8
)

diana_k_silhouette <- select_best_k(diana_silhouettes)
diana_k_silhouette

plot_silhouette_widths(
  silhouette_results = diana_silhouettes,
  title = "Silhouette width by cluster count (DIANA)"
)

diana_clusters <- cutree(
  diana_hc,
  k = diana_k_silhouette
)

plot_pca_clusters(
  pca_result = pca_result,
  cluster_membership = diana_clusters,
  title = "DIANA day clusters in PCA space"
)

plot_colored_dendrogram(
  hc = diana_hc,
  cluster_membership = diana_clusters,
  title = "Divisive clustering (DIANA)"
)
```

```{r}
# Keep an index so you can join cluster labels back, then pull the numeric DO profile and drop incomplete days
do_clean <- do_metrics_daily |>
  mutate(row_id = row_number()) |>
  filter(if_all(
    c(avgDO.pctsat, DOpctsat.amp, DOSat.exc, probHypSat, minDO.pctsat),
    ~ !is.na(.)
  ))

feat <- do_clean |>
  select(avgDO.pctsat, DOpctsat.amp, DOSat.exc, probHypSat, minDO.pctsat)

# Standardize
feat_scaled <- scale(feat)

# Choose eps from the k-nearest-neighbor distance plot.
# Rule of thumb: minPts = (number of dimensions) + 1, often a bit higher.
# Look for the "knee" / elbow in the curve; that y-value is your eps.
minPts <- 6
dbscan::kNNdistplot(feat_scaled, k = minPts - 1)
abline(h = 0.8, lty = 2) # move this line to wherever the elbow sits

# Run the three clusterings
res_08 <- suppressMessages(dbscan::dbscan(feat_scaled, eps = 0.8, minPts = 6))
res_05 <- suppressMessages(dbscan::dbscan(feat_scaled, eps = 0.5, minPts = 6))
res_h <- suppressMessages(dbscan::hdbscan(feat_scaled, minPts = 6))
res_h50 <- suppressMessages(dbscan::hdbscan(feat_scaled, minPts = 50))
# PCA on the same scaled features
# feat_scaled is already centered + scaled, so center/scale = FALSE here
pca <- prcomp(feat_scaled, center = FALSE, scale. = FALSE)

var_explained <- pca$sdev^2 / sum(pca$sdev^2)
pc1_lab <- sprintf("PC1 (%.1f%%)", 100 * var_explained[1])
pc2_lab <- sprintf("PC2 (%.1f%%)", 100 * var_explained[2])

# Assemble a plotting frame with all three label sets
plot_df <- do_clean |>
  mutate(
    PC1 = pca$x[, 1],
    PC2 = pca$x[, 2],
    `DBSCAN eps=0.8` = factor(res_08$cluster),
    `DBSCAN eps=0.5` = factor(res_05$cluster),
    HDBSCAN = factor(res_h$cluster)
  )

# One reusable panel builder. Cluster 0 (noise) drawn grey
make_panel <- function(df, cluster_col, title) {
  df <- df |>
    mutate(.clust = .data[[cluster_col]], .noise = .clust == "0")

  ggplot(df, aes(PC1, PC2)) +
    geom_point(
      data = ~ filter(.x, .noise),
      color = "grey75",
      size = 0.7,
      alpha = 0.6
    ) +
    geom_point(
      data = ~ filter(.x, !.noise),
      aes(color = .clust),
      size = 0.9,
      alpha = 0.8
    ) +
    labs(title = title, x = pc1_lab, y = pc2_lab, color = "Cluster") +
    theme_minimal(base_size = 11) +
    theme(legend.position = "right")
}

p1 <- make_panel(plot_df, "DBSCAN eps=0.8", "DBSCAN, eps = 0.8")
p2 <- make_panel(plot_df, "DBSCAN eps=0.5", "DBSCAN, eps = 0.5")
p3 <- make_panel(plot_df, "HDBSCAN", "HDBSCAN, minPts = 6")
p4 <- make_panel(plot_df, "HDBSCAN", "HDBSCAN, minPts = 50")
# Stack the three panels
combined <- p1 /
  p2 /
  p3 /
  p4 +
  plot_annotation(
    title = "DO daily profiles in PCA space, by clustering method"
  )
combined
# Quick numeric comparison of how each method partitions the days
cat("DBSCAN eps = 0.8\n")
print(table(res_08$cluster))
cat("\nDBSCAN eps = 0.5\n")
print(table(res_05$cluster))
cat("\nHDBSCAN\n")
print(table(res_h$cluster))
```

```{r}
# suppressPackageStartupMessages({
#   library(mclust)
#   library(diptest)
# })

set.seed(42)

X <- do_metrics_daily |>
  select(all_of(metric_names)) |>
  as.matrix()

# --- Part A: Gaussian mixture, model selection by BIC and ICL ------------
# BIC can prefer G = 1; ICL penalises overlapping components. If multi-
# component models win, the question becomes whether they win by being
# genuinely separated or just flexible. The separation diagnostic answers that.
mc_bic <- suppressMessages(
  mclust::mclustBIC(X, G = 1:9)
)
mclustBIC <- suppressMessages(
  mclust::mclustBIC
)
mc_icl <- suppressMessages(
  mclust::mclustICL(X, G = 1:9)
)
print(summary(mc_bic))
print(summary(mc_icl))

mc <- suppressMessages(
  mclust::Mclust(X, x = mc_bic)
)
cat("Best G by BIC:", mc$G, "\n")

# Separation: how far apart are component centroids relative to within spread?
# Computed from MAP labels so it is independent of the covariance parameterisation.
lab <- mc$classification
centroids <- X |>
  as.data.frame() |>
  mutate(.lab = lab) |>
  group_by(.lab) |>
  summarise(across(everything(), mean), .groups = "drop") |>
  select(-.lab) |>
  as.matrix()

within_sd <- X |>
  as.data.frame() |>
  mutate(.lab = lab) |>
  group_by(.lab) |>
  summarise(across(everything(), sd), .groups = "drop") |>
  select(-.lab) |>
  as.matrix() |>
  mean()

min_centroid_gap <- min(dist(centroids))
cat("Min centroid distance:", round(min_centroid_gap, 2), "\n")
cat("Typical within-component sd:", round(within_sd, 2), "\n")
cat("Separation ratio:", round(min_centroid_gap / within_sd, 2), "\n")
# Ratio well above ~2 with a visible density gap argues for discrete states;
# a ratio near or below 1 means components overlap -> flexible fit of a continuum.

# --- Part B: Gap statistic (supports k = 1) ------------------------------
gap <- cluster::clusGap(X, FUN = kmeans, K.max = 9, B = 100, nstart = 25)
print(gap, method = "Tibs2001SEmax")
# If this selects k = 1, that is direct evidence against discrete states.

# --- Part C: Decorrelation-aware dip test of unimodality -----------------
acf_decay_lag <- function(x, threshold = 1 / exp(1), lag_max = 90) {
  a <- as.numeric(acf(x, plot = FALSE, lag.max = lag_max)$acf)[-1]
  hit <- which(a < threshold)
  if (length(hit) == 0) lag_max else hit[1]
}

site_stride <- do_metrics_daily |>
  group_by(kg_name) |>
  summarise(across(all_of(metric_names), acf_decay_lag), .groups = "drop") |>
  rowwise() |>
  mutate(stride = max(c_across(all_of(metric_names)))) |>
  ungroup() |>
  select(kg_name, stride)

thinned <- do_metrics_daily |>
  arrange(kg_name, Date) |>
  left_join(site_stride, by = "kg_name") |>
  group_by(kg_name) |>
  filter((row_number() - 1) %% first(stride) == 0) |>
  ungroup()

cat("Rows after thinning:", nrow(thinned), "of", nrow(do_metrics_daily), "\n")

Xt <- thinned |>
  select(all_of(metric_names)) |>
  as.matrix()

# Dip on the dominant unsupervised axis (PC1), full vs thinned.
pc1_full <- prcomp(X)$x[, 1]
pc1_thin <- prcomp(Xt)$x[, 1]
print(diptest::dip.test(pc1_full))
print(diptest::dip.test(pc1_thin))
# If multimodality on PC1 disappears after thinning, it was pseudoreplication.

# Per-metric dip on thinned data (remember probHypSat's zero point mass).
for (m in metric_names) {
  d <- diptest::dip.test(Xt[, m])
  cat(sprintf("%-14s dip = %.4f  p = %.4f\n", m, d$statistic, d$p.value))
}

# --- Part D: Cluster stability as corroboration --------------------------
# Real states recur under resampling; boundaries on a continuum do not.
cb <- suppressMessages(
  fpc::clusterboot(
    Xt,
    B = 100,
    clustermethod = fpc::kmeansCBI,
    krange = 3,
    seed = 42
  )
)
cat("Jaccard stability per cluster:\n")
print(round(cb$bootmean, 2))
# Values above ~0.85 = stable/real; 0.6-0.75 = ambiguous; below ~0.6 = not meaningful.
```

# K-Means & PCA with & w/o Mean DO
```{r}
majority_rule <- 3

run_pca_kmeans <- function(data, predictors, k) {
  pca_data <- data |>
    select(kg_name, Date, all_of(predictors)) |>
    mutate(probHypSat = case_when(is.nan(probHypSat) ~ 0, TRUE ~ probHypSat))

  pca_recipe <- recipe(~., data = pca_data) |>
    update_role(kg_name, Date, new_role = "Id") |>
    step_naomit(all_predictors()) |>
    step_normalize(all_predictors()) |>
    step_pca(all_predictors(), id = "PCA")

  pca_fit <- prep(pca_recipe)
  scores <- juice(pca_fit)

  km <- scores |>
    select(PC1:PC2) |>
    as.matrix() |>
    eclust(
      FUNcluster = "kmeans",
      k = k,
      nstart = 25,
      graph = FALSE,
      nboot = 500
    )

  scores |>
    mutate(Cluster = as_factor(km$cluster)) |>
    left_join(
      data |> select(kg_name, Date, avgDO.pctsat),
      by = c("kg_name", "Date")
    ) |>
    group_by(Cluster) |>
    mutate(cluster_mean_DO = mean(avgDO.pctsat, na.rm = TRUE)) |>
    ungroup() |>
    mutate(Cluster = as_factor(dense_rank(desc(cluster_mean_DO)))) |>
    select(-cluster_mean_DO)
}

# Predictor sets
vars_no_mean <- c(
  "DOpctsat.amp",
  "DOSat.exc",
  "probHypSat",
  "minDO.pctsat",
  "maxDO.pctsat"
)
vars_with_mean <- c(vars_no_mean, "avgDO.pctsat")

# Run both versions (same k so cluster counts are comparable)
scores_no_mean <- run_pca_kmeans(dailyDat, vars_no_mean, majority_rule)
scores_with_mean <- run_pca_kmeans(dailyDat, vars_with_mean, majority_rule)

# Compare cluster assignments row by row
comparison <- scores_no_mean |>
  select(kg_name, Date, Cluster_no_mean = Cluster) |>
  inner_join(
    scores_with_mean |> select(kg_name, Date, Cluster_with_mean = Cluster),
    by = c("kg_name", "Date")
  ) |>
  mutate(changed = Cluster_no_mean != Cluster_with_mean)

n_changed <- sum(comparison$changed)
pct_changed <- mean(comparison$changed) * 100

# Confusion-style table of how assignments map between the two runs
table(comparison$Cluster_no_mean, comparison$Cluster_with_mean)
```

# K-Means on Raw DO with & w/o Mean DO
```{r}
run_kmeans_raw <- function(data, predictors, k) {
  raw_data <- data |>
    select(kg_name, Date, all_of(predictors)) |>
    mutate(probHypSat = case_when(is.nan(probHypSat) ~ 0, TRUE ~ probHypSat)) |>
    drop_na()

  z_scores <- raw_data |>
    select(all_of(predictors)) |>
    scale() |>
    as.data.frame()

  km <- z_scores |>
    as.matrix() |>
    eclust(
      FUNcluster = "kmeans",
      k = k,
      nstart = 25,
      graph = FALSE,
      nboot = 500
    )

  raw_data |>
    select(kg_name, Date) |>
    mutate(Cluster = as_factor(km$cluster)) |>
    left_join(
      data |> select(kg_name, Date, avgDO.pctsat),
      by = c("kg_name", "Date")
    ) |>
    group_by(Cluster) |>
    mutate(cluster_mean_DO = mean(avgDO.pctsat, na.rm = TRUE)) |>
    ungroup() |>
    mutate(Cluster = as_factor(dense_rank(desc(cluster_mean_DO)))) |>
    select(-cluster_mean_DO)
}

# Run raw z-score k-means
scores_raw <- run_kmeans_raw(dailyDat, vars_no_mean, majority_rule)

# You could also run it with avgDO.pctsat included
scores_raw_with_mean <- run_kmeans_raw(dailyDat, vars_with_mean, majority_rule)
```

# Compare K-Means with & w/o PCA and with & w/o Mean DO
```{r}
all_clusters <- scores_no_mean |>
  select(kg_name, Date, PCA_no_mean = Cluster) |>
  inner_join(
    scores_with_mean |> select(kg_name, Date, PCA_with_mean = Cluster),
    by = c("kg_name", "Date")
  ) |>
  inner_join(
    scores_raw |> select(kg_name, Date, Raw_no_mean = Cluster),
    by = c("kg_name", "Date")
  ) |>
  inner_join(
    scores_raw_with_mean |> select(kg_name, Date, Raw_with_mean = Cluster),
    by = c("kg_name", "Date")
  )

# Pairwise agreement rates
pairwise_agreement <- function(col_a, col_b) {
  mean(all_clusters[[col_a]] == all_clusters[[col_b]])
}

comparisons <- combn(
  c("PCA_no_mean", "PCA_with_mean", "Raw_no_mean", "Raw_with_mean"),
  2,
  simplify = FALSE
)

agreement_summary <- map_dfr(comparisons, function(pair) {
  tibble(
    comparison = paste(pair[1], "vs", pair[2]),
    pct_agree = pairwise_agreement(pair[1], pair[2]) * 100
  )
})

agreement_summary

# Confusion matrix for the specific PCA vs raw comparison
table(all_clusters$PCA_no_mean, all_clusters$Raw_no_mean)
```
```{r}
library(ggalluvial)

alluvial_data <- all_clusters |>
  count(PCA_with_mean, PCA_no_mean, Raw_with_mean, Raw_no_mean) |>
  mutate(origin_cluster = PCA_with_mean) |>
  to_lodes_form(
    axes = 1:4,
    id = "site_day"
  )

n_total <- nrow(all_clusters)


# ---- Plot 1: PCA mean-DO effect only -------------------------------
changed_pca_only <- all_clusters |>
  filter(PCA_with_mean != PCA_no_mean)

n_changed_pca_only <- nrow(changed_pca_only)

alluvial_data_pca <- changed_pca_only |>
  count(PCA_with_mean, PCA_no_mean) |>
  mutate(origin_cluster = PCA_with_mean) |>
  to_lodes_form(
    axes = 1:2,
    id = "site_day"
  )

subtitle_pca <- sprintf(
  "Showing %d of %d days (%.1f%%) reassigned when mean DO is removed",
  n_changed_pca_only,
  n_total,
  n_changed_pca_only / n_total * 100
)

alluvial_plt_pca <- alluvial_data_pca |>
  ggplot(aes(x = x, stratum = stratum, alluvium = site_day, y = n)) +
  geom_flow(
    aes(fill = factor(origin_cluster)),
    alpha = 0.55,
    curve_type = "sigmoid",
    width = 0.3
  ) +
  geom_stratum(width = 0.3, fill = "grey88", color = "white", linewidth = 0.6) +
  geom_text(
    stat = "stratum",
    aes(label = paste0(after_stat(stratum), "\n(n = ", after_stat(count), ")")),
    size = 3.5,
    fontface = "bold",
    color = "grey20"
  ) +
  scale_x_discrete(
    limits = c("PCA_with_mean", "PCA_no_mean"),
    labels = c("PCA\n(with mean DO)", "PCA\n(no mean DO)"),
    expand = c(0.15, 0.15)
  ) +
  scale_fill_manual(
    values = cluster_colors,
    name = "Cluster"
  ) +
  labs(
    title = "Days Reassigned When Mean DO Is Removed (PCA)",
    subtitle = subtitle_pca,
    x = NULL,
    y = "Number of Days"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(color = "grey30", size = 10.5),
    legend.position = "bottom",
    panel.grid = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 11, color = "grey20")
  )

alluvial_plt_pca

ggsave(
  filename = "cluster_alluvial_pca_mean_do_effect.png",
  plot = alluvial_plt_pca,
  path = here("Figs/Supplement/"),
  device = "png",
  width = 18,
  height = 15,
  units = "cm",
  dpi = 600
)

# ---- Plot 2: PCA vs. raw disagreement, no mean DO -------------------
changed_method_only <- all_clusters |>
  filter(PCA_no_mean != Raw_no_mean)

n_changed_method_only <- nrow(changed_method_only)

alluvial_data_method <- changed_method_only |>
  count(PCA_no_mean, Raw_no_mean) |>
  mutate(origin_cluster = PCA_no_mean) |>
  to_lodes_form(
    axes = 1:2,
    id = "site_day"
  )

subtitle_method <- sprintf(
  "Showing %d of %d days (%.1f%%) assigned to different clusters by PCA vs. raw z-scores",
  n_changed_method_only,
  n_total,
  n_changed_method_only / n_total * 100
)

alluvial_plt_method <- alluvial_data_method |>
  ggplot(aes(x = x, stratum = stratum, alluvium = site_day, y = n)) +
  geom_flow(
    aes(fill = factor(origin_cluster)),
    alpha = 0.55,
    curve_type = "sigmoid",
    width = 0.3
  ) +
  geom_stratum(width = 0.3, fill = "grey88", color = "white", linewidth = 0.6) +
  geom_text(
    stat = "stratum",
    aes(label = paste0(after_stat(stratum), "\n(n = ", after_stat(count), ")")),
    size = 3.5,
    fontface = "bold",
    color = "grey20"
  ) +
  scale_x_discrete(
    limits = c("PCA_no_mean", "Raw_no_mean"),
    labels = c("PCA\n(no mean DO)", "Raw z-score\n(no mean DO)"),
    expand = c(0.15, 0.15)
  ) +
  scale_fill_manual(
    values = cluster_colors,
    name = "Cluster"
  ) +
  labs(
    title = "Days Where PCA and Raw Z-Score Clustering Disagree",
    subtitle = subtitle_method,
    x = NULL,
    y = "Number of Days"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(color = "grey30", size = 10.5),
    legend.position = "bottom",
    panel.grid = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_text(size = 11, color = "grey20")
  )

alluvial_plt_method

ggsave(
  filename = "cluster_alluvial_pca_vs_raw_disagreement.png",
  plot = alluvial_plt_method,
  path = here("Figs/Supplement/"),
  device = "png",
  width = 18,
  height = 15,
  units = "cm",
  dpi = 600
)
```

# K-Means on Raw DO

```{r}
# Prepare daily dissolved oxygen metrics
do_metrics_daily <- dailyDat |>
  select(
    kg_name,
    Date,
    avgDO.pctsat,
    DOpctsat.amp,
    DOSat.exc,
    probHypSat,
    minDO.pctsat
  ) |>
  mutate(
    # Treat undefined hypoxia probabilities as zero.
    probHypSat = replace(probHypSat, is.nan(probHypSat), 0)
  ) |>
  drop_na() |>
  as_tibble() |>
  mutate(across(where(is.numeric), \(x) (x - mean(x)) / sd(x)))

# Build and standardize the clustering feature matrix
metric_names <- c(
  "DOpctsat.amp",
  "DOSat.exc",
  "probHypSat",
  "avgDO.pctsat"
)

do_features <- do_metrics_daily |>
  mutate(obs_id = paste(kg_name, Date, sep = "_")) |>
  select(obs_id, all_of(metric_names))

# Convert the metric columns to a numeric matrix and preserve the observation
# identifier as the row name.
feature_matrix <- do_features |>
  select(all_of(metric_names)) |>
  as.matrix()

rownames(feature_matrix) <- do_features$obs_id

# Clustering tendency: Hopkins statistic + ordered dissimilarity (VAT).
# If the value of Hopkins statistic is close to 1 (far above 0.5), then we can conclude that the dataset is significantly clusterable.
clust_tendency <- feature_matrix |>
  factoextra::get_clust_tendency(
    n = nrow(feature_matrix) - 1,
    graph = FALSE,
    seed = 42
  )
clust_tendency$hopkins_stat

# Unimodality tests per metric: Hartigan dip test + Silverman critical-bandwidth
# For both dip.test and modetest, the null is unimodality, so a small p-value is evidence for multiple modes (it supports clustering)
modality_tests <- metric_names |>
  purrr::map(\(m) {
    x <- feature_matrix[, m]
    tibble(
      metric = m,
      dip_p = diptest::dip.test(x)$p.value,
      silverman_p = multimode::modetest(x, mod0 = 1, method = "SI")$p.value
    )
  }) |>
  purrr::list_rbind()

modality_tests

nbclust_res <- NbClust::NbClust(
  data = feature_matrix,
  distance = "euclidean",
  min.nc = 2,
  max.nc = 9,
  method = "kmeans",
  index = "all"
)

majority_rule <- nbclust_res |>
  (\(x) {
    tbl <- table(x$Best.nc["Number_clusters", ])
    as.numeric(names(tbl)[which.max(tbl)])
  })()

# elbow method
res_ch <- NbClust::NbClust(
  data = feature_matrix,
  distance = "euclidean",
  min.nc = 3,
  max.nc = 15,
  method = "kmeans",
  index = "ch"
)
# Silhouette index
res_sil <- NbClust::NbClust(
  data = feature_matrix,
  distance = "euclidean",
  min.nc = 3,
  max.nc = 15,
  method = "kmeans",
  index = "silhouette"
)
majority_rule <- tibble(
  method = c("Elbow", "Silhouette"),
  num_clust = c(
    res_ch$Best.nc[["Number_clusters"]],
    res_sil$Best.nc[["Number_clusters"]]
  )
) |>
  (\(df) {
    floor(mean(df$num_clust))
  })()

# Cluster stability via bootstrap (Jaccard similarity per cluster)
set.seed(42)
suppressMessages({
  clusterboot_res <- feature_matrix |>
    fpc::clusterboot(
      B = 500,
      bootmethod = "boot",
      clustermethod = fpc::kmeansCBI,
      k = majority_rule,
      runs = 100,
      seed = 42,
      count = FALSE
    )
})


#K-means clustering
kmeans_result <- feature_matrix |>
  eclust(
    FUNcluster = "kmeans", # Use K-means clustering
    k = majority_rule, # Number of clusters
    nstart = 25, # Number of random starts for K-means
    graph = FALSE,
    nboot = 500
  )

performance::model_performance(kmeans_result)
# Davies-Bouldin and WCSS lower-is-better, Silhouette and Calinski-Harabasz higher-is-better.
internal_summary <- tibble(
  metric = c("Davies-Bouldin", "Silhouette", "Calinski-Harabasz", "WCSS"),
  value = c(
    clusterSim::index.DB(feature_matrix, kmeans_result$cluster)$DB,
    kmeans_result$silinfo$avg.width,
    fpc::calinhara(feature_matrix, kmeans_result$cluster),
    kmeans_result$tot.withinss
  )
)
internal_summary
# Rebuild the raw-label -> DO-ordered-label mapping used in clustered_dat
cluster_relabel <- do_metrics_daily |>
  mutate(raw_cluster = kmeans_result$cluster) |>
  group_by(raw_cluster) |>
  summarise(
    cluster_mean_DO = mean(avgDO.pctsat, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(cluster = dense_rank(desc(cluster_mean_DO))) |>
  select(raw_cluster, cluster)

# Mean Jaccard stability per cluster, aligned to the plotted cluster labels
cluster_stability <- tibble(
  raw_cluster = seq_along(clusterboot_res$bootmean),
  jaccard_mean = clusterboot_res$bootmean,
  n_dissolved = clusterboot_res$bootbrd,
  n_recovered = clusterboot_res$bootrecover
) |>
  left_join(cluster_relabel, by = "raw_cluster") |>
  arrange(cluster) |>
  select(cluster, jaccard_mean, n_dissolved, n_recovered)

# Values above about 0.85 indicate highly stable clusters, 0.6 to 0.75 suggests the cluster reflects a real pattern but with uncertainty about which points belong, and below 0.6 means the cluster shouldn't be trusted as a distinct group.
cluster_stability

#Add cluster information back to the original dataset
clustered_dat <- do_metrics_daily |>
  mutate(Cluster = as_factor(kmeans_result$cluster)) |>
  mutate(kg_name = fct_relevel(kg_name, levels(site_meta$kg_name))) |>
  group_by(Cluster) |>
  mutate(cluster_mean_DO = mean(avgDO.pctsat, na.rm = TRUE)) |>
  ungroup() |>
  mutate(Cluster = as_factor(dense_rank(desc(cluster_mean_DO)))) |>
  select(-cluster_mean_DO)


do_metrics <- c(
  "avgDO.pctsat" = "Avg.",
  "DOpctsat.amp" = "Amplitude",
  "DOSat.exc" = "Exceedence\nProbability",
  "probHypSat" = "Probability of\nHypoxia",
  "minDO.pctsat" = "Min",
  "maxDO.pctsat" = "Max"
)
# Reshape data to long format for faceting
metrics <- c(
  "avgDO.pctsat",
  "DOpctsat.amp",
  "DOSat.exc",
  "probHypSat"
)

plot_data <- clustered_dat |>
  filter(!is.na(Cluster)) |>
  select(kg_name, Cluster, any_of(names(do_metrics))) |>
  pivot_longer(
    cols = any_of(names(do_metrics)),
    names_to = "Metric",
    values_to = "Value"
  ) |>
  mutate(Metric = recode(Metric, !!!do_metrics))

# Create faceted plot
ggplot(plot_data, aes(x = Cluster, y = Value, fill = Cluster)) +
  geom_boxplot(outliers = FALSE) +
  scale_fill_manual(values = cluster_colors) +
  scale_color_manual(values = cluster_colors) +
  geom_jitter(
    alpha = 0.1,
    aes(color = Cluster),
    position = position_jitterdodge(jitter.width = 0.5, dodge.width = 0.75)
  ) +
  facet_wrap(~Metric, ncol = 3, scales = "free") +
  labs(title = "DO Metrics by Cluster", y = "Value", x = "Cluster") +
  theme_minimal() +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 12, face = "bold")
  )

# Create faceted plot
site_cluster_plt <- plot_data |>
  ggplot(aes(x = kg_name, y = Value, fill = Cluster)) +
  geom_boxplot(outliers = FALSE) +
  scale_fill_manual(values = cluster_colors) +
  scale_color_manual(values = cluster_colors) +
  geom_jitter(
    alpha = 0.1,
    aes(color = Cluster),
    position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.75)
  ) +
  facet_wrap(~Metric, ncol = 3, scales = "free") +
  labs(title = "DO Metrics by Cluster", y = "Value", x = "") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
site_cluster_plt

# PCA on the already-standardized feature matrix
pca_res <- prcomp(feature_matrix)

var_explained <- summary(pca_res)$importance["Proportion of Variance", ]

# Bind PC scores to the relabeled clusters (row order matches)
pca_plot_data <- clustered_dat |>
  mutate(
    PC1 = pca_res$x[, 1],
    PC2 = pca_res$x[, 2]
  )

ggplot(pca_plot_data, aes(x = PC1, y = PC2, color = Cluster)) +
  geom_point(alpha = 0.3, size = 1) +
  stat_ellipse(
    aes(color = Cluster),
    geom = "polygon",
    level = 0.95,
    linewidth = 0.8,
    alpha = 0.2,
    fill = NA
  ) +
  scale_color_manual(values = cluster_colors) +
  labs(
    title = "Clusters in PC space",
    x = paste0("PC1 (", round(100 * var_explained[1], 1), "%)"),
    y = paste0("PC2 (", round(100 * var_explained[2], 1), "%)")
  ) +
  theme_minimal()

# t-SNE on the already-standardized feature matrix
set.seed(42)
tsne_res <- feature_matrix |>
  Rtsne::Rtsne(
    dims = 2,
    perplexity = 30,
    pca = FALSE,
    normalize = FALSE,
    check_duplicates = FALSE,
    num_threads = 0
  )

# Bind t-SNE coordinates to the relabeled clusters (row order matches)
tsne_plot_data <- clustered_dat |>
  mutate(
    tSNE1 = tsne_res$Y[, 1],
    tSNE2 = tsne_res$Y[, 2]
  )

ggplot(tsne_plot_data, aes(x = tSNE1, y = tSNE2, color = Cluster)) +
  geom_point(alpha = 0.3, size = 1) +
  stat_ellipse(
    aes(color = Cluster),
    geom = "polygon",
    level = 0.95,
    linewidth = 0.8,
    alpha = 0.2,
    fill = NA
  ) +
  scale_color_manual(values = cluster_colors) +
  labs(
    title = "Clusters in t-SNE space",
    x = "t-SNE 1",
    y = "t-SNE 2"
  ) +
  theme_minimal()

sil_data <- kmeans_result$silinfo$widths |>
  rownames_to_column("obs_id") |>
  rename(raw_cluster = cluster) |>
  mutate(raw_cluster = as.integer(as.character(raw_cluster))) |>
  left_join(cluster_relabel, by = "raw_cluster") |>
  mutate(Cluster = as_factor(cluster)) |>
  arrange(Cluster, desc(sil_width)) |>
  mutate(obs_order = row_number())

sil_plt <- ggplot(sil_data, aes(x = obs_order, y = sil_width, fill = Cluster)) +
  geom_col(width = 1) +
  geom_hline(
    yintercept = mean(sil_data$sil_width),
    linetype = "dashed"
  ) +
  scale_fill_manual(values = cluster_colors) +
  coord_flip() +
  labs(
    title = "Silhouette plot",
    x = "",
    y = "Silhouette width"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
sil_plt
```

# K-Means on PCA scores

## Optimal Number of Cluster Selection

```{r}
# nbclust_res <- NbClust::NbClust(data = pca_scores |> select(PC1:PC2) |> as.matrix(),
#                                   distance = "euclidean",
#                                   min.nc = 2, max.nc = 9,
#                                   method = "kmeans",
#                                   index = "all")
#
# majority_rule <- nbclust_res |>
#   (\(x) {
#     tbl <- table(x$Best.nc["Number_clusters", ])
#     as.numeric(names(tbl)[which.max(tbl)])
#   })()
#
# # elbow method
# res_ch <- NbClust::NbClust(
#   data = pca_scores |> select(PC1:PC2) |> as.matrix(),
#   distance = "euclidean",
#   min.nc = 3,
#   max.nc = 15,
#   method = "kmeans",
#   index = "ch"
# )
# # Silhouette index
# res_sil <- NbClust::NbClust(
#   data = pca_scores |> select(PC1:PC2) |> as.matrix(),
#   distance = "euclidean",
#   min.nc = 3,
#   max.nc = 15,
#   method = "kmeans",
#   index = "silhouette"
# )
# majority_rule <- tibble(
#   method = c("Elbow", "Silhouette"),
#   num_clust = c(res_ch$Best.nc[["Number_clusters"]], res_sil$Best.nc[["Number_clusters"]])) |>
#   (\(df){floor(mean(df$num_clust))})()
majority_rule <- 3
```

```{r}
#K-means clustering
kmeans_result <- pca_scores |>
  select(PC1:PC2) |>
  as.matrix() |>
  eclust(
    FUNcluster = "kmeans", # Use K-means clustering
    k = majority_rule, # Number of clusters
    nstart = 25, # Number of random starts for K-means
    graph = F,
    nboot = 500
  )
pca_scores <- pca_scores |>
  mutate(Cluster = as_factor(kmeans_result$cluster)) |>
  left_join(
    dailyDat |> select(kg_name, Date, avgDO.pctsat),
    by = c("kg_name", "Date")
  ) |>
  group_by(Cluster) |>
  mutate(cluster_mean_DO = mean(avgDO.pctsat, na.rm = TRUE)) |>
  ungroup() |>
  mutate(Cluster = as_factor(dense_rank(desc(cluster_mean_DO)))) |>
  select(-cluster_mean_DO)

#Add cluster information back to the original dataset
clustered_dat <- dailyDat |>
  left_join(
    pca_scores |> select(kg_name, Date, Cluster),
    by = c("kg_name", "Date")
  ) |>
  mutate(kg_name = fct_relevel(kg_name, levels(site_meta$kg_name))) |>
  write_csv(here("Data/rf_data.csv"))
```

## Plot K-Means

```{r}
# Summary statistics by cluster
cluster_summary <- clustered_dat |>
  drop_na(Cluster) |>
  group_by(Cluster) |>
  summarise(
    n_sites = n_distinct(Site),
    n_observations = n(),
    avg_DO_pctsat = mean(avgDO.pctsat, na.rm = TRUE),
    avg_DO_amp = mean(DOpctsat.amp, na.rm = TRUE),
    avg_DO_exc = mean(DOSat.exc, na.rm = TRUE),
    avg_ProbHyp = mean(probHypSat, na.rm = TRUE),
    .groups = "drop"
  )

cli::cli_inform("Cluster summary statistics:")
print(cluster_summary)

kmeans_plt <- ggplot(pca_scores, aes(x = PC1, y = PC2)) +
  geom_point(
    aes(fill = Cluster, shape = kg_name, color = Cluster),
    size = 3,
    alpha = 0.6,
    stroke = 0.3
  ) +
  stat_ellipse(
    aes(color = Cluster),
    geom = "polygon",
    level = 0.95,
    alpha = 0.2,
    fill = NA
  ) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_vline(xintercept = 0, linetype = "dashed") +
  # Add loading arrows
  geom_segment(
    data = loadings,
    aes(x = 0, y = 0, xend = PC1, yend = PC2),
    arrow = arrow(length = unit(0.2, "cm")),
    color = "black"
  ) +
  # Use ggrepel for smart label positioning
  geom_text_repel(
    data = loadings,
    aes(x = PC1 * 1.1, y = PC2 * 1.1, label = var_label),
    size = 5,
    direction = "both"
  ) +
  scale_color_manual(values = cluster_colors) +
  scale_fill_manual(values = cluster_colors) +
  #scale_shape(solid = FALSE) +
  scale_shape_manual(
    values = setNames(site_meta$shape, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  labs(
    x = pc1_label,
    y = pc2_label,
    shape = "Watershed",
    color = "State",
    fill = "State"
  ) +
  theme(legend.box = "vertical")
kmeans_plt
# Save the plot
ggsave(
  filename = "kmeansPCA.png",
  plot = kmeans_plt,
  path = here("Figs/"),
  device = "png",
  width = 20,
  height = 20,
  units = "cm",
  dpi = "retina"
)
```

```{r}
do_metrics <- c(
  "DOpctsat.amp" = "Amplitude",
  "DOSat.exc" = "Exceedence\nProbability",
  "probHypSat" = "Probability of\nHypoxia",
  "minDO.pctsat" = "Min",
  # "avgDO.pctsat" = "Avg.",
  "maxDO.pctsat" = "Max"
)
# Reshape data to long format for faceting
metrics <- c(
  "DOpctsat.amp",
  "DOSat.exc",
  "probHypSat",
  "minDO.pctsat",
  # "avgDO.pctsat",
  "maxDO.pctsat"
)

plot_data <- clustered_dat |>
  filter(!is.na(Cluster)) |>
  select(kg_name, Cluster, all_of(names(do_metrics))) |>
  pivot_longer(
    cols = all_of(names(do_metrics)),
    names_to = "Metric",
    values_to = "Value"
  ) |>
  mutate(Metric = recode(Metric, !!!do_metrics))

# Create faceted plot
ggplot(plot_data, aes(x = Cluster, y = Value, fill = Cluster)) +
  geom_boxplot(outliers = FALSE) +
  scale_fill_manual(values = cluster_colors) +
  scale_color_manual(values = cluster_colors) +
  geom_jitter(
    alpha = 0.1,
    aes(color = Cluster),
    position = position_jitterdodge(jitter.width = 0.5, dodge.width = 0.75)
  ) +
  facet_wrap(~Metric, ncol = 3, scales = "free") +
  labs(title = "DO Metrics by Cluster", y = "Value", x = "Cluster") +
  theme_minimal() +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 12, face = "bold")
  )

# Create faceted plot
site_cluster_plt <- plot_data |>
  ggplot(aes(x = kg_name, y = Value, fill = Cluster)) +
  geom_boxplot(outliers = FALSE) +
  scale_fill_manual(values = cluster_colors) +
  scale_color_manual(values = cluster_colors) +
  geom_jitter(
    alpha = 0.1,
    aes(color = Cluster),
    position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.75)
  ) +
  facet_wrap(~Metric, ncol = 3, scales = "free") +
  labs(title = "DO Metrics by Cluster", y = "Value", x = "") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
site_cluster_plt
ggsave(
  filename = "site_cluster_metric.png",
  plot = site_cluster_plt,
  path = here("Figs/Supplement"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = "retina"
)
```

# K-Means Validation

```{r}
library(vegan)
# MRPP Validation
mrpp_data <- pca_scores |> select(PC1:PC2) |> as.matrix()

# Run MRPP test
mrpp_result <- mrpp(
  dat = mrpp_data,
  grouping = pca_scores$Cluster,
  permutations = 1000,
  distance = "euclidean"
)

# Group by Site instead of clusters
site_mrpp <- mrpp(
  dat = mrpp_data,
  grouping = pca_scores$kg_name,
  permutations = 1000,
  distance = "euclidean"
)

# Display results
cli::cli_inform(c(
  "v" = "MRPP Cluster Validation Results:",
  " " = "",
  "i" = "Number of clusters: {.val {majority_rule}}",
  "i" = "A-statistic: {.val {round(mrpp_result$A, 4)}}",
  "i" = "P-value: {.val {round(mrpp_result$Pvalue, 4)}}",
  "i" = "Delta (observed): {.val {round(mrpp_result$delta, 4)}}",
  "i" = "Expected delta: {.val {round(mrpp_result$E.delta, 4)}}"
))

# Interpretation
if (mrpp_result$Pvalue < 0.05) {
  if (mrpp_result$A > 0.3) {
    cli::cli_inform(c(
      "v" = "Interpretation: {.strong STRONG} evidence of meaningful cluster separation"
    ))
  } else if (mrpp_result$A > 0.1) {
    cli::cli_inform(c(
      "v" = "Interpretation: {.strong MODERATE} evidence of meaningful cluster separation"
    ))
  } else if (mrpp_result$A > 0.05) {
    cli::cli_inform(c(
      "!" = "Interpretation: {.strong WEAK} but significant cluster separation"
    ))
  } else {
    cli::cli_warn(c(
      "!" = "Interpretation: {.strong SIGNIFICANT} but very weak cluster separation"
    ))
  }
} else {
  cli::cli_warn(c(
    "x" = "Interpretation: {.strong NO} significant evidence of cluster separation"
  ))
}

cli::cli_inform(c(
  "i" = "Testing Site Effect:",
  "i" = "Site-based A-statistic: {.val {round(site_mrpp$A, 4)}}",
  "i" = "Site-based P-value: {.val {round(site_mrpp$Pvalue, 4)}}",
  " " = "",
  "i" = "Cluster-based A-statistic: {.val {round(mrpp_result$A, 4)}}",
  "i" = "Cluster-based P-value: {.val {round(mrpp_result$Pvalue, 4)}}"
))

# Compare effect sizes
if (mrpp_result$A > site_mrpp$A) {
  cli::cli_inform(c("v" = "Clusters explain MORE variance than sites alone"))
} else {
  cli::cli_inform(c("!" = "Sites may be driving the clustering patterns"))
}
```

### Cluster Time-series

```{r}
doDates <- clustered_dat |>
  # filter(Site %in% c("KNZ", "SHN", "YMR")) |>
  mutate(Year = year(Date)) |>
  filter(!is.na(avgDO.pctsat)) |>
  group_by(kg_name, Year) |>
  summarise(
    first_date = min(Date),
    last_date = max(Date),
    .groups = "drop"
  ) |>
  mutate(
    WinterStart_preY = as.Date(paste0(
      as.integer(as.character(Year)) - 1,
      "-12-01"
    )),
    WinterStart = as.Date(paste0(as.integer(as.character(Year)), "-12-01"))
  ) |>
  mutate(
    WinterStart = case_when(
      kg_name == "CFA-1" & Year == "2021" ~ as_date("2021-12-20"),
      T ~ WinterStart
    ),
    WinterStart_preY = case_when(
      kg_name == "CFA-1" & Year == "2022" ~ as_date("2021-12-20"),
      T ~ WinterStart_preY
    )
  ) |>
  mutate(
    WinterStart_preY = case_when(
      kg_name == "DFA-2" & Year == "2023" ~ NA,
      T ~ WinterStart_preY
    )
  )


# Create separate data frame for the SHN pre-first-date grey box
preFirstDate_SHN <- doDates |>
  filter(kg_name == "DFA-2") |>
  mutate(
    xmin = as.Date("2021-06-01"), # or your actual xlim min
    xmax = first_date,
    ymin = -Inf,
    ymax = Inf
  ) |>
  filter(Year == "2023")


# Create the plot
clustered_dat |>
  drop_na(Cluster) |>
  ggplot(aes(x = Date, y = avgDO.pctsat, color = Cluster)) +
  geom_point(size = 3, alpha = 0.8, na.rm = TRUE) +
  facet_wrap(~kg_name, scales = "free_y", ncol = 1) +
  scale_color_manual(values = cluster_colors, name = "Cluster") +
  # geom_rect(data = doDates,
  #           aes(xmin = WinterStart_preY, xmax = first_date, ymin = -Inf, ymax = Inf),
  #           inherit.aes = FALSE,
  #           fill = "#42d4f4", alpha = 0.2) +
  # geom_rect(data = doDates,
  #           aes(xmin = last_date, xmax = WinterStart, ymin = -Inf, ymax = Inf),
  #           inherit.aes = FALSE,
  #           fill = "#9A6324", alpha = 0.2) +
  # geom_rect(data = preFirstDate_SHN,
  #         aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
  #         inherit.aes = FALSE,
  #         fill = "grey5", alpha = 0.2) +
  ylim(0, 100) +
  scale_x_date(labels = date_format("%b '%y")) +
  xlim(as_date("2021-06-01"), as_date("2024-12-01")) +
  labs(
    x = "",
    y = expression("Daily Average DO"["% Sat"])
  ) +
  theme(axis.text = element_text(size = 15)) -> p
p
ggsave(
  filename = "Average.png",
  plot = p,
  path = here("Figs/ClusterTimeseries"),
  device = "png",
  width = 25.4,
  height = 25.4,
  units = "cm",
  dpi = "retina"
)

doDates <- doDates |>
  filter(Year == "2024") |>
  mutate(WinterStart_preY = as_date("2024-01-01"))

clustered_dat |>
  filter(year(Date) == 2024) |>
  filter(!(kg_name == "CFA-4" & Date == as_date("2024-01-01"))) |>
  ggplot(aes(x = Date)) +
  geom_point(
    aes(y = avgDO.pctsat, color = Cluster),
    size = 3,
    alpha = 0.8,
    na.rm = TRUE
  ) +
  facet_wrap(~kg_name, scales = "free_y", ncol = 1) +
  geom_rect(
    data = doDates,
    aes(xmin = WinterStart_preY, xmax = first_date, ymin = -Inf, ymax = Inf),
    inherit.aes = FALSE,
    fill = "#42d4f4",
    alpha = 0.2
  ) +
  geom_rect(
    data = doDates,
    aes(xmin = last_date, xmax = WinterStart, ymin = -Inf, ymax = Inf),
    inherit.aes = FALSE,
    fill = "#9A6324",
    alpha = 0.2
  ) +
  scale_color_manual(
    values = cluster_colors,
    name = "Cluster",
    na.translate = FALSE
  ) +
  scale_x_date(labels = date_format("%b"), breaks = "2 months") +
  ylim(0, 100) +
  xlim(as_date("2024-01-01"), as_date("2024-12-01")) +
  labs(x = "", y = expression("Daily Average DO"["% Sat"])) +
  theme(axis.text = element_text(size = 15)) -> p
p
ggsave(
  filename = "Average_2024.png",
  plot = p,
  path = here("Figs/ClusterTimeseries"),
  device = "png",
  width = 25.4,
  height = 25.4,
  units = "cm",
  dpi = "retina"
)

clustered_dat |>
  filter(!(kg_name == "CFA-4" & Date == as_date("2024-01-01"))) |>
  filter(year(Date) == 2024) |>
  ggplot(aes(x = Date)) +
  geom_point(
    aes(y = avgDO.pctsat, color = Cluster),
    size = 3,
    alpha = 0.8,
    na.rm = TRUE
  ) +
  geom_line(
    aes(y = percent_network_wet),
    color = "black",
    linewidth = 1,
    na.rm = TRUE
  ) +
  facet_wrap(~kg_name, scales = "free_y", ncol = 1, drop = FALSE) +
  geom_rect(
    data = doDates,
    aes(xmin = WinterStart_preY, xmax = first_date, ymin = -Inf, ymax = Inf),
    inherit.aes = FALSE,
    fill = "#42d4f4",
    alpha = 0.2
  ) +
  geom_rect(
    data = doDates,
    aes(xmin = last_date, xmax = WinterStart, ymin = -Inf, ymax = Inf),
    inherit.aes = FALSE,
    fill = "#9A6324",
    alpha = 0.2
  ) +
  scale_color_manual(
    values = cluster_colors,
    name = "Cluster",
    na.translate = FALSE
  ) +
  scale_x_date(labels = date_format("%b"), breaks = "2 months") +
  scale_y_continuous(
    limits = c(0, 100),
    name = expression("Daily Average DO"["% Sat"]),
    sec.axis = sec_axis(~., name = "Network Wet (%)")
  ) +
  xlim(as_date("2024-01-01"), as_date("2024-12-01")) +
  labs(x = "") +
  theme(axis.text = element_text(size = 15)) -> p
p
ggsave(
  filename = "Average_2024_PercentWet.png",
  plot = p,
  path = here("Figs/ClusterTimeseries"),
  device = "png",
  width = 25.4,
  height = 25.4,
  units = "cm",
  dpi = "retina"
)
```

## Cluster Time Series for Selected Year

```{r}
clustered_dat |>
  filter(!(kg_name == "CFA-4" & Date == as_date("2024-01-01"))) |>
  filter(
    (kg_name %in% c("DFA-1", "DFA-2", "CFA-1", "CFA-4") & year(Date) == 2024) |
      (kg_name %in% c("DSB-1", "CFA-2", "CFA-3") & year(Date) == 2023) |
      (kg_name %in% c("DFB-1") & year(Date) == 2022)
  ) |>
  mutate(
    Year = year(Date),
    SiteLabel = paste(kg_name, Year),
    fakeDate = as.Date(paste0("2000-", month(Date), "-", day(Date)))
  ) |>
  ggplot(aes(x = fakeDate)) +
  geom_point(
    aes(y = avgDO.pctsat, color = Cluster),
    size = 3,
    alpha = 0.8,
    na.rm = TRUE
  ) +
  geom_line(
    aes(y = percent_network_wet),
    color = "black",
    linewidth = 1,
    na.rm = TRUE
  ) +
  facet_wrap(~SiteLabel, scales = "fixed", ncol = 2, drop = FALSE) +
  scale_color_manual(
    values = cluster_colors,
    name = "State",
    na.translate = FALSE
  ) +
  xlim(as_date("2000-01-01"), as_date("2000-12-01")) +
  scale_x_date(
    labels = date_format("%b"),
    breaks = "1 months",
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    #limits = c(0, 100),
    name = expression(bold("Daily Average DO"["% Sat"])),
    sec.axis = sec_axis(~., name = "Network Wet (%)")
  ) +
  labs(x = "") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(size = 12),
    text = element_text(size = 20), # overall text size
    legend.text = element_text(size = 20), # legend entries
    legend.title = element_text(size = 20), # legend title
    axis.title = element_text(size = 20), # axis titles
    axis.text = element_text(size = 20), # axis tick labels
    plot.title = element_text(size = 20, hjust = 0.5),
    legend.position = "bottom"
  ) -> p

p
ggsave(
  filename = "Average_YEAR_PercentWet.png",
  plot = p,
  path = here("Figs/ClusterTimeseries"),
  device = "png",
  width = 40,
  height = 25.4,
  units = "cm",
  dpi = "retina"
)


# year_map <- c(
#   "GP-1" = 2024,
#   "GP-2" = 2023,
#   "GP-3" = 2024,
#   "EF-1" = 2024,
#   "EF-2" = 2023,
#   "EF-3" = 2023,
#   "ND-2" = 2023,
#   "ND-1" = 2022
# )
year_map <- c(
  "DFA-1" = 2024,
  "DFA-2" = 2023,
  "CFA-1" = 2024,
  "CFA-4" = 2024,
  "CFA-3" = 2023,
  "CFA-2" = 2023,
  "DSB-1" = 2023,
  "DFB-1" = 2022
)
clustered_dat |>
  group_by(kg_name) |>
  mutate(
    active_network_length_m = active_network_length_m /
      max(active_network_length_m, na.rm = T) *
      100
  ) |>
  ungroup() |>
  filter(!(kg_name == "CFA-4" & Date == as_date("2024-01-01"))) |>
  mutate(target_year = year_map[as.character(kg_name)]) |>
  filter(year(Date) == target_year) |>
  mutate(
    Date = update(Date, year = 2000),
    kg_name = paste0(kg_name, " (", target_year, ")")
  ) |>
  ggplot(aes(x = Date)) +
  see::geom_point2(
    aes(y = avgDO.pctsat, color = Cluster),
    size = 4,
    alpha = 0.7,
    na.rm = TRUE
  ) +
  geom_line(
    aes(y = active_network_length_m),
    color = "black",
    linewidth = 1,
    alpha = 0.7,
    na.rm = TRUE
  ) +
  facet_wrap(~kg_name, scales = "fixed", ncol = 2, drop = FALSE) +
  scale_color_manual(
    values = cluster_colors,
    name = "State",
    na.translate = FALSE
  ) +
  scale_x_date(
    labels = date_format("%b"),
    date_breaks = "1 month",
    limits = c(as_date("2000-01-01"), as_date("2000-09-30")),
    expand = expansion(0, 0)
  ) +
  scale_y_continuous(
    limits = c(0, 100),
    name = expression(bold("Daily Average DO"["% Sat"])),
    sec.axis = sec_axis(~., name = "ASDN (%)")
  ) +
  labs(x = "") +
  guides(color = guide_legend(override.aes = list(size = 4))) +
  theme(
    strip.text = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) -> p
ggsave(
  filename = "Average_YEAR_ASDN.png",
  plot = p,
  path = here("Figs/ClusterTimeseries"),
  device = "png",
  width = 40,
  height = 25.4,
  units = "cm",
  dpi = "retina"
)
```

```{r}
clustered_dat |>
  drop_na(Cluster) |>
  ggplot(aes(x = Date, y = DOpctsat.amp, color = Cluster)) +
  geom_point(size = 3, alpha = 0.8, na.rm = TRUE) +
  facet_wrap(~kg_name, scales = "free_y", ncol = 1) +
  scale_color_manual(values = cluster_colors, name = "Cluster") +
  ylim(0, 100) +
  scale_x_date(labels = date_format("%b '%y")) +
  xlim(as_date("2021-06-01"), as_date("2024-12-01")) +
  labs(
    x = "",
    y = expression("Daily DO"["% Sat"] * " Amplitude")
  ) +
  theme(axis.text = element_text(size = 15)) -> p
p
ggsave(
  filename = "Amplitude.png",
  plot = p,
  path = here("Figs/ClusterTimeseries"),
  device = "png",
  width = 25.4,
  height = 25.4,
  units = "cm",
  dpi = "retina"
)

clustered_dat |>
  drop_na(Cluster) |>
  ggplot(aes(x = Date, y = minDO.pctsat, color = Cluster)) +
  geom_point(size = 3, alpha = 0.8, na.rm = TRUE) +
  facet_wrap(~kg_name, scales = "free_y", ncol = 1) +
  scale_color_manual(values = cluster_colors, name = "Cluster") +
  ylim(0, 100) +
  scale_x_date(labels = date_format("%b '%y")) +
  xlim(as_date("2021-06-01"), as_date("2024-12-01")) +
  labs(
    x = "",
    y = expression("Daily DO"["% Sat"] * " Minimum")
  ) +
  theme(axis.text = element_text(size = 15)) -> p
p
ggsave(
  filename = "Minimum.png",
  plot = p,
  path = here("Figs/ClusterTimeseries"),
  device = "png",
  width = 25.4,
  height = 25.4,
  units = "cm",
  dpi = "retina"
)

clustered_dat |>
  drop_na(Cluster) |>
  ggplot(aes(x = Date, y = maxDO.pctsat, color = Cluster)) +
  geom_point(size = 3, alpha = 0.8, na.rm = TRUE) +
  facet_wrap(~kg_name, scales = "free_y", ncol = 1) +
  scale_color_manual(values = cluster_colors, name = "Cluster") +
  ylim(0, 100) +
  scale_x_date(labels = date_format("%b '%y")) +
  xlim(as_date("2021-06-01"), as_date("2024-12-01")) +
  labs(
    x = "",
    y = expression("Daily DO"["% Sat"] * " Maximum")
  ) +
  theme(axis.text = element_text(size = 15)) -> p
p
ggsave(
  filename = "Maximum.png",
  plot = p,
  path = here("Figs/ClusterTimeseries"),
  device = "png",
  width = 25.4,
  height = 25.4,
  units = "cm",
  dpi = "retina"
)


clustered_dat |>
  drop_na(Cluster) |>
  ggplot(aes(x = Date, y = DOSat.exc, color = Cluster)) +
  geom_point(size = 3, alpha = 0.8, na.rm = TRUE) +
  facet_wrap(~kg_name, scales = "free_y", ncol = 1) +
  scale_color_manual(values = cluster_colors, name = "Cluster") +
  ylim(0, 1) +
  scale_x_date(labels = date_format("%b '%y")) +
  xlim(as_date("2021-06-01"), as_date("2024-12-01")) +
  labs(
    x = "",
    y = expression("DO"["% Sat"] * " Exceedance Probability")
  ) +
  theme(axis.text = element_text(size = 15)) -> p
p
ggsave(
  filename = "ExcProb.png",
  plot = p,
  path = here("Figs/ClusterTimeseries"),
  device = "png",
  width = 25.4,
  height = 25.4,
  units = "cm",
  dpi = "retina"
)

clustered_dat |>
  drop_na(Cluster) |>
  ggplot(aes(x = Date, y = probHypSat, color = Cluster)) +
  geom_point(size = 3, alpha = 0.8, na.rm = TRUE) +
  facet_wrap(~kg_name, scales = "free_y", ncol = 1) +
  scale_color_manual(values = cluster_colors, name = "Cluster") +
  ylim(0, 1) +
  scale_x_date(labels = date_format("%b '%y")) +
  xlim(as_date("2021-06-01"), as_date("2024-12-01")) +
  labs(
    x = "",
    y = "Probability of Hypoxia"
  ) +
  theme(axis.text = element_text(size = 15)) -> p
p
ggsave(
  filename = "ProbHyp.png",
  plot = p,
  path = here("Figs/ClusterTimeseries"),
  device = "png",
  width = 25.4,
  height = 25.4,
  units = "cm",
  dpi = "retina"
)
```

```{r}

site_colors <- tibble(
  kg_name = c(
    "DSB-1",
    "DFB-1",
    "DFA-1",
    "DFA-2",
    "CFA-1",
    "CFA-2",
    "CFA-3",
    "CFA-4"
  ),
  color = c(
    "#006475",
    "#56CFD8",
    "#613921",
    "#F2C27B",
    "#A77652",
    "#345022",
    "#5A6C0A",
    "#83A102"
  )
) |>
  tibble::deframe()
clustered_dat |>
  filter(!(kg_name %in% c("DSB-1", "DFB-1"))) |>
  group_by(kg_name) |>
  mutate(
    headGradient_scaled = scales::rescale(headGradient_1d, to = c(0, 100))
  ) |>
  ungroup() |>
  ggplot(aes(x = Date)) +
  geom_point(
    aes(y = avgDO.pctsat, color = Cluster),
    size = 1,
    alpha = 0.8,
    na.rm = TRUE
  ) +
  geom_line(
    aes(y = headGradient_scaled),
    color = "black",
    linewidth = 1,
    na.rm = TRUE
  ) +
  facet_wrap(~kg_name, scales = "free_y", ncol = 1, drop = FALSE) +
  scale_color_manual(
    values = cluster_colors,
    name = "Cluster",
    na.translate = FALSE
  ) +
  scale_x_date(labels = date_format("%b '%y")) +
  scale_y_continuous(
    limits = c(0, 100),
    name = expression("Daily Average DO"["% Sat"]),
    sec.axis = sec_axis(~., name = "Head Greadient (Scaled)")
  ) +
  labs(x = "") +
  theme(axis.text = element_text(size = 15)) -> p
p
ggsave(
  filename = "HeadGradient.png",
  plot = p,
  path = here("Figs/ClusterTimeseries"),
  device = "png",
  width = 25.4,
  height = 25.4,
  units = "cm",
  dpi = "retina"
)

clustered_dat |>
  group_by(kg_name) |>
  mutate(
    active_upstream_length = active_upstream_length /
      max(active_upstream_length, na.rm = T) *
      100
  ) |>
  ungroup() |>
  ggplot(aes(x = Date)) +
  geom_point(
    aes(y = avgDO.pctsat, color = Cluster),
    size = 1,
    alpha = 0.8,
    na.rm = TRUE
  ) +
  geom_line(
    aes(y = active_upstream_length),
    color = "black",
    linewidth = 1,
    na.rm = TRUE
  ) +
  facet_wrap(~kg_name, scales = "free_y", ncol = 1, drop = FALSE) +
  scale_color_manual(
    values = cluster_colors,
    name = "Cluster",
    na.translate = FALSE
  ) +
  scale_x_date(labels = date_format("%b '%y")) +
  scale_y_continuous(
    limits = c(0, 100),
    name = expression("Daily Average DO"["% Sat"]),
    sec.axis = sec_axis(~., name = "Active Upstream Length")
  ) +
  labs(x = "") +
  theme(axis.text = element_text(size = 15)) -> p
p
ggsave(
  filename = "ActiveUpstreamLength.png",
  plot = p,
  path = here("Figs/ClusterTimeseries"),
  device = "png",
  width = 25.4,
  height = 25.4,
  units = "cm",
  dpi = "retina"
)


clustered_dat |>
  group_by(kg_name) |>
  mutate(
    active_network_length_m = active_network_length_m /
      max(active_network_length_m, na.rm = T) *
      100
  ) |>
  ungroup() |>
  ggplot(aes(x = Date)) +
  geom_point(
    aes(y = avgDO.pctsat, color = Cluster),
    size = 1,
    alpha = 0.8,
    na.rm = TRUE
  ) +
  geom_line(
    aes(y = active_network_length_m),
    color = "black",
    linewidth = 1,
    alpha = 0.7,
    na.rm = TRUE
  ) +
  facet_wrap(~kg_name, scales = "fixed", ncol = 1, drop = FALSE) +
  scale_color_manual(
    values = cluster_colors,
    name = "State",
    na.translate = FALSE
  ) +
  scale_x_date(
    labels = date_format("%b '%y"),
    date_breaks = "3 months",
    limits = c(as_date(NA), as_date("2024-12-31")),
    expand = expansion(0, 0)
  ) +
  scale_y_continuous(
    limits = c(0, 100),
    name = expression(bold("Daily Average DO"["% Sat"])),
    sec.axis = sec_axis(~., name = "ASDN (%)")
  ) +
  labs(x = "") +
  guides(color = guide_legend(override.aes = list(size = 4))) +
  theme(
    axis.text = element_text(size = 15),
    strip.text = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) -> p
p
ggsave(
  filename = "ASDN_Timeseries.png",
  plot = p,
  path = here("Figs/ClusterTimeseries"),
  device = "png",
  width = 25.4,
  height = 25.4,
  units = "cm",
  dpi = "retina"
)

clustered_dat |>
  ggplot(aes(x = Date)) +
  geom_point(
    aes(y = avgDO.pctsat, color = Cluster),
    size = 1,
    alpha = 0.8,
    na.rm = TRUE
  ) +
  geom_line(
    aes(y = percent_network_wet),
    color = "black",
    linewidth = 1,
    na.rm = TRUE
  ) +
  facet_wrap(~kg_name, scales = "free_y", ncol = 1, drop = FALSE) +
  scale_color_manual(
    values = cluster_colors,
    name = "Cluster",
    na.translate = FALSE
  ) +
  scale_x_date(labels = date_format("%b '%y")) +
  scale_y_continuous(
    limits = c(0, 100),
    name = expression("Daily Average DO"["% Sat"]),
    sec.axis = sec_axis(~., name = "Network Wet (%)")
  ) +
  labs(x = "") +
  theme(axis.text = element_text(size = 15)) -> p
p
ggsave(
  filename = "PercentNetworkWet.png",
  plot = p,
  path = here("Figs/ClusterTimeseries"),
  device = "png",
  width = 25.4,
  height = 25.4,
  units = "cm",
  dpi = "retina"
)

clustered_dat |>
  ggplot(aes(x = Date)) +
  geom_point(
    aes(y = avgDO.pctsat, color = kg_name),
    size = 1,
    alpha = 0.8,
    na.rm = TRUE
  ) +
  geom_line(
    aes(y = percent_network_wet),
    color = "black",
    linewidth = 1,
    na.rm = TRUE
  ) +
  facet_wrap(~kg_name, scales = "free_y", ncol = 1, drop = FALSE) +
  scale_color_manual(
    values = site_colors,
    name = "Cluster",
    na.translate = FALSE
  ) +
  scale_x_date(labels = date_format("%b '%y")) +
  scale_y_continuous(
    limits = c(0, 100),
    name = expression("Daily Average DO"["% Sat"]),
    sec.axis = sec_axis(~., name = "Network Wet (%)")
  ) +
  labs(x = "") +
  theme(axis.text = element_text(size = 15), legend.position = "none") -> p
p
ggsave(
  filename = "PercentNetworkWet_FinalReport.png",
  plot = p,
  path = here("Figs/ClusterTimeseries"),
  device = "png",
  width = 25.4,
  height = 25.4,
  units = "cm",
  dpi = "retina"
)

clustered_dat |>
  group_by(kg_name) |>
  mutate(a_cent = scales::rescale(a_cent, to = c(0, 100))) |>
  ungroup() |>
  ggplot(aes(x = Date)) +
  geom_point(
    aes(y = avgDO.pctsat, color = Cluster),
    size = 1,
    alpha = 0.8,
    na.rm = TRUE
  ) +
  geom_line(aes(y = a_cent), color = "black", linewidth = 1, na.rm = TRUE) +
  facet_wrap(~kg_name, scales = "free_y", ncol = 1, drop = FALSE) +
  scale_color_manual(
    values = cluster_colors,
    name = "Cluster",
    na.translate = FALSE
  ) +
  scale_x_date(labels = date_format("%b '%y")) +
  scale_y_continuous(
    limits = c(0, 100),
    name = expression("Daily Average DO"["% Sat"]),
    sec.axis = sec_axis(~., name = "Alpha Centrality")
  ) +
  labs(x = "") +
  theme(axis.text = element_text(size = 15)) -> p
p
ggsave(
  filename = "AlphaCentrality.png",
  plot = p,
  path = here("Figs/ClusterTimeseries"),
  device = "png",
  width = 25.4,
  height = 25.4,
  units = "cm",
  dpi = "retina"
)
```

## Head Gradient

```{r}
site_plot_data <- clustered_dat |>
  filter(!(kg_name %in% c("DSB-1", "DFB-1"))) |>
  select(kg_name, Date, avgDO.pctsat, headGradient_1d, Cluster) |>
  pivot_longer(
    cols = c(avgDO.pctsat, headGradient_1d),
    names_to = "Variable",
    values_to = "Value"
  ) |>
  mutate(
    Variable = recode(
      Variable,
      "avgDO.pctsat" = "Daily Avg DO (%Sat)",
      "headGradient_1d" = "Head Gradient"
    )
  )

custom_labeller <- labeller(
  Variable = as_labeller(
    c(
      "Daily Avg DO (%Sat)" = "Daily~Avg.~DO['% Sat']",
      "Head Gradient" = "Head~Gradient~(m)"
    ),
    default = label_parsed
  ),
  name = label_value # Leave Site labels unchanged
)

site_faceted_plot <- site_plot_data |>
  ggplot(aes(x = Date, y = Value)) +
  geom_point(
    data = filter(site_plot_data, Variable == "Daily Avg DO (%Sat)"),
    aes(color = Cluster),
    size = 1,
    alpha = 0.8,
    na.rm = TRUE
  ) +
  geom_line(
    data = filter(site_plot_data, Variable == "Head Gradient"),
    color = "black",
    linewidth = 1,
    na.rm = TRUE
  ) +
  geom_hline(
    data = filter(site_plot_data, Variable == "Head Gradient") |>
      distinct(Variable, kg_name),
    aes(yintercept = 0),
    linetype = "dashed",
    color = "grey50"
  ) +
  facet_grid(
    Variable ~ kg_name,
    scales = "free_y",
    labeller = custom_labeller
  ) +
  scale_color_manual(
    values = cluster_colors,
    name = "Cluster",
    na.translate = FALSE
  ) +
  scale_x_date() +
  labs(x = "", y = "") +
  theme(
    strip.placement = "outside",
    strip.text.y.left = element_text(angle = 0),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )

site_faceted_plot

ggsave(
  filename = "HeadGradient.png",
  plot = site_faceted_plot,
  path = here::here("Figs/Supplement/"),
  width = 40,
  height = 20,
  units = "cm",
  dpi = "retina"
)
```

## Stacked Bars

```{r}
# Create month column and calculate frequencies
cluster_freq <- clustered_dat |>
  mutate(
    Month = factor(lubridate::month(Date, label = T))
  ) |>
  drop_na(Cluster) |>
  group_by(kg_name, Month, Cluster) |>
  summarise(count = n(), .groups = 'drop')

monthlyPercWet <- allDat |>
  left_join(site_meta |> select(Site, kg_name), by = "Site") |>
  mutate(,
    Month = factor(lubridate::month(Date, label = TRUE))
  ) |>
  group_by(kg_name, Month) |>
  summarise(
    meanPercWet = mean(percent_network_wet, na.rm = TRUE),
    .groups = "drop"
  )

cluster_freq |>
  filter(!Month %in% c("Oct", "Nov", "Dec", "Jan")) |>
  ggplot(aes(x = Month, y = count, fill = Cluster)) +
  geom_bar(stat = "identity", position = "fill") +
  geom_point(
    data = monthlyPercWet |> filter(!Month %in% c("Oct", "Nov", "Dec", "Jan")),
    aes(x = Month, y = meanPercWet / 100), # scale to proportion
    inherit.aes = FALSE,
    size = 3,
    shape = 21,
    fill = "black"
  ) +
  facet_wrap(~kg_name, ncol = 3) +
  scale_fill_manual(values = cluster_colors, name = "Cluster") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  labs(
    title = "",
    x = "",
    y = "Proportion"
  ) +
  theme_few() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    text = element_text(size = 18),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 16),
    plot.title = element_text(size = 20, hjust = 0.5),
    plot.margin = grid::unit(c(0, 0, 0, 0), "lines"),
    legend.position = "bottom"
  ) -> p
p
ggsave(
  filename = "clusterDist_PercWet.png",
  plot = p,
  path = here("Figs/"),
  device = "png",
  width = 25.4,
  height = 25.4,
  units = "cm",
  dpi = "retina"
)


cluster_freq |>
  #filter(!Month %in% c("Oct", "Nov", "Dec")) |>
  ggplot(aes(x = Month, y = count, fill = Cluster)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~kg_name, nrow = 2) +
  scale_fill_manual(values = cluster_colors, name = "Cluster") +
  scale_y_continuous(labels = scales::percent, expand = c(0, 0)) +
  theme_minimal() +
  labs(
    title = "",
    x = "",
    y = "Proportion"
  ) +
  theme_few() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    text = element_text(size = 20), # overall text size
    legend.text = element_text(size = 20), # legend entries
    legend.title = element_text(size = 20), # legend title
    axis.title = element_text(size = 20, face = "bold"), # axis titles
    axis.text = element_text(size = 20), # axis tick labels
    plot.title = element_text(size = 20, hjust = 0.5),
    plot.margin = grid::unit(c(0, 0, 0, 0), "lines"),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  ) -> p
p

# Find month/name combos that have no data and add a "No Data" row
all_months <- unique(cluster_freq$Month) # or use factor levels to preserve order

cluster_freq_plot <- cluster_freq |>
  complete(Month = all_months, kg_name) |>
  group_by(Month, kg_name) |>
  mutate(has_data = any(!is.na(count) & count > 0)) |>
  ungroup() |>
  filter(!is.na(Cluster)) |> # drop NAs introduced by complete()
  bind_rows(
    cluster_freq |>
      complete(Month = all_months, kg_name) |>
      group_by(Month, kg_name) |>
      filter(all(is.na(count) | count == 0)) |>
      ungroup() |>
      distinct(Month, kg_name) |>
      mutate(Cluster = "No Data", count = 1)
  )

# Add grey to color scale
cluster_colors_plot <- c(cluster_colors, "No Data" = "grey80")
cluster_freq_plot |>
  ggplot(aes(x = Month, y = count, fill = Cluster)) +
  geom_bar(stat = "identity", position = "fill") +
  facet_wrap(~kg_name, nrow = 2) +
  scale_fill_manual(values = cluster_colors_plot, name = "State") +
  scale_x_discrete(drop = FALSE) + # keep all months even if still missing
  scale_y_continuous(
    expand = c(0, 0),
    labels = function(x) x * 100,
    breaks = seq(0, 1, by = 0.25)
  ) +
  labs(x = "", y = "Proportion (%)") +
  theme_few() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
    text = element_text(size = 20),
    legend.text = element_text(size = 20),
    legend.title = element_text(size = 20),
    axis.title = element_text(size = 20, face = "bold"),
    axis.text = element_text(size = 20),
    plot.title = element_text(size = 20, hjust = 0.5),
    plot.margin = grid::unit(c(0, 0, 0, 0), "lines"),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  ) -> p
p
ggsave(
  filename = "clusterDist.png",
  plot = p,
  path = here("Figs/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = "retina"
)
```

## Correlation

```{r}
rf_dat <- clustered_dat |>
  group_by(Site) |>
  mutate(
    active_network_length_m = active_network_length_m /
      max(active_network_length_m, na.rm = T) *
      100,
    active_upstream_length = active_upstream_length /
      max(active_upstream_length, na.rm = T) *
      100
  ) |>
  ungroup() |>
  select(
    -c(
      avgDO.pctsat,
      DOSat.exc,
      DOpctsat.amp,
      maxDO.pctsat,
      probHypSat,
      minDO.pctsat,
      DOY,
      waterYear,
      maxPar
    )
  ) |>
  arrange(Date) |>
  mutate(
    across(
      where(is.numeric),
      ~ na.approx(.x, maxgap = 7, na.rm = FALSE),
      .names = "{.col}"
    ),
    across(.cols = matches("^(Q_)"), ~ scale(.)[, 1]),
    .by = "Site"
  ) |>
  #mutate(across(.cols = matches("^(Depth_)"), ~ scale(.)[, 1])) |>
  drop_na(Cluster)


png(here("Figs/Corr_00.png"), width = 2160, height = 2160)
rf_dat |>
  group_by(Site) |>
  mutate(across(where(is.numeric), ~ scale(.)[, 1])) |>
  ungroup() |>
  select(where(is.numeric)) |>
  cor(use = "complete.obs") |>
  (\(cor_mat) {
    corrplot(
      cor_mat,
      method = "circle",
      type = "upper",
      order = "alphabet",
      tl.col = "black",
      tl.pos = "lt",
      tl.cex = 2,
      cl.pos = "n",
      na.label = " ",
      diag = TRUE
    )
    corrplot(
      cor_mat,
      method = "number",
      type = "lower",
      order = "alphabet",
      tl.pos = "n",
      add = TRUE,
      number.cex = 1.5,
      na.label = " ",
      diag = FALSE
    )
  })()
dev.off()

png(here("Figs/Corr_50.png"), width = 2160, height = 2160)
rf_dat |>
  group_by(Site) |>
  mutate(across(where(is.numeric), ~ scale(.)[, 1])) |>
  ungroup() |>
  select(where(is.numeric)) |>
  cor(use = "complete.obs") |>
  (\(cor_mat) {
    cor_mat[abs(cor_mat) < 0.5] <- NA
    corrplot(
      cor_mat,
      method = "circle",
      type = "upper",
      order = "alphabet",
      tl.col = "black",
      tl.pos = "lt",
      tl.cex = 2,
      cl.pos = "n",
      na.label = " ",
      diag = TRUE
    )
    corrplot(
      cor_mat,
      method = "number",
      type = "lower",
      order = "alphabet",
      tl.pos = "n",
      add = TRUE,
      number.cex = 1.5,
      na.label = " ",
      diag = FALSE
    )
  })()
dev.off()

png(here("Figs/Corr_75.png"), width = 2160, height = 2160)
rf_dat |>
  group_by(Site) |>
  mutate(across(where(is.numeric), ~ scale(.)[, 1])) |>
  ungroup() |>
  select(where(is.numeric)) |>
  cor(use = "complete.obs") |>
  (\(cor_mat) {
    cor_mat[abs(cor_mat) < 0.75] <- NA
    corrplot(
      cor_mat,
      method = "circle",
      type = "upper",
      order = "alphabet",
      tl.col = "black",
      tl.pos = "lt",
      tl.cex = 2,
      cl.pos = "n",
      na.label = " ",
      diag = TRUE
    )
    corrplot(
      cor_mat,
      method = "number",
      type = "lower",
      order = "alphabet",
      tl.pos = "n",
      add = TRUE,
      number.cex = 1.5,
      na.label = " ",
      diag = FALSE
    )
  })()
dev.off()


png(here("Figs/Corr_70.png"), width = 2160, height = 2160)
rf_dat |>
  group_by(Site) |>
  mutate(across(where(is.numeric), ~ scale(.)[, 1])) |>
  ungroup() |>
  select(where(is.numeric)) |>
  cor(use = "complete.obs") |>
  (\(cor_mat) {
    cor_mat[abs(cor_mat) < 0.70] <- NA
    corrplot(
      cor_mat,
      method = "circle",
      type = "upper",
      order = "alphabet",
      tl.col = "black",
      tl.pos = "lt",
      tl.cex = 2,
      cl.pos = "n",
      na.label = " ",
      diag = TRUE
    )
    corrplot(
      cor_mat,
      method = "number",
      type = "lower",
      order = "alphabet",
      tl.pos = "n",
      add = TRUE,
      number.cex = 1.5,
      na.label = " ",
      diag = FALSE
    )
  })()
dev.off()
```

## Look at vars to standardize

```{r}
# density plots with means
dat_dens <- clustered_dat |>
  filter(PPT_1d < 1, Q_1d < 5) |>
  select(
    kg_name,
    Date,
    active_network_length_m,
    PPT_1d,
    Q_1d,
    Depth_1d,
    ET_1d,
    avgPar,
    headGradient_1d
  ) |>
  pivot_longer(
    cols = -c(kg_name, Date),
    names_to = "variable",
    values_to = "value"
  )

means_dens <- dat_dens |>
  group_by(kg_name, variable) |>
  summarise(mean_val = mean(value, na.rm = TRUE), .groups = "drop")

ggplot(dat_dens, aes(x = value, fill = kg_name)) +
  geom_density(position = "identity", alpha = 0.5, na.rm = TRUE) +
  geom_vline(
    data = means_dens,
    aes(xintercept = mean_val, color = kg_name),
    linetype = "dashed",
    linewidth = 0.6,
    show.legend = FALSE,
    na.rm = TRUE
  ) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  labs(title = "RF Variables") +
  facet_wrap(~variable, scales = "free") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) -> p
p
ggsave(
  filename = "vars_dens.png",
  plot = p,
  path = here::here("Figs/Supplement/"),
  width = 40,
  height = 20,
  units = "cm",
  dpi = "retina"
)


# Z-Score transformed density plots
dat_dens_z <- clustered_dat |>
  group_by(kg_name) |>
  mutate(across(where(is.numeric), ~ scale(.)[, 1])) |>
  ungroup() |>
  select(
    kg_name,
    Date,
    active_network_length_m,
    PPT_1d,
    Q_1d,
    Depth_1d,
    ET_1d,
    avgPar,
    headGradient_1d
  ) |>
  pivot_longer(
    cols = -c(kg_name, Date), # keep Site and Date, melt everything else
    names_to = "variable",
    values_to = "value"
  )
means_dens_z <- dat_dens_z |>
  group_by(kg_name, variable) |>
  summarise(mean_val = mean(value, na.rm = TRUE), .groups = "drop")

ggplot(
  dat_dens_z |>
    filter(
      (variable == "Q_1d" & value < 4) |
        (variable == "PPT_1d" & value < 2.5) |
        !(variable %in% c("Q_1d", "PPT_1d"))
    ),
  aes(x = value, fill = kg_name)
) +
  geom_density(position = "identity", alpha = 0.5, na.rm = TRUE) +
  geom_vline(
    data = means_dens_z,
    aes(xintercept = mean_val, color = kg_name),
    linetype = "dashed",
    linewidth = 0.6,
    show.legend = FALSE,
    na.rm = TRUE
  ) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  labs(title = "RF Variables Z-Score Transformed") +
  facet_wrap(~variable, scales = "free") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) -> p
p
ggsave(
  filename = "vars_dens_zscore.png",
  plot = p,
  path = here::here("Figs/Supplement/"),
  width = 40,
  height = 20,
  units = "cm",
  dpi = "retina"
)

# 0-100 transformed density plots
dat_dens_scaled <- clustered_dat |>
  select(
    kg_name,
    Date,
    active_network_length_m,
    PPT_1d,
    Q_1d,
    Depth_1d,
    ET_1d,
    avgPar,
    headGradient_1d
  ) |>
  pivot_longer(
    cols = -c(kg_name, Date),
    names_to = "variable",
    values_to = "value"
  ) |>
  group_by(kg_name, variable) |>
  mutate(
    scaled_value = (value - min(value, na.rm = TRUE)) /
      (max(value, na.rm = TRUE) - min(value, na.rm = TRUE)) *
      100
  ) |>
  ungroup()

means_dens_scaled <- dat_dens_scaled |>
  group_by(kg_name, variable) |>
  summarise(mean_val = mean(scaled_value, na.rm = TRUE), .groups = "drop")

ggplot(dat_dens_scaled, aes(x = scaled_value, fill = kg_name)) +
  geom_density(position = "identity", alpha = 0.5, na.rm = TRUE) +
  geom_vline(
    data = means_dens_scaled,
    aes(xintercept = mean_val, color = kg_name),
    linetype = "dashed",
    linewidth = 0.6,
    show.legend = FALSE,
    na.rm = TRUE
  ) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  labs(title = "RF Variables 0-100 Scale") +
  facet_wrap(~variable, scales = "free") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) -> p
p
ggsave(
  filename = "vars_dens_0-100.png",
  plot = p,
  path = here("Figs/Supplement/"),
  width = 40,
  height = 20,
  units = "cm",
  dpi = "retina"
)

# Site centered density plots
dat_dens_cent <- clustered_dat |>
  select(
    kg_name,
    Date,
    active_network_length_m,
    PPT_1d,
    Q_1d,
    Depth_1d,
    ET_1d,
    avgPar,
    headGradient_1d
  ) |>
  pivot_longer(
    cols = -c(kg_name, Date),
    names_to = "variable",
    values_to = "value"
  ) |>
  group_by(kg_name, variable) |>
  mutate(
    value = value - mean(value, na.rm = TRUE)
  ) |>
  ungroup()
means_dens_cent <- dat_dens_z |>
  group_by(kg_name, variable) |>
  summarise(mean_val = mean(value, na.rm = TRUE), .groups = "drop")

ggplot(
  dat_dens_cent |>
    filter(
      (variable == "Q_1d" & value < 4) |
        (variable == "PPT_1d" & value < 2.5) |
        !(variable %in% c("Q_1d", "PPT_1d"))
    ),
  aes(x = value, fill = kg_name)
) +
  geom_density(position = "identity", alpha = 0.5, na.rm = TRUE) +
  geom_vline(
    data = means_dens_cent,
    aes(xintercept = mean_val, color = kg_name),
    linetype = "dashed",
    linewidth = 0.6,
    show.legend = FALSE,
    na.rm = TRUE
  ) +
  scale_fill_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  scale_color_manual(
    values = setNames(site_meta$color, site_meta$kg_name),
    breaks = levels(site_meta$kg_name)
  ) +
  labs(title = "RF Variables Centered") +
  facet_wrap(~variable, scales = "free") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) -> p
p
ggsave(
  filename = "vars_dens_mean_centered.png",
  plot = p,
  path = here::here("Figs/Supplement/"),
  width = 40,
  height = 20,
  units = "cm",
  dpi = "retina"
)
```

```{r}
metrics <- c(
  "PPT_7d",
  "Q_1d",
  "Q_7d",
  "Q_60d",
  "Depth_1d",
  "Depth_60d",
  "headGradient_1d",
  "headGradient_30d",
  "ET_7d",
  "avgPar",
  "active_network_length_m",
  "water.temp"
)

metric_names <- c(
  PPT_7d = "Precip (7 Day Avg.)",
  Q_1d = "Discharge",
  Q_7d = "Discharge (7 Day Avg.)",
  Q_60d = "Discharge (60 Day Avg.)",
  Depth_1d = "Depth",
  Depth_60d = "Depth (60 Day Avg.)",
  headGradient_1d = "Head Gradient",
  headGradient_30d = "Head Gradient (30 Day Avg.)",
  ET_7d = "ET (7 Day Avg.)",
  avgPar = "Light",
  active_network_length_m = "ASDN",
  water.temp = "Water Temp"
)

# Pivot data long
plot_data <- rf_dat |>
  filter(!is.na(Cluster)) |>
  select(kg_name, Cluster, any_of(metrics)) |>
  pivot_longer(
    cols = any_of(metrics),
    names_to = "Metric",
    values_to = "Value"
  ) |>
  mutate(
    Metric_label = metric_names[Metric]
  )


p <- ggplot(plot_data, aes(x = kg_name, y = Value, fill = Cluster)) +
  geom_boxplot(
    outlier.shape = NA,
    position = position_dodge(width = 0.75),
    na.rm = TRUE
  ) +
  geom_jitter(
    alpha = 0.1,
    aes(color = Cluster),
    position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.75),
    na.rm = TRUE
  ) +
  facet_wrap(~Metric_label, scales = "free_y") + # facet by metric
  scale_fill_manual(values = cluster_colors) +
  scale_color_manual(values = cluster_colors) +
  labs(title = "Random Forest Data by Cluster", x = "", y = "Value") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  )
p
ggsave(
  filename = "rf_data.png",
  plot = p,
  path = here::here("Figs/Supplement/"),
  width = 40,
  height = 20,
  units = "cm",
  dpi = "retina"
)
```

# Find Best Lag

```{r}
source(here("Code/functions/rf_model.R"))

plan(multisession, workers = parallel::detectCores() - 1)

evaluate_lag <- function(dat, var, lag_k) {
  # Case weights
  cluster_counts <- table(dat$Cluster)
  #cluster_weights <- sqrt(max(cluster_counts) / cluster_counts)
  cluster_weights <- c("1" = .2, "2" = .3, "3" = .5)
  roll_var <- paste0(var, "_roll", lag_k)

  dat_roll <- dat |>
    group_by(Site, kg_name) |>
    arrange(Date) |>
    mutate(
      "{roll_var}" := slider::slide_dbl(
        .data[[var]],
        .f = mean,
        .before = lag_k,
        .after = 0,
        .complete = FALSE
      )
    ) |>
    ungroup() |>
    filter(!is.na(.data[[roll_var]]), !is.na(Cluster)) |>
    mutate(
      case_wt = case_when(
        Cluster == "1" ~ cluster_weights["1"],
        Cluster == "2" ~ cluster_weights["2"],
        Cluster == "3" ~ cluster_weights["3"]
      ),
      case_wt = importance_weights(case_wt)
    ) |>
    select(Cluster, all_of(roll_var), case_wt)

  folds <- vfold_cv(dat_roll, v = 5, strata = "Cluster")

  # Recipe
  rec <- recipe(Cluster ~ ., data = dat_roll) |>
    step_zv()

  # RF Spec
  model_spec <- rand_forest(trees = 500) |>
    set_engine("ranger", seed = 42, num.threads = 1) |>
    set_mode("classification")
  # Workflow
  wf <- workflow() |>
    add_model(model_spec) |>
    add_recipe(rec) |>
    add_case_weights(case_wt)

  # Results
  res <- wf |>
    fit_resamples(
      folds,
      metrics = metric_set(kap, f_meas, accuracy, roc_auc),
      control = control_resamples(save_pred = TRUE)
    )

  res |>
    collect_metrics() |>
    mutate(variable = var, lag = lag_k)
}


vars <- c(
  "water.temp",
  "Q",
  "Depth",
  "headGradient",
  "PPT",
  "avgPar",
  "ET",
  "active_network_length_m"
)

max_lag <- 60

grid <- tidyr::expand_grid(
  variable = vars,
  lag = 0:max_lag
)

lagged_data <- dailyDat |>
  as_tibble() |>
  # group_by(Site) |>
  # mutate(active_network_length_m = active_network_length_m/max(active_network_length_m, na.rm = T)*100,
  #        active_upstream_length = active_upstream_length/max(active_upstream_length, na.rm = T)*100) |>
  # ungroup() |>
  select(-ends_with(c("_7d", "_14d", "_30d", "_60d"))) |>
  rename_with(~ gsub("_1d$", "", .x), ends_with("_1d")) |>
  select(
    -c(
      "active_upstream_length",
      "a_cent",
      "a_cent_weighted",
      "Season",
      "DOY",
      "excProb",
      "DOSat.exc",
      "air.temp",
      "minDO.pctsat",
      "maxDO.pctsat",
      "avgDO.pctsat",
      "DOpctsat.amp",
      "probHypSat",
      "wetQf",
      "maxPar",
      "waterYear",
      "percent_network_wet"
    )
  ) |>
  left_join(
    pca_scores |> select(kg_name, Date, Cluster),
    by = c("kg_name", "Date")
  ) |>
  relocate(c("kg_name", "Cluster"), .after = Site) |>
  arrange(Site, Date)


# scale data
lagged_data_scaled <- lagged_data |>
  group_by(Site) |>
  # # mutate(across(
  # #   all_of(c("active_network_length_m", "avgPar", "Depth", "Q", "ET", "PPT")),
  # #   ~ ( (.x - min(.x, na.rm = TRUE)) /
  # #       (max(.x, na.rm = TRUE) - min(.x, na.rm = TRUE)) ) * 100
  # # )) |>
  mutate(
    active_network_length_m = active_network_length_m /
      max(active_network_length_m, na.rm = T) *
      100
  ) # |>
# mutate(
#   across(all_of(c("avgPar", "Depth", "Q", "ET", "PPT")),
#          ~ .x - mean(.x, na.rm = TRUE))
# ) |>
#  # mutate(across(all_of(c("active_network_length_m", "avgPar", "Depth", "Q", "ET", "PPT")), ~ scale(.)[, 1])) |>
# ungroup() |>
# group_by(Site) |>
# mutate(across(where(is.numeric), ~ as.numeric(scale(.x)))) |>
# ungroup()

lag_results <- future_map_dfr(
  seq_len(nrow(grid)),
  function(i) {
    evaluate_lag(
      dat = lagged_data_scaled,
      var = grid$variable[i],
      lag_k = grid$lag[i]
    )
  },
  .options = furrr_options(seed = TRUE)
)
plan(sequential)
```


```{r}
best_points <- lag_results |>
  group_by(variable, .metric) |>
  #mutate(mean = round(mean, 2)) |>
  mutate(
    best_mean = max(mean, na.rm = TRUE)
  ) |>
  filter(mean >= best_mean * .97) |>
  slice_min(lag, n = 1) |>
  ungroup() |>
  select(-best_mean) |>
  mutate(
    variable = case_when(
      variable %in% names(var_names) ~ var_names[variable],
      TRUE ~ variable
    ),
    .metric = case_when(
      .metric == "accuracy" ~ "Accuracy",
      .metric == "f_meas" ~ "F1",
      .metric == "kap" ~ "Kappa",
      .metric == "roc_auc" ~ "ROC AUC"
    )
  )


lag_results |>
  mutate(
    variable = case_when(
      variable %in% names(var_names) ~ var_names[variable],
      TRUE ~ variable
    ),
    .metric = case_when(
      .metric == "accuracy" ~ "Accuracy",
      .metric == "f_meas" ~ "F1",
      .metric == "kap" ~ "Kappa",
      .metric == "roc_auc" ~ "ROC AUC"
    )
  ) |>
  #filter(.metric %in% c("F1")) |>
  ggplot(aes(x = lag, y = mean, color = variable)) +
  geom_line() +
  see::scale_color_okabeito() +
  # geom_point(
  #   data = best_points,#|> filter(.metric %in% c("F1")),
  #   aes(x = lag, y = mean),
  #   color = "red",
  #   alpha = 0.5,
  #   show.legend = FALSE
  # ) +
  #   geom_text_repel(
  #   data = best_points,#|> filter(.metric %in% c("F1")),
  #   aes(x = lag, y = mean, label = lag),
  #   vjust = -0.6,
  #   color = "red",
  #   size = 4,
  #   show.legend = FALSE
  # ) +
  facet_grid(.metric ~ variable, scales = "free_y") +
  labs(
    title = "Performance Across Metrics",
    subtitle = "Higher is better",
    x = "Days Prior",
    y = "Performance",
    color = ""
  ) +
  theme_bw() +
  theme(
    text = element_text(size = 18), # overall text size
    legend.text = element_text(size = 16), # legend entries
    legend.title = element_text(size = 18), # legend title
    axis.title = element_text(size = 18), # axis titles
    axis.text = element_text(size = 16), # axis tick labels
    plot.title = element_text(size = 20),
    legend.position = "none"
  ) -> p
p


ggsave(
  filename = "best_lags.png",
  plot = p,
  path = here("Figs/Supplement/"),
  width = 40,
  height = 20,
  units = "cm",
  dpi = "retina"
)
beepr::beep(2)
```

```{r}
# Find best lag
best_lag <- lag_results |>
  filter(.metric == c("f_meas")) |>
  group_by(variable) |>
  mutate(
    best_mean = max(mean, na.rm = TRUE)
  ) |>
  filter(mean >= best_mean * 0.97) |>
  slice_min(lag, n = 1) |>
  ungroup() |>
  select(-best_mean)

# - Day 0: ASDN, depth, discharge
# - Day 14: PPT, water temp
# - Day 30: ET, head gradient (?), light
# - Day 60: head gradient (?)
day0_lag <- best_lag |>
  mutate(lag = 0)
best_lag <- best_lag |>
  mutate(lag = c(0, 30, 14, 0, 0, 30, 30, 14))
#mutate(lag = 0)

# best_lag <- rbind(best_lag, day0_lag) |>
#  distinct(variable, lag)

best_lags_data <- lagged_data_scaled

for (i in seq_len(nrow(best_lag))) {
  var <- best_lag$variable[i]
  lagn <- best_lag$lag[i]
  col_name <- paste0(var, "_", lagn, "d")

  best_lags_data <- best_lags_data |>
    group_by(Site, name, kg_name) |>
    arrange(Date) |>
    mutate(
      !!col_name := slider::slide_dbl(
        .data[[var]],
        mean,
        .before = lagn,
        .after = 0,
        .complete = FALSE
      )
    ) |>
    ungroup()
}

# Drop the original (un-lagged) variable columns
best_lags_data <- best_lags_data |>
  select(-any_of(best_lag$variable)) |>
  drop_na(Cluster) |>
  group_by(Site) |>
  mutate(across(
    where(is.numeric) & !active_network_length_m_0d,
    ~ scale(.) |> as.numeric()
  )) |>
  ungroup()

png(here("Figs/best_lag_corr.png"), width = 2160, height = 2160)
best_lags_data |>
  select(where(is.numeric)) |>
  cor(use = "complete.obs") |>
  (\(cor_mat) {
    pretty_names <- var_names[colnames(cor_mat)] |>
      stringr::str_wrap(width = 10)
    dimnames(cor_mat) <- list(pretty_names, pretty_names)
    corrplot(
      cor_mat,
      method = "number",
      type = "lower",
      order = "hclust",
      tl.pos = "d",
      tl.col = "black",
      diag = TRUE,
      tl.cex = 3,
      number.cex = 4,
      cl.cex = 3,
      col = colorRampPalette(c("#e8000d", "#dde5ed", "#0051ba"))(200),
      col.lim = c(-1, 1),
    )
  })()

dev.off()
```

## Autocorrelation
```{r}
site_order <- site_meta |>
  arrange(lon) |>
  distinct(Site) |>
  pull(Site) |>
  as.character()

site_colors <- site_meta |>
  distinct(Site, color) |>
  mutate(Site = as.character(Site)) |>
  deframe()

acf_by_site <- best_lags_data |>
  arrange(Site, name, Date) |>
  pivot_longer(
    cols = where(is.numeric),
    names_to = "variable",
    values_to = "value"
  ) |>
  group_by(Site, name, variable) |>
  filter(sum(!is.na(value)) >= 3) |>
  summarise(
    acf_result = list(
      acf(
        value[!is.na(value)],
        lag.max = 120,
        plot = FALSE,
        na.action = na.pass
      )
    ),
    .groups = "drop"
  ) |>
  mutate(
    acf_data = map(
      acf_result,
      \(x) {
        tibble(
          lag = as.numeric(x$lag),
          autocorrelation = as.numeric(x$acf)
        )
      }
    )
  ) |>
  select(-acf_result) |>
  unnest(acf_data) |>
  mutate(
    Site = factor(
      Site,
      levels = site_order
    )
  ) |>
  mutate(
    variable = recode(variable, !!!var_names),
    Site = factor(
      Site,
      levels = site_order
    )
  )
first_nonpositive <- acf_by_site |>
  filter(lag > 0, autocorrelation <= 0) |>
  group_by(Site, variable) |>
  summarise(
    first_nonpositive_lag = min(lag),
    .groups = "drop"
  )
acf_by_site |>
  ggplot(
    aes(
      x = lag,
      y = autocorrelation,
      group = name,
      color = Site
    )
  ) +
  geom_hline(
    yintercept = 0,
    linewidth = 0.3
  ) +
  geom_vline(
    data = first_nonpositive,
    aes(xintercept = first_nonpositive_lag),
    inherit.aes = FALSE,
    linetype = "dashed",
    linewidth = 0.4
  ) +
  geom_line(alpha = 0.5) +
  facet_grid(
    variable ~ Site,
    scales = "free_y",
    labeller = labeller(
      variable = label_wrap_gen(width = 16)
    )
  ) +
  scale_color_manual(
    values = site_colors
  ) +
  labs(
    x = "Lag in days",
    y = "Autocorrelation",
    color = "Site"
  ) +
  theme_few() +
  theme(
    legend.position = "none"
  )

ggsave(
  filename = "temporal_autocorr.png",
  path = here("Figs/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = "retina"
)

autocorr_length_by_series <- acf_by_site |>
  filter(lag > 0) |>
  group_by(Site, name, variable) |>
  summarise(
    autocorr_length = if (any(autocorrelation <= 0, na.rm = TRUE)) {
      min(lag[autocorrelation <= 0], na.rm = TRUE)
    } else {
      NA_real_
    },
    .groups = "drop"
  )

avg_autocorr_length_by_site <- autocorr_length_by_series |>
  group_by(Site) |>
  summarise(
    avg_autocorr_length = mean(
      autocorr_length,
      na.rm = TRUE
    ),
    n_series = sum(!is.na(autocorr_length)),
    .groups = "drop"
  )
```


# Random Forest

```{r}
source(here("Code/Functions/rf_testing.R"))

rf_lagged_mod <- fit_full_model_new(
  df = best_lags_data |> select(-kg_name),
  label = "best_lag_rounded_scaled",
  class_weights = c("1" = 0.2, "2" = 0.6, "3" = 0.5),
  quiet = TRUE
)

fitted_model <- extract_fit_parsnip(rf_lagged_mod$final_fit)$fit
train_data <- rf_lagged_mod$train_data
test_data <- rf_lagged_mod$test_data
selected_vars <- rf_lagged_mod$selected_vars
tune_res <- rf_lagged_mod$tune_results
best_params <- rf_lagged_mod$best_params

shap_bootstrap_input <- make_shap_bootstrap_input(
  rf_lagged_mod,
  best_lags_data
)
load(here("Results/rf/model_objects_lag_rounded_scaled.RData"))
save(
  rf_lagged_mod,
  fitted_model,
  train_data,
  test_data,
  selected_vars,
  tune_res,
  best_params,
  shap_bootstrap_input,
  fit_full_model_new,
  file = here("Results/rf/model_objects_lag_rounded_scaled.RData")
)
beep(2)
```

```{r}
# How distinct is class 2 in the selected feature space?
train_data |>
  select(all_of(rf_lagged_mod$selected_vars), Cluster) |>
  pivot_longer(-Cluster) |>
  ggplot(aes(value, fill = Cluster)) +
  geom_density(alpha = 0.4) +
  facet_wrap(~name, scales = "free")

set.seed(42)
random_split <- initial_split(
  best_lags_data |> select(all_of(selected_vars)),
  prop = 0.8
)
random_train <- training(random_split)
random_test <- testing(random_split)

tb_split_plt <- train_data |>
  mutate(source = "Train") |>
  bind_rows(test_data |> mutate(source = "Test")) |>
  select(source, all_of(selected_vars)) |>
  pivot_longer(-source, names_to = "variable", values_to = "value") |>
  mutate(
    variable = case_when(
      variable %in% names(var_names) ~ var_names[variable],
      TRUE ~ variable # Keep original if not found in var_names
    )
  ) |>
  ggplot(aes(x = value, fill = source)) +
  geom_density(alpha = 0.5, na.rm = TRUE) +
  facet_wrap(~variable, scales = "free") +
  labs(title = "Time Block Split") +
  scale_fill_manual(values = c(Train = "#CB7223", Test = "#289A84")) +
  theme(legend.title = element_blank())


rnd_split_plt <- random_train |>
  mutate(source = "Train") |>
  bind_rows(random_test |> mutate(source = "Test")) |>
  select(source, where(is.numeric)) |>
  pivot_longer(-source, names_to = "variable", values_to = "value") |>
  mutate(
    variable = case_when(
      variable %in% names(var_names) ~ var_names[variable],
      TRUE ~ variable # Keep original if not found in var_names
    )
  ) |>
  ggplot(aes(x = value, fill = source)) +
  geom_density(alpha = 0.5, na.rm = TRUE) +
  facet_wrap(~variable, scales = "free") +
  labs(title = "Random 80/20 split") +
  scale_fill_manual(values = c(Train = "#CB7223", Test = "#289A84")) +
  theme(legend.title = element_blank())

tb_split_plt
rnd_split_plt
ggsave(
  filename = "tb_split_plt.png",
  plot = tb_split_plt,
  path = here("Figs/Supplement/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = 600
)
ggsave(
  filename = "rnd_split_plt.png",
  plot = rnd_split_plt,
  path = here("Figs/Supplement/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = 600
)
```


```{r}
# List of cluster labels
labels <- c("all_clusters", "cluster_1", "cluster_2", "cluster_3")
cluster_colors <- c("1" = "#0072B2", "2" = "#009E73", "3" = "#E69F00")
# Read all importance files into one long dataframe
importance_df <- read_csv(here(
  "Results/rf/best_lag_rounded_scaled_importance.csv"
)) |>
  #mutate(Cluster = label) |>
  mutate(
    predictor = case_when(
      predictor %in% names(var_names) ~ var_names[predictor],
      TRUE ~ predictor # Keep original if not found in var_names
    )
  ) |>
  mutate(norm_importance = importance / max(importance, na.rm = TRUE)) |>
  ungroup()


impVars <- importance_df |>
  mutate(predictor = str_wrap(predictor, width = 15)) |>
  ggplot(aes(x = reorder(predictor, importance), y = importance)) +
  geom_col(position = position_dodge(width = 0.8), fill = "#2C3E50") +
  coord_flip() +
  labs(
    x = "Predictor",
    y = "Scaled Variable Importance",
    title = "Random Forest Variable Importance"
  ) +
  theme_few() +
  theme(
    text = element_text(size = 18), # overall text size
    legend.text = element_text(size = 16), # legend entries
    legend.title = element_text(size = 18), # legend title
    axis.title = element_text(size = 18), # axis titles
    axis.text = element_text(size = 16), # axis tick labels
    plot.title = element_text(size = 20, hjust = 0.5),
    legend.position = "none"
  )
impVars
conf_mat <- read_csv(here(
  "Results/rf/best_lag_rounded_scaled_temporal_confusion_matrix.csv"
)) |>
  group_by(Prediction) |>
  mutate(precision = n / sum(n), freq.Perc = round(n / sum(n), 3)) |>
  ungroup() |>
  group_by(Reference) |>
  mutate(recall = n / sum(n)) |>
  ungroup() |>
  mutate(
    label = paste0("(", round(precision, 2), " / ", round(recall, 2), ")")
  ) |>
  mutate(label = if_else(Prediction == Reference, label, NA_character_))

conf_mat_plt <- ggplot(conf_mat) +
  geom_tile(aes(x = Reference, y = Prediction, fill = freq.Perc)) +
  scale_fill_viridis_c(option = "viridis", direction = 1, limits = c(0, 1)) +
  geom_text(
    aes(
      x = Reference,
      y = Prediction,
      label = n,
      color = Prediction == Reference
    ),
    size = 8
  ) +
  scale_color_manual(
    values = c("TRUE" = "black", "FALSE" = "white"),
    guide = "none"
  ) +
  geom_text(
    aes(x = Reference, y = Prediction + 0.2, label = label),
    color = "black",
    size = 6
  ) +
  scale_y_continuous(
    trans = "reverse",
    expand = c(0, 0),
    breaks = 1:3,
    labels = paste("State", 1:3)
  ) +
  scale_x_continuous(
    position = "top",
    expand = c(0, 0),
    breaks = 1:3,
    labels = paste("State", 1:3)
  ) +
  labs(
    fill = "Fraction Predicted",
    x = "Actual",
    y = "Predicted"
  ) +
  theme(
    panel.background = element_rect(fill = NA, colour = "black", linewidth = 1),
    axis.line = element_line(colour = "black"),
    legend.position = "bottom",
    legend.key.width = unit(1.5, "cm"),
    legend.key.height = unit(0.3, "cm"),
    legend.title = element_text(vjust = 1),
    axis.text.y = element_text(angle = 90, hjust = .5)
  )
conf_mat_plt
```

```{r}
cor_mtest <- function(mat, conf.level = 0.95) {
  mat <- as.matrix(mat)
  n <- ncol(mat)
  p.mat <- matrix(NA_real_, n, n)
  lowCI <- matrix(NA_real_, n, n)
  uppCI <- matrix(NA_real_, n, n)

  diag(p.mat) <- 0
  diag(lowCI) <- 1
  diag(uppCI) <- 1

  for (i in seq_len(n - 1)) {
    for (j in (i + 1):n) {
      tmp <- cor.test(
        mat[, i],
        mat[, j],
        use = "complete.obs",
        conf.level = conf.level
      )

      p.mat[i, j] <- tmp$p.value
      p.mat[j, i] <- tmp$p.value

      lowCI[i, j] <- tmp$conf.int[1]
      lowCI[j, i] <- tmp$conf.int[1]

      uppCI[i, j] <- tmp$conf.int[2]
      uppCI[j, i] <- tmp$conf.int[2]
    }
  }

  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  colnames(lowCI) <- rownames(lowCI) <- colnames(mat)
  colnames(uppCI) <- rownames(uppCI) <- colnames(mat)

  list(p = p.mat, lowCI = lowCI, uppCI = uppCI)
}

num_data <- best_lags_data |>
  dplyr::select(where(is.numeric))

cor_mat <- cor(num_data, use = "complete.obs")
test_res <- cor_mtest(num_data)

pretty_names <- var_names[colnames(cor_mat)] |>
  stringr::str_wrap(width = 10)

dimnames(cor_mat) <- list(pretty_names, pretty_names)
dimnames(test_res$p) <- list(pretty_names, pretty_names)

png(here("Figs/best_lag_corr.png"), width = 2160, height = 2160, res = 300)

corrplot(
  cor_mat,
  method = "number",
  type = "lower",
  order = "hclust",
  number.cex = 0.9,
  tl.pos = "d",
  tl.col = "black",
  tl.cex = 0.75,
  diag = T,
  p.mat = test_res$p,
  sig.level = 0.05,
  insig = "blank",
  col = colorRampPalette(c("#e8000d", "#dde5ed", "#0051ba"))(200),
  cl.lim = c(-1, 1)
)

dev.off()
```

# SHAP
05_calc_bootstrapped_shap.R

```{r}
# Read all saved kernelshap objects
files <- list.files(
  path = here("Results/SHAP_Rounded_Lags_Scaled"),
  pattern = "kernelshap_iter_\\d+\\.rds",
  full.names = TRUE
)
shap_list <- lapply(files, readRDS)

# Overall Mean |SHAP| with 95% CI
overall_mean_abs_shap <- files |>
  purrr::map(\(file) {
    ks <- readRDS(file)$kernelshap

    class_means <- purrr::map(
      ks$S,
      \(shap_values) colMeans(abs(shap_values))
    )

    mean_abs_shap <- Reduce(`+`, class_means) / length(class_means)

    tibble(
      predictor = names(mean_abs_shap),
      mean_abs_shap = unname(mean_abs_shap)
    )
  }) |>
  purrr::list_rbind() |>
  summarise(
    lo = quantile(mean_abs_shap, 0.025),
    hi = quantile(mean_abs_shap, 0.975),
    mean_abs_shap = mean(mean_abs_shap),
    .by = predictor
  ) |>
  mutate(
    predictor = case_when(
      predictor %in% names(var_names) ~ var_names[predictor],
      TRUE ~ predictor
    )
  )

shap_imp <- overall_mean_abs_shap |>
  ggplot(aes(x = mean_abs_shap, y = reorder(predictor, mean_abs_shap))) +
  geom_col(fill = "#2C3E50", width = 0.7) +
  geom_errorbar(aes(xmin = lo, xmax = hi), width = 0.25, orientation = "y") +
  scale_x_continuous(expand = expansion(mult = c(0, 0.05))) +
  labs(
    title = "SHAP Variable Importance",
    x = "Mean |SHAP|",
    y = "Predictor"
  )
shap_imp

# State-level SHAP with 95% CI
clust_shap <- files |>
  purrr::map(\(file) {
    bootstrap <- readRDS(file)

    bootstrap$kernelshap$S |>
      purrr::imap(\(shap_mat, cluster) {
        mean_abs_shap <- colMeans(abs(shap_mat))

        tibble(
          Boot = bootstrap$metadata$iteration,
          Cluster = cluster,
          predictor = names(mean_abs_shap),
          shap = unname(mean_abs_shap)
        )
      }) |>
      purrr::list_rbind() |>
      mutate(
        norm_shap = shap / sum(shap),
        .by = c(Boot, Cluster)
      )
  }) |>
  purrr::list_rbind() |>
  summarise(
    lo = quantile(norm_shap, 0.025),
    hi = quantile(norm_shap, 0.975),
    norm_shap = mean(norm_shap),
    .by = c(Cluster, predictor)
  ) |>
  mutate(
    ClusterLabel = paste("State", Cluster),
    predictor = case_when(
      predictor %in% names(var_names) ~ var_names[predictor],
      TRUE ~ predictor
    )
  )

cluster_var_prop <- clust_shap |>
  ggplot(aes(
    x = reorder(predictor, -norm_shap),
    y = norm_shap,
    fill = Cluster
  )) +
  geom_col(position = position_dodge(width = 0.8), alpha = 0.8, width = 0.75) +
  geom_errorbar(
    aes(ymin = lo, ymax = hi),
    position = position_dodge(width = 0.8),
    width = 0.25
  ) +
  scale_x_discrete(labels = label_wrap(15)) +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "Relative SHAP Importance by Regime",
    x = "Predictor",
    y = "Proportion of Total |SHAP|",
    fill = "State"
  )
cluster_var_prop


ggsave(
  filename = "rounded_lags_cluster_var_prop.png",
  plot = cluster_var_prop,
  path = here("Figs/RF_Models/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = "retina"
)

# Region-level SHAP with 95% CI
region_shap <- files |>
  purrr::map(\(file) {
    bootstrap <- readRDS(file)

    if (is.null(bootstrap$shap_context)) {
      stop(
        "This result has no `shap_context`; regenerate it with the updated bootstrap script.",
        call. = FALSE
      )
    }

    context <- bootstrap$shap_context |>
      left_join(
        site_info |> select(Site, Region),
        by = join_by(Site),
        relationship = "many-to-one"
      )

    if (anyNA(context$Region)) {
      stop("At least one explained site has no matching region.", call. = FALSE)
    }

    bootstrap$kernelshap$S |>
      purrr::imap(\(shap_mat, cluster) {
        as_tibble(shap_mat) |>
          bind_cols(context) |>
          pivot_longer(
            -c(Site, Region),
            names_to = "predictor",
            values_to = "shap"
          ) |>
          summarise(
            shap = mean(abs(shap)),
            .by = c(Region, predictor)
          ) |>
          mutate(
            Boot = bootstrap$metadata$iteration,
            Cluster = cluster,
            .before = Region
          )
      }) |>
      purrr::list_rbind() |>
      mutate(
        norm_shap = shap / sum(shap),
        .by = c(Boot, Cluster, Region)
      )
  }) |>
  purrr::list_rbind() |>
  summarise(
    lo = quantile(norm_shap, 0.025),
    hi = quantile(norm_shap, 0.975),
    norm_shap = mean(norm_shap),
    .by = c(Cluster, Region, predictor)
  ) |>
  mutate(
    ClusterLabel = paste("State", Cluster),
    predictor = case_when(
      predictor %in% names(var_names) ~ var_names[predictor],
      TRUE ~ predictor
    )
  ) |>
  mutate(
    pred_mean_region = mean(norm_shap, na.rm = TRUE),
    .by = c(Region, predictor)
  ) |>
  mutate(
    predictor = reorder(predictor, pred_mean_region),
    Cluster = factor(Cluster, levels = c("1", "2", "3"))
  )


zero_trim <- function(x) {
  ifelse(x == 0, "0", label_number(accuracy = 0.01)(x))
}

region_cluster_var_imp <- ggplot(
  region_shap,
  aes(x = predictor, y = norm_shap, fill = Cluster)
) +
  geom_col(
    position = position_dodge(width = 0.8, reverse = TRUE),
    width = 0.75,
    alpha = 0.9
  ) +
  geom_errorbar(
    aes(ymin = lo, ymax = hi),
    position = position_dodge(width = 0.8, reverse = TRUE),
    width = 0.25
  ) +
  facet_wrap(~Region, scales = "fixed") +
  coord_flip() +
  scale_fill_manual(values = cluster_colors) +
  scale_y_continuous(
    labels = zero_trim,
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(x = NULL, y = "Proportion of Total |SHAP|", fill = "State") +
  theme(
    panel.spacing.x = unit(2.0, "lines"),
    strip.text = element_text(face = "bold", size = 20)
  )

region_cluster_var_imp

ggsave(
  filename = "rounded_lags_region_shap.png",
  plot = region_cluster_var_imp,
  path = here("Figs/RF_Models/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = "retina"
)
```


```{r}
# Individual-observation SHAP values from a representative bootstrap draw
beeswarm_file <- files[[1L]]
bootstrap <- readRDS(beeswarm_file)
ks <- bootstrap$kernelshap

if (is.null(bootstrap$shap_context)) {
  stop("This result does not contain saved site context.", call. = FALSE)
}

context <- bootstrap$shap_context |>
  mutate(.observation = row_number())

shap_beeswarm <- ks$S |>
  purrr::imap(\(shap_mat, cluster) {
    shap_long <- shap_mat |>
      as_tibble() |>
      mutate(.observation = row_number()) |>
      pivot_longer(
        -.observation,
        names_to = "predictor",
        values_to = "shap"
      )

    feature_long <- ks$X |>
      as_tibble() |>
      mutate(.observation = row_number()) |>
      pivot_longer(
        -.observation,
        names_to = "predictor",
        values_to = "feat_value"
      )

    shap_long |>
      left_join(
        feature_long,
        by = join_by(.observation, predictor),
        relationship = "many-to-one"
      ) |>
      left_join(
        context,
        by = join_by(.observation),
        relationship = "many-to-one"
      ) |>
      filter(!is.na(feat_value)) |>
      mutate(
        Cluster = cluster,
        Boot = bootstrap$metadata$iteration
      )
  }) |>
  purrr::list_rbind() |>
  mutate(
    feat_scaled = {
      feature_range <- range(feat_value, na.rm = TRUE)

      if (diff(feature_range) == 0) {
        rep(0.5, n())
      } else {
        (feat_value - feature_range[[1]]) / diff(feature_range)
      }
    },
    .by = predictor
  ) |>
  mutate(
    predictor = case_when(
      predictor %in% names(var_names) ~ var_names[predictor],
      TRUE ~ predictor
    ),
    ClusterLabel = paste("State", Cluster),
    Cluster = factor(Cluster, levels = c("1", "2", "3"))
  ) |>
  mutate(
    pred_order = mean(abs(shap), na.rm = TRUE),
    .by = predictor
  ) |>
  mutate(predictor = reorder(predictor, pred_order))

shap_bee <- ggplot(
  shap_beeswarm,
  aes(x = shap, y = predictor, colour = feat_scaled)
) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
  ggbeeswarm::geom_quasirandom(
    groupOnX = FALSE,
    alpha = 0.6,
  ) +
  facet_wrap(~ClusterLabel) +
  scale_color_viridis_c(
    option = "plasma",
    breaks = c(0, 1),
    labels = c("Low", "High")
  ) +
  #scale_x_continuous(labels = scales::label_number(accuracy = 0.001)) +
  labs(
    title = "",
    x = "SHAP Value",
    y = NULL,
    colour = "Feature\nValue"
  ) +
  coord_cartesian(clip = "off") +
  theme(
    plot.margin = ggplot2::margin(5, 20, 5, 5, "mm"),
    panel.spacing = unit(1.5, "cm"),
    legend.key.height = unit(1.5, "cm"),
    strip.text = element_text(face = "bold", size = 14)
  )

shap_bee

ggsave(
  filename = "shap_beeswarm.png",
  plot = shap_bee,
  path = here("Figs/"),
  device = "png",
  width = 45,
  height = 25,
  units = "cm",
  dpi = "retina"
)

shap_beeswarm_region <- shap_beeswarm |>
  left_join(site_info |> select(Site, Region), by = "Site")

shap_bee_region <- ggplot(
  shap_beeswarm_region,
  aes(x = shap, y = predictor, colour = feat_scaled)
) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey40") +
  ggbeeswarm::geom_quasirandom(
    groupOnX = FALSE,
    alpha = 0.4,
    size = 0.6
  ) +
  facet_grid(Region ~ Cluster, labeller = labeller(Cluster = label_both)) +
  scale_color_viridis_c(
    option = "plasma",
    breaks = c(0, 1),
    labels = c("Low", "High")
  ) +
  # scale_x_continuous(labels = scales::label_number(accuracy = 0.001)) +
  labs(
    title = "SHAP Summary Plot by Region and State",
    x = "SHAP Value",
    y = NULL,
    colour = "Feature\nValue"
  ) +
  theme(
    legend.key.height = unit(1.5, "cm"),
    strip.text = element_text(face = "bold", size = 14),
    strip.text.y = element_text(angle = 0), # readable region labels
    panel.spacing = unit(1.2, "lines"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

shap_bee_region

ggsave(
  filename = "shap_beeswarm_region.png",
  plot = shap_bee_region,
  path = here("Figs/"),
  device = "png",
  width = 45,
  height = 40, # taller to fit region rows
  units = "cm",
  dpi = "retina"
)
```

## SHAP + Confusion Matrix

```{r}
# List of cluster labels
labels <- c("all_clusters", "cluster_1", "cluster_2", "cluster_3")
cluster_colors <- c("1" = "#0072B2", "2" = "#009E73", "3" = "#E69F00")

# Read all importance files into one long dataframe
importance_df <- read_csv(here(
  "Results/rf/best_lag_rounded_scaled_importance.csv"
)) |>
  #mutate(Cluster = label) |>
  mutate(
    predictor = case_when(
      predictor %in% names(var_names) ~ var_names[predictor],
      TRUE ~ predictor # Keep original if not found in var_names
    )
  ) |>
  mutate(norm_importance = importance / max(importance, na.rm = TRUE)) |>
  ungroup()


impVars <- importance_df |>
  mutate(predictor = str_wrap(predictor, width = 15)) |>
  ggplot(aes(x = reorder(predictor, importance), y = importance)) +
  geom_col(position = position_dodge(width = 0.8), fill = "#2C3E50") +
  coord_flip() +
  labs(
    x = "Predictor",
    y = "Scaled Variable Importance",
    title = "Random Forest Variable Importance"
  ) +
  theme_few() +
  theme(
    text = element_text(size = 18), # overall text size
    legend.text = element_text(size = 16), # legend entries
    legend.title = element_text(size = 18), # legend title
    axis.title = element_text(size = 18), # axis titles
    axis.text = element_text(size = 16), # axis tick labels
    plot.title = element_text(size = 20, hjust = 0.5),
    legend.position = "none"
  )
impVars


conf_mat <- read_csv(here(
  "Results/rf/best_lag_rounded_scaled_temporal_confusion_matrix.csv"
)) |>
  group_by(Prediction) |>
  mutate(precision = n / sum(n), freq.Perc = round(n / sum(n), 3)) |>
  ungroup() |>
  group_by(Reference) |>
  mutate(recall = n / sum(n)) |>
  ungroup() |>
  mutate(
    label = paste0("(", round(precision, 2), " / ", round(recall, 2), ")")
  ) |>
  mutate(label = if_else(Prediction == Reference, label, NA_character_))

conf_mat_plt <- ggplot(conf_mat) +
  geom_tile(aes(x = Reference, y = Prediction, fill = freq.Perc)) +
  scale_fill_viridis_c(option = "viridis", direction = 1, limits = c(0, 1)) +
  geom_text(
    aes(
      x = Reference,
      y = Prediction,
      label = n,
      color = Prediction == Reference
    ),
    size = 8
  ) +
  scale_color_manual(
    values = c("TRUE" = "black", "FALSE" = "white"),
    guide = "none"
  ) +
  geom_text(
    aes(x = Reference, y = Prediction + 0.2, label = label),
    color = "black",
    size = 6
  ) +
  scale_y_continuous(
    trans = "reverse",
    expand = c(0, 0),
    breaks = 1:3,
    labels = paste("State", 1:3)
  ) +
  scale_x_continuous(
    position = "top",
    expand = c(0, 0),
    breaks = 1:3,
    labels = paste("State", 1:3)
  ) +
  labs(
    fill = "Fraction Predicted",
    x = "Actual",
    y = "Predicted"
  ) +
  theme(
    panel.background = element_rect(fill = NA, colour = "black", linewidth = 1),
    axis.line = element_line(colour = "black"),
    legend.position = "bottom",
    legend.key.width = unit(1.5, "cm"),
    legend.key.height = unit(0.3, "cm"),
    legend.title = element_text(vjust = 1),
    axis.text.y = element_text(angle = 90, hjust = .5)
  )
conf_mat_plt
shap_imp

shap_conf <- plot_grid(conf_mat_plt, shap_imp, label_size = 0) +
  theme(plot.tag = element_text(face = "bold", size = 20))
shap_conf
ggsave(
  filename = "rounded_lags_shap_confmat.png",
  plot = shap_conf,
  path = here("Figs/RF_Models/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = 600
)

ggsave(
  filename = "conf_mat.png",
  plot = conf_mat_plt,
  path = here("Figs/RF_Models/"),
  device = "png",
  width = 20,
  height = 20,
  units = "cm",
  dpi = 600
)
```

### Temporal LOSO

```{r}
conf_mat_loso <- read_csv(here(
  "Results/rf/best_lag_rounded_scaled_site_loso_confusion_matrix.csv"
)) |>
  group_by(Prediction) |>
  mutate(precision = n / sum(n), freq.Perc = round(n / sum(n), 3)) |>
  ungroup() |>
  group_by(Reference) |>
  mutate(recall = n / sum(n)) |>
  ungroup() |>
  mutate(
    label = paste0("(", round(precision, 2), " / ", round(recall, 2), ")")
  ) |>
  mutate(label = if_else(Prediction == Reference, label, NA_character_))

conf_mat_loso_plt <- ggplot(conf_mat_loso) +
  geom_tile(aes(x = Reference, y = Prediction, fill = freq.Perc)) +
  scale_fill_viridis_c(option = "viridis", direction = 1, limits = c(0, 1)) +
  geom_text(
    aes(
      x = Reference,
      y = Prediction,
      label = n,
      color = Prediction == Reference
    ),
    size = 8
  ) +
  scale_color_manual(
    values = c("TRUE" = "black", "FALSE" = "white"),
    guide = "none"
  ) +
  geom_text(
    aes(x = Reference, y = Prediction + 0.2, label = label),
    color = "black",
    size = 6
  ) +
  scale_y_continuous(
    trans = "reverse",
    expand = c(0, 0),
    breaks = 1:3,
    labels = paste("State", 1:3)
  ) +
  scale_x_continuous(
    position = "top",
    expand = c(0, 0),
    breaks = 1:3,
    labels = paste("State", 1:3)
  ) +
  labs(
    fill = "Fraction Predicted",
    x = "Actual",
    y = "Predicted"
  ) +
  theme(
    panel.background = element_rect(fill = NA, colour = "black", linewidth = 1),
    axis.line = element_line(colour = "black"),
    legend.position = "bottom",
    legend.key.width = unit(1.5, "cm"),
    legend.key.height = unit(0.3, "cm"),
    legend.title = element_text(vjust = 1),
    axis.text.y = element_text(angle = 90, hjust = .5)
  )

ggsave(
  filename = "conf_mat_loso.png",
  plot = conf_mat_loso_plt,
  path = here("Figs/RF_Models/"),
  device = "png",
  width = 20,
  height = 20,
  units = "cm",
  dpi = 600
)
```

# ALE
```{r}
library(ale)

rf_pred_fun <- function(object, newdata, type = "response") {
  predictions <- predict(object, data = newdata)$predictions

  # A probability forest already returns an observation × class matrix.
  if (is.matrix(predictions) && is.numeric(predictions)) {
    return(predictions)
  }

  # For a standard classification forest, convert tree votes to
  # per-class vote proportions.
  tree_votes <- predict(
    object,
    data = newdata,
    predict.all = TRUE
  )$predictions

  class_labels <- object$forest$levels
  if (is.null(class_labels)) {
    stop("Could not determine ranger class labels.", call. = FALSE)
  }

  probabilities <- vapply(
    seq_along(class_labels),
    \(class_index) rowMeans(tree_votes == class_index),
    numeric(nrow(tree_votes))
  )
  colnames(probabilities) <- class_labels

  probabilities
}

ale_file <- here("Results/rf/ale_rounded_lags_scaled.rds")
dir.create(dirname(ale_file), recursive = TRUE, showWarnings = FALSE)

ale_input <- train_data |>
  select(all_of(c(selected_vars, "Cluster"))) |>
  drop_na() |>
  mutate(Cluster = factor(Cluster))

if (file.exists(ale_file)) {
  ale_rf <- readRDS(ale_file)
} else {
  old_max_size <- getOption("future.globals.maxSize")
  on.exit(options(future.globals.maxSize = old_max_size), add = TRUE)

  options(future.globals.maxSize = 2 * 1024^3)
  plan(multisession, workers = parallelly::availableCores(omit = 1L))
  ale_rf <- ale::ALE(
    model = fitted_model,
    data = ale_input,
    y_col = "Cluster",
    #y_type = "categorical",
    pred_fun = rf_pred_fun,
    x_cols = list(d1 = TRUE),
    boot_it = 1000,
    p_values = NULL,
    parallel = "all but one",
    model_packages = "ranger",
    seed = 42
  )

  saveRDS(ale_rf, ale_file)
  plan(sequential)
}

ale_plots <- plot(ale_rf)
print(ale_plots, ncol = 2)

rescale_against_ale <- function(x, ale_values) {
  limits <- range(ale_values, na.rm = TRUE)

  if (!all(is.finite(limits)) || diff(limits) == 0) {
    return(rep(0.5, length(x)))
  }

  (x - limits[[1]]) / diff(limits)
}

var_order <- overall_mean_abs_shap |>
  arrange(desc(mean_abs_shap)) |>
  pull(predictor)
```


```{r}
#| label: Ale Plot
# Extract ALE data for all clusters and variables
vars <- c(
  "Depth_0d",
  "ET_30d",
  "PPT_14d",
  "Q_0d",
  "active_network_length_m_0d",
  "avgPar_30d",
  "headGradient_30d",
  "water.temp_14d"
)

ale_data <- map_dfr(1:3, function(cat) {
  map_dfr(vars, function(var) {
    df <- ale_rf@effect[[cat]]$ale$d1[[var]]
    x_col <- names(df)[1]
    df |>
      rename(x_ceil = all_of(x_col)) |>
      select(x_ceil, .y, .y_lo, .y_hi, .n) |>
      mutate(variable = var, cluster = factor(cat))
  })
}) |>
  mutate(
    variable = case_when(
      variable %in% names(var_names) ~ var_names[variable],
      TRUE ~ variable
    )
  )

vars_ordered <- var_order[var_order %in% unique(ale_data$variable)]

# Apply factor ordering
ale_data <- ale_data |>
  mutate(variable = factor(variable, levels = vars_ordered))
# Get y reference lines from y_summary
y_refs <- as.data.frame(ale_rf@params$y_summary) |>
  rownames_to_column("quantile") |>
  filter(quantile %in% c("25%", "50%", "75%")) |>
  pivot_longer(-quantile, names_to = "cluster", values_to = "value") |>
  filter(cluster != "Cluster") |>
  mutate(
    cluster = factor(cluster),
    label = case_when(
      quantile == "25%" ~ "25%",
      quantile == "50%" ~ "median",
      quantile == "75%" ~ "75%"
    )
  )

# Extract raw data for rugs
raw_data <- ale_rf@params$data$data_sample
raw_data_cols <- names(raw_data)[names(raw_data) != "Cluster"]
vars_rename <- var_names[names(var_names) %in% raw_data_cols]
reverse_var_names <- setNames(names(vars_rename), vars_rename)

# Build rug data for all variables at once
rug_data <- map_dfr(vars_ordered, function(var) {
  orig_var <- unname(reverse_var_names[var])
  if (is.na(orig_var)) {
    orig_var <- var
  }
  raw_data |>
    select(x = all_of(orig_var)) |>
    mutate(variable = factor(var, levels = vars_ordered))
})

# ALE Plot
ale_plt <- ale_data |>
  filter(!is.na(x_ceil), !is.na(.y)) |>
  ggplot(aes(x = x_ceil, color = cluster, fill = cluster)) +
  geom_ribbon(
    aes(ymin = .y_lo, ymax = .y_hi),
    alpha = 0.15,
    color = NA
  ) +
  geom_line(aes(y = .y), linewidth = 0.8) +
  geom_rug(
    data = rug_data,
    aes(x = x, y = NULL),
    inherit.aes = FALSE,
    sides = "b",
    alpha = 0.2,
    linewidth = 0.3,
    color = "grey30"
  ) +
  geom_hline(
    yintercept = 0,
    linetype = "dashed",
    color = "grey40",
    linewidth = 0.5
  ) +
  scale_color_manual(
    values = cluster_colors,
    name = "State"
  ) +
  scale_fill_manual(
    values = cluster_colors,
    name = "State"
  ) +
  facet_wrap(
    ~variable,
    scales = "free_x",
    ncol = 4
  ) +
  labs(
    x = "Feature values",
    y = "ALE effect"
  ) +
  theme(
    strip.text = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.margin = ggplot2::margin(5, 6, 5, 5, "mm")
  )
ale_plt
ggsave(
  filename = "ale_plt.png",
  plot = ale_plt,
  here("Figs/"),
  device = "png",
  width = 40,
  height = 20,
  units = "cm",
  dpi = "retina"
)
```
