---
title: "Untitled"
format: html
editor: visual
---

```{r}
#| include: false
#| echo: false

library(tidyverse)
library(tidygraph)
library(data.table)
library(ggthemes)
library(here)
library(sf)
library(streamDAG)
library(furrr)
library(future)
library(riverdist)
library(sfnetworks)
source(here("Code/paths+packages.R"))
library(mapview)
gpGeopkg <- "~/Dropbox/AIMS/GIS/Results/AIMS_GreatPlains.gpkg"
# EPSG for NAD83 / UTM zone 14N
UTM_14N <- 26914
source(here("Code/AIMS_DAGs.R"))
source(here("Code/Functions/aims.R"))
source(here("Code/Functions/stic_StreamDAG.R"))
```

```{r}
stream <- st_read(gpGeopkg, layer = "Streams", quiet = T) |> 
  filter(watershed %in% c("KNZ", "SHN", "YMR"))

# Load sites
coords <- st_read(gpGeopkg, layer = "Sites", quiet = TRUE) |> 
  filter(watershed %in% c("KNZ", "SHN", "YMR")) |> 
  st_transform(crs = 4326) |> 
  mutate(
    lon = st_coordinates(geom)[, "X"],
    lat  = st_coordinates(geom)[, "Y"]
  ) |> 
  # Transform back to original CRS (NAD83 / UTM zone 14N)
  st_transform(crs = UTM_14N) 

shnStream <- stream |> filter(watershed == "SHN")
shnCoords <- coords |> filter(watershed == "SHN")

kmzStream <- stream |> filter(watershed == "KNZ")
kmzCoords <- coords |> filter(watershed == "KNZ")

ymrStream <- stream |> filter(watershed == "YMR")
ymrCoords <- coords |> filter(watershed == "YMR")
map <- mapview(
    coords,
    map.types = "Esri.WorldImagery",
    col.regions = "red",
    cex = 5,
    layer.name = "Sampling Sites"
  ) +
  mapview(
    stream,
    color = "blue",
    lwd = 2,
    layer.name = "Streams"
  ) 
map
```

# STICs

```{r}

sticsRaw <- list.files(here("Data/STICs"), full.names = TRUE, recursive = TRUE) |>
  # Only keep files in folders that end with _combined
  keep(\(x) basename(dirname(x)) |> str_detect("_combined$")) |>
  lapply(\(file) {
    read_csv(file, col_types = cols(.default = col_character()))
  }) |>
  bind_rows() |>
  rename(dateTime = datetime) |>
  filter(qual_rating != "poor") |>
  #filter(siteId %in% seSites$siteId & qual_rating != "poor") |>
  mutate(
    dateTime = ymd_hms(dateTime),
    wetdry = case_when(
    wetdry %in% c("dry", "0") ~ 0,
    wetdry %in% c("wet", "1") ~ 1,
    TRUE ~ NA_real_),
    site_subloc = coalesce(siteID, siteId)) |> 
  select(dateTime, site_subloc, wetdry)


# Step 1: Extract base ID and set priority
sticsRaw <- sticsRaw |>
  mutate(
    siteId = str_remove(site_subloc, "_(1|2|DUP)$"),
    priority = case_when(
      str_ends(site_subloc, "_1")   ~ 2,
      str_ends(site_subloc, "_DUP") ~ 3,
      str_ends(site_subloc, "_2")   ~ 4,
      TRUE                     ~ 1  # fallback for others
    )
  )

# Step 2: Convert to data.table
setDT(sticsRaw)

# Step 3: Sort so that higher priority comes first
setorder(sticsRaw, siteId, dateTime, priority)

# Step 4: Coalesce wetdry using first non-NA value with priority
sticsRaw <- sticsRaw[
  !is.na(wetdry), 
  .SD[1], 
  by = .(siteId, dateTime)
][, priority := NULL]


#sticsRaw[, sublocation := ifelse(is.na(rep), "HS", rep)]

sticsRaw <- sticsRaw |> 
  drop_na(siteId) |> 
  select(dateTime, siteId, wetdry) |> 
  filter(siteId %in% coords$siteId) |> 
  group_by(siteId) |>
  complete(dateTime = seq(min(dateTime), max(dateTime), by = "15 min")) |>
  arrange(siteId, dateTime) |>
  # Fill missing wetdry values with the last observation carried forward;
  # maxgap = 4 means only fill if the gap is no longer than 4 intervals (i.e. 1 hour)
  mutate(wetdry = zoo::na.locf(wetdry, na.rm = FALSE, maxgap = 8)) |>
  ungroup() |> 
  filter(minute(dateTime) %% 15 == 0) |> 
  pivot_wider(
    names_from = siteId,
    values_from = wetdry
    )

sticsRaw |> 
  ggplot(aes(y = SFM01, x = dateTime)) +
  geom_line(linewidth = 0.75) +
  theme_few()+
  labs(title = "SFM01",
       x = "")

sticsRaw |> 
  ggplot(aes(y = ENM01, x = dateTime)) +
  geom_line(linewidth = 0.75) +
  theme_few()+
  labs(title = "ENM01",
       x = "")

sticsRaw |> 
  ggplot(aes(y = SHM01, x = dateTime)) +
  geom_line(linewidth = 0.75) +
  theme_few()+
  labs(title = "SHM01",
       x = "")
```

# Fake STICs

```{r}
#Read in PTs for LTMs, since there are no STICs
pts <- aims_read(here("Data/STICs/PTs"), rType = "PRES", sublocation = "SW", fill = T) |> 
  select(siteId, dateTime = datetime, wetdry) |> 
  mutate(wetdry = ifelse(wetdry == "wet", 1, 
                          ifelse(wetdry == "dry", 0, NA))) |> 
  group_by(siteId) |>
  complete(dateTime = seq(min(dateTime), max(dateTime), by = "15 min")) |>
  arrange(siteId, dateTime) |>
  # Fill missing wetdry values with the last observation carried forward;
  # maxgap = 4 means only fill if the gap is no longer than 4 intervals (i.e. 1 hour)
  mutate(wetdry = zoo::na.locf(wetdry, na.rm = FALSE, maxgap = 4)) |>
  ungroup() |> 
  filter(minute(dateTime) %% 15 == 0) |> 
  mutate(siteId = paste0(siteId, "_PT")) |> 
  pivot_wider(
    names_from = siteId,
    values_from = wetdry
    )


pts |> 
  ggplot(aes(y = SFM01_PT, x = dateTime)) +
  geom_line(linewidth = 0.75) +
  theme_few()+
  labs(title = "SFM01",
       x = "")

pts |> 
  ggplot(aes(y = ENM01_PT, x = dateTime)) +
  geom_line(linewidth = 0.75) +
  theme_few()+
  labs(title = "ENM01",
       x = "")

pts |> 
  ggplot(aes(y = SHM01_PT, x = dateTime)) +
  geom_line(linewidth = 0.75) +
  theme_few()+
  labs(title = "SHM01",
       x = "")
```

# Join STICs and PTs

```{r}
# 1. Pivot both to long format
pts <- pts |> 
  pivot_longer(-dateTime, names_to = "siteId", values_to = "wet_PT") |>
  mutate(siteId = str_remove(siteId, "_PT$"))

sticsRaw <- sticsRaw |>
  pivot_longer(-dateTime, names_to = "siteId", values_to = "wet_STIC")

# 2. Join and coalesce values
stics_wide <- full_join(sticsRaw, pts, by = c("dateTime", "siteId")) |>
  mutate(wetdry = coalesce(wet_STIC, wet_PT)) |>
  select(dateTime, siteId, wetdry) |> 
  pivot_wider(names_from = siteId, values_from = wetdry) |> 
  mutate(ENM10 = coalesce(ENM10, ENM11),
         EN202 = coalesce(EN202, EN201)) |> 
    select(-ENM10, -EN201)


stics_wide |> 
  ggplot(aes(y = SFM01, x = dateTime)) +
  geom_line(linewidth = 0.75) +
  theme_few()+
  labs(title = "SFM01",
       x = "")

stics_wide |> 
  ggplot(aes(y = ENM01, x = dateTime)) +
  geom_line(linewidth = 0.75) +
  theme_few()+
  labs(title = "ENM01",
       x = "")

stics_wide |> 
  ggplot(aes(y = SHM01, x = dateTime)) +
  geom_line(linewidth = 0.75) +
  theme_few()+
  labs(title = "SHM01",
       x = "")


stics_wide |> 
  ggplot(aes(y = EN302, x = dateTime)) +
  geom_line(linewidth = 0.75) +
  theme_few()+
  labs(title = "EN302",
       x = "")
```

# Daily % Wet

```{r}
dailyPercWet <- stics_wide |>
  pivot_longer(-dateTime, names_to = "siteId", values_to = "wetdry") |> 
  left_join(coords |> select(siteId, watershed) |> st_drop_geometry(), by = "siteId") |> 
  mutate(Date = as.Date(dateTime)) |> 
  group_by(watershed, Date) |>                                
  summarize(
    PercWet_1d = sum(wetdry == 1, na.rm = TRUE) / sum(!is.na(wetdry)),
    n_stics = n_distinct(siteId[!is.na(wetdry)]),
    tot_stics = n_distinct(siteId),  # ← counted from all sites present on that day in this watershed
    pct_coverage = n_stics / tot_stics,
    .groups = "drop"
  ) |> 
  group_by(watershed) |> 
  mutate(dPercWet_1d = PercWet_1d - lag(PercWet_1d)) |> 
  ungroup() |> 
  filter(!is.na(PercWet_1d) & !is.na(dPercWet_1d)) |> 
  write_csv(here("Data/gpDailyPercWet.csv"))

percWet_15 <- stics_wide |>
  pivot_longer(-dateTime, names_to = "siteId", values_to = "wetdry") |> 
  left_join(coords |> select(siteId, watershed) |> st_drop_geometry(), by = "siteId") |> 
  group_by(watershed, dateTime) |>                                
  summarize(
    PercWet = sum(wetdry == 1, na.rm = TRUE) / sum(!is.na(wetdry)),
    n_stics = n_distinct(siteId[!is.na(wetdry)]),
    tot_stics = n_distinct(siteId),  # ← counted from all sites present on that day in this watershed
    pct_coverage = n_stics / tot_stics,
    .groups = "drop"
  ) |> 
  mutate(PercWet = case_when(is.na(PercWet)~0, T~PercWet)) |> 
  #filter(! & !is.na(dPercWet_1d)) |> 
  write_csv(here("Data/gp15minPercWet.csv"))

percWet_15 |>
  ggplot(aes(dateTime, PercWet)) +
  geom_point() +
  theme_few() +
    theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~watershed, scales = "free", ncol = 1) 

dailyPercWet |>
  ggplot(aes(Date, PercWet_1d)) +
  geom_point() +
  theme_few() +
    theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~watershed, scales = "free", ncol = 1) 

dailyPercWet |>
  ggplot(aes(Date, dPercWet_1d)) +
  geom_point() +
  theme_few() +
    theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~watershed, scales = "free", ncol = 1)

dailyPercWet |>
  ggplot(aes(Date, n_stics)) +
  geom_point() +
  theme_few() +
    theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))+
  facet_wrap(~watershed, scales = "free", ncol = 1)
```

# Site % Wet

```{r}
# Set up parallel processing - adjust workers as needed
plan(multisession, workers = availableCores() - 1)

# Process data
daily_mode <- stics_wide |>
  mutate(date = as_date(dateTime),
         DOY = lubridate::yday(date)) |>
  # Keep original date for gap calculation
  group_by(DOY) |>
  summarise(across(
    where(is.numeric),
    ~ preMetabolizer::calc_mode(.x, na.rm = TRUE)[1],
    .names = "{.col}"
  ), 
  date = first(date)) |>  # Preserve date for gap calculation
  ungroup() |>
  arrange(date) |>  # Ensure chronological order
  mutate(across(
    where(is.numeric) & !matches("DOY|date"),
    ~ case_when(.x == 1 ~ "Wet", .x == 0 ~ "Dry", TRUE ~ NA_character_)
  )) |>
  mutate(across(
    where(is.character) & !matches("DOY|date"),
    ~ {
      # Convert to numeric index for gap calculation
      rle <- rle(is.na(.x))
      cum_length <- cumsum(rle$lengths)
      # Identify fillable gaps (<10 days)
      fillable <- rle$values & 
        (rle$lengths < 10) & 
        (lag(rle$lengths, default = 0) > 0) & 
        (lead(rle$lengths, default = 0) > 0
      # Create fill vector
      fill_vec <- rep(NA, length(.x))
      for(i in which(fillable)) {
        start_idx <- ifelse(i == 1, 1, cum_length[i-1] + 1)
        end_idx <- cum_length[i]
        fill_vec[start_idx:end_idx] <- "Filled"
      }
      # Perform the fill
      filled <- .x
      filled[fill_vec == "Filled"] <- na.locf(filled, na.rm = FALSE)[fill_vec == "Filled"]
      filled[fill_vec == "Filled"] <- na.locf(filled, fromLast = TRUE, na.rm = FALSE)[fill_vec == "Filled"]
      filled
    }
  )) |>
  select(-date) |>  # Remove temporary date column
  pivot_longer(
    cols = -DOY,
    names_to = "siteId",
    values_to = "wet_dry"
  )

# Join with site coordinates
daily_map <- daily_mode |>
  left_join(coords, by = "siteId") |>
  sf::st_as_sf() |>  # Convert to sf before filtering to keep geometry for all sites
  mutate(watershed = case_when(
    watershed == "KNZ" ~ "Kings Creek",
    watershed == "SHN" ~ "Shane Creek",
    watershed == "YMR" ~ "Youngmeyer",
    TRUE ~ watershed
  ))

# Get all unique watersheds
unique_watersheds <- unique(daily_map$watershed)

# Create temporary directory for all watershed frames
tmp_dir <- tempfile("frames_")
dir.create(tmp_dir)

# Function to process a single watershed with device management
process_watershed <- function(watershed_name, daily_map, stream, tmp_dir) {
  # Explicitly open and close devices for each plot
  on.exit(while (dev.cur() > 1) dev.off(), add = TRUE)
  
  # Filter data for current watershed
  daily_map_ws <- daily_map |> 
    filter(watershed == watershed_name) |> 
    mutate(DateLabel = as.Date(DOY, origin = "1999-12-31"),
           MonthDay = format(DateLabel, "%B %d"))
  
  stream_ws <- stream |> 
    filter(watershed == case_when(
      watershed_name == "Kings Creek" ~ "KNZ",
      watershed_name == "Shane Creek" ~ "SHN",
      watershed_name == "Youngmeyer" ~ "YMR",
      TRUE ~ watershed_name
    ))
  
  # Create subdirectory for this watershed's frames
  ws_dir <- file.path(tmp_dir, watershed_name)
  dir.create(ws_dir, showWarnings = FALSE)
  
  # Identify all unique sites for this watershed
  all_sites <- daily_map_ws |> 
    st_drop_geometry() |> 
    distinct(siteId) |> 
    pull(siteId)
  
  # Generate individual frames
  unique_days <- unique(daily_map_ws$DateLabel)
  png_files <- file.path(ws_dir, sprintf("frame_%03d.png", seq_along(unique_days)))
  
  for (i in seq_along(unique_days)) {
    day <- unique_days[i]
    
    day_data <- daily_map_ws |> 
      filter(DateLabel == day) |> 
      complete(siteId = all_sites) |> 
      sf::st_as_sf()
    
    day_label <- unique(na.omit(day_data$MonthDay))
    
    # Explicit device management
    png(png_files[i], width = 1080, height = 1080)
    
    p <- ggplot() +
      geom_sf(data = stream_ws, color = "#0084A8", size = 1.5) +
      geom_sf(data = day_data |> filter(!is.na(wet_dry)),
              aes(color = wet_dry), size = 6) +
      geom_sf(data = day_data |> filter(is.na(wet_dry)),
              shape = 1, size = 6, color = "gray50") +  # Hollow circles for NA sites
      scale_color_manual(values = c("Dry" = "#D55E00", "Wet" = "#009E73"),
                        drop = FALSE) +
      labs(title = watershed_name,
           subtitle = day_label) +
      ggthemes::theme_few() +
      theme(
        plot.title = element_text(size = 30, face = "bold", hjust = 0.5),
        plot.subtitle = element_text(size = 25, hjust = 0.5),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.text = element_text(size = 20),
        axis.text = element_text(size = 20),  
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm")
      )
    
    print(p)
    dev.off()
  }
  
  # Combine frames into GIF
  output_gif <- paste0("~/Desktop/wet_dry_animation_", 
                      gsub(" ", "_", watershed_name), ".gif")
  gifski(png_files, 
         output_gif,
         width = 1080, 
         height = 1080, 
         delay = 0.2)
  
  return(output_gif)
}

# Process watersheds in parallel with progress bar
message("Processing watersheds in parallel...")
output_gifs <- future_map(unique_watersheds, 
                         ~process_watershed(.x, daily_map, stream, tmp_dir),
                         .options = furrr_options(seed = TRUE, 
                                                packages = c("tidyverse", "sf", "gifski")),
                         .progress = TRUE)

# Clean up
unlink(tmp_dir, recursive = TRUE)

# Print results
message("\nSuccessfully created animations:")
walk(output_gifs, ~message("- ", .x))
```

# StreamDAG

```{r}
source(here("Code/AIMS_DAGs.R"))
```

## Clean STICs

```{r}
stics <- stics_wide |>
  pivot_longer(-dateTime, names_to = "siteId", values_to = "wetdry") |> 
  left_join(coords |> select(siteId, watershed) |> st_drop_geometry(), by = "siteId")

daily_presAbs <- stics |>
  mutate(
    date = as_date(dateTime)
  ) |>
  filter(!is.na(watershed)) |>
  group_by(watershed, siteId, date) |>
  summarise(wetdry = calc_mode(wetdry), .groups = "drop") |>
  split(~watershed, drop = TRUE) |> 
  lapply(\(df) {
    df |>
      select(date, siteId, wetdry) |>
      pivot_wider(
        names_from = siteId,
        values_from = wetdry
      )
  })

# Convert the list to a single dataframe with watershed column
gpPresAbs <- bind_rows(daily_presAbs, .id = "watershed")

# Write to CSV
write_csv(gpPresAbs, here("Data/STICs/GP_DailyPresAbs.csv"))

daily_presAbs <- daily_presAbs |> imap(\(df, ws) {
  site_data <- df |> select(-date)
  
  # Logical vector: TRUE if any site has non-NA on that row
  non_na_rows <- site_data |> 
    pmap_lgl(\(...) any(!is.na(c(...))))
  
  # Find the first and last row with any data
  row_start <- which(non_na_rows) |> first()
  row_end   <- which(non_na_rows) |> last()
  
  # Slice to keep only rows between first and last data
  df |> slice(row_start:row_end)
})

daily_presAbs_clean <- map(daily_presAbs, stics_not_na)
doParallel::registerDoParallel(cores = nCores)
daily_presAbs_filled <- map(daily_presAbs_clean, STIC.RFimpute_parallel)

# Trim and rebuild networks based on daily_presAbs_filled
aimsNets <- imap(aimsNets, \(net, wshed) {
  keep_nodes <- names(daily_presAbs_filled[[wshed]])[-1]
  reconnect_and_remove_nodes(net, keep_nodes)
})
```

```{r}
shnDailyPresAbs <- daily_presAbs_clean[["SHN"]]
kmzDailyPresAbs <- daily_presAbs_clean[["KNZ"]]
ymrDailyPresAbs <- daily_presAbs_clean[["YMR"]]
```

```{r}

global_metrics <- c(
  "Size", "Diameter", "Graph.order", "n.Sources", "Mean.a.centrality", 
  "n.Paths.to.sink", "Path.length.mean", "Path.length.var", "Path.length.skew", 
  "Path.length.kurt", "Degree.mean", "Degree.var", "Degree.skew", "Degree.kurt", 
  "Shreve.number", "Strahler.number", "First.Zagreb", "Second.Zagreb", "ABC", 
  "Harary", "Global.efficiency", "Assort.in.out", "Assort.in.in"
)
```

# Shane

## Network Graph

```{r}

# Check which site IDs are not in the graph
missing_graph <- setdiff(shnCoords$siteId, names(V(aimsNets[["SHN"]])))
# Check which graph nodes are not in the coordinate data
missing_coords <- setdiff(names(V(aimsNets[["SHN"]])), shnCoords$siteId)
# Combine all sites from coords or graph
all_sites <- union(names(V(aimsNets[["SHN"]])), shnCoords$siteId)
# Find sites missing from the presence/absence data
missing_from_presabs <- setdiff(all_sites, colnames(shnDailyPresAbs)[-1])

spatial.plot(aimsNets[["SHN"]], shnCoords$long, shnCoords$lat, shnCoords$siteId, cex.text = .7)
```

## Unweighted

### Local

```{r}
localMetrics <- local.summary(aimsNets[["SHN"]]) |> 
  as_tibble(rownames = "metrics") |> 
  drop_na() |> 
  column_to_rownames("metrics") |>
  t() |>
  scale() |>
  t() |>
  as_tibble(rownames = "metrics") |> 
  pivot_longer(-metrics, names_to = "siteId", values_to = "value") |> 
  mutate(siteId = factor(siteId, levels = topo_sort(aimsNets[["SHN"]], mode = "out")$name))


localMetrics |> 
  ggplot(aes(x = siteId, y = value)) +
   geom_bar(stat = "identity") +
   theme_few() + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
   ylab("Standardized local measures") +
   xlab("\nNode")+
  facet_wrap(~metrics, ncol = 2)
```

### Global

```{r}
#global_metrics <- rownames(global.summary(aimsNets[["SHN"]], sink = "SHM01"))

# Set up parallel processing
plan(multisession, workers = 6)

# Process all rows in the original dataset
# No filtering by date range - use all dates in shnPresAbs$dateTime
global_list <- future_map(1:nrow(shnDailyPresAbs), function(i) {
  global.summary(
    delete.nodes.pa(aimsNets[["SHN"]], shnDailyPresAbs[i, -1], na.response = "treat.as.1"),
    sink = "SHM01"
  )
}, .options = furrr_options(seed = TRUE), .progress = TRUE)


global_df <- bind_rows(lapply(global_list, function(x) {
  as_tibble_row(setNames(as.list(x), global_metrics))
})) |> 
  mutate(dateTime = shnDailyPresAbs[[1]]) |> 
  relocate(dateTime, .before = 1) |> 
  mutate(across(
    where(is.numeric) & !all_of(c("Strahler.number", "Shreve.number")),
    ~ scale(.x) |> as.vector()
  )) |>
  pivot_longer(
    cols = -dateTime,           # All columns except date
    names_to = "metric",    # New column for metric names
    values_to = "value"     # New column for the values
  )

global_df |> 
  ggplot(aes(y = value, x = dateTime)) +
  geom_line(linewidth = 0.75, aes(color = metric)) +
  facet_wrap(~metric) +
  theme_few()+
    scale_color_manual(values = rep(okabeIto, length.out = length(unique(global_df$metric))))+
  labs(title = "Unweighted",
    x = "",
       y = "Stand. Global Measures") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")


```

## Calculate Stream Distance

```{r}
# Calculate the distance along the stream network
streamDist <- calc_stream_dist(shnStream, shnCoords)
# Extract edges from aimsNets[["SHN"]] and format them as "From -> To"
shnEdges <- aimsNets[["SHN"]] |> 
  as_edgelist() |> 
  apply(1, paste, collapse = " -> ")

# Filter streamDist to keep only arcs found in shnEdges
shn_lengths <- streamDist |> 
  filter(Arcs %in% shnEdges) |> 
  mutate(Lengths = dist_m / 1000) |> 
  select(Arcs, Lengths)

library(viridis)

# Map lengths to colors
shn_lengths$col <- viridis(nrow(shn_lengths))[rank(shn_lengths$Lengths)]
# match edge colors by name
edge_colors <- shn_lengths$col[match(shnEdges, shn_lengths$Arcs)]

spatial.plot(aimsNets[["SHN"]], shnCoords$lon, shnCoords$lat, shnCoords$siteId,
             arrow.col = edge_colors, arrow.lwd = 1.5, 
             col = "lightblue", cex.text = .7, plot.bg = "white")

```

## Weighted

### Local

```{r}
# Add lengths as weights to aimsNets[["SHN"]]
aimsNets[["SHN"]] <- aimsNets[["SHN"]] |> 
  as_tbl_graph() |> 
  activate(edges) |> 
  mutate(weight = shn_lengths$Lengths)

# Calculate strength and alpha centrality
weighted_local_scaled <- aimsNets[["SHN"]] |> 
  activate(nodes) |> 
  mutate(
    topological_order = node_topo_order(),  # Correct function for topological sorting
    StrengthLength = centrality_degree(weights = weight),
    AlphaCentrality = centrality_alpha(weights = weight)
  ) |> 
  as_tibble() |> 
  arrange(topological_order) |> 
  mutate(across(c(StrengthLength, AlphaCentrality), scale)) |> 
  pivot_longer(cols = c(StrengthLength, AlphaCentrality), names_to = "Measure", values_to = "Value") |> 
  mutate(name = factor(name, levels = unique(name)))

# Plot the results
ggplot(weighted_local_scaled, aes(x = name, y = Value)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Measure, scales = "free_y") +
  theme_few() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 4.3)) +
  labs(
    title = "Weighted", 
    y = "Stand. Local Measures",
    x = "Node"
  )
```

### Global

```{r}
# Define the metrics names
metric_names <- c(
  icsl = "ICSL",
  intact_to_sink = "IntactStreamLength",
  a_cent = "AlphaCentrality",
  harary = "Harary"
)

# Set up parallel processing
plan(multisession)

# Compute global metrics for each row of the dataframe
gWeightedList <- future_map(1:nrow(shnDailyPresAbs), function(i) {
  
  temp_graph <- delete.nodes.pa(aimsNets[["SHN"]], shnDailyPresAbs[i, -1], na.response = "treat.as.1")
  
  # Match arcs to lengths explicitly
  edges_attr <- attributes(E(temp_graph))$vname
  namelv <- gsub(" -> ", "|", shn_lengths$Arcs)
  length_sub <- shn_lengths$Lengths[match(edges_attr, namelv)]
  
  E(temp_graph)$weights <- length_sub
  
  tibble(
    icsl = ICSL(temp_graph, lengths = length_sub),
    intact_to_sink = size.intact.to.sink(temp_graph, sink = "SHM01"),
    a_cent = mean(alpha_centrality(temp_graph), na.rm = TRUE),
    harary = harary(temp_graph)
  )
}, .options = furrr_options(seed = TRUE), .progress = TRUE)
plan(sequential)

# Combine the list into a dataframe
gWeighted <- bind_rows(gWeightedList) |> 
  mutate(dateTime = shnDailyPresAbs[[1]]) |> 
  relocate(dateTime, .before = 1) |> 
  pivot_longer(cols = -dateTime, names_to = "metric", values_to = "value") |> 
  mutate(metric = recode(metric, !!!metric_names))

shnGlobalWeighted <- gWeighted |> 
  pivot_wider(names_from = metric, values_from = value) |> 
  mutate(siteId = "SHN")

gWeightedScaled <- bind_rows(gWeightedList) |> 
  mutate(dateTime = shnDailyPresAbs[[1]]) |> 
  relocate(dateTime, .before = 1) |> 
  mutate(across(where(is.numeric), ~ scale(.x) |> as.vector())) |> 
  pivot_longer(cols = -dateTime, names_to = "metric", values_to = "value") |> 
  mutate(metric = recode(metric, !!!metric_names))

# Plot using ggplot2
gWeighted |> 
  ggplot(aes(x = dateTime, y = value)) +
  geom_line(linewidth = 0.75, aes(color = metric)) +
  facet_wrap(~factor(metric, levels = metric_names), scales = "free_y") +
  theme_few() +
  scale_color_manual(values = okabeIto)+
  labs(
    title = "Weighted",
    x = "",
    y = "Global Measures",
    color = "Metric"
  ) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")

gWeightedScaled |> 
  ggplot(aes(x = dateTime, y = value)) +
  geom_line(linewidth = 0.75, aes(color = metric)) +
  facet_wrap(~factor(metric, levels = metric_names)) +
  theme_few() +
  scale_color_manual(values = okabeIto)+
  labs(
    title = "Shane Weighted STIC Metrics",
    x = "",
    y = "Stand. Global Measures",
    color = "Metric"
  ) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") -> p
  ggsave(filename = "SHN_weightedGlobal.png", plot = p, path = here("Figs/STICs/"), device = "png", width = 25.4, height = 14.2875, units = "cm", dpi = "retina")



```

## Arc Probabilities

```{r}
shnArcDailyPresAbs <- arc.pa.from.nodes(aimsNets[["SHN"]], shnDailyPresAbs[, -1])
shnMargProb  <- colMeans(shnArcDailyPresAbs, na.rm = T)
shnArcProb <- cor(shnArcDailyPresAbs, use = "pairwise.complete.obs")
shnArcCorProb <- R.bounds(shnMargProb, shnArcProb)
shnProb <- apply(shnArcCorProb, 2, mean)


ordProb <- order(shnProb)
ordProb <- order(ordProb)

col <- hcl.colors(27,  palette = "Inferno", rev = T)[ordProb]

spatial.plot(aimsNets[["SHN"]], shnCoords$lon, shnCoords$lat, shnCoords$siteId, arrow.col = col, arrow.lwd = 1.5,
             col = "lightblue", cex.text = .7, plot.bg = "white")
```

# Konza

## Site Info

```{r}
# Load stream network
kmzStream <- st_read(gpGeopkg, layer = "Streams", quiet = T) |> 
  filter(watershed == "KNZ") 

# Load sites
kmzCoords <- st_read(gpGeopkg, layer = "Sites", quiet = TRUE) |> 
  filter(watershed == "KNZ") |> 
  st_transform(crs = 4326) |> 
  mutate(
    longitude = st_coordinates(geom)[, "X"],
    latitude  = st_coordinates(geom)[, "Y"]
  ) |> 
  # Transform back to original CRS (NAD83 / UTM zone 14N)
  st_transform(crs = UTM_14N) 
# 
# knzKings <- mapview(
#     sfkcStreams, 
#     zcol = "strahler_order",  # Color by stream order
#     legend = TRUE, 
#     col.regions = viridis::viridis(3),  # Use Viridis color palette
#     map.types = "Esri.WorldImagery"
#   )
# 
# knzKings
# #shnS <- pointshp2segvert(path = here("Data/STICs/SHN_Coords.csv"), rivers = shnS)
```

## Network Graph

```{r}
# Check which site IDs are not in the graph
missing_graph <- setdiff(kmzCoords$siteId, names(V(aimsNets[["KNZ"]])))
# Check which graph nodes are not in the coordinate data
missing_coords <- setdiff(names(V(aimsNets[["KNZ"]])), kmzCoords$siteId)
# Combine all sites from coords or graph
all_sites <- union(names(V(aimsNets[["KNZ"]])), kmzCoords$siteId)
# Find sites missing from the presence/absence data
missing_from_presabs <- setdiff(all_sites, colnames(kmzDailyPresAbs)[-1])

spatial.plot(aimsNets[["KNZ"]], kmzCoords$long, kmzCoords$lat, kmzCoords$siteId, cex.text = .7)
```

## Unweighted

### Local

```{r}
localMetrics <- local.summary(aimsNets[["KNZ"]]) |> 
  as_tibble(rownames = "metrics") |> 
  drop_na() |> 
  column_to_rownames("metrics") |>
  t() |>
  scale() |>
  t() |>
  as_tibble(rownames = "metrics") |> 
  pivot_longer(-metrics, names_to = "siteId", values_to = "value") |> 
  mutate(siteId = factor(siteId, levels = topo_sort(aimsNets[["KNZ"]], mode = "out")$name))


localMetrics |> 
  ggplot(aes(x = siteId, y = value)) +
   geom_bar(stat = "identity") +
   theme_few() + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
   ylab("Standardized local measures") +
   xlab("\nNode")+
  facet_wrap(~metrics, ncol = 2)
```

### Global

```{r}
global_metrics <- rownames(global.summary(aimsNets[["KNZ"]], sink = "SFM01"))

# Set up parallel processing
plan(multisession, workers = 6)

# Process all rows in the original dataset
# No filtering by date range - use all dates in shnPresAbs$dateTime
global_list <- future_map(1:nrow(kmzDailyPresAbs), function(i) {
  global.summary(
    delete.nodes.pa(aimsNets[["KNZ"]], kmzDailyPresAbs[57, -1], na.response = "treat.as.1"),
    sink = "SFM01"
  )
}, .options = furrr_options(seed = TRUE), .progress = TRUE)
plan(sequential)

global_df <- bind_rows(lapply(global_list, function(x) {
  as_tibble_row(setNames(as.list(x), global_metrics))
})) |> 
  mutate(dateTime = kmzDailyPresAbs[[1]]) |> 
  relocate(dateTime, .before = 1) |> 
  # mutate(across(
  #   where(is.numeric) & !all_of(c("Strahler.number", "Shreve.number")),
  #   ~ scale(.x) |> as.vector()
  # )) |>
  pivot_longer(
    cols = -dateTime,           # All columns except date
    names_to = "metric",    # New column for metric names
    values_to = "value"     # New column for the values
  )

global_df |> 
  ggplot(aes(y = value, x = dateTime)) +
  geom_line(linewidth = 0.75, aes(color = metric)) +
  facet_wrap(~metric, scales = "free") +
  theme_few()+
    scale_color_manual(values = rep(okabeIto, length.out = length(unique(global_df$metric))))+
  labs(title = "Unweighted",
    x = "",
       y = "Stand. Global Measures") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")


```

```{r}
# Calculate the distance along the stream network
streamDist <- kmzStream |>
  st_combine() |> # Combine all stream lines into a single geometry
  st_distance(kmzCoords, along = _) |> # Calculate the distance along the stream network between snapped points
  # Set the row and column names of the distances matrix
  `rownames<-`(kmzCoords$siteId) |>
  `colnames<-`(kmzCoords$siteId) |>
  as_tibble(rownames = "From") |> 
  pivot_longer(-From, names_to = "To", values_to = "dist_m") |>
  mutate(Arcs = paste(From, "->", To),
         dist_m = as.numeric(dist_m))

# Extract edges from aimsNets[["SHN"]] and format them as "From -> To"
kmzEdges <- aimsNets[["KNZ"]] |> 
  as_edgelist() |> 
  apply(1, paste, collapse = " -> ")

# Filter streamDist to keep only arcs found in shnEdges
kmz_lengths <- streamDist |> 
  filter(Arcs %in% kmzEdges) |> 
  mutate(Lengths = dist_m / 1000) |> 
  select(Arcs, Lengths)

library(viridis)

# Map lengths to colors
kmz_lengths$col <- viridis(nrow(kmz_lengths))[rank(kmz_lengths$Lengths)]
# match edge colors by name
edge_colors <- kmz_lengths$col[match(kmzEdges, kmz_lengths$Arcs)]

spatial.plot(aimsNets[["KNZ"]], kmzCoords$long, kmzCoords$lat, kmzCoords$siteId,
             arrow.col = edge_colors, arrow.lwd = 1.5, 
             col = "lightblue", cex.text = .7, plot.bg = "white")

```

## Weighted

### Local

```{r}
# Add lengths as weights to aimsNets[["SHN"]]
aimsNets[["KNZ"]] <- aimsNets[["KNZ"]] |> 
  as_tbl_graph() |> 
  activate(edges) |> 
  mutate(weight = kmz_lengths$Lengths)

# Calculate strength and alpha centrality
weighted_local_scaled <- aimsNets[["KNZ"]] |> 
  activate(nodes) |> 
  mutate(
    topological_order = node_topo_order(),  # Correct function for topological sorting
    StrengthLength = centrality_degree(weights = weight),
    AlphaCentrality = centrality_alpha(weights = weight)
  ) |> 
  as_tibble() |> 
  arrange(topological_order) |> 
  mutate(across(c(StrengthLength, AlphaCentrality), scale)) |> 
  pivot_longer(cols = c(StrengthLength, AlphaCentrality), names_to = "Measure", values_to = "Value") |> 
  mutate(name = factor(name, levels = unique(name)))

# Plot the results
ggplot(weighted_local_scaled, aes(x = name, y = Value)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Measure, scales = "free_y") +
  theme_few() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 4.3)) +
  labs(
    title = "Weighted", 
    y = "Stand. Local Measures",
    x = "Node"
  )
```

### Global

```{r}
# Define the metrics names
metric_names <- c(
  icsl = "ICSL",
  intact_to_sink = "IntactStreamLength",
  a_cent = "AlphaCentrality",
  harary = "Harary"
)

# Set up parallel processing
plan(multisession)
# Compute global metrics for each row of the dataframe
gWeightedList <- future_map(1:nrow(kmzDailyPresAbs), function(i) {

  temp_graph <- delete.nodes.pa(aimsNets[["KNZ"]], kmzDailyPresAbs[i, -1], na.response = "treat.as.1")

  # Match arcs to lengths explicitly
  edges_attr <- attributes(E(temp_graph))$vname
  namelv <- gsub(" -> ", "|", kmz_lengths$Arcs)
  length_sub <- kmz_lengths$Lengths[match(edges_attr, namelv)]
  
  E(temp_graph)$weights <- length_sub
  
  tibble(
    icsl = ICSL(temp_graph, lengths = length_sub),
    intact_to_sink = size.intact.to.sink(temp_graph, sink = "SFM01"),
    a_cent = mean(alpha_centrality(temp_graph), na.rm = TRUE),
    harary = harary(temp_graph)
  )
}, .options = furrr_options(seed = TRUE), .progress = TRUE)
plan(sequential)

# Combine the list into a dataframe
gWeighted <- bind_rows(gWeightedList) |> 
  mutate(dateTime = kmzDailyPresAbs[[1]]) |> 
  relocate(dateTime, .before = 1) |> 
  pivot_longer(cols = -dateTime, names_to = "metric", values_to = "value") |> 
  mutate(metric = recode(metric, !!!metric_names))

kmzGlobalWeighted <- gWeighted |> 
  pivot_wider(names_from = metric, values_from = value) |> 
  mutate(siteId = "KNZ")

gWeightedScaled <- bind_rows(gWeightedList) |> 
  mutate(dateTime = kmzDailyPresAbs[[1]]) |> 
  relocate(dateTime, .before = 1) |> 
  mutate(across(where(is.numeric), ~ scale(.x) |> as.vector())) |> 
  pivot_longer(cols = -dateTime, names_to = "metric", values_to = "value") |> 
  mutate(metric = recode(metric, !!!metric_names))

# Plot using ggplot2
gWeighted |> 
  ggplot(aes(x = dateTime, y = value)) +
  geom_line(linewidth = 0.75, aes(color = metric)) +
  facet_wrap(~factor(metric, levels = metric_names), scales = "free_y") +
  theme_few() +
  scale_color_manual(values = okabeIto)+
  labs(
    title = "Weighted",
    x = "",
    y = "Global Measures",
    color = "Metric"
  ) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")

gWeightedScaled |> 
  ggplot(aes(x = dateTime, y = value)) +
  geom_line(linewidth = 0.75, aes(color = metric)) +
  facet_wrap(~factor(metric, levels = metric_names)) +
  theme_few() +
  scale_color_manual(values = okabeIto)+
  labs(
    title = "Konza Weighted STIC Metrics",
    x = "",
    y = "Stand. Global Measures",
    color = "Metric"
  ) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") -> p
p
ggsave(filename = "KNZ_weightedGlobal.png", plot = p, path = here("Figs/STICs/"), device = "png", width = 25.4, height = 14.2875, units = "cm", dpi = "retina")
```

## Arc Probabilities

```{r}
kmzArcDailyPresAbs <- arc.pa.from.nodes(aimsNets[["KNZ"]], kmzDailyPresAbs[, -1])
kmzMargProb  <- colMeans(kmzArcDailyPresAbs, na.rm = T)
kmzArcProb <- cor(kmzArcDailyPresAbs, use = "pairwise.complete.obs")
kmzArcCorProb <- R.bounds(kmzMargProb, kmzArcProb)
kmzProb <- apply(kmzArcCorProb, 2, mean)


ordProb <- order(kmzProb)
ordProb <- order(ordProb)

col <- hcl.colors(27,  palette = "Inferno", rev = T)[ordProb]

spatial.plot(aimsNets[["KNZ"]], kmzCoords$long, kmzCoords$lat, kmzCoords$siteId, arrow.col = col, arrow.lwd = 1.5,
             col = "lightblue", cex.text = .7, plot.bg = "white")
```

# YMR

## Site Info

```{r}
# Load stream network
ymrStream <- st_read(gpGeopkg, layer = "Streams", quiet = T) |> 
  filter(watershed == "YMR") 

# Load sites
ymrCoords <- st_read(gpGeopkg, layer = "Sites", quiet = TRUE) |> 
  filter(watershed == "YMR") |> 
  st_transform(crs = 4326) |> 
  mutate(
    longitude = st_coordinates(geom)[, "X"],
    latitude  = st_coordinates(geom)[, "Y"]
  ) |> 
  # Transform back to original CRS (NAD83 / UTM zone 14N)
  st_transform(crs = UTM_14N) 
```

## STICs

```{r}
# Check which site IDs are not in the graph
missing_graph <- setdiff(ymrCoords$siteId, names(V(aimsNets[["YMR"]])))
# Check which graph nodes are not in the coordinate data
missing_coords <- setdiff(names(V(aimsNets[["YMR"]])), ymrCoords$siteId)
# Combine all sites from coords or graph
all_sites <- union(names(V(aimsNets[["YMR"]])), ymrCoords$siteId)
# Find sites missing from the presence/absence data
missing_from_presabs <- setdiff(all_sites, colnames(ymrDailyPresAbs)[-1])

spatial.plot(aimsNets[["YMR"]], ymrCoords$long, ymrCoords$lat, ymrCoords$siteId, cex.text = .7)
```

## Unweighted

### Local

```{r}
localMetrics <- local.summary(aimsNets[["YMR"]]) |> 
  as_tibble(rownames = "metrics") |> 
  drop_na() |> 
  column_to_rownames("metrics") |>
  t() |>
  scale() |>
  t() |>
  as_tibble(rownames = "metrics") |> 
  pivot_longer(-metrics, names_to = "siteId", values_to = "value") |> 
  mutate(siteId = factor(siteId, levels = topo_sort(aimsNets[["YMR"]], mode = "out")$name))


localMetrics |> 
  ggplot(aes(x = siteId, y = value)) +
   geom_bar(stat = "identity") +
   theme_few() + 
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
   ylab("Standardized local measures") +
   xlab("\nNode")+
  facet_wrap(~metrics, ncol = 2)
```

### Global

```{r}
global_metrics <- rownames(global.summary(aimsNets[["YMR"]], sink = "ENM01"))

# Set up parallel processing
plan(multisession, workers = 6)

# Process all rows in the original dataset
# No filtering by date range - use all dates in shnPresAbs$dateTime
global_list <- future_map(1:nrow(ymrDailyPresAbs), function(i) {
  global.summary(
    delete.nodes.pa(aimsNets[["YMR"]], ymrDailyPresAbs[i, -1], na.response = "treat.as.1"),
    sink = "ENM01"
  )
}, .options = furrr_options(seed = TRUE), .progress = TRUE)


global_df <- bind_rows(lapply(global_list, function(x) {
  as_tibble_row(setNames(as.list(x), global_metrics))
})) |> 
  mutate(dateTime = ymrDailyPresAbs[[1]]) |> 
  relocate(dateTime, .before = 1) |> 
  mutate(across(
    where(is.numeric) & !all_of(c("Strahler.number", "Shreve.number")),
    ~ scale(.x) |> as.vector()
  )) |>
  pivot_longer(
    cols = -dateTime,           # All columns except date
    names_to = "metric",    # New column for metric names
    values_to = "value"     # New column for the values
  )

global_df |> 
  ggplot(aes(y = value, x = dateTime)) +
  geom_line(linewidth = 0.75, aes(color = metric)) +
  facet_wrap(~metric) +
  theme_few()+
    scale_color_manual(values = rep(okabeIto, length.out = length(unique(global_df$metric))))+
  labs(title = "Unweighted",
    x = "",
       y = "Stand. Global Measures") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")


```

## Calculate Stream Length

```{r}
# Calculate the distance along the stream network
streamDist <- ymrStream |>
  st_combine() |> # Combine all stream lines into a single geometry
  st_distance(ymrCoords, along = _) |> # Calculate the distance along the stream network between snapped points
  # Set the row and column names of the distances matrix
  `rownames<-`(ymrCoords$siteId) |>
  `colnames<-`(ymrCoords$siteId) |>
  as_tibble(rownames = "From") |> 
  pivot_longer(-From, names_to = "To", values_to = "dist_m") |>
  mutate(Arcs = paste(From, "->", To),
         dist_m = as.numeric(dist_m))

# Extract edges from aimsNets[["SHN"]] and format them as "From -> To"
ymrEdges <- aimsNets[["YMR"]] |> 
  as_edgelist() |> 
  apply(1, paste, collapse = " -> ")

# Filter streamDist to keep only arcs found in shnEdges
ymr_lengths <- streamDist |> 
  filter(Arcs %in% ymrEdges) |> 
  mutate(Lengths = dist_m/1000) |> 
  select(Arcs, Lengths)

library(viridis)

# Map lengths to colors
ymr_lengths$col <- viridis(nrow(ymr_lengths))[rank(ymr_lengths$Lengths)]
# match edge colors by name
edge_colors <- ymr_lengths$col[match(ymrEdges, ymr_lengths$Arcs)]

spatial.plot(aimsNets[["YMR"]], ymrCoords$long, ymrCoords$lat, ymrCoords$siteId,
             arrow.col = edge_colors, arrow.lwd = 1.5, 
             col = "lightblue", cex.text = .7, plot.bg = "white")

```

## Weighted

### Local

```{r}
# Add lengths as weights to aimsNets[["SHN"]]
aimsNets[["YMR"]] <- aimsNets[["YMR"]] |> 
  as_tbl_graph() |> 
  activate(edges) |> 
  mutate(weight = ymr_lengths$Lengths)

# Calculate strength and alpha centrality
weighted_local_scaled <- aimsNets[["YMR"]] |> 
  activate(nodes) |> 
  mutate(
    topological_order = node_topo_order(),  # Correct function for topological sorting
    StrengthLength = centrality_degree(weights = weight),
    AlphaCentrality = centrality_alpha(weights = weight)
  ) |> 
  as_tibble() |> 
  arrange(topological_order) |> 
  mutate(across(c(StrengthLength, AlphaCentrality), scale)) |> 
  pivot_longer(cols = c(StrengthLength, AlphaCentrality), names_to = "Measure", values_to = "Value") |> 
  mutate(name = factor(name, levels = unique(name)))

# Plot the results
ggplot(weighted_local_scaled, aes(x = name, y = Value)) +
  geom_bar(stat = "identity") +
  facet_wrap(~Measure, scales = "free_y") +
  theme_few() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, size = 4.3)) +
  labs(
    title = "Weighted", 
    y = "Stand. Local Measures",
    x = "Node"
  )
```

### Global

```{r}
# Define the metrics names
metric_names <- c(
  icsl = "ICSL",
  intact_to_sink = "IntactStreamLength",
  a_cent = "AlphaCentrality",
  harary = "Harary"
)

# Set up parallel processing
plan(multisession)

# Compute global metrics for each row of the dataframe
gWeightedList <- future_map(1:nrow(ymrDailyPresAbs), function(i) {

  temp_graph <- delete.nodes.pa(aimsNets[["YMR"]], ymrDailyPresAbs[i, -1], na.response = "treat.as.1")

  # Match arcs to lengths explicitly
  edges_attr <- attributes(E(temp_graph))$vname
  namelv <- gsub(" -> ", "|", ymr_lengths$Arcs)
  length_sub <- ymr_lengths$Lengths[match(edges_attr, namelv)]
  
  E(temp_graph)$weights <- length_sub
  
  tibble(
    icsl = ICSL(temp_graph, lengths = length_sub),
    intact_to_sink = size.intact.to.sink(temp_graph, sink = "ENM01"),
    a_cent = mean(alpha_centrality(temp_graph), na.rm = TRUE),
    harary = harary(temp_graph)
  )
}, .options = furrr_options(seed = TRUE), .progress = TRUE)
plan(sequential)

# Combine the list into a dataframe
gWeighted <- bind_rows(gWeightedList) |> 
  mutate(dateTime = ymrDailyPresAbs[[1]]) |> 
  relocate(dateTime, .before = 1) |> 
  pivot_longer(cols = -dateTime, names_to = "metric", values_to = "value") |> 
  mutate(metric = recode(metric, !!!metric_names))

ymrGlobalWeighted <- gWeighted |> 
  pivot_wider(names_from = metric, values_from = value) |> 
  mutate(siteId = "YMR")

gWeightedScaled <- bind_rows(gWeightedList) |> 
  mutate(dateTime = ymrDailyPresAbs[[1]]) |> 
  relocate(dateTime, .before = 1) |> 
  mutate(across(where(is.numeric), ~ scale(.x) |> as.vector())) |> 
  pivot_longer(cols = -dateTime, names_to = "metric", values_to = "value") |> 
  mutate(metric = recode(metric, !!!metric_names))

# Plot using ggplot2
gWeighted |> 
  ggplot(aes(x = dateTime, y = value)) +
  geom_line(linewidth = 0.75, aes(color = metric)) +
  facet_wrap(~factor(metric, levels = metric_names), scales = "free_y") +
  theme_few() +
  scale_color_manual(values = okabeIto)+
  labs(
    title = "Weighted",
    x = "",
    y = "Global Measures",
    color = "Metric"
  ) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")

gWeightedScaled |> 
  ggplot(aes(x = dateTime, y = value)) +
  geom_line(linewidth = 0.75, aes(color = metric)) +
  facet_wrap(~factor(metric, levels = metric_names)) +
  theme_few() +
  scale_color_manual(values = okabeIto)+
  labs(
    title = "Youngmeyer Weighted STIC Metrics",
    x = "",
    y = "Stand. Global Measures",
    color = "Metric"
  ) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") -> p
p
ggsave(filename = "YMR_weightedGlobal.png", plot = p, path = here("Figs/STICs/"), device = "png", width = 25.4, height = 14.2875, units = "cm", dpi = "retina")
```

# Save Global Metrics

```{r}
gWeightedMet <- shnGlobalWeighted |> 
  full_join(kmzGlobalWeighted) |> 
  full_join(ymrGlobalWeighted) |> 
  write_csv(here("Data/sticGlobalMetricsUnscaled.csv"))

gWeightedMet |>
  group_by(siteId) |>
  mutate(across(where(is.numeric), ~ scale(.x) |> as.vector())) |> 
  ggplot(aes(x = dateTime, y = AlphaCentrality, color = siteId, group = siteId)) +
  geom_line() +
  scale_color_manual(values = okabeIto) +  
  labs(title = "",
       x = "Date",
       y = "Alpha-Centrality",
       color = "") +
  theme_few() +
  theme(legend.position = "bottom") -> p
p
ggsave(filename = "AC.png", plot = p, path = here("Figs/STICs/"), device = "png", width = 25.4, height = 14.2875, units = "cm", dpi = "retina")


gWeightedMet |>
  group_by(siteId) |> 
  mutate(across(where(is.numeric), ~ scale(.x) |> as.vector())) |> 
  ggplot(aes(x = dateTime, y = ICSL, color = siteId, group = siteId)) +
  geom_line() +
  scale_color_manual(values = okabeIto) +  
  labs(title = "",
       x = "Date",
       y = "Integral Connectivity Scale Length",
       color = "") +
  theme_few() +
  theme(legend.position = "bottom") -> p
p
ggsave(filename = "ICSL.png", plot = p, path = here("Figs/STICs/"), device = "png", width = 25.4, height = 14.2875, units = "cm", dpi = "retina")


gWeightedMet |>
  group_by(siteId) |>
  mutate(across(where(is.numeric), ~ scale(.x) |> as.vector())) |> 
  ggplot(aes(x = dateTime, y = Harary, color = siteId, group = siteId)) +
  geom_line() +
  scale_color_manual(values = okabeIto) +  
  labs(title = "",
       x = "Date",
       y = "Harary Index",
       color = "") +
  theme_few() +
  theme(legend.position = "bottom") -> p
p
ggsave(filename = "Harary.png", plot = p, path = here("Figs/STICs/"), device = "png", width = 25.4, height = 14.2875, units = "cm", dpi = "retina")

gWeightedMet |>
  group_by(siteId) |>
  mutate(across(where(is.numeric), ~ scale(.x) |> as.vector())) |> 
  ggplot(aes(x = dateTime, y = IntactStreamLength, color = siteId, group = siteId)) +
  geom_line() +
  scale_color_manual(values = okabeIto) +  
  labs(title = "",
       x = "Date",
       y = "Intact Stream Length",
       color = "") +
  theme_few() +
  theme(legend.position = "bottom") -> p
p
ggsave(filename = "IntactStreamLength.png", plot = p, path = here("Figs/STICs/"), device = "png", width = 25.4, height = 14.2875, units = "cm", dpi = "retina")

```

## Arc Probabilities

```{r}
ymrArcDailyPresAbs <- arc.pa.from.nodes(aimsNets[["YMR"]], ymrDailyPresAbs[, -1])
ymrMargProb  <- colMeans(ymrArcDailyPresAbs, na.rm = T)
ymrArcProb <- cor(ymrArcDailyPresAbs, use = "pairwise.complete.obs")
ymrArcCorProb <- R.bounds(ymrMargProb, ymrArcProb)
ymrProb <- apply(ymrArcCorProb, 2, mean)


ordProb <- order(ymrProb)
ordProb <- order(ordProb)

col <- hcl.colors(27,  palette = "Inferno", rev = T)[ordProb]

spatial.plot(aimsNets[["YMR"]], ymrCoords$long, ymrCoords$lat, ymrCoords$siteId, arrow.col = col, arrow.lwd = 1.5, col = "lightblue", cex.text = .7, plot.bg = "white")
title("Youngmeyer Ranch")
```

# Calculate Upstream Active Network Length

```{r}
# KMZ
kmzUpstreamNodes <- calculate_network_metrics(
  network = aimsNets[["KNZ"]],
  lengths_data = kmz_lengths,
  pres_abs_data = kmzDailyPresAbs,
  site_id = "SFM01"
)

# SHN
shnUpstreamNodes <- calculate_network_metrics(
  network = aimsNets[["SHN"]],
  lengths_data = shn_lengths,
  pres_abs_data = shnDailyPresAbs,
  site_id = "SHM01"
)

# YMR
ymrUpstreamNodes <- calculate_network_metrics(
  network = aimsNets[["YMR"]],
  lengths_data = ymr_lengths,
  pres_abs_data = ymrDailyPresAbs,
  site_id = "ENM01"
)



kmzUpstreamNodes |>
  ggplot(aes(x = date, y = upstream_length)) +
  geom_line() +
  labs(x = "Date", y = "Upstream Connected Length (m)", 
       title = "Konza")

shnUpstreamNodes |>
  ggplot(aes(x = date, y = upstream_length)) +
  geom_line() +
  labs(x = "Date", y = "Upstream Connected Length (m)", 
       title = "Shane")

ymrUpstreamNodes |>
  ggplot(aes(x = date, y = upstream_length)) +
  geom_line() +
  labs(x = "Date", y = "Upstream Connected Length (m)", 
       title = "Youngmeyer")

kmzUpstreamNodes |>
  ggplot(aes(x = date, y = percentConnected)) +
  geom_line() +
  labs(x = "Date", y = "% Upstream Connected", 
       title = "Konza")

shnUpstreamNodes |>
  ggplot(aes(x = date, y = percentConnected)) +
  geom_line() +
  labs(x = "Date", y = "% Upstream Connected", 
       title = "Shane")

ymrUpstreamNodes |>
  ggplot(aes(x = date, y = percentConnected)) +
  geom_line() +
  labs(x = "Date", y = "% Upstream Connected", 
       title = "Youngmeyer")
```

```{r}
upstreamNodes <- kmzUpstreamNodes |> 
  mutate(siteId = "KNZ") |> 
  full_join(shnUpstreamNodes |> mutate(siteId = "SHN")) |> 
  full_join(ymrUpstreamNodes |>  mutate(siteId = "YMR")) |> 
  group_by(siteId) |> 
  mutate(dPercCon = percentConnected - lag(percentConnected)) |> 
  ungroup()

upstreamNodes |> 
  select(-data) |> 
  rename(Date = date) |> 
  write_csv(here("Data/sticUpstream.csv"))

upstreamNodes |>
 # mutate(upstream_length = ifelse(is.infinite(upstream_length), NA, upstream_length)) |> 
 # group_by(siteId) |>
 # mutate(across(where(is.numeric), ~ scale(.x) |> as.vector())) |> 
  ggplot(aes(x = date, y = upstream_length, color = siteId, group = siteId)) +
  geom_line() +
  scale_color_manual(values = okabeIto) +  
  labs(title = "",
       x = "Date",
       y = "Upstream Connected Length (m)",
       color = "") +
  theme_few() +
  theme(legend.position = "bottom")


upstreamNodes |>
 # mutate(upstream_length = ifelse(is.infinite(upstream_length), NA, upstream_length)) |> 
 # group_by(siteId) |>
 # mutate(across(where(is.numeric), ~ scale(.x) |> as.vector())) |> 
  ggplot(aes(x = date, y = percentConnected, color = siteId, group = siteId)) +
  geom_line() +
  scale_color_manual(values = okabeIto) +  
  labs(title = "",
       x = "Date",
       y = "% Upstream Connected",
       color = "") +
  theme_few() +
  theme(legend.position = "bottom")

upstreamNodes |>
 # mutate(upstream_length = ifelse(is.infinite(upstream_length), NA, upstream_length)) |> 
 # group_by(siteId) |>
 # mutate(across(where(is.numeric), ~ scale(.x) |> as.vector())) |> 
  ggplot(aes(x = date, y = dPercCon, color = siteId, group = siteId)) +
  geom_line() +
  lims(y = c(-50,50))+
  scale_color_manual(values = okabeIto) +  
  labs(title = "",
       x = "Date",
       y = "\u394 % Upstream Connected",
       color = "") +
  theme_few() +
  theme(legend.position = "bottom")+
  facet_wrap(~siteId)
```

\
