---
title: "STICs"
author: "CLB"
format: html
editor: visual
---

```{r}
#| include: false
#| echo: false

library(tidyverse)
library(tidygraph)
library(data.table)
library(ggthemes)
library(here)
library(sf)
library(streamDAG)
library(furrr)
library(future)
library(riverdist)
library(sfnetworks)
library(gganimate)
source(here("Code/paths+packages.R"))
library(mapview)
library(cli)
library(whitebox)

# EPSG for NAD83 / UTM zone 14N
UTM_14N <- 26914
source(here("Code/AIMS_DAGs.R"))
source(here("Code/Functions/aims.R"))
source(here("Code/Functions/stic_StreamDAG.R"))

theme_set(
  theme_few() +
  theme(text = element_text(size = 20),          # overall text size
    legend.text = element_text(size = 20),   # legend entries
    legend.title = element_text(size = 20),  # legend title
    axis.title = element_text(size = 20),    # axis titles
    axis.text = element_text(size = 20),     # axis tick labels
    plot.title = element_text(size = 20, hjust = 0.5),
    legend.position = "bottom")
)
```

# Streams and Sites

```{r}
# knzS <- st_read("~/Desktop/AIMS_GreatPlains.gpkg", layer = "Streams") |> 
#   st_transform(crs = 4326)

gpStreams <- st_read("~/Dropbox/AIMS/GIS/Results/AIMS_GreatPlains.gpkg", layer = "Streams", quiet = T) |> 
  filter(watershed %in% c("KNZ", "SHN", "YMR")) |> 
  st_transform(crs = 4326)
gpSheds <- st_read("~/Dropbox/AIMS/GIS/Results/AIMS_GreatPlains.gpkg", layer = "Watersheds", quiet = T) |> 
  filter(watershed %in% c("KNZ", "SHN", "YMR")) |> 
  st_transform(crs = 4326)
# Load sites
gpSites <- st_read("~/Dropbox/AIMS/GIS/Results/AIMS_GreatPlains.gpkg", layer = "Sites") |> 
  filter(watershed %in% c("KNZ", "SHN", "YMR")) |> 
  st_transform(crs = 4326) |> 
  mutate(
    lon = st_coordinates(geom)[, "X"],
    lat  = st_coordinates(geom)[, "Y"]
  ) 

seSites <- st_read("~/Dropbox/AIMS/GIS/Results/AIMS_SouthEast.gpkg", layer = "Sites") |> 
  filter(!sublocation == "HW") |> 
  st_transform(4326) |>
  mutate(
    lon = st_coordinates(geom)[, 1],
    lat = st_coordinates(geom)[, 2]
  ) 

seStreams <- st_read("~/Dropbox/AIMS/GIS/Results/AIMS_SouthEast.gpkg", layer = "Streams") |> 
  st_transform(4326)
seSheds <- st_read("~/Dropbox/AIMS/GIS/Results/AIMS_SouthEast.gpkg", layer = "Watersheds") |> 
  st_transform(4326)

mwSites <- st_read("~/Dropbox/AIMS/GIS/Results/AIMS_MountainWest.gpkg", layer = "Sites") |> 
  st_transform(4326) |> 
  mutate(
    lon = st_coordinates(geom)[, 1],
    lat = st_coordinates(geom)[, 2]
  ) 
mwStreams <- st_read("~/Dropbox/AIMS/GIS/Results/AIMS_MountainWest.gpkg", layer = "Streams") |> 
  st_transform(4326)
mwSheds <- st_read("~/Dropbox/AIMS/GIS/Results/AIMS_MountainWest.gpkg", layer = "Watersheds") |> 
  st_transform(4326)

allSheds <- bind_rows(gpSheds, mwSheds, seSheds)
```

```{r}
#### Get common cols ####
common_cols <- c("siteId", "slope", "twi", "drainage_area_sqkm", "outlet_dist_m", "elev_m", "watershed", "lon", "lat")

# SE
seSites <- seSites |>
  mutate(
    elev_m = elevation,
    slope = slope_point,
    outlet_dist_m = distance_from_outlet,
    drainage_area_sqkm = drainage_area_m / 1000000  # Convert m² to km²
  ) |>
  # Select only the columns we want to keep
  select(all_of(common_cols))
  

# Merge GP and MW, all have same cols, then drop and merge with SE
allSites <- gpSites |>
  st_drop_geometry() |> 
  full_join(mwSites |> st_drop_geometry()) |> 
  select(all_of(common_cols)) |> 
  full_join(seSites |> st_drop_geometry()) |> 
  st_as_sf(coords = c("lon", "lat"), crs = 4326)


# Combine streams
allStreams <- list(
  seStreams |> select(watershed),
  gpStreams |> select(watershed),
  mwStreams |> select(watershed)) |> 
  bind_rows()

allStreams |> 
  st_transform(5070) |> #NAD83 / Conus Albers
  mutate(length_m = st_length(geom)) |>
  group_by(watershed) |>
  summarize(total_length_m = sum(length_m)) |>
  mutate(total_length_km = as.numeric(total_length_m) / 1000) |>
  st_drop_geometry() #|> 
  #write_csv("~/Desktop/AIMS_StreamLengths.csv")
```

# STICs

```{r}
gpPresAbs <- read_csv(here("Data/STICs/GP_DailyPresAbs.csv"))
sePresAbs <- read_csv(here("Data/STICs/SE_DailyPresAbs.csv"))

presAbs <- bind_rows(
  gpPresAbs,
  sePresAbs)  |> 
  arrange(watershed, date)

suppressWarnings(rm(gpSites, gpStreams, common_cols, seSites, seStreams, gpPresAbs, sePresAbs, gpGeopkg, UTM_14N))
```

## MW STICs

```{r}

mwStics <- list.files(here("Data/MW/STICs"), full.names = TRUE, recursive = TRUE) |>
  # Only keep files in folders that end with _combined
  lapply(\(file) {
    read_csv(file, col_types = cols(.default = col_character()))
  }) |> 
  bind_rows() |>
  rename(dateTime = datetime) |>
  filter(qual_rating != "poor") |>
  select(dateTime, siteId = SiteID, wetdry) |>
  mutate(
    wetdry = case_when(
    wetdry %in% c("dry", "0") ~ 0,
    wetdry %in% c("wet", "1") ~ 1,
    TRUE ~ NA_real_),
    watershed = case_when(
    startsWith(siteId, "D") ~ "DRC",
    startsWith(siteId, "G") ~ "GBJ",
    startsWith(siteId, "J") ~ "JDR",
    TRUE ~ NA_character_  
  )) |> 
  mutate(dateTime = parse_date_time(dateTime, orders = c("ymd HM", "ymd HMS"), tz = "UTC")) |> 
  drop_na(dateTime)

daily_presAbsMW <- mwStics |>
  mutate(
    date = as_date(dateTime)
  ) |>
  filter(!is.na(watershed)) |>
  group_by(watershed, siteId, date) |>
  summarise(wetdry = calc_mode(wetdry), .groups = "drop") |>
  split(~watershed, drop = T) |> 
  lapply(\(df) {
    df |>
      select(date, siteId, wetdry) |>
      pivot_wider(
        names_from = siteId,
        values_from = wetdry
      )
  }) 
daily_presAbsMW <- bind_rows(daily_presAbsMW, .id = "watershed") |> 
  drop_na(date) |> 
  mutate(GJ8 = coalesce(GJ8, GJ8b)) |> 
  select(-GJ8b) |> 
  write_csv(here("Data/STICs/MW_DailyPresAbs.csv"))


presAbs <- presAbs |> 
  bind_rows(daily_presAbsMW) |> 
  arrange(watershed, date)

avgPercWet <- presAbs |>
  pivot_longer(
    cols = -c(watershed, date),
    names_to = "siteId",
    values_to = "wet"
  ) |>
  group_by(watershed, siteId) |>
  summarise(
    avg_pct_wet = mean(wet == 1, na.rm = TRUE) * 100,
    avg_pct_dry = mean(wet == 0, na.rm = TRUE) * 100,
    total = avg_pct_wet + avg_pct_dry,
    .groups = "drop"
  )
  
allSites <- allSites |> 
  left_join(avgPercWet |> select(watershed, siteId, avg_pct_wet), by = c("watershed", "siteId"))

n_stics <- allSites |>
  st_drop_geometry() |> 
  distinct(siteId, watershed) |> 
  group_by(watershed) |>
  summarise(n_sites = n()) |> 
  write_csv(here("Data/n_stics.csv"))
# allSites <- allSites |> 
#   filter(!(watershed %in% c("DRC")))
# 
# allStreams <- allStreams |> 
#   filter(!(watershed %in% c("DRC")))
# 
# presAbs <- presAbs |> 
#   filter(!(watershed %in% c("DRC")))
```

# Map

```{r}
aims_map <- mapview(allSheds,
  layer.name = "Watersheds",
  color = "#444444",
  col.regions = "#91D1C280",
  lwd = 2,
  alpha.regions = 0.4,
  legend = FALSE,
  zcol = "watershed",
  map.types = "Esri.WorldImagery",
  popup = leafpop::popupTable(allSheds, zcol = c("watershed", "area_sqkm"), row.numbers = FALSE, feature.id = FALSE)
) +
  mapview(allStreams,
    layer.name = "Streams",
    legend = FALSE,
    lwd = 3,
    color = "#1F305E"
  ) +
  mapview(allSites,
    layer.name = "Sites",
    zcol = "siteId",
    legend = FALSE,
    col.regions = "#AB4433",
    popup = leafpop::popupTable(allSites, zcol = c("siteId", "avg_pct_wet", "elev_m", "drainage_area_sqkm", "outlet_dist_m"), row.numbers = FALSE, feature.id = FALSE)
  )

aims_map
mapshot(aims_map, url = "~/Dropbox/AIMS/GIS/aims_map.html", title = "AIMS Map")
```

# Stream Distance

```{r}
#| include: false
# Function to compute arc lengths for a single watershed with UTM zone handling
get_watershed_lengths <- function(watershed_name, sites, streams) {
  # Subset data
  ws_sites <- sites |> filter(watershed == watershed_name) 
  ws_streams <- streams |> filter(watershed == watershed_name) 
     
  # Skip if fewer than 2 sites or no streams
  if (nrow(ws_sites) < 2 || nrow(ws_streams) < 1) return(NULL)
  
  # Determine UTM zone EPSG
  utm_epsg <- case_when(
    watershed_name %in% c("SHN", "KNZ", "YMR") ~ 26914, # NAD83/UTM zone 14N
    watershed_name %in% c("PRF", "TAL", "WHR") ~ 26916, # NAD83/UTM zone 16N
    watershed_name %in% c("GBJ") ~ 26912, # NAD83/UTM zone 12N
    watershed_name %in% c("JDR", "DRC") ~ 26911, # NAD83/UTM zone 11N
    TRUE ~ NA_integer_
  )
  
  if (is.na(utm_epsg)) {
    warning(paste("No UTM zone defined for watershed:", watershed_name))
    return(NULL)
  }
  
  # Transform to appropriate UTM zone if not already
  if (st_crs(ws_sites)$epsg != utm_epsg) {
    ws_sites <- ws_sites |> st_transform(crs = utm_epsg)
  }
  
  if (st_crs(ws_streams)$epsg != utm_epsg) {
    ws_streams <- ws_streams |> st_transform(crs = utm_epsg)
  }
  

  # Build network
  net <- as_sfnetwork(ws_streams, directed = F) |>
    st_network_blend(ws_sites, tolerance = 50) |>
    activate("edges") |> 
    mutate(weight = edge_length())
  
  # Get cost matrix
  cost_mat <- net |>
    st_network_cost(from = ws_sites, to = ws_sites, direction = "out", weights = "weight") |> 
    units::drop_units()
  
  rownames(cost_mat) <- ws_sites$siteId
  colnames(cost_mat) <- ws_sites$siteId
  
  # Format into long tibble
  as.data.frame(as.table(cost_mat)) |>
    rename(from = Var1, to = Var2, Length = Freq) |>
    filter(is.finite(Length), Length > 0) |>
    mutate(
      Arc = paste(from, to, sep = "->"),
      Length = as.numeric(Length),
      watershed = watershed_name,
      UTM_EPSG = utm_epsg
    ) |>
    select(watershed, Arc, Length) |>
    as_tibble()
}

# Get list of unique watersheds
watersheds <- intersect(unique(allSites$watershed), unique(allStreams$watershed))

# Apply function across watersheds
all_arc_lengths <- watersheds |> 
  set_names() |>  # Name list by watershed
  map(\(w) get_watershed_lengths(w, allSites, allStreams)) |> 
  compact()  # Drop NULLs

stream_lengths <- all_arc_lengths |>
  lapply(\(df) df |>
    separate(Arc, into = c("from", "to"), sep = "->", remove = TRUE)
  )
write_rds(stream_lengths, here("Data/all_stream_lengths.rds"))
```

```{r}
#| include: false


# Clean stic data
nestPresAbs <- presAbs |> 
  nest(data = -watershed) |> 
  mutate(
    data = map2(data, watershed, ~ {
      # Trim leading/trailing NA rows
      trimmed <- trim_na_rows(.x)
      # Then drop columns that are all NA (except date)
      cleaned <- select(trimmed, date, where(~ !all(is.na(.))))
      # Add watershed column to the cleaned data
      cleaned$watershed <- .y
      # Process STICs and automatically print the summary
      result <- stics_not_na(presAbs = cleaned, percThresh = 0.5, watershed = .y)
      # Ensure watershed column is preserved in the result
      if (!"watershed" %in% names(result)) {
        result$watershed <- .y
      }
      return(result)
    })
  )

doParallel::registerDoParallel(cores = nCores)

# Gap fill 
nestPresAbs_filled <- nestPresAbs |> 
  mutate(
    data = map(data, ~ {
      # Remove watershed column if it exists
      if ("watershed" %in% names(.)) {
        . <- select(., -watershed)
      }
      STIC.RFimpute_parallel(.)
    })
  )

# Remove dropped STICs from network and reconnect
aimsNets <- imap(aimsNets_full, \(net, wshed) {
  # Start watershed processing
  cli_h1("Processing watershed: {.field {wshed}}")
  # Extract watershed data
  wshed_data <- nestPresAbs_filled |> 
    filter(watershed == wshed) |> 
    pull(data) |> 
    first()
  # Get nodes to keep (excluding date)
  keep_nodes <- names(wshed_data)[names(wshed_data) != "date"]
  original_nodes <- igraph::V(net)$name
  # 1. Check for missing expected nodes
  missing_in_net <- setdiff(keep_nodes, original_nodes)
  if (length(missing_in_net) > 0) {
    cli_alert_warning(paste(
      "{.val {length(missing_in_net)}} nodes in data but not in network:",
      "{.val {missing_in_net}}"
    ))
  }
  # 2. Check for extra nodes in network
  extra_in_net <- setdiff(original_nodes, keep_nodes)
  cli_ul()
  cli_li("Nodes being removed: {.val {length(extra_in_net)}}")
  # 3. Actual nodes that will be kept
  actual_keep <- intersect(keep_nodes, original_nodes)
  cli_li("Nodes being kept: {.val {length(actual_keep)}}/{.val {length(keep_nodes)}}")
  # 4. Edge counts
  cli_li("Original edges: {.val {igraph::ecount(net)}}")
  # Process network
  filtered_net <- reconnect_and_remove_nodes(net, actual_keep)
  cli_li("Filtered edges: {.val {igraph::ecount(filtered_net)}}")
  cli_end()
  # 5. Verify final node set
  final_nodes <- igraph::V(filtered_net)$name
  if (!setequal(final_nodes, actual_keep)) {
    cli_alert_danger("Critical error: Final nodes don't match expected set")
    stop("Node mismatch in watershed ", wshed)
  }
  cli_alert_success("Successfully processed {.field {wshed}}")
  return(filtered_net)
})

scaled_arc_lengths <-list()
for (site in unique(presAbs$watershed)) {
  tempNet <- aimsNets[[site]]
  edge_keys <- attributes(E(tempNet))$vname
  arc_lengths <- all_arc_lengths[[site]]
  
  # Prepare lookup for edge weights
      arc_lengths <- arc_lengths |>
        mutate(Arc_key = gsub("->", "|", Arc)) 
      
  length_sub <- arc_lengths$Length[match(edge_keys, arc_lengths$Arc_key)]
  
  scaled_lengths <- scale(length_sub/1000, center = FALSE) #meters to km
  
  scaled_arc_lengths[[site]][["Arc"]] <- arc_lengths$Arc[match(edge_keys, arc_lengths$Arc_key)]
  scaled_arc_lengths[[site]][["scaled_lengths"]] <- scaled_lengths
  scaled_arc_lengths[[site]][["Length"]] <- length_sub
  scaled_arc_lengths[[site]][["watershed"]] <- site
}
# Convert to a nested tibble by watershed
scaled_arc_lengths <- scaled_arc_lengths |>
  imap(\(x, watershed) {
    tibble(
      Arc = x$Arc,
      Length = x$Length,
      scaled_lengths = as.numeric(x$scaled_lengths)
    ) |>
      mutate(watershed = watershed)
  }) |>
  bind_rows() |>
  group_by(watershed) |>
  nest()

all_stics <- map(aimsNets, ~ vertex_attr(.x, "name")) |> 
  unlist() |> 
  unique() |> 
  (\(x) tibble(siteId = x))()
all_stics |> write_csv(here("Data/STICs/stics_for_doRegimes.csv"))
```

## Check for dupes

```{r}
date_dups <- nestPresAbs_filled |>
  mutate(
    dup_dates = map(data, \(df) {
      df |>
        count(date) |>
        filter(n > 1)
    })
  )

# View duplicate date counts per watershed
date_dups |>
  select(watershed, dup_dates)

date_dups |>
  mutate(n_dups = map_int(dup_dates, nrow)) |>
  select(watershed, n_dups)
```

# All StreamDAG Metrics

Have to use "weight" not "weights", Maggi used weights in the KNZ code, Ken uses weight everywhere except in global metrics example

```{r}
sink_nodes <- list(
  KNZ = "SFM01",
  SHN = "SHM01",
  YMR = "ENM01",
  PRF = "PRM01",
  TAL = "TLM01",
  WHR = "WHM01",
  GBJ = "GJ1",
  JDR = "JD1",
  DRC = "DC6"
)

# Set up parallel processing
plan(multisession)

# Compute metrics per watershed with upstream metrics
streamDAG_Metrics <- nestPresAbs_filled |> 
  mutate(
    metrics = future_map2(watershed, data, \(wshed, df) {
      # Get network and sink node for this watershed
      net <- aimsNets[[wshed]]
      sink <- sink_nodes[[wshed]]
      arc_lengths <- scaled_arc_lengths |> filter(watershed == wshed) |> unnest(col = data) 
      #arc_lengths <- all_arc_lengths[[wshed]]
      
      # Get all node names from the network
      all_nodes <- V(net)$name
      total_nodes <- length(all_nodes)
      
      # Prepare lookup for edge weights
      arc_lengths <- arc_lengths |>
        mutate(Arc_key = gsub("->", "|", Arc))
      
      # Compute metrics for each day
      metric_df <- future_map(1:nrow(df), \(i) {
        tryCatch({
          # Get wet nodes (value = 1) and present nodes (non-NA)
          node_values <- df[i, -which(names(df) == "date")]
          wet_nodes_count <- sum(node_values == 1, na.rm = TRUE)
          present_nodes_count <- sum(!is.na(node_values))
          
          # Calculate percent_network_wet
          percent_network_wet <- ifelse(present_nodes_count > 0, 
                              (wet_nodes_count/present_nodes_count)*100, 
                              0)
          # Convert from factors to characters to numeric
          pa_vect <- node_values[1, ] |>
            as.character() |>
            as.numeric()
          
          # Set the names
          names(pa_vect) <- colnames(node_values)
          pa_vect <- pa_vect - 1
          # Create temporary graph for this day
          temp_graph <- delete.nodes.pa(
            net, 
            node_values,
            na.response = "treat.as.1"
          )
          
          # Skip if graph is empty
          if (vcount(temp_graph) == 0) {
            return(tibble(
              date = df$date[i],
              icsl = NA_real_,
              #intact_to_sink = NA_real_,
              a_cent_weighted = NA_real_,
              a_cent = NA_real_,
              harary = NA_real_,
              active_upstream_length = 0,
              present_nodes = present_nodes_count,
              total_nodes = total_nodes,
              #percent_upstream_wet = 0,
              percent_network_wet = percent_network_wet
            ))
          }
          
          edge_keys <- attributes(E(temp_graph))$vname
          length_scaled <- arc_lengths$scaled_lengths[match(edge_keys, arc_lengths$Arc_key)]
          lengths_m <- arc_lengths$Length[match(edge_keys, arc_lengths$Arc_key)]
          
          # Total active network length
          active_network_length <- sum(lengths_m)
          
          # Calculate upstream metrics
          if (sink %in% V(temp_graph)$name) {
            E(temp_graph)$weight <- lengths_m
            upstream_nodes <- subcomponent(temp_graph, sink, mode = "in")
            upstream_subgraph <- induced_subgraph(temp_graph, upstream_nodes)
            active_upstream_length <- size.intact.to.sink(temp_graph, sink = sink)
            #active_upstream_length <- sum(E(upstream_subgraph)$weight, na.rm = TRUE)
            wet_nodes_upstream <- vcount(upstream_subgraph)
          } else {
            #active_upstream_length <- 0
            wet_nodes_upstream <- 0
            active_upstream_length <- 0
          }
          # Set edge weights
          E(temp_graph)$weight <- length_scaled
          wet_nodes_all <- vcount(temp_graph)
          # Calculate all metrics
          tibble(
            date = df$date[i],
            icsl = ICSL(temp_graph, lengths = lengths_m),
            active_upstream_length = active_upstream_length,
            a_cent_weighted = mean(alpha_centrality(temp_graph), na.rm = TRUE),
            a_cent = mean(alpha_centrality(temp_graph, weights = NA), na.rm = TRUE),
            harary = harary(temp_graph),
            wet_upstream_nodes = wet_nodes_upstream,
            wet_network_nodes = wet_nodes_all,
            present_nodes = present_nodes_count,
            total_nodes = total_nodes,
            percent_network_wet = percent_network_wet,
            active_network_length_m = active_network_length
          )
          
        }, error = function(e) {
          message("Error for ", wshed, " on ", df$date[i], ": ", e$message)
          tibble(
            date = df$date[i],
            icsl = NA_real_,
            #intact_to_sink = NA_real_,
            a_cent_weighted = NA_real_,
            a_cent = NA_real_,
            harary = NA_real_,
            active_upstream_length = NA_real_,
            present_nodes = NA_real_,
            total_nodes = total_nodes,
            percent_network_wet = NA_real_
          )
        })
      }, .options = furrr_options(seed = TRUE), .progress = TRUE) |> 
        bind_rows()
      
      # Format final output
      metric_df |>
        pivot_longer(cols = -c(date, wet_upstream_nodes, wet_network_nodes, present_nodes, 
                             total_nodes, percent_network_wet), 
                     names_to = "metric", values_to = "value") |>
        #mutate(metric = recode(metric, !!!metric_names)) |> 
        pivot_wider(names_from = metric, values_from = value) 
    }, .options = furrr_options(seed = TRUE), .progress = TRUE)
  ) |> 
  select(-data) |> 
  unnest(metrics)


streamDAG_Metrics <- streamDAG_Metrics |> 
  group_by(watershed) |>
  mutate(
    a_cent_scaled = as.numeric(scale(a_cent)),
    a_cent_weighted_scaled = as.numeric(scale(a_cent_weighted))
  ) |>
  ungroup() |> 
  select(watershed, date, total_nodes, present_nodes, wet_upstream_nodes, wet_network_nodes, 
         percent_network_wet, active_upstream_length, active_network_length_m, icsl, harary, a_cent, a_cent_scaled, 
         a_cent_weighted, a_cent_weighted_scaled) |> 
  write_csv(here("Data/STICs/streamDAG_metrics.csv"))

streamDAG_Metrics |> 
  filter(watershed %in% c("WHR", "TAL", "PRF")) |> 
  write_csv(here("Data/STICs/streamDAG_metrics_SouthEast.csv"))
  #write_csv(here("Data/STICs/streamDAG_metrics_connectivityPaper.csv"))
# Reset parallel processing
plan(sequential)
```

```{r}
# First reshape the data to long format for plotting
streamDAG_Metrics_long <- streamDAG_Metrics |>
  pivot_longer(
    cols = -c(watershed, date),
    names_to = "Metric",
    values_to = "Value"
  ) |> 
  mutate(watershed = forcats::fct_relevel(watershed, 
                               "KNZ", "SHN", "YMR",  # Left column
                               "TAL", "WHR", "PRF", 
                               "GBJ", "DRC", "JDR")) # Right column

# Create faceted plot
streamDAG_Metrics_long |> 
  filter(Metric == "a_cent") |> 
ggplot(aes(x = date, y = Value, color = watershed)) +
  geom_line(alpha = 0.8, linewidth = 0.7) +
  facet_wrap(~watershed, ncol = 2, dir = "v", scales = "free") +
  scale_color_manual(values = okabeIto) +
  labs(
    title = "Global Unweighted Alpha Centrality",
    x = "",
    y = "Alpha Centrality",
    color = ""
  ) +
  theme_few() +
  theme(
    legend.position = "none") +
  scale_x_date(date_labels = "%b %Y") +
  scale_y_continuous(labels = scales::label_number(accuracy = 0.1)) -> p
p
ggsave("gu_ac.png", p, here("Figs/STICs"), device = "png", width = 25.4, height = 25.4, units = "cm", dpi = "retina")

streamDAG_Metrics_long |> 
  filter(Metric == "a_cent_weighted") |> 
ggplot(aes(x = date, y = Value, color = watershed)) +
  geom_line(alpha = 0.8, linewidth = 0.7) +
  facet_wrap(~watershed, ncol = 2, dir = "v", scales = "free") +
  scale_color_manual(values = okabeIto) +
  labs(
    title = "Global Weighted Alpha Centrality",
    x = "",
    y = "Alpha Centrality",
    color = ""
  ) +
  theme_few() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "none",
    panel.grid.minor = element_blank()
  ) +
  scale_x_date(date_labels = "%b %Y") +
  scale_y_continuous(labels = scales::label_number(accuracy = 0.1)) ->p 
p
ggsave("gw_ac.png", p, here("Figs/STICs"), device = "png", width = 25.4, height = 25.4, units = "cm", dpi = "retina")

streamDAG_Metrics_long |> 
  filter(Metric == "a_cent_scaled") |> 
ggplot(aes(x = date, y = Value, color = watershed)) +
  geom_line(alpha = 0.8, linewidth = 0.7) +
  facet_wrap(~watershed, ncol = 2, dir = "v", scales = "free") +
  scale_color_manual(values = okabeIto) +
  labs(
    title = "Global Unweighted Alpha Centrality (Scaled)",
    x = "",
    y = "Alpha Centrality",
    color = ""
  ) +
  theme_few() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "none",
    panel.grid.minor = element_blank()
  ) +
  scale_x_date(date_labels = "%b %Y") +
  scale_y_continuous(labels = scales::label_number(accuracy = 0.1)) -> p
p
ggsave("gus_ac.png", p, here("Figs/STICs"), device = "png", width = 25.4, height = 25.4, units = "cm", dpi = "retina")

streamDAG_Metrics_long |> 
  filter(Metric == "a_cent_weighted_scaled") |> 
ggplot(aes(x = date, y = Value, color = watershed)) +
  geom_line(alpha = 0.8, linewidth = 0.7) +
  facet_wrap(~watershed, ncol = 2, dir = "v", scales = "free") +
  scale_color_manual(values = okabeIto) +
  labs(
    title = "Global Weighted Alpha Centrality (Scaled)",
    x = "",
    y = "Alpha Centrality",
    color = ""
  ) +
  theme_few() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "none",
    panel.grid.minor = element_blank()
  ) +
  scale_x_date(date_labels = "%b %Y") +
  scale_y_continuous(labels = scales::label_number(accuracy = 0.1)) -> p
p
ggsave("gws_ac.png", p, here("Figs/STICs"), device = "png", width = 25.4, height = 25.4, units = "cm", dpi = "retina")

streamDAG_Metrics_long |> 
  filter(Metric %in% c("icsl", "a_cent_weighted", "a_cent", "harary")) |> 
  # group_by(watershed, Metric) |>
  #   dplyr::mutate(across(where(is.numeric), ~ (. - mean(., na.rm = TRUE)) / sd(., na.rm = TRUE), .names = "{.col}")) |>
  #   ungroup() |>
ggplot(aes(x = date, y = Value, color = Metric)) +
  geom_line(alpha = 0.8, linewidth = 0.7) +
  facet_grid(Metric~watershed, scales = "free") +
  scale_color_manual(values = okabeIto) +
  labs(
    x = "",
    y = "",
    color = ""
  ) +
  theme_few() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  ) +
  scale_x_date(date_labels = "%b %Y") +
  scale_y_continuous(labels = scales::label_number(accuracy = 0.1)) 

streamDAG_Metrics_long |> 
  filter(Metric %in% c("present_nodes", "wet_nodes", "total_nodes", "active_upstream_length", "active_network_length_m", "percent_upstream_wet", "percent_network_wet")) |> 
ggplot(aes(x = date, y = Value, color = Metric)) +
  geom_line(alpha = 0.8, linewidth = 0.7) +
  facet_grid(Metric~watershed, scales = "free") +
  scale_color_manual(values = okabeIto) +
  labs(
    x = "",
    y = "",
    color = ""
  ) +
  theme_few() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  ) +
  scale_x_date(date_labels = "%b %Y") +
  scale_y_continuous(labels = scales::comma) -> p
p
ggsave("Nodes.png", p, here("Figs/STICs"), device = "png", width = 25.4, height = 25.4, units = "cm", dpi = "retina")
```

# StreamDAG Metrics - Connectivity Paper

```{r}
# Define the date ranges for each watershed
connectivity_dates <- data.frame(
  watershed = c("KNZ", "TAL", "GBJ", "PRF", "YMR", "JDR"),
  start_date = as.Date(c("2022-06-22", "2023-02-01", "2022-04-22", "2023-03-09", "2022-03-25", "2023-03-22")),
  end_date = as.Date(c("2022-09-01", "2023-12-31", "2022-09-07", "2023-06-12", "2022-08-04", "2023-07-17"))
) |> 
  mutate(start_wy = calcWaterYear(start_date),
         end_wy = calcWaterYear(end_date),
         Year = year(start_date))

# Filter the data
filt_presAbs <- presAbs  |> 
  left_join(connectivity_dates, by = "watershed") |> 
  filter(year(date) == Year) |> 
  select(-start_date, -end_date, -start_wy, -end_wy, -Year)

# Clean stic data
nestPresAbs_filt <- filt_presAbs |> 
  nest(data = -watershed) |> 
  mutate(
    data = map2(data, watershed, ~ {
      # Trim leading/trailing NA rows
      trimmed <- trim_na_rows(.x)
      # Then drop columns that are all NA (except date)
      cleaned <- select(trimmed, date, where(~ !all(is.na(.))))
      # Add watershed column to the cleaned data
      cleaned$watershed <- .y
      # Process STICs and automatically print the summary
      result <- stics_not_na(presAbs = cleaned, percThresh = 0.5, watershed = .y)
      # Ensure watershed column is preserved in the result
      if (!"watershed" %in% names(result)) {
        result$watershed <- .y
      }
      return(result)
    })
  )

doParallel::registerDoParallel(cores = nCores)

# Gap fill 
nestPresAbs_filt_filled <- nestPresAbs_filt |> 
  mutate(
    data = map(data, ~ {
      # Remove watershed column if it exists
      if ("watershed" %in% names(.)) {
        . <- select(., -watershed)
      }
      STIC.RFimpute_parallel(.)
    })
  )

# Remove dropped STICs from network and reconnect
aimsNets <- imap(aimsNets_full, \(net, wshed) {
  # Start watershed processing
  cli_h1("Processing watershed: {.field {wshed}}")
  # Extract watershed data
  wshed_data <- nestPresAbs_filt_filled |> 
    filter(watershed == wshed) |> 
    pull(data) |> 
    first()
  # Get nodes to keep (excluding date)
  keep_nodes <- names(wshed_data)[names(wshed_data) != "date"]
  original_nodes <- igraph::V(net)$name
  # 1. Check for missing expected nodes
  missing_in_net <- setdiff(keep_nodes, original_nodes)
  if (length(missing_in_net) > 0) {
    cli_alert_warning(paste(
      "{.val {length(missing_in_net)}} nodes in data but not in network:",
      "{.val {missing_in_net}}"
    ))
  }
  # 2. Check for extra nodes in network
  extra_in_net <- setdiff(original_nodes, keep_nodes)
  cli_ul()
  cli_li("Nodes being removed: {.val {length(extra_in_net)}}")
  # 3. Actual nodes that will be kept
  actual_keep <- intersect(keep_nodes, original_nodes)
  cli_li("Nodes being kept: {.val {length(actual_keep)}}/{.val {length(keep_nodes)}}")
  # 4. Edge counts
  cli_li("Original edges: {.val {igraph::ecount(net)}}")
  # Process network
  filtered_net <- reconnect_and_remove_nodes(net, actual_keep)
  cli_li("Filtered edges: {.val {igraph::ecount(filtered_net)}}")
  cli_end()
  # 5. Verify final node set
  final_nodes <- igraph::V(filtered_net)$name
  if (!setequal(final_nodes, actual_keep)) {
    cli_alert_danger("Critical error: Final nodes don't match expected set")
    stop("Node mismatch in watershed ", wshed)
  }
  cli_alert_success("Successfully processed {.field {wshed}}")
  return(filtered_net)
})

scaled_arc_lengths <-list()
for (site in unique(presAbs$watershed)) {
  tempNet <- aimsNets[[site]]
  edge_keys <- attributes(E(tempNet))$vname
  arc_lengths <- all_arc_lengths[[site]]
  
  # Prepare lookup for edge weights
      arc_lengths <- arc_lengths |>
        mutate(Arc_key = gsub("->", "|", Arc)) 
      
  length_sub <- arc_lengths$Length[match(edge_keys, arc_lengths$Arc_key)]
  
  scaled_lengths <- scale(length_sub/1000, center = FALSE) #meters to km
  
  scaled_arc_lengths[[site]][["Arc"]] <- arc_lengths$Arc[match(edge_keys, arc_lengths$Arc_key)]
  scaled_arc_lengths[[site]][["scaled_lengths"]] <- scaled_lengths
  scaled_arc_lengths[[site]][["Length"]] <- length_sub
  scaled_arc_lengths[[site]][["watershed"]] <- site
}
# Convert to a nested tibble by watershed
scaled_arc_lengths <- scaled_arc_lengths |>
  imap(\(x, watershed) {
    tibble(
      Arc = x$Arc,
      Length = x$Length,
      scaled_lengths = as.numeric(x$scaled_lengths)
    ) |>
      mutate(watershed = watershed)
  }) |>
  bind_rows() |>
  group_by(watershed) |>
  nest()

all_stics_connectivity <- map(aimsNets, ~ vertex_attr(.x, "name")) |> 
  unlist() |> 
  unique() |> 
  (\(x) tibble(siteId = x))()
all_stics_connectivity |> write_csv(here("Data/STICs/stics_for_connectivity.csv"))
```

```{r}
date_dups <- nestPresAbs_filt_filled |>
  mutate(
    dup_dates = map(data, \(df) {
      df |>
        count(date) |>
        filter(n > 1)
    })
  )

# View duplicate date counts per watershed
date_dups |>
  select(watershed, dup_dates)

date_dups |>
  mutate(n_dups = map_int(dup_dates, nrow)) |>
  select(watershed, n_dups)
```

```{r}
sink_nodes <- list(
  KNZ = "SFM01",
  SHN = "SHM01",
  YMR = "ENM01",
  PRF = "PRM01",
  TAL = "TLM01",
  WHR = "WHM01",
  GBJ = "GJ1",
  JDR = "JD1",
  DRC = "DC6"
)
# Set up parallel processing
plan(multisession)

# Compute metrics per watershed with upstream metrics
streamDAG_Metrics_Connectivity <- nestPresAbs_filt_filled |> 
  mutate(
    metrics = future_map2(watershed, data, \(wshed, df) {
      # Get network and sink node for this watershed
      net <- aimsNets[[wshed]]
      sink <- sink_nodes[[wshed]]
      arc_lengths <- scaled_arc_lengths |> filter(watershed == wshed) |> unnest(col = data) 
      #arc_lengths <- all_arc_lengths[[wshed]]
      
      # Get all node names from the network
      all_nodes <- V(net)$name
      total_nodes <- length(all_nodes)
      
      # Prepare lookup for edge weights
      arc_lengths <- arc_lengths |>
        mutate(Arc_key = gsub("->", "|", Arc))
      
      # Compute metrics for each day
      metric_df <- future_map(1:nrow(df), \(i) {
        tryCatch({
          # Get wet nodes (value = 1) and present nodes (non-NA)
          node_values <- df[i, -which(names(df) == "date")]
          wet_nodes_count <- sum(node_values == 1, na.rm = TRUE)
          present_nodes_count <- sum(!is.na(node_values))
          
          # Calculate percent_network_wet
          percent_network_wet <- ifelse(present_nodes_count > 0, 
                              (wet_nodes_count/present_nodes_count)*100, 
                              0)
          # Convert from factors to characters to numeric
          pa_vect <- node_values[1, ] |>
            as.character() |>
            as.numeric()
          
          # Set the names
          names(pa_vect) <- colnames(node_values)
          pa_vect <- pa_vect - 1
          # Create temporary graph for this day
          temp_graph <- delete.nodes.pa(
            net, 
            node_values,
            na.response = "treat.as.1"
          )
          
          # Skip if graph is empty
          if (vcount(temp_graph) == 0) {
            return(tibble(
              date = df$date[i],
              icsl = NA_real_,
              #intact_to_sink = NA_real_,
              a_cent_weighted = NA_real_,
              a_cent = NA_real_,
              harary = NA_real_,
              active_upstream_length = 0,
              present_nodes = present_nodes_count,
              total_nodes = total_nodes,
              #percent_upstream_wet = 0,
              percent_network_wet = percent_network_wet
            ))
          }
          
          edge_keys <- attributes(E(temp_graph))$vname
          length_scaled <- arc_lengths$scaled_lengths[match(edge_keys, arc_lengths$Arc_key)]
          lengths_m <- arc_lengths$Length[match(edge_keys, arc_lengths$Arc_key)]
          
          # Total active network length
          active_network_length <- sum(lengths_m)
          
          # Calculate upstream metrics
          if (sink %in% V(temp_graph)$name) {
            E(temp_graph)$weight <- lengths_m
            upstream_nodes <- subcomponent(temp_graph, sink, mode = "in")
            upstream_subgraph <- induced_subgraph(temp_graph, upstream_nodes)
            active_upstream_length <- size.intact.to.sink(temp_graph, sink = sink)
            #active_upstream_length <- sum(E(upstream_subgraph)$weight, na.rm = TRUE)
            wet_nodes_upstream <- vcount(upstream_subgraph)
          } else {
            #active_upstream_length <- 0
            wet_nodes_upstream <- 0
            active_upstream_length <- 0
          }
          # Set edge weights
          E(temp_graph)$weight <- length_scaled
          wet_nodes_all <- vcount(temp_graph)
          # Calculate all metrics
          tibble(
            date = df$date[i],
            icsl = ICSL(temp_graph, lengths = lengths_m),
            active_upstream_length = active_upstream_length,
            a_cent_weighted = mean(alpha_centrality(temp_graph), na.rm = TRUE),
            a_cent = mean(alpha_centrality(temp_graph, weights = NA), na.rm = TRUE),
            harary = harary(temp_graph),
            wet_upstream_nodes = wet_nodes_upstream,
            wet_network_nodes = wet_nodes_all,
            present_nodes = present_nodes_count,
            total_nodes = total_nodes,
            percent_network_wet = percent_network_wet,
            active_network_length_m = active_network_length
          )
          
        }, error = function(e) {
          message("Error for ", wshed, " on ", df$date[i], ": ", e$message)
          tibble(
            date = df$date[i],
            icsl = NA_real_,
            #intact_to_sink = NA_real_,
            a_cent_weighted = NA_real_,
            a_cent = NA_real_,
            harary = NA_real_,
            active_upstream_length = NA_real_,
            present_nodes = NA_real_,
            total_nodes = total_nodes,
            percent_network_wet = NA_real_
          )
        })
      }, .options = furrr_options(seed = TRUE), .progress = TRUE) |> 
        bind_rows()
      
      # Format final output
      metric_df |>
        pivot_longer(cols = -c(date, wet_upstream_nodes, wet_network_nodes, present_nodes, 
                             total_nodes, percent_network_wet), 
                     names_to = "metric", values_to = "value") |>
        #mutate(metric = recode(metric, !!!metric_names)) |> 
        pivot_wider(names_from = metric, values_from = value) 
    }, .options = furrr_options(seed = TRUE), .progress = TRUE)
  ) |> 
  select(-data) |> 
  unnest(metrics)

streamDAG_Metrics_Connectivity <- streamDAG_Metrics_Connectivity |> 
  group_by(watershed) |>
  mutate(
    a_cent_scaled = as.numeric(scale(a_cent)),
    a_cent_weighted_scaled = as.numeric(scale(a_cent_weighted))
  ) |>
  ungroup() |> 
  select(watershed, date, total_nodes, present_nodes, wet_upstream_nodes, wet_network_nodes, 
         percent_network_wet, active_upstream_length, icsl, harary, a_cent, a_cent_scaled, 
         a_cent_weighted, a_cent_weighted_scaled) |> 
  write_csv(here("Data/STICs/streamDAG_metrics_connectivityPaper.csv"))
# Reset parallel processing
plan(sequential)
```

# Graveyard

## TidyNetworks

```{r}
# reconnect_and_remove_nodes_tidy <- function(graph, keep_nodes) {
#   g <- as.igraph(graph)
# 
#   # Set node names to siteId before manipulating
#   site_ids <- graph |> activate("nodes") |> pull(siteId)
#   V(g)$name <- site_ids
# 
#   remove_nodes <- setdiff(V(g)$name, keep_nodes)
# 
#   for (v in remove_nodes) {
#     if (!v %in% V(g)$name) next
# 
#     upstream <- igraph::neighbors(g, v, mode = "in")$name
#     downstream <- igraph::neighbors(g, v, mode = "out")$name
# 
#     if (length(upstream) > 0 && length(downstream) > 0) {
#       new_edges <- expand.grid(from = upstream, to = downstream, stringsAsFactors = FALSE)
#       for (i in seq_len(nrow(new_edges))) {
#         from <- new_edges$from[i]
#         to <- new_edges$to[i]
#         if (from != to && !are_adjacent(g, from, to)) {
#           g <- add_edges(g, c(from, to))
#         }
#       }
#     }
# 
#     g <- delete_vertices(g, v)
#   }
#   graph_out <- as_sfnetwork(g)
#   return(graph_out)
# }

# sink_nodes <- list(
#   KNZ = "SFM01",
#   SHN = "SHM01",
#   YMR = "ENM01",
#   PRF = "PRM01",
#   TAL = "TLM01",
#   WHR = "WHM01",
#   GBJ = "GJ1"
# )
# # Step 1: Build the input tibble
# network_inputs <- aims_edges |>
#   mutate(
#     crs = case_when(
#       watershed %in% c("KNZ", "SHN", "YMR") ~ 26914,
#       watershed %in% c("PRF", "TAL", "WHR") ~ 26916,
#       watershed %in% c("GBJ") ~ 26912,
#       watershed %in% c("JDR", "DRC") ~ 26911,
#     ),
#     sink_node = map_chr(watershed, \(ws) sink_nodes[[ws]]),
#     sites = map2(
#       watershed,
#       crs,
#       ~ allSites |> filter(watershed == .x) |> st_transform(.y)
#     ),
#     streams = map2(
#       watershed,
#       crs,
#       ~ allStreams |> filter(watershed == .x) |> st_transform(.y)
#     ),
#     lengths = map(
#   watershed,
#   ~ {
#     arc_df <- all_arc_lengths[[.x]]
#     if (is.null(arc_df)) return(NULL)
# 
#     arc_df |>
#       select(-watershed) |>
#       separate(Arc, into = c("from", "to"), sep = "->") |>
#       mutate(across(c(from, to), trimws), Length_km = Length / 1000) |>
#       select(-Length)
#   }
#   ),
#     presAbs = map(
#       watershed,
#       ~ nestPresAbs_filled |>
#         filter(watershed == .x) |>
#         unnest(cols = c(data)) |>
#         select(c("watershed", "date"), any_of(
#           allSites |> filter(watershed == .x) |> pull(siteId)
#         ))
#     ),
#     marginal_probs = map2(presAbs, edges, \(node_df, edge_df) {
#       node_df <- node_df |>
#         mutate(across(-c(watershed, date), ~ as.numeric(as.character(.))))
#       
#       node_long <- node_df |>
#         pivot_longer(-c(watershed, date),
#                      names_to = "siteId",
#                      values_to = "present")
#       
#       edge_cross <- expand_grid(date = unique(node_df$date), edge = seq_len(nrow(edge_df))) |>
#         mutate(from_site = edge_df$from[edge], to_site = edge_df$to[edge]) |>
#         select(-edge)
#       
#       arc_presence <- edge_cross |>
#         left_join(node_long, by = c("date", "from_site" = "siteId")) |>
#         rename(p_from = present) |>
#         left_join(node_long, by = c("date", "to_site" = "siteId")) |>
#         rename(p_to = present) |>
#         mutate(
#           arc_presence = case_when(
#             p_from == 1 & p_to == 1 ~ 1,
#             p_from == 0 & p_to == 0 ~ 0,
#             is.na(p_from) | is.na(p_to) ~ NA_real_,
#             TRUE ~ 0.5
#           )
#         )
#       
#       arc_presence |>
#         group_by(from_site, to_site) |>
#         summarise(prob_presence = mean(arc_presence, na.rm = TRUE),
#                   .groups = "drop") |>
#         mutate(
#           arc_id = paste(from_site, to_site, sep = "->"),
#           prob_presence = coalesce(prob_presence, min(prob_presence, na.rm = TRUE))
#         )
#     })
#   )

# plan(multisession)
# 
# jdrDat <- get.AIMS.data("jd_full")
# 
# # Full workflow per watershed
# network_results <- network_inputs |> 
#   mutate(
#     metrics = pmap(
#       list(sites, edges, lengths, presAbs, watershed, sink_node),
#       \(sites, edges, lengths, presAbs, net_name, sink_id) {
#         message(glue::glue("Processing {net_name}..."))
#         
#         net <- tryCatch({
#           # Check for bad edges
#           valid_site_ids <- sites$siteId
#           bad_edges <- edges |> filter(!(from %in% valid_site_ids) | !(to %in% valid_site_ids))
#           
#           # Warn if any bad edges found
#           if (nrow(bad_edges) > 0) {
#             message(glue::glue(
#               "Invalid edges in {net_name}: ",
#               paste0(bad_edges$from, "->", bad_edges$to, collapse = ", ")
#             ))
#           }
#           
#           # Filter out invalid edges
#           edges <- edges |> filter(from %in% valid_site_ids, to %in% valid_site_ids)
#           
#           # Build the sfnetwork
#           sfnetwork(
#             nodes = sites,
#             edges = edges,
#             node_key = "siteId",
#             directed = TRUE
#           )
#         }, error = function(e) {
#           message("❌ Failed to create network for ", net_name, ": ", e$message)
#           if (exists("bad_edges") && nrow(bad_edges) > 0) {
#             message("Bad edges:\n", paste0(bad_edges$from, " -> ", bad_edges$to, collapse = "\n"))
#           }
#           return(NULL)
#         })
#         
#         
#         pa_nodes <- presAbs |> names() |> setdiff(c("watershed", "date"))
#         
#         pruned_net <- reconnect_and_remove_nodes_tidy(net, pa_nodes)
#         
#         node_table <- pruned_net |> 
#           activate(nodes) |> 
#           as_tibble() |> 
#           mutate(index = row_number()) |> 
#           select(index, siteId)
#         
#         edge_table <- pruned_net |> 
#           activate(edges) |> 
#           as_tibble() |> 
#           mutate(edge_id = row_number()) |> 
#           left_join(node_table, by = c("from" = "index")) |> 
#           rename(from_site = siteId) |>
#           left_join(node_table, by = c("to" = "index")) |> 
#           rename(to_site = siteId)
#         
#         marg_probs <- network_inputs |> 
#           filter(watershed == net_name) |> 
#           pull(marginal_probs) |> 
#           pluck(1)
#         
#         edge_table_weighted <- edge_table |> 
#           left_join(lengths, by = c("from_site" = "from", "to_site" = "to")) |> 
#           left_join(marg_probs, by = c("from_site", "to_site"))
#         
#         pruned_net <- pruned_net |> 
#           activate(edges) |> 
#           mutate(
#             reach_length_km = edge_table_weighted$Length_km,
#             weight = reach_length_km,
#             prob_presence = edge_table_weighted$prob_presence
#           ) |>
#           activate(nodes) |> 
#           mutate(
#             weighted_ac = centrality_alpha(weights = reach_length_km),
#             ac = centrality_alpha(weights = NA)
#           )
#         
#         long_data <- presAbs |>
#           pivot_longer(cols = -c(watershed, date), 
#                        names_to = "siteId", 
#                        values_to = "present") |>
#           filter(present == 1) |>
#           group_by(date)
#         
#         data_list <- group_split(long_data)
#         key_df <- group_keys(long_data)
#         key_list <- split(key_df, seq_len(nrow(key_df)))
#         
#         future_map2_dfr(
#           data_list,
#           key_list,
#           \(df, key) {
#             present_nodes <- df$siteId
#             
#             pruned_subgraph <- pruned_net |>
#               activate(nodes) |>
#               morph(to_subgraph, siteId %in% present_nodes, subset_by = "nodes")
#             
#             node_metrics <- pruned_subgraph |>
#               mutate(
#                 alpha_centrality = centrality_alpha(),
#                 alpha_centrality_weighted = centrality_alpha(weights = reach_length_km),
#                 arcProbAC = centrality_alpha(weights = prob_presence),
#                 upstream_length = size.intact.to.node(.G(), node = "all"),
#                 icsl = ICSL(.G()),
#                 harary = harary(.G())
#               ) |>
#               unmorph() |>
#               activate(nodes) |>
#               st_drop_geometry() |>
#               as_tibble()
#             
#             
#             node_metrics |>
#               filter(!is.na(alpha_centrality)) |>
#               mutate(
#                 date = key[[1]]
#               )
#           },
#           .options = furrr_options(seed = TRUE),
#           .progress = TRUE
#         ) |> 
#           st_drop_geometry() |> 
#           select(siteId, date, alpha_centrality_weighted, alpha_centrality, arcProbAC, upstream_length, icsl, harary)
#       }
#     )
#   )
# 
# # Step 3: Combine all watershed results
# all_metrics <- network_results |> 
#   select(watershed, metrics) |> 
#   unnest(metrics) |> 
#   left_join(
#     allSites |> 
#       st_drop_geometry() |> 
#       select(siteId, watershed, drainage_area_sqkm, outlet_dist_m),
#     by = c("siteId", "watershed")
#   ) #|> 
# #write_csv(here("Data/STICs/alpha_centrality_GP-SE.csv"))
# plan(sequential)

# 
# all_metrics |> 
#   group_by(date, watershed) |> 
#   summarise(alpha_centrality = mean(alpha_centrality, na.rm = T)) |> 
#   ggplot(aes(x = date, y = alpha_centrality, color = watershed)) +
#   geom_line(alpha = 0.8, linewidth = 0.7) +
#   scale_color_manual(values = okabeIto) +
#   theme_few() +
#   facet_wrap(~watershed) +
#   theme(
#     axis.text.x = element_text(angle = 45, hjust = 1),
#     strip.text = element_text(face = "bold", size = 10),
#     legend.position = "bottom",
#     panel.grid.minor = element_blank()
#   )
# 
# all_metrics |> 
#   group_by(date, watershed) |> 
#   summarise(alpha_centrality_weighted = mean(alpha_centrality_weighted, na.rm = T)) |> 
#   ggplot(aes(x = date, y = alpha_centrality_weighted, color = watershed)) +
#   geom_line(alpha = 0.8, linewidth = 0.7) +
#   scale_color_manual(values = okabeIto) +
#   theme_few() +
#   facet_wrap(~watershed) +
#   theme(
#     axis.text.x = element_text(angle = 45, hjust = 1),
#     strip.text = element_text(face = "bold", size = 10),
#     legend.position = "bottom",
#     panel.grid.minor = element_blank()
#   )
# 
# out_nodes <- c("SFM01", "SHM01", "ENM01", "PRM01", "TLM01", "WHM01")
# 
# # Reshape and scale
# all_metrics_long <- all_metrics |>
#   pivot_longer(
#     cols = c(alpha_centrality, alpha_centrality_weighted, upstream_length, icsl, harary),
#     names_to = "Metric",
#     values_to = "Value"
#   ) |>
#   mutate(
#     Metric = case_when(
#       Metric == "icsl" ~ "ICSL",
#       Metric == "upstream_length" ~ "UpstreamLength",
#       Metric == "alpha_centrality" ~ "AlphaCentrality",
#       Metric == "alpha_centrality_weighted" ~ "WeightedAlphaCentrality",
#       Metric == "harary" ~ "Harary"
#     )
#   ) |>
#   group_by(watershed, Metric, date) |>
#   mutate(Value = mean(Value, na.rm = T)) |>
#   ungroup()
# 
# # Plot
# ggplot(all_metrics_long, aes(x = date, y = Value, color = Metric)) +
#   geom_line(alpha = 0.8, linewidth = 0.7) +
#   facet_grid(Metric ~ watershed, scales = "free_y") +
#   scale_color_manual(values = okabeIto) +
#   labs(
#     x = "",
#     y = "",
#     color = ""
#   ) +
#   theme_few() +
#   theme(
#     axis.text.x = element_text(angle = 45, hjust = 1),
#     strip.text = element_text(face = "bold", size = 10),
#     legend.position = "bottom",
#     panel.grid.minor = element_blank()
#   ) +
#   scale_x_date(date_labels = "%b %Y") +
#   scale_y_continuous(labels = scales::label_number(accuracy = 0.1))

```

```{r}
sink_nodes <- list(
  KNZ = "SFM01",
  SHN = "SHM01",
  YMR = "ENM01",
  PRF = "PRM01",
  TAL = "TLM01",
  WHR = "WHM01",
  GBJ = "GJ1"
)

# Set up parallel processing
plan(multisession)

# Compute metrics per watershed with upstream metrics
streamDAG_Metrics <- nestPresAbs_filled |> 
  mutate(
    metrics = future_map2(watershed, data, \(wshed, df) {
      # Get network and sink node for this watershed
      net <- aimsNets[[wshed]]
      sink <- sink_nodes[[wshed]]
      arc_lengths <- all_arc_lengths[[wshed]]
      
      # Get all node names from the network
      all_nodes <- V(net)$name
      total_nodes <- length(all_nodes)
      
      # Prepare lookup for edge weights
      arc_lengths <- arc_lengths |>
        mutate(Arc_key = gsub("->", "|", Arc))
      
      # Compute metrics for each day
      metric_df <- future_map(1:nrow(df), \(i) {
        tryCatch({
          # Get wet nodes (value = 1) and present nodes (non-NA)
          node_values <- df[i, -which(names(df) == "date")]
          wet_nodes_count <- sum(node_values == 1, na.rm = TRUE)
          present_nodes_count <- sum(!is.na(node_values))
          
          # Calculate percent_network_wet
          percent_network_wet <- ifelse(present_nodes_count > 0, 
                              (wet_nodes_count/present_nodes_count)*100, 
                              0)
          # Convert from factors to characters to numeric
          pa_vect <- node_values[1, ] |>
            as.character() |>
            as.numeric()
          
          # Set the names
          names(pa_vect) <- colnames(node_values)
          pa_vect <- pa_vect - 1
          
          # Get Edge Weights
          edge_keys <- attributes(E(net))$vname
          length_sub <- arc_lengths$Length[match(edge_keys, arc_lengths$Arc_key)]
          # Z transform lengths/weights
          length_scaled <- (length_sub - mean(length_sub, na.rn = T)) / sd(length_sub, na.rm = T)
                   
          # Set weights
          E(net)$weight <- length_scaled
          
          
          # Create temporary graph for this day
          temp_graph <- delete.nodes.pa(
            net, 
            node_values,
            na.response = "treat.as.1"
          )

          # Skip if graph is empty
          if (vcount(temp_graph) == 0) {
            return(tibble(
              date = df$date[i],
              icsl = NA_real_,
              #intact_to_sink = NA_real_,
              a_cent_weighted = NA_real_,
              a_cent = NA_real_,
              harary = NA_real_,
              active_upstream_length = 0,
              present_nodes = present_nodes_count,
              total_nodes = total_nodes,
              #percent_upstream_wet = 0,
              percent_network_wet = percent_network_wet
            ))
          }
          
          
          
          # Calculate upstream metrics
          if (sink %in% V(temp_graph)$name) {
            upstream_nodes <- subcomponent(temp_graph, sink, mode = "in")
            upstream_subgraph <- induced_subgraph(temp_graph, upstream_nodes)
            #active_upstream_length <- sum(E(upstream_subgraph)$weight, na.rm = TRUE)
            wet_nodes_upstream <- vcount(upstream_subgraph)
          } else {
            #active_upstream_length <- 0
            wet_nodes_upstream <- 0
          }
          wet_nodes_all <- vcount(temp_graph)
          # Calculate all metrics
          tibble(
            date = df$date[i],
            icsl = ICSL(temp_graph, lengths = length_sub),
            active_upstream_length = size.intact.to.sink(temp_graph, sink = sink),
            a_cent_weighted = mean(alpha_centrality(temp_graph), na.rm = TRUE),
            a_cent = mean(alpha_centrality(temp_graph, weights = NA), na.rm = TRUE),
            harary = harary(temp_graph),
            #active_upstream_length = active_upstream_length,
            wet_upstream_nodes = wet_nodes_upstream,
            wet_network_nodes = wet_nodes_all,
            present_nodes = present_nodes_count,
            total_nodes = total_nodes,
            # percent_upstream_wet = ifelse(present_nodes_count > 0, 
            #                         (wet_nodes_upstream/present_nodes_count)*100, 
            #                         0),
            percent_network_wet = percent_network_wet
          )
          
        }, error = function(e) {
          message("Error for ", wshed, " on ", df$date[i], ": ", e$message)
          tibble(
            date = df$date[i],
            icsl = NA_real_,
            #intact_to_sink = NA_real_,
            a_cent_weighted = NA_real_,
            a_cent = NA_real_,
            harary = NA_real_,
            active_upstream_length = NA_real_,
            present_nodes = NA_real_,
            total_nodes = total_nodes,
            #percent_upstream_wet = NA_real_,
            percent_network_wet = NA_real_
          )
        })
      }, .options = furrr_options(seed = TRUE), .progress = TRUE) |> 
        bind_rows()
      
      # Format final output
      metric_df |>
        pivot_longer(cols = -c(date, wet_upstream_nodes, wet_network_nodes, present_nodes, 
                             total_nodes, percent_network_wet), 
                     names_to = "metric", values_to = "value") |>
        #mutate(metric = recode(metric, !!!metric_names)) |> 
        pivot_wider(names_from = metric, values_from = value) 
    }, .options = furrr_options(seed = TRUE), .progress = TRUE)
  ) |> 
  select(-data) |> 
  unnest(metrics)


streamDAG_Metrics <- streamDAG_Metrics |> 
  group_by(watershed) |>
  mutate(
    a_cent_scaled = as.numeric(scale(a_cent)),
    a_cent_weighted_scaled = as.numeric(scale(a_cent_weighted))
  ) |>
  ungroup() |> 
  select(watershed, date, total_nodes, present_nodes, wet_upstream_nodes, wet_network_nodes, 
         percent_network_wet, active_upstream_length, icsl, harary, a_cent, a_cent_scaled, 
         a_cent_weighted, a_cent_weighted_scaled) |> 
  write_csv(here("Data/STICs/streamDAG_metrics.csv"))
# Reset parallel processing
plan(sequential)
```

```{r}
# Step 1: Summarize all_metrics by watershed and date
alpha_summary <- all_metrics |>
  group_by(watershed, date) |>
  summarise(
    AlphaCentrality_cb = mean(alpha_centrality, na.rm = TRUE),
    WeightedAlphaCentrality_cb = mean(alpha_centrality_weighted, na.rm = TRUE),
    max_upstream = max(upstream_length, na.rm = T),
    upstream_area_sqkm = max(drainage_area_sqkm, na.rm = T),
    icsl_cb = max(icsl, na.rm = T),
    harary_cb = max(harary, na.rm = T),
    .groups = "drop"
  )

# Step 2: Join with streamDAG_Metrics
comparison_df <- streamDAG_Metrics |>
  left_join(alpha_summary, by = c("watershed", "date")) |>
  mutate(
    ac_diff = AlphaCentrality - AlphaCentrality_cb,
    weightedAC_diff = WeightedAlphaCentrality - WeightedAlphaCentrality_cb
  )

comparison_df |> 
ggplot(aes(x = date, y = ac_diff, color = watershed)) +
  geom_line(alpha = 0.8, linewidth = 0.7) +
  scale_color_manual(values = okabeIto) +
  labs(
    x = "",
    y = "",
    color = ""
  ) +
  theme_few() +
  facet_wrap(~watershed)+
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  ) +
  scale_x_date(date_labels = "%b %Y")

comparison_df |> 
ggplot(aes(x = date, y = weightedAC_diff, color = watershed)) +
  geom_line(alpha = 0.8, linewidth = 0.7) +
  scale_color_manual(values = okabeIto) +
  labs(
    x = "",
    y = "",
    color = ""
  ) +
  theme_few() +
  facet_wrap(~watershed)+
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  ) +
  scale_x_date(date_labels = "%b %Y")
```

```{r}
plan(mu)
walk(unique(all_metrics$watershed), \(ws) {

  message(glue::glue("Rendering GIF for {ws}..."))

  crs <- case_when(
  ws %in% c("KNZ", "SHN", "YMR") ~ 26914,
  ws %in% c("PRF", "TAL", "WHR") ~ 26916,
  TRUE ~ 4326  # default to WGS 84
  )

  ws_sites <- allSites |> 
    filter(watershed == ws) |> 
    st_transform(crs)

  ws_streams <- allStreams |> 
    filter(watershed == ws) |> 
    st_transform(crs)

  ws_results <- all_metrics |> 
    filter(watershed == ws)

  # Calculate color scale per watershed
  ac_min <- floor(min(ws_results$alpha_centrality, na.rm = TRUE))
  ac_max <- ceiling(max(ws_results$alpha_centrality, na.rm = TRUE))

  unique_days <- unique(ws_results$date)
  # Create temp dir for individual plots
  tmp_dir <- tempfile(paste0("frames_", ws, "_"))
  dir.create(tmp_dir)
  tmp_dir <- normalizePath(tmp_dir)
  message("temporary files: ", tmp_dir)
  png_files <- file.path(tmp_dir, sprintf("frame_%03d.png", seq_along(unique_days)))

  future_walk2(
    unique_days,
    png_files,
    \(day, png_file) {
      day_results <- ws_results |> 
        filter(date == day) |> 
        st_drop_geometry()

      day_data <- ws_sites |>
        select(siteId) |>
        left_join(day_results, by = "siteId")

      day_label <- unique(na.omit(day_data$date))

      png(png_file, width = 1080, height = 1080)

      p <- ggplot() +
        geom_sf(data = ws_streams, color = "#0084A8", size = 1.5) +
        geom_sf(data = day_data |> filter(!is.na(alpha_centrality)),
                aes(color = alpha_centrality), size = 6) +
        geom_sf(data = day_data |> filter(is.na(alpha_centrality)),
                size = 6, color = "gray") +
        scale_color_viridis_c(option = "magma", limits = c(ac_min, ac_max)) +
        labs(title = ws, subtitle = day_label) +
        guides(color = guide_colorbar(barwidth = 20, barheight = 1.5)) +
        theme_void() +
        theme(
          plot.title = element_text(size = 30, face = "bold", hjust = 0.5),
          plot.subtitle = element_text(size = 25, hjust = 0.5),
          legend.position = "bottom",
          legend.title = element_blank(),
          legend.text = element_text(size = 20),
          axis.text = element_blank()
        )

      print(p)
      dev.off()
    },
    .options = furrr_options(seed = TRUE),
    .progress = TRUE
  )

  # Combine frames into GIF
  output_gif <- here(glue::glue("Figs/STICs/gifs/{ws}_AC.gif"))
  gifski::gifski(png_files, output_gif, width = 1080, height = 1080, delay = 0.05)

  unlink(tmp_dir, recursive = TRUE)
  }
  )
```

```{r}

plan(multisession)  

# Generate individual frames
tmp_dir <- tempfile("combined_frames_")
dir.create(tmp_dir)
unique_days <- unique(results$date)
png_files <- file.path(tmp_dir, sprintf("frame_%03d.png", seq_along(unique_days)))

all_sites <- knz_sites |> 
    st_drop_geometry() |>
    distinct(siteId) |>
    pull(siteId)
 
# Calculate global min and max values (excluding NAs)
ac_min <- floor(min(results$alpha_centrality, na.rm = TRUE))
ac_max <- ceiling(max(results$alpha_centrality, na.rm = TRUE))

 
future_walk2(
  unique_days,
  png_files,
  \(day, png_file) {
    day_results <- results |> filter(date == day) |> 
      st_drop_geometry()

    day_data <- knz_sites |>
      select(siteId) |>
      left_join(day_results, by = "siteId")

    day_label <- unique(na.omit(day_data$date))

    png(png_file, width = 1080, height = 1080)

    p <- ggplot() +
    geom_sf(data = knz_streams, color = "#0084A8", size = 1.5) +
    geom_sf(data = day_data |> filter(!is.na(alpha_centrality)),
            aes(color = alpha_centrality), size = 6) +
    geom_sf(data = day_data |> filter(is.na(alpha_centrality)),
            size = 6, color = "gray") +  # Hollow circles for NA sites
    scale_color_viridis_c(option = "magma", limits = c(ac_min, ac_max)) +  # Set fixed limits
    labs(title = "KNZ",
         subtitle = day_label) +
    guides(color = guide_colorbar(barwidth = 20, barheight = 1.5)) +
    theme_void() +
    theme(
      plot.title = element_text(size = 30, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 25, hjust = 0.5),
      legend.position = "bottom",
      legend.title = element_blank(),
      legend.text = element_text(size = 20),
      axis.text = element_blank()
    )

    print(p)
    dev.off()
  },
  .options = furrr_options(seed = TRUE), .progress = TRUE
)

# Combine frames into GIF
output_gif <- paste0(here("Figs/STICs/gifs/KNZ_AC.gif"))
gifski::gifski(png_files,
       output_gif,
       width = 1080,
       height = 1080,
       delay = 0.05)
unlink(tmp_dir, recursive = TRUE)


tmp_dir <- tempfile("combined_frames_")
dir.create(tmp_dir)
unique_days <- unique(results$date)
png_files <- file.path(tmp_dir, sprintf("frame_%03d.png", seq_along(unique_days)))

all_sites <- knz_sites |> 
    st_drop_geometry() |>
    distinct(siteId) |>
    pull(siteId)
 
# Calculate global min and max values (excluding NAs)
ac_min <- floor(min(results$alpha_centrality_weighted, na.rm = TRUE))
ac_max <- ceiling(max(results$alpha_centrality_weighted, na.rm = TRUE))

 
future_walk2(
  unique_days,
  png_files,
  \(day, png_file) {
    day_results <- results |> filter(date == day) |> 
      st_drop_geometry()

    day_data <- knz_sites |>
      select(siteId) |>
      left_join(day_results, by = "siteId")

    day_label <- unique(na.omit(day_data$date))

    png(png_file, width = 1080, height = 1080)

    p <- ggplot() +
    geom_sf(data = knz_streams, color = "#0084A8", size = 1.5) +
    geom_sf(data = day_data |> filter(!is.na(alpha_centrality_weighted)),
            aes(color = alpha_centrality_weighted), size = 6) +
    geom_sf(data = day_data |> filter(is.na(alpha_centrality_weighted)),
            size = 6, color = "gray") +  # Hollow circles for NA sites
    scale_color_viridis_c(option = "magma", limits = c(ac_min, ac_max)) +  # Set fixed limits
    labs(title = "KNZ",
         subtitle = day_label) +
    guides(color = guide_colorbar(barwidth = 20, barheight = 1.5)) +
    theme_void() +
    theme(
      plot.title = element_text(size = 30, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 25, hjust = 0.5),
      legend.position = "bottom",
      legend.title = element_blank(),
      legend.text = element_text(size = 20),
      axis.text = element_blank()
    )

    print(p)
    dev.off()
  },
  .options = furrr_options(seed = TRUE), .progress = TRUE
)

# Combine frames into GIF
output_gif <- paste0(here("Figs/STICs/gifs/KNZ_ACweighted.gif"))
gifski::gifski(png_files,
       output_gif,
       width = 1080,
       height = 1080,
       delay = 0.05)
unlink(tmp_dir, recursive = TRUE)
plan(sequential)
```

```{r}
plan(multisession)  

# Calc Network Measures
plan(multisession)  
long_data <- knz_presAbs |>
  pivot_longer(cols = -c(watershed, date), 
               names_to = "siteId", 
               values_to = "present") |>
  mutate(DOY = yday(date)) |> 
  group_by(DOY, siteId) |> 
  summarise(present = calc_mode(present)) |> 
  filter(present == 1) |>
  group_by(DOY)

data_list <- group_split(long_data)
key_df <- group_keys(long_data)
key_list <- split(key_df, seq_len(nrow(key_df)))

doy_res <- future_map2_dfr(
  data_list,
  key_list,
  \(df, key) {
    present_nodes <- df$siteId

    pruned_net |>
      activate(nodes) |>
      morph(to_subgraph, siteId %in% present_nodes, subset_by = "nodes") |>
      mutate(
        alpha_centrality = centrality_alpha(),
        alpha_centrality_weighted = centrality_alpha(weights = reach_length_km),
        upstream_length = size.intact.to.node(.G(), node = "all"),
        icsl = ICSL(.G()),
        harary = harary(.G())
      ) |>
      unmorph() |>
      activate(nodes) |>
      st_drop_geometry() |>
      as_tibble() |>
      filter(!is.na(alpha_centrality)) |>
      mutate(DOY = key[[1]])
  },
  .options = furrr_options(seed = TRUE), .progress = TRUE
) |> 
  st_drop_geometry() |> 
  select(siteId, DOY, alpha_centrality_weighted, alpha_centrality, upstream_length, icsl, harary) |> 
  mutate(DateLabel = as.Date(DOY, origin = "1999-12-31"),
            MonthDay = format(DateLabel, "%B %d"))

# Generate individual frames
tmp_dir <- tempfile("combined_frames_")
dir.create(tmp_dir)

      
unique_days <- unique(doy_res$DOY)
png_files <- file.path(tmp_dir, sprintf("frame_%03d.png", seq_along(unique_days)))

all_sites <- knz_sites |> 
    st_drop_geometry() |>
    distinct(siteId) |>
    pull(siteId)
 
# Calculate global min and max values (excluding NAs)
ac_min <- floor(min(doy_res$alpha_centrality, na.rm = TRUE))
ac_max <- ceiling(max(doy_res$alpha_centrality, na.rm = TRUE))

 
future_walk2(
  unique_days,
  png_files,
  \(d, png_file) {
    day_results <- doy_res |> filter(DOY == d) |> 
      st_drop_geometry()
      

    day_data <- knz_sites |>
      select(siteId) |>
      left_join(day_results, by = "siteId")

    day_label <- unique(na.omit(day_data$MonthDay))

    png(png_file, width = 1080, height = 1080)

    p <- ggplot() +
    geom_sf(data = knz_streams, color = "#0084A8", size = 1.5) +
    geom_sf(data = day_data |> filter(!is.na(alpha_centrality)),
            aes(color = alpha_centrality), size = 6) +
    geom_sf(data = day_data |> filter(is.na(alpha_centrality)),
            size = 6, color = "gray") +  # Hollow circles for NA sites
    scale_color_viridis_c(option = "magma", limits = c(ac_min, ac_max)) +  # Set fixed limits
    labs(title = "KNZ",
         subtitle = day_label) +
    guides(color = guide_colorbar(barwidth = 20, barheight = 1.5)) +
    theme_void() +
    theme(
      plot.title = element_text(size = 30, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 25, hjust = 0.5),
      legend.position = "bottom",
      legend.title = element_blank(),
      legend.text = element_text(size = 20),
      axis.text = element_blank()
    )

    print(p)
    dev.off()
  },
  .options = furrr_options(seed = TRUE), .progress = TRUE
)

# Combine frames into GIF
output_gif <- paste0(here("Figs/STICs/gifs/KNZ_AC_DOY.gif"))
gifski::gifski(png_files,
       output_gif,
       width = 1080,
       height = 1080,
       delay = 0.05)
unlink(tmp_dir, recursive = TRUE)



tmp_dir <- tempfile("combined_frames_")
dir.create(tmp_dir)

unique_days <- unique(doy_res$DOY)
png_files <- file.path(tmp_dir, sprintf("frame_%03d.png", seq_along(unique_days)))

# Calculate global min and max values (excluding NAs)
ac_min <- floor(min(doy_res$alpha_centrality_weighted, na.rm = TRUE))
ac_max <- ceiling(max(doy_res$alpha_centrality_weighted, na.rm = TRUE))

plan(multisession)  
future_walk2(
  unique_days,
  png_files,
  \(d, png_file) {
    day_results <- doy_res |> filter(DOY == d) |> 
      st_drop_geometry()
      

    day_data <- knz_sites |>
      select(siteId) |>
      left_join(day_results, by = "siteId")

    day_label <- unique(na.omit(day_data$MonthDay))

    png(png_file, width = 1080, height = 1080)

    p <- ggplot() +
    geom_sf(data = knz_streams, color = "#0084A8", size = 1.5) +
    geom_sf(data = day_data |> filter(!is.na(alpha_centrality_weighted)),
            aes(color = alpha_centrality_weighted), size = 6) +
    geom_sf(data = day_data |> filter(is.na(alpha_centrality_weighted)),
            size = 6, color = "gray") +  # Hollow circles for NA sites
    scale_color_viridis_c(option = "magma", limits = c(ac_min, ac_max)) +  # Set fixed limits
    labs(title = "KNZ",
         subtitle = day_label) +
    guides(color = guide_colorbar(barwidth = 20, barheight = 1.5)) +
    theme_void() +
    theme(
      plot.title = element_text(size = 30, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 25, hjust = 0.5),
      legend.position = "bottom",
      legend.title = element_blank(),
      legend.text = element_text(size = 20),
      axis.text = element_blank()
    )

    print(p)
    dev.off()
  },
  .options = furrr_options(seed = TRUE), .progress = TRUE
)

# Combine frames into GIF
output_gif <- paste0(here("Figs/STICs/gifs/KNZ_ACweighted_DOY.gif"))
gifski::gifski(png_files,
       output_gif,
       width = 1080,
       height = 1080,
       delay = 0.05)
unlink(tmp_dir, recursive = TRUE)

```

```{r}

```

## Plot

```{r}

```

## Exc Prob

```{r}
excProb <- streamDAG_Metrics |>
  group_by(watershed) |> 
  mutate(acUnweighted_expProb = calc_exceedance_prob(AlphaCentrality, rm.zero = F),
         acWeighted_excProb = calc_exceedance_prob(WeightedAlphaCentrality)) |> 
  ungroup() |> 
  mutate(watershed = forcats::fct_relevel(watershed, 
                               "KNZ", "SHN", "YMR",  # Left column
                               "TAL", "WHR", "PRF")) # Right column


excProb |> 
  ggplot(aes(x = acUnweighted_expProb, y = AlphaCentrality, color = watershed)) +
  geom_line(alpha = 0.8, linewidth = 0.7) +
  scale_color_manual(values = okabeIto) +
  labs(
    x = "Exceedance probability",
    y = "Unweighted Alpha Centrality",
    color = ""
  ) +
  theme_few() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )+
  scale_x_continuous(labels = scales::label_percent())

excProb |> 
  ggplot(aes(x = acWeighted_excProb, y = WeightedAlphaCentrality, color = watershed)) +
  geom_line(alpha = 0.8, linewidth = 0.7) +
  scale_color_manual(values = okabeIto) +
  labs(
    x = "Exceedance probability",
    y = "Weighted Alpha Centrality",
    color = ""
  ) +
  theme_few() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )+
  scale_x_continuous(labels = scales::label_percent())

excProb |> 
  ggplot(aes(x = percent_network_wet, y = acUnweighted_expProb, color = watershed)) +
  geom_point(alpha = 0.8, size = 2) +
  scale_color_manual(values = okabeIto) +
  labs(
    x = "% Wet",
    y = "Unweighted Alpha Centrality (Exc Prob)",
    color = ""
  ) +
  theme_few() +
  facet_wrap(~watershed) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

excProb |> 
  ggplot(aes(x = percent_network_wet, y = acWeighted_excProb, color = watershed)) +
  geom_point(alpha = 0.8, size = 2) +
  scale_color_manual(values = okabeIto) +
  labs(
    x = "% Wet",
    y = "Weighted Alpha Centrality (Exc Prob)",
    color = ""
  ) +
  theme_few() +
  facet_wrap(~watershed) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

excProb |> 
  ggplot(aes(x = percent_network_wet, y = WeightedAlphaCentrality, color = watershed)) +
  geom_point(alpha = 0.8, size = 2) +
  scale_color_manual(values = okabeIto) +
  labs(
    x = "% Wet",
    y = "Weighted Alpha Centrality",
    color = ""
  ) +
  theme_few() +
  facet_wrap(~watershed) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )

excProbWY <- streamDAG_Metrics |>
  mutate(waterYear = calcWaterYear(date)) |> 
  group_by(watershed, waterYear) |> 
  mutate(acUnweighted_expProb = calc_exceedance_prob(AlphaCentrality),
         acWeighted_excProb = calc_exceedance_prob(WeightedAlphaCentrality)) |> 
  ungroup() |> 
  mutate(watershed = forcats::fct_relevel(watershed, 
                               "KNZ", "SHN", "YMR",  # Left column
                               "TAL", "WHR", "PRF")) # Right column


excProbWY |> 
  ggplot(aes(x = acUnweighted_expProb, y = AlphaCentrality, color = factor(waterYear))) +
  geom_line(alpha = 0.8, linewidth = 0.7) +
  scale_color_manual(values = okabeIto) +
  labs(
    x = "Exceedance Probability",
    y = "Unweighted Alpha Centrality",
    color = "Water Year"
  ) +
  theme_few() +
  facet_wrap(~watershed) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )+
  scale_x_continuous(labels = scales::label_percent())

excProbWY |> 
  ggplot(aes(x = acWeighted_excProb, y = WeightedAlphaCentrality, color = factor(waterYear))) +
  geom_line(alpha = 0.8, linewidth = 0.7) +
  scale_color_manual(values = okabeIto) +
  labs(
    x = "Exceedance Probability",
    y = "Weighted Alpha Centrality",
    color = "Water Year"
  ) +
  theme_few() +
  facet_wrap(~watershed) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(face = "bold", size = 10),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )+
  scale_x_continuous(labels = scales::label_percent())
```

# Wet/Dry Through Time

```{r}
# #| include: false
# 
# # Process data
# daily_mode <- nestPresAbs_filled |> 
#   unnest(cols = data) |> 
#   mutate(across(
#     .cols = -c(watershed, date),
#     .fns = ~ as.numeric(as.character(.))
#   )) |> 
#   # Convert date and create DOY
#   mutate(date = as_date(date),
#          DOY = yday(date)) |>
#   # Group by DOY to calculate modes
#   group_by(DOY) |>
#   summarise(
#     across(
#       where(is.numeric),
#       ~ calc_mode(., na.rm = TRUE)[1],  # Take first mode if multiple exist
#       .names = "{.col}"
#     ),
#     date = first(date)  # Keep date for ordering
#   ) |>
#   ungroup() |>
#   arrange(date) |>
#   # Convert numeric values to Wet/Dry
#   mutate(across(
#     where(is.numeric) & !matches("DOY|date"),
#     ~ case_when(. == 1 ~ "Wet", 
#                 . == 0 ~ "Dry", 
#                 TRUE ~ NA_character_)
#   )) |>
#   # Pivot to long format
#   select(-date) |>
#   pivot_longer(
#     cols = -DOY,
#     names_to = "siteId",
#     values_to = "wet_dry"
#   )
# 
# # Join with site coordinates
# daily_map <- daily_mode |>
#   left_join(allSites, by = "siteId") |>
#   sf::st_as_sf() |>  # Convert to sf before filtering to keep geometry for all sites
#   mutate(watershed = case_when(
#     watershed == "KNZ" ~ "Kings Creek",
#     watershed == "SHN" ~ "Shane Creek",
#     watershed == "YMR" ~ "Youngmeyer",
#     watershed == "TAL" ~ "Talladega",
#     watershed == "WHR" ~ "Weyerhaeuser",
#     watershed == "PRF" ~ "Paint Rock",
#     TRUE ~ watershed
#   ))
# 
# allStreams <- allStreams |> 
#   mutate(watershed = case_when(
#     watershed == "KNZ" ~ "Kings Creek",
#     watershed == "SHN" ~ "Shane Creek",
#     watershed == "YMR" ~ "Youngmeyer",
#     watershed == "TAL" ~ "Talladega",
#     watershed == "WHR" ~ "Weyerhaeuser",
#     watershed == "PRF" ~ "Paint Rock",
#     TRUE ~ watershed
#   ))
# 
# # Get all unique watersheds
# unique_watersheds <- unique(daily_map$watershed)
# 
# # Create temporary directory for all watershed frames
# tmp_dir <- tempfile("frames_")
# dir.create(tmp_dir)
# 
# # Function to process a single watershed with device management
# process_watershed <- function(watershed_name, daily_map, allStreams, tmp_dir) {
#   # Explicitly open and close devices for each plot
#   on.exit(while (dev.cur() > 1) dev.off(), add = TRUE)
#   
#   # Filter data for current watershed
#   daily_map_ws <- daily_map |> 
#     filter(watershed == watershed_name) |> 
#     mutate(DateLabel = as.Date(DOY, origin = "1999-12-31"),
#            MonthDay = format(DateLabel, "%B %d"))
#   
#   stream_ws <- allStreams |> 
#     filter(watershed == case_when(
#       watershed_name == "KNZ" ~ "Kings Creek",
#     watershed_name == "SHN" ~ "Shane Creek",
#     watershed_name == "YMR" ~ "Youngmeyer",
#     watershed_name == "TAL" ~ "Talladega",
#     watershed_name == "WHR" ~ "Weyerhaeuser",
#     watershed_name == "PRF" ~ "Paint Rock",
#       TRUE ~ watershed_name
#     ))
#   
#   # Create subdirectory for this watershed's frames
#   ws_dir <- file.path(tmp_dir, watershed_name)
#   dir.create(ws_dir, showWarnings = FALSE)
#   
#   # Identify all unique sites for this watershed
#   all_sites <- daily_map_ws |> 
#     st_drop_geometry() |> 
#     distinct(siteId) |> 
#     pull(siteId)
#   
#   # Generate individual frames
#   unique_days <- unique(daily_map_ws$DateLabel)
#   png_files <- file.path(ws_dir, sprintf("frame_%03d.png", seq_along(unique_days)))
#   
#   for (i in seq_along(unique_days)) {
#     day <- unique_days[i]
#     
#     day_data <- daily_map_ws |> 
#       filter(DateLabel == day) |> 
#       complete(siteId = all_sites) |> 
#       sf::st_as_sf()
#     
#     day_label <- unique(na.omit(day_data$MonthDay))
#     
#     # Explicit device management
#     png(png_files[i], width = 1080, height = 1080)
#     
#     p <- ggplot() +
#       geom_sf(data = stream_ws, color = "#0084A8", size = 1.5) +
#       geom_sf(data = day_data |> filter(!is.na(wet_dry)),
#               aes(color = wet_dry), size = 6) +
#       geom_sf(data = day_data |> filter(is.na(wet_dry)),
#               shape = 1, size = 6, color = "gray50") +  # Hollow circles for NA sites
#       scale_color_manual(values = c("Dry" = "#D55E00", "Wet" = "#009E73"),
#                         drop = FALSE) +
#       labs(title = watershed_name,
#            subtitle = day_label) +
#       theme_void() +
#       theme(
#         plot.title = element_text(size = 30, face = "bold", hjust = 0.5),
#         plot.subtitle = element_text(size = 25, hjust = 0.5),
#         legend.position = "bottom",
#         legend.title = element_blank(),
#         legend.text = element_text(size = 20),
#         axis.text = element_blank()
#       )
#     
#     print(p)
#     dev.off()
#   }
#   
#   # Combine frames into GIF
#   output_gif <- paste0(here("Figs/STICs/gifs/"), 
#                       gsub(" ", "_", watershed_name), ".gif")
#   gifski(png_files, 
#          output_gif,
#          width = 1080, 
#          height = 1080, 
#          delay = 0.1)
#   
#   return(output_gif)
# }
# # Set up parallel processing - adjust workers as needed
# plan(multisession, workers = availableCores() - 1)
# # Process watersheds in parallel with progress bar
# message("Processing watersheds in parallel...")
# output_gifs <- suppressWarnings(future_map(unique_watersheds, 
#                          ~process_watershed(.x, daily_map, allStreams, tmp_dir),
#                          .options = furrr_options(seed = TRUE, 
#                                                 packages = c("tidyverse", "sf", "gifski")),
#                          .progress = TRUE))
# 
# # Clean up
# unlink(tmp_dir, recursive = TRUE)
# 
# # Print results
# message("\nSuccessfully created animations:")
# walk(output_gifs, ~message("- ", .x))
```

## All Sheds

```{r}
# #| include: false
# 
# # Set up parallel processing
# plan(multisession, workers = availableCores() - 1)  # Leave one core free
# 
# # First, ensure all watersheds have the same DOY range
# all_days <- daily_map |> 
#   st_drop_geometry() |> 
#   mutate(
#     DateLabel = as.Date(DOY, origin = as.Date("1999-12-31")),  # Adjust origin year as needed
#     MonthDay = format(DateLabel, "%B %d")
#   ) |> 
#   distinct(DOY, DateLabel, MonthDay) |> 
#   arrange(DOY)
# 
# # Create a temporary directory for frames
# tmp_dir <- tempfile("combined_frames_")
# dir.create(tmp_dir)
# 
# # Get all unique sites for each watershed
# watershed_sites <- daily_map |> 
#   st_drop_geometry() |> 
#   distinct(watershed, siteId) |> 
#   group_by(watershed) |> 
#   summarise(sites = list(siteId))
# 
# # Function to create a single combined frame with one shared legend
# create_single_frame <- function(i, all_days, daily_map, allStreams, watershed_sites, tmp_dir) {
#   library(tidyverse)
#   library(sf)
#   library(patchwork)
#   library(grid)
#   
#   current_day <- all_days$DOY[i]
#   day_label <- all_days$MonthDay[i]
#   
#   day_data <- daily_map |> 
#     filter(DOY == current_day)
#   
#   # Create a list of plots for each watershed (WITHOUT LEGENDS)
#   plot_list <- map(watershed_sites$watershed, function(ws) {
#     ws_day_data <- day_data |> 
#       filter(watershed == ws) |> 
#       complete(siteId = watershed_sites$sites[[which(watershed_sites$watershed == ws)]]) |> 
#       sf::st_as_sf()
#     
#     ws_stream <- allStreams |> 
#       filter(watershed == ws)
#     
#     ggplot() +
#       geom_sf(data = ws_stream, color = "#0084A8", size = 0.8) +
#       geom_sf(data = ws_day_data |> filter(!is.na(wet_dry)),
#               aes(color = wet_dry), size = 3, show.legend = TRUE) +  # Ensure color mapping is consistent
#       geom_sf(data = ws_day_data |> filter(is.na(wet_dry)),
#               shape = 1, size = 3, color = "gray50") +
#       scale_color_manual(values = c("Dry" = "#D55E00", "Wet" = "#009E73"),
#                         drop = FALSE, name = "Status") +  # Name for the legend
#       labs(title = ws) +
#       theme_void() +
#       theme(
#         plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
#         legend.position = "none"  # Remove individual legends
#       )
#   })
#   
#   # Combine plots into a 2x3 grid and collect legends
#   combined_plot <- wrap_plots(plot_list, nrow = 2, ncol = 3, guides = "collect") +
#     plot_annotation(
#       subtitle = day_label,
#       theme = theme(
#         plot.subtitle = element_text(size = 20, hjust = 0.5),
#         legend.position = "bottom",  # Single legend at bottom
#         legend.text = element_text(size = 14),
#         legend.title = element_text(size = 14)
#       )
#     )
#   
#   # Save the frame
#   png_file <- file.path(tmp_dir, sprintf("frame_%03d.png", i))
#   ggsave(png_file, combined_plot, width = 18, height = 12, dpi = 310, units = "in")
#   
#   return(png_file)
# }
# 
# # Generate all frames in parallel with progress bar
# message("Creating combined frames in parallel...")
# png_files <- suppressWarnings(future_map(seq_len(nrow(all_days)),
#                        ~create_single_frame(.x, all_days, daily_map, allStreams, watershed_sites, tmp_dir),
#                        .options = furrr_options(
#                          seed = TRUE,
#                          packages = c("tidyverse", "sf", "patchwork", "ggthemes")
#                        ),
#                        .progress = TRUE)
#                        )
# 
# # Combine frames into GIF
# output_gif <- here("Figs/STICs/gifs/allSheds.gif")
# gifski::gifski(unlist(png_files), 
#        output_gif,
#        width = 1800, 
#        height = 1200, 
#        delay = 0.1)
# 
# # Clean up
# unlink(tmp_dir, recursive = TRUE)
# 
# # Reset parallel plan
# plan(sequential)
# 
# message("\nSuccessfully created combined animation:")
# message("- ", output_gif)
```

## Great Plains

```{r}
# #| include: false
# 
# # Set up parallel processing
# plan(multisession, workers = availableCores() - 1)  # Leave one core free
# 
# # Define the watersheds to include
# selected_watersheds <- c("Kings Creek", "Shane Creek", "Youngmeyer")
# 
# # First, ensure all watersheds have the same DOY range
# all_days <- daily_map |> 
#   st_drop_geometry() |> 
#   mutate(
#     DateLabel = as.Date(DOY, origin = as.Date("1999-12-31")),  # Adjust origin year as needed
#     MonthDay = format(DateLabel, "%B %d")
#   ) |> 
#   distinct(DOY, DateLabel, MonthDay) |> 
#   arrange(DOY)
# 
# # Create a temporary directory for frames
# tmp_dir <- tempfile("combined_frames_")
# dir.create(tmp_dir)
# 
# # Get all unique sites for each selected watershed
# watershed_sites <- daily_map |> 
#   st_drop_geometry() |> 
#   filter(watershed %in% selected_watersheds) |>
#   distinct(watershed, siteId) |> 
#   group_by(watershed) |> 
#   summarise(sites = list(siteId), .groups = "drop")
# 
# # Function to create a single combined frame with one shared legend
# create_single_frame <- function(i, all_days, daily_map, allStreams, watershed_sites, tmp_dir) {
#   library(tidyverse)
#   library(sf)
#   library(patchwork)
#   library(grid)
#   
#   current_day <- all_days$DOY[i]
#   day_label <- all_days$MonthDay[i]
#   
#   day_data <- daily_map |> 
#     filter(DOY == current_day, watershed %in% selected_watersheds)
#   
#   # Create a list of plots for each selected watershed
#   plot_list <- map(watershed_sites$watershed, function(ws) {
#     ws_day_data <- day_data |> 
#       filter(watershed == ws) |> 
#       complete(siteId = watershed_sites$sites[[which(watershed_sites$watershed == ws)]]) |> 
#       sf::st_as_sf()
#     
#     ws_stream <- allStreams |> 
#       filter(watershed == ws)
#     
#     ggplot() +
#       geom_sf(data = ws_stream, color = "#0084A8", size = 0.8) +
#       geom_sf(data = ws_day_data |> filter(!is.na(wet_dry)),
#               aes(color = wet_dry), size = 3, show.legend = TRUE) +
#       geom_sf(data = ws_day_data |> filter(is.na(wet_dry)),
#               shape = 1, size = 3, color = "gray50") +
#       scale_color_manual(values = c("Dry" = "#D55E00", "Wet" = "#009E73"),
#                          drop = FALSE, name = "Status") +
#       labs(title = ws) +
#       theme_void() +
#       theme(
#         plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
#         legend.position = "none"
#       )
#   })
#   # Create a named list of plots
#   names(plot_list) <- watershed_sites$watershed
#   p1 <- plot_list[["Kings Creek"]] +
#   theme(plot.margin = ggplot2::margin(1, 1, 1, 1, unit = "pt"))
# p2 <- plot_list[["Shane Creek"]] +
#   theme(plot.margin = ggplot2::margin(1, 1, 1, 1, unit = "pt"))
# p3 <- plot_list[["Youngmeyer"]] +
#   theme(plot.margin = ggplot2::margin(1, 1, 1, 1, unit = "pt"))
#   
#   # Combine plots with reduced spacing
#   combined_plot <- (
#     p1 | (p2 / p3 + plot_layout(guides = "auto", heights = c(1, 1)))
#   ) +
#     plot_layout(
#       guides = "collect",
#       widths = c(1, 1),
#       design = NULL
#     ) +
#     plot_annotation(
#       subtitle = day_label,
#       theme = theme(
#         plot.subtitle = element_text(size = 20, hjust = 0.5),
#         legend.position = "bottom",
#         legend.text = element_text(size = 14),
#         legend.title = element_text(size = 14),
#         plot.margin = ggplot2::margin(2, 2, 2, 2, unit = "pt")  # Global margin
#       )
#     )
#   
#   # Save the frame
#   png_file <- file.path(tmp_dir, sprintf("frame_%03d.png", i))
#   ggsave(png_file, combined_plot, width = 12, height = 12, dpi = 310, units = "in")
#   
#   return(png_file)
# }
# 
# # Generate all frames in parallel with progress bar
# message("Creating combined frames in parallel...")
# png_files <- suppressWarnings(future_map(seq_len(nrow(all_days)),
#                                          ~create_single_frame(.x, all_days, daily_map, allStreams, watershed_sites, tmp_dir),
#                                          .options = furrr_options(
#                                            seed = TRUE,
#                                            packages = c("tidyverse", "sf", "patchwork", "ggthemes")
#                                          ),
#                                          .progress = TRUE)
# )
# 
# # Combine frames into GIF
# output_gif <- here("Figs/STICs/gifs/gpSheds.gif")
# gifski::gifski(unlist(png_files), 
#                output_gif,
#                width = 1200, 
#                height = 1200,
#                delay = 0.1)
# 
# # Clean up
# unlink(tmp_dir, recursive = TRUE)
# 
# # Reset parallel plan
# plan(sequential)
# 
# message("\nSuccessfully created combined animation:")
# message("- ", output_gif)

```

# Check streamDAG coords and lengths

```{r}
data(kon_coords)
kon_coords <- kon_coords |>
  mutate(Object.ID = sub("_[0-9]+$", "", Object.ID),
         watershed = "KNZ") |> 
  dplyr::rename(siteId = Object.ID) |> 
  st_as_sf(coords = c("long", "lat"), crs = 4326, remove = F) |> 
  st_transform(crs = 26914) 

data("kon_lengths")
kon_lengths <- kon_lengths |>
  within(Arcs <- gsub("(_[0-9]+)(?=->|$)", "", Arcs, perl = TRUE)) |> 
  separate(Arcs, into = c("from", "to"), sep = "->") |> 
  rename(streamDAG_lengths = Lengths)


kon_full <- streamDAGs("konza_full") |>
  set_vertex_attr("name", value = sub("_[0-9]+$", "", V(kon_full)$name))

knzCoords <- allSites |> 
  filter(watershed == "KNZ") |> 
  st_transform(crs = 26914)
knzLengths <- all_arc_lengths[["KNZ"]] |> 
  select(-watershed) |> 
  separate(Arc, into = c("from", "to"), sep = "->") |> 
  mutate(
    from = trimws(from),
    to   = trimws(to),
    Length_km = Length/1000
  ) |> 
  select(-Length) |> 
  rename(Con_length_km = Length_km)

compLength <- kon_lengths |> 
  left_join(knzLengths) |> 
  mutate(diff_m = (Con_length_km - streamDAG_lengths)*1000,
         Arc = paste(from, to, sep = "->")) 
ggplot(compLength, aes(x = Arc, y = diff_m)) +
  geom_point(color = "steelblue") +
  geom_hline(yintercept = 0) +
  labs(x = "Segment", y = "Difference (m)") +
  theme_few()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) -> p
p
ggsave(filename = "arcDif_2.png", plot = p,
         path = "~/Desktop/", device = "png", create.dir = TRUE,
         width = 25.4, height = 14.2875, units = "cm", dpi = "retina")

spatial.plot.sf(kon_coords$long, kon_coords$lat, kon_coords$siteId, shapefile = allStreams |> filter(watershed == "KNZ")) +
  labs(title = "streamDAG")
spatial.plot.sf(knzCoords$lon, knzCoords$lat, knzCoords$siteId, shapefile = allStreams |> filter(watershed == "KNZ")) +
  labs(title = "Connor")


kon_coords_length <- get_watershed_lengths("KNZ", kon_coords, allStreams) 

```

```{r}
# Subset data
knz_sites <- allSites |> filter(watershed == "KNZ") |> 
  st_transform(26914)
knz_streams <- allStreams |> filter(watershed == "KNZ") |> 
  st_transform(26914)
knzLengths <- all_arc_lengths[["KNZ"]] |> 
  select(-watershed) |> 
  separate(Arc, into = c("from", "to"), sep = "->") |> 
  mutate(
    from = trimws(from),
    to   = trimws(to),
    Length_km = Length/1000
  ) |> 
  select(-Length) |> 
  rename(Con_length_km = Length_km)
knz_presAbs <- nestPresAbs_filled |> 
  unnest(cols = c(data)) |>
  select(
    c("watershed", "date"),
    any_of(knz_sites$siteId)
  )

# Build network
sites <- tribble(
  ~from,   ~to,
  "04M13", "04M12",
  "04M12", "04M11",
  "04M11", "04M10",
  "04M10", "04M09",
  "04M09", "04M08",
  "04T02", "04M08",
  "04M08", "04M07",
  "04M07", "04M06",
  "04T01", "04M06",
  "04M06", "04M05",
  "04M05", "04M04",
  "04M04", "04M03",
  "04M03", "04M02",
  "04M02", "04M01",
  "04M01", "SFM02",
  "04W04", "04W03",
  "04W03", "04W02",
  "04W02", "04W01",
  "04W01", "04M01",
  "02M11", "02M10",
  "02M10", "02M09",
  "02M09", "02M08",
  "02M08", "02M07",
  "02M07", "02M06",
  "02M06", "02M05",
  "02M05", "02M04",
  "02M04", "02M03",
  "02M03", "02M02",
  "02M02", "02M01",
  "02M01", "SFM01",
  "20M05", "20M04",
  "20M04", "20M03",
  "20M03", "20M02",
  "20M02", "20M01",
  "20M01", "SFM07",
  "01M06", "01M05",
  "01M05", "01M04",
  "01M04", "01M03",
  "01M03", "01M02",
  "01M02", "01M01",
  "01M01", "SFM07",
  "SFM07", "SFM06",
  "SFM06", "SFM05",
  "SFT02", "SFM05",
  "SFM05", "SFM04",
  "SFM04", "SFM03",
  "SFT01", "SFM03",
  "SFM03", "SFM02",
  "SFM02", "SFM01"
)
knz_nodes <- knz_sites |>
  dplyr::mutate(siteId = as.character(siteId)) |>
  dplyr::distinct(siteId, .keep_all = TRUE)

# Build the sfnetwork
knz_net <- sfnetwork(
  nodes = knz_nodes,
  edges = edges,
  node_key = "siteId",
  directed = TRUE
)
plot(knz_net)

knz_net <- as_sfnetwork(knz_streams, directed = TRUE) |> 
  st_network_blend(knz_sites) |> 
  activate("edges") |> 
  mutate(weight = st_length(geom))
plot(knz_net)
# Pull node data to get siteId info
node_data <- knz_net |> 
  activate("nodes") |> 
  as_tibble() |> 
  mutate(node_id = row_number()) |> 
  st_drop_geometry()

# Pull edge data and add from/to node siteIds
edge_data <- knz_net |> 
  activate("edges") |> 
  as_tibble() |>
  mutate(edge_id = row_number()) |>
  left_join(node_data |> select(from = node_id, siteId_from = siteId), by = "from") |>
  left_join(node_data |> select(to = node_id, siteId_to = siteId), by = "to")

# Get edge geometries and compute length
edge_geom <- knz_net |> 
  activate("edges") |> 
  st_geometry()

edge_data <- edge_data |> 
  mutate(length_m = st_length(edge_geom) |> as.numeric(),
         weight = ifelse(!is.na(siteId_from) & !is.na(siteId_to), length_m, NA_real_))

# Add the weights back into the sfnetwork
knz_net <- knz_net |> 
  activate("edges") |> 
  mutate(weight = edge_data$weight) |> 
  activate("nodes") |> 
  mutate(is_terminal = node_is_leaf(),
         has_siteId = !is.na(siteId)) |>
  filter(!(is_terminal & !has_siteId)) |>
  select(-is_terminal, -has_siteId) |> 
  activate("nodes") |> 
  mutate(is_terminal = node_is_leaf(),
         has_siteId = !is.na(siteId)) |>
  filter(!(is_terminal & !has_siteId)) |>
  select(-is_terminal, -has_siteId) |> 
  filter(!(node_is_root() & is.na(siteId)))




tm_shape(knz_net |> activate(edges) |> as_tibble() |> st_as_sf()) +
  tm_lines() +
tm_shape(knz_net |> activate(nodes) |> as_tibble() |> st_as_sf()) +
  tm_dots(size = 0.5, fill = "red") +
    tm_basemap("Esri.WorldImagery")+
  tmap_mode("view")

day_i <- knz_presAbs[1,]
# Get vector of site IDs with a value of 0 on that day
sites_to_remove <- day_i |>
  select(-watershed, -date) |>
  purrr::keep(~ .x == "0") |>
  names()

# Remove those nodes from the network
knz_net_subset <- knz_net |>
  activate("nodes") |>
  filter(!(siteId %in% sites_to_remove)) |> 
  mutate(is_terminal = node_is_leaf(),
         has_siteId = !is.na(siteId),
         is_sink = node_is_sink())

library(tmap)
tm_shape(knz_net_subset |> activate(edges) |> as_tibble() |> st_as_sf()) +
  tm_lines() +
tm_shape(knz_net_subset |> activate(nodes) |> as_tibble() |> st_as_sf()) +
  tm_dots(size = 0.5, fill = "red") +
    tm_basemap("Esri.WorldImagery")+
  tmap_mode("view")



plot(knznet)
# Extract coordinates from node geometries
coords <- knz_net_subset |> 
  sf::st_as_sf() |>
  sf::st_coordinates() |>
  as.data.frame()

# Combine coordinates with node data for manual layout
layout_df <- knz_net_subset |> 
  sf::st_as_sf() |> 
  dplyr::mutate(x = coords$X, y = coords$Y)

# Now plot using ggraph with manual layout
ggraph(knz_net_subset, layout = "manual", x = layout_df$x, y = layout_df$y) +
  geom_edge_link(
    edge_width = 0.8,
    arrow = arrow(length = unit(4, "mm")), 
    end_cap = circle(3, 'mm')
  ) +
  geom_node_point(color = "black") +
  geom_node_text(aes(label = siteId), size = 2.5, repel = TRUE) +
  theme_void()
```
